```{r eval=FALSE}

# conda activate voyager
library(Seurat)
library(tidyverse)
library(cowplot)
library(patchwork)
library(RColorBrewer)
library(EnsDb.Hsapiens.v86)
library(GenomicRanges)
library(ensembldb)
library(viridis)
library(ggpubr)
library(igraph)
library(harmony)
library(WGCNA)
library(dendextend)
library(circlize)
library(RColorBrewer)
library(magrittr)


#devtools::install_github('smorabit/hdWGCNA', ref='dev')
library(hdWGCNA)

colfunc <- colorRampPalette(rev(brewer.pal(11, 'Spectral' )))
theme_set(theme_cowplot())

setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/hdWGCNA")

fig_dir <- "figures/"
data_dir <- "data/"

source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

# re-load hdWGCNA seurat object
seurat_human <- readRDS(file=paste0(data_dir, "ADDS_seurat_processed_annotated_hdWGCNA_all.rds"))
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')

# re-load seurat obj & BayesSpace object:
# seurat_human <- readRDS(paste0(data_dir,'ADDS_seurat_processed.rds'))

# seurat_human <- readRDS("/dfs7/swaruplab/emiyoshi/Visium_ADDS/ADDS_seurat_processed_annotated.rds")

# seurat_human$Diagnosis <- factor(
#   as.character(seurat_human$Diagnosis),
#   levels = c("Control", "earlyAD", "AD", "AD_DS")
# )

# # 5x data:
# seurat_human <- readRDS('/dfs7/swaruplab/emiyoshi/Visium_5X/5XFAD_seurat_processed_annotated.rds')

# seurat_human$Age <- factor(
#   as.character(seurat_human$Age),
#   levels = c("4mo", "6mo", "8mo", "12mo")
# )
œ
human_cp <- c(
      "L1" = "#8B3D5A", "L2-3" = "#E7BDE1", "L3-4" = "#E6A4CD",
      "L3-4-5" = "#CF8BA3", "L5-6" = "#9E6D7F", "L6b" = "#CDAEB9", "WM1" = "#64BCDB", "WM2" = "#62A7D7", "WM3" = "#99C8D7")

mouse_cp <- c(
      "ctx-deep-layers" = "#5b4468", "ctx-upper-layers" = "#9581a4", "ctx-olfactory" = "#8073ab",
      "hippocampus" = "#d6a5b2", "hippocampus-pyramidal" = "#e193aa",
      "lateral-ventricle" = "#d2c3b2",
      "striatum" = "#F0DE8C",
      "thalamus1" = "#00CBA7", "thalamus2" = "#abcc94",
      "hypothalamus-amygdala" = "#72c9b1",
      "WM1" = "#325ea8", "WM2" = "#64bcdb", "WM-cerebral-peduncle" = "#62a7d7",
      "erythrocytes-neurons" = "#d07fc4", "unknown" = "#c47c4d"
    )


# set factor levels for human clusters:
seurat_human$annotation <- factor(
  as.character(seurat_human$annotation),
  levels = names(human_cp)
)

# human ADDS + AD integrated and processed with hdWGCNA
seurat_obj <- readRDS(file=paste0(data_dir, 'ADDS_AD_integrated_hdWGCNA.rds'))


# human ADDS + AD snRNA-seq
# seurat_obj <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_AD_integrated.rds" )

group_levels <- c(
  'EX L2', 'EX L2-3', 'EX L3-5', 'EX L5', 'EX L5-6', 'EX L6',
  'INH VIP+', 'INH', 'INH LAMP5+', 'INH PVALB+', 'INH SST+',
  'ODC1', 'ODC2', 'ODC3',
  'OPC1', 'OPC2', 'OPC3',
  'MG1', 'MG2',
  'ASC1', 'ASC2', 'ASC3', 'ASC4',
  'END Arterial', 'END Capillary',
  'T-Pericyte', 'M-Pericyte', 'SMC',
  'Perivascular Fibroblast', 'Meningeal Fibroblast'
)
color_df$group <- factor(as.character(color_df$group), levels=group_levels)
color_df <- arrange(color_df, group)

seurat_obj$cell_identity <- factor(
  as.character(seurat_obj$cell_identity),
  levels = group_levels
)


# re-load dendrogram from module clustering:
load(paste0(data_dir, 'dendro.rda'))


modules <- GetModules(seurat_human)
#write.csv(modules, file=paste0(data_dir, 'metamodules.csv'), row.names=FALSE)

```

Run hdWGCNA on each cortical layer using pseudo-bulk replicated

```{r eval=FALSE}


# list of groups to run hdWGCNA
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')

# get list of genes to use 
seurat_human <- SetupForWGCNA(
  seurat_human,
  gene_select = "fraction",
  fraction = 0.05,
  group.by = 'annotation',
  wgcna_name = groups[1]
)
length(GetWGCNAGenes(seurat_human))
genes_use <- GetWGCNAGenes(seurat_human)

# compute pseudo-bulk replicates 
datExpr <- to_pseudobulk(
    seurat_human,
    replicate_col = 'Sample',
    cell_type_col = 'annotation',
    label_col = 'Diagnosis'
)
datExpr <- Reduce(cbind, lapply(names(datExpr), function(x){
    cur <- datExpr[[x]]
    colnames(cur) <- paste0(x, '_', colnames(cur))
    cur
}))
datExpr <- t(datExpr[GetWGCNAGenes(seurat_human),])

# compute CPM:
cpm <- t(apply(datExpr, 1, function(x){
    y <- (x) / sum(x) * 1000000
    log2(y + 1)
}))

# run WGCNA separately for each group
for(cur_group in groups){
    print(cur_group)

    # setup WGCNA in seurat object
    wgcna_name <- cur_group
    seurat_human <- SetupForWGCNA(
        seurat_human,
        features = genes_use,
        wgcna_name = cur_group
    )

    # take thee pseudobulk expression for this group 
    cur_cpm <- cpm[grepl(cur_group, rownames(cpm)),]
    seurat_human@misc[[wgcna_name]]$datExpr <- cur_cpm

    seurat_human <- TestSoftPowers(seurat_human)
    plot_list <- PlotSoftPowers(seurat_human)

    # assemble with patchwork
    pdf(paste0(fig_dir, cur_group, '_softpower.pdf'), width=12, height=8)
    print(wrap_plots(plot_list, ncol=2))
    dev.off()

    # construct wgcna network:
    seurat_human <- ConstructNetwork(
        seurat_human,
        #PamStage = T,
            minModuleSize=50,
            #detectCutHeight=0.995,
            mergeCutHeight=0.1,
        tom_name=cur_group,
        overwrite_tom=TRUE
    )

    # plot the dendrogram
    pdf(paste0(fig_dir, cur_group, "_pb_dendro.pdf"),height=3, width=6)
    print(PlotDendrogram(seurat_human, main='hdWGCNA Dendrogram'))
    dev.off()

    seurat_human <- ModuleEigengenes(
        seurat_human,
        group.by.vars="Sample",
    )

    seurat_human <- ModuleConnectivity(
        seurat_human,
        reassign_modules=FALSE
    )

}

saveRDS(seurat_human, file=paste0(data_dir, "ADDS_seurat_processed_annotated_hdWGCNA_all.rds"))


# plot the hub genes 

# individual module networks
# run WGCNA separately for each group
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
for(cur_group in groups){
    print(cur_group)

    seurat_human <- SetActiveWGCNA(seurat_human, cur_group)

  ModuleNetworkPlot(
    seurat_human,
    outdir = paste0(fig_dir, 'module_hubs/')
  )
}



# seurat_human <- SetupForWGCNA(
#   seurat_human,
#   gene_select = "fraction",
#   fraction = 0.05,
#   group.by = 'annotation',
#   wgcna_name = cur_group
# )
# length(GetWGCNAGenes(seurat_human))









# net <- GetNetworkData(seurat_human)

# MEs <- net$multiMEs[[cur_group]]$data

# # Calculate consensus dissimilarity of consensus module eigengenes
# consMEDiss = consensusMEDissimilarity(net$multiMEs);
# # Cluster consensus modules
# consMETree = hclust(as.dist(consMEDiss), method = "average");

# # Plot the result
# pdf(paste0(fig_dir, cur_group, "_pb_ME_dendro.pdf"),height=6, width=6)
# plot(consMETree, main = "Consensus clustering of consensus module eigengenes",
# xlab = "", sub = "")
# abline(h=0.25, col = "red")
# dev.off()










```


Overlap the L3-4-5 and the L5-6 modules

```{r eval=FALSE}


# comparison 3:
group1 <- 'L3-4-5'
group2 <- 'L5-6'
plotname <- paste0(group1, '_vs_', group2)
high_color = 'navy'
w=6


m1 <- GetModules(seurat_human, wgcna_name=group1)
m2 <- GetModules(seurat_human, wgcna_name=group2)

mods1 <- levels(m1$module); mods1 <- mods1[mods1 != 'grey']
mods2 <- levels(m2$module); mods2 <- mods2[mods2 != 'grey']

genome.size <- nrow(seurat_human)


# run overlaps between module gene lists and DEG lists:
overlap_df <- do.call(rbind, lapply(mods1, function(cur_mod1){
  cur_m1_genes <- m1 %>% subset(module == cur_mod1) %>% .$gene_name
  cur_overlap_df <- do.call(rbind, lapply(mods2, function(cur_mod2){
    print(paste0(cur_mod1, ' ', cur_mod2))
    cur_m2_genes <- m2 %>% subset(module == cur_mod2) %>% .$gene_name
    cur_overlap <- testGeneOverlap(newGeneOverlap(
        cur_m1_genes,
        cur_m2_genes,
        genome.size=genome.size
    ))
    c(cur_overlap@odds.ratio, cur_overlap@pval, cur_overlap@Jaccard, length(cur_overlap@intersection))
  })) %>% as.data.frame
  colnames(cur_overlap_df) <- c('odds_ratio', 'pval', 'Jaccard', 'size_intersection')
  cur_overlap_df$m1 <- cur_mod1
  cur_overlap_df$m2 <- mods2

  # module color:
  #cur_overlap_df$color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique
  cur_overlap_df
}))

overlap_df$m1 <- factor(as.character(overlap_df$m1), levels=mods1)
overlap_df$m2 <- factor(as.character(overlap_df$m2), levels=mods2)

# re-order
tmp <- unlist(lapply(mods2, function(cur_mod2){
  cur <- subset(overlap_df, m2==cur_mod2)
  which(cur$odds_ratio == max(cur$odds_ratio))
}))

overlap_df$m2 <- factor(as.character(overlap_df$m2), levels=rev(mods2[order(tmp)]))

# adjust for multiple comparisons:
overlap_df$fdr <- p.adjust(overlap_df$pval, method='fdr')

# significance level:
overlap_df$Significance <- gtools::stars.pval(overlap_df$fdr)
overlap_df$Significance <- ifelse(
  overlap_df$Significance == '.', '',
  overlap_df$Significance
)

# plot the results as a heatmap:
maxval <- 50
plot_df <- overlap_df
plot_df$odds_ratio <- ifelse(plot_df$odds_ratio > maxval, maxval, plot_df$odds_ratio)
plot_df$textcolor <- ifelse(plot_df$odds_ratio > 0.7*maxval, 'white', 'black')


p <- plot_df %>%
  ggplot(aes(x=m1, y=m2, fill=odds_ratio)) +
  geom_tile() +
  geom_text(label=plot_df$Significance, vjust = 0.72, color=plot_df$textcolor) +
  scale_fill_gradient(low='white', high=high_color) +
  RotatedAxis() +
  labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    #axis.title.x = element_blank(),
    #axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
   # axis.ticks.x = element_blank(),
   # axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() + xlab(group1) + ylab(group2)


# Plot the result
pdf(paste0(fig_dir, plotname, "_overlap.pdf"),height=10, width=10)
p
dev.off()

```

Make a sankey plot to show the co-expression remodeling across cortical layers 


```{r eval=FALSE}


test <- lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    cur_mods 

})


all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(module, color)) %>% 
    distinct()

    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') %>% distinct
mod_colors <- all_mods$color; names(mod_colors) <- all_mods$module


names(test) <- groups

df <- test[[1]]
df <- dplyr::rename(df, module1=module, color1=color)
df$module1 <- df$module
df$color1 <- df$color
for(i in 2:length(groups)){
    df[,paste0('module', i)] <- test[[i]]$module
    df[,paste0('color', i)] <- test[[i]]$color
}




library(ggsankey)

df <- df%>% 
  ggsankey::make_long(module1, module2, module3, module4, module5, module6, module7)



p <- ggplot(df, aes(x = x, 
               next_x = next_x, 
               node = node, 
               next_node = next_node,
               fill = factor(node),
               label = node)) +
  geom_sankey() +
  geom_sankey_label(size=2) +
  theme_sankey(base_size = 16) + NoLegend() +
  scale_fill_manual(values=mod_colors) + 
  scale_x_discrete(labels=groups) 

pdf(paste0(fig_dir, 'module_sankey.pdf'), width=12, height=6)
p
dev.off()


```

Compute module eigengenes for all 166 of the modules from the 7 co-ex nets

```{r eval=FALSE}

groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
genes_use <- GetWGCNAGenes(seurat_human)

# run RenameModules
for(group in groups){
    print(group)
    seurat_human <- ResetModuleNames(
        seurat_human,
        new_name = paste0(gsub('-', '', group), '-M'),
        wgcna_name = group
    )
}

# make a table containing the modules for all of the nets
all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(gene_name, module, color)) 
    cur_mods 

}))

# write the all_mods table to file:
write.csv(all_mods, row.names=FALSE, file=paste0(data_dir, 'all_spatial_modules.csv'))


# manually checking for some genes
subset(all_mods, gene_name == 'CSF1R')
subset(all_mods, gene_name == 'CX3CR1')
subset(all_mods, gene_name == 'AQP4')
subset(all_mods, gene_name == 'GAD2')
subset(all_mods, gene_name == 'APP')


# only keep genes that are in the single-cell object
all_mods <- subset(all_mods, gene_name %in% rownames(seurat_obj) & module != 'grey')

# run the module projection function
seurat_obj <- ScaleData(seurat_obj, features=rownames(seurat_obj)[1:100])
seurat_obj <- ProjectModules(
  seurat_obj,
  modules = all_mods,
  group.by.vars = "Batch",
  seurat_ref = NULL,
  wgcna_name = "None",
  wgcna_name_proj = 'spatial'
)


# save the single-cell dataset
saveRDS(seurat_obj, file=paste0(data_dir, 'ADDS_AD_integrated_hdWGCNA.rds'))
seurat_obj <- readRDS(file=paste0(data_dir, 'ADDS_AD_integrated_hdWGCNA.rds'))


##########################################################
# DotPlot
##########################################################

MEs <- GetMEs(seurat_obj)
modules <- GetModules(seurat_obj)
mods <- levels(modules$module)
mods <- mods[mods!='grey']

# add region column:
modules$region <- do.call(rbind, strsplit(as.character(modules$module), '-'))[,1]

meta <- seurat_obj@meta.data
seurat_obj@meta.data <- cbind(meta, MEs)

plot_list <- list()
for(cur_region in unique(modules$region)){
    print(cur_region)
    cur_modules <- subset(modules, region == cur_region)

    mods <- paste0(cur_region, '-M', 1:length(unique(cur_modules$module)))
    cur_modules$module <- factor(as.character(cur_modules$module), levels=mods)

    # make dotplot
    plot_list[[cur_region]] <- DotPlot(
        seurat_obj,
        group.by='cell_type',
        features = mods
        )  + RotatedAxis() +
        scale_color_gradient2(high='red', mid='grey95', low='blue') + xlab('') + ylab('') +
        theme(
            plot.title = element_text(hjust = 0.5),
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(colour = "black", fill=NA, size=1)
        ) + ggtitle(cur_region)

}


pdf(paste0(fig_dir, 'projected_MEs_dotplot_all.pdf'), width=9, height=5)
for(p in plot_list){
    print(p)
}
dev.off()

# remove from seurat obj
#seurat_obj@meta.data <- cbind(seurat_obj@meta.data, MEs)

seurat_obj@meta.data <- meta


```


Try to hierarchically cluster modules similarly to the Hartl et al Nature Neuro 2021 paper


Compute module-module overlaps for all of the PFC modules with GeneOverlap

```{r eval=FALSE}

library(GeneOverlap)

groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
genes_use <- GetWGCNAGenes(seurat_human, 'L1')

all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(gene_name, module, color)) 
    cur_mods 

}))
all_mods <- subset(all_mods, gene_name %in% rownames(seurat_obj) & module != 'grey')
all_mods$region <- do.call(rbind, strsplit(as.character(all_mods$module), '-'))[,1]


# set factor levels:
mod_levels <- c()
for(cur_group in groups){
    cur_group <- gsub('-', '', cur_group)
    cur_modules <- subset(all_mods, region == cur_group)
    cur_mods <- paste0(cur_group, '-M', 1:length(unique(cur_modules$module)))
    mod_levels <- c(mod_levels, cur_mods)
}
all_mods$module <- factor(as.character(all_mods$module), levels=mod_levels)


modules <- all_mods


m1 <- modules
m2 <- modules

mods1 <- levels(m1$module); mods1 <- mods1[mods1 != 'grey']
mods2 <- levels(m2$module); mods2 <- mods2[mods2 != 'grey']

genome.size <- nrow(seurat_obj)


# run overlaps between module gene lists and DEG lists:
overlap_df <- do.call(rbind, lapply(mods1, function(cur_mod1){
  cur_m1_genes <- m1 %>% subset(module == cur_mod1) %>% .$gene_name
  cur_overlap_df <- do.call(rbind, lapply(mods2, function(cur_mod2){
    print(paste0(cur_mod1, ' ', cur_mod2))
    cur_m2_genes <- m2 %>% subset(module == cur_mod2) %>% .$gene_name
    cur_overlap <- testGeneOverlap(newGeneOverlap(
        cur_m1_genes,
        cur_m2_genes,
        genome.size=genome.size
    ))
    c(cur_overlap@odds.ratio, cur_overlap@pval, cur_overlap@Jaccard, length(cur_overlap@intersection))
  })) %>% as.data.frame
  colnames(cur_overlap_df) <- c('odds_ratio', 'pval', 'Jaccard', 'size_intersection')
  cur_overlap_df$m1 <- cur_mod1
  cur_overlap_df$m2 <- mods2

  # module color:
  #cur_overlap_df$color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique
  cur_overlap_df
}))


maxval <- 75
test <- overlap_df %>% dplyr::select(c(m1,m2,odds_ratio)) %>% spread(m1,odds_ratio)
test <- as.matrix(test[,-1])
rownames(test) <- colnames(test)
test[upper.tri(test)] <- 0
test <- as.data.frame(test)
test$m1 <- colnames(test)
plot_df <- test %>% gather(colnames(test)[1:ncol(test)-1], key=m2, value=odds_ratio )
#plot_df$odds_ratio <- ifelse(plot_df$odds_ratio == Inf, 0, plot_df$odds_ratio)
plot_df$odds_ratio <- ifelse(plot_df$odds_ratio > maxval, maxval, plot_df$odds_ratio)

p <- plot_df %>%
  ggplot(aes(x=m1, y=m2, fill=odds_ratio)) +
  geom_tile() +
  #geom_text(label=plot_df$Significance, vjust = 0.72, color=plot_df$textcolor) +
  scale_fill_gradient2(mid = 'grey90', low='grey90', high='darkorchid3', midpoint=5) +
  RotatedAxis() +
  labs(fill = 'OR') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size=6),
    #axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=6),
   # axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() + RotatedAxis()


pdf(paste0(fig_dir, 'mod-mod_overlap.pdf'), width=15, height=15)
p 
dev.off()


maxval <- 0.5
test <- overlap_df %>% dplyr::select(c(m1,m2,Jaccard)) %>% spread(m1,Jaccard)
test <- as.matrix(test[,-1])
rownames(test) <- colnames(test)
test[lower.tri(test)] <- 0
test <- as.data.frame(test)
test$m1 <- colnames(test)
plot_df2 <- test %>% gather(colnames(test)[1:ncol(test)-1], key=m2, value=Jaccard )
#plot_df$odds_ratio <- ifelse(plot_df$odds_ratio == Inf, 0, plot_df$odds_ratio)
plot_df2$Jaccard <- ifelse(plot_df2$Jaccard > maxval, maxval, plot_df2$Jaccard)


p <- plot_df2 %>%
  ggplot(aes(x=m1, y=m2, fill=Jaccard)) +
  geom_tile() +
  #geom_tile(data=plot_df, aes(m1,m2,fill=odds_ratio), inherit.aes=FALSE) +
  #geom_text(label=plot_df$Significance, vjust = 0.72, color=plot_df$textcolor) +
  scale_fill_gradient(low='grey90', high='seagreen', limits=c(0,0.5)) +
  RotatedAxis() +
  labs(fill = 'JC') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size=6),
    #axis.ticks.x = element_blank(),
    axis.text.y = element_text(size=6),
   #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() + RotatedAxis()


pdf(paste0(fig_dir, 'mod-mod_overlap_Jaccard.pdf'), width=15, height=15)
p 
dev.off()

# calculate the median jaccard index:
overlap_df$ct1 <- do.call(rbind, strsplit(overlap_df$m1, '-'))[,1]
overlap_df$ct2 <- do.call(rbind, strsplit(overlap_df$m2, '-'))[,1]
subset(overlap_df, ct1 != ct2) %>% .$Jaccard %>% quantile


########################################################################################
# Correlation of module eigengenes 
########################################################################################

# get the modules
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
MEs <- Reduce(cbind, lapply(groups, function(x){
    z <- GetMEs(seurat_human, x, harmonized=TRUE)
    z[,colnames(z) != 'grey']
}))


# correlate MEs in each of the brain regions
groups <- levels(seurat_human$annotation)
mm_cor <- Reduce(rbind, lapply(groups, function(x){
    cur_spots <- seurat_human@meta.data %>% subset(annotation == x) %>% rownames
    z <- qlcMatrix::corSparse(
        X = as.matrix(MEs[cur_spots,]),
        Y = as.matrix(MEs[cur_spots,])
    ) 
    colnames(z) <- colnames(MEs)
    rownames(z) <- colnames(MEs)
    z <-  reshape2::melt(z)
    z$group <- x
    z$pair <- paste0(z$Var1, '_', z$Var2)
    z
}))
head(mm_cor)

# take the component-wise maximum correlation across each of the regions
test <- mm_cor %>% 
    group_by(pair) %>% 
    mutate(val = max(value)) %>% 
    ungroup %>% 
    select(c(Var1, Var2, val)) %>% 
    distinct %>% 
    spread(Var1, val) 
test <- as.matrix(test[,-1])
rownames(test) <- colnames(test)

# re-order:
test <- test[levels(modules$module), levels(modules$module)]
E <- test

# set up Jaccard matrix:
J <- overlap_df %>% select(c(m1,m2,Jaccard)) %>% spread(m1,Jaccard)
J <- as.matrix(J[,-1])
rownames(J) <- colnames(J)
J <- J[rownames(E), colnames(E)]

# compute the dissimilarity matrix:
D = 1 - ((E + (3*J))/4)
D_st <- D

dist_mat <- dist(D, method = 'euclidean')
hclust_avg <- hclust(dist_mat, method = 'average')

pdf(paste0(fig_dir, 'test_module_hierarchy.pdf'), width=15, height=5)
plot(hclust_avg, cex=0.5, hang=-1)
dev.off()


mod_cut <- cutree(hclust_avg, h=seq(1,2,0.1)) 

# range(metamodules[,4])

# make the metamodules:
meta_assignments <- mod_cut[,6]
#meta_assignments <- mod_cut[,6]
table(meta_assignments)

mod_cut <- cutree(hclust_avg, h=1.45) 
table(mod_cut)

dend <- as.dendrogram(hclust_avg)
library(dendextend)
# labels(dend) <- x[order.dendrogram(dend)]

pdf(paste0(fig_dir, 'test_module_hierarchy.pdf'), width=15, height=5)
for(cut_height in seq(1,2,0.1)){
    dend1 <- color_branches(dend, h=cut_height)
    dend1 <- color_labels(dend1, h=cut_height)
    print(plot(dend1, main = paste0('cut height: ', cut_height), cex=0.3))
}
dev.off()



cut_height=1.4
dend1 <- color_branches(dend, h=cut_height)
dend1 <- color_labels(dend1, h=cut_height)

mod_cut <- cutree(dend1, h=cut_height)
mod_cut_full <- mod_cut 
table(mod_cut)

mod_cut_full[mod_cut_full == 8] <- 7



##########################################
# take this and sub-cluster cuts 1 
##########################################

D_sub <- D[mod_cut_full == 3, mod_cut_full == 3]
dist_mat_sub <- dist(D_sub, method = 'euclidean')
hclust_sub <- hclust(dist_mat_sub, method = 'average')
dend_sub <- as.dendrogram(hclust_sub)

pdf(paste0(fig_dir, 'test_module_hierarchy_1.pdf'), width=10, height=5)
for(cut_height in seq(0.8,1.3,0.05)){
    dend1 <- color_branches(dend_sub, h=cut_height)
    dend1 <- color_labels(dend1, h=cut_height)
    print(plot(dend1, main = paste0('cut height: ', cut_height), cex=0.3))
}
dev.off()

# need to manually merge the singletons to their nearest branch
cut_height=1.15
dend1 <- color_branches(dend_sub, h=cut_height)
dend1 <- color_labels(dend1, h=cut_height)

mod_cut <- cutree(dend1, h=cut_height)
mod_cut1 <- mod_cut
table(mod_cut1)

# 5,8,9,10,11

mod_cut1[names(mod_cut1) == 'L1-M20']
mod_cut1[names(mod_cut1) == 'L6b-M19']

# 5 <- 4
mod_cut1[mod_cut1 == 5]

mod_cut1[mod_cut1 == 8]

# 9 <- 1
mod_cut1[mod_cut1 == 9]


mod_cut1[mod_cut1 == 10]
mod_cut1[mod_cut1 == 11]

# merge 8,10 into a new one
mod_cut1[mod_cut1 %in% c(8,10)] <- 1

mod_cut1[mod_cut1 == 9] <- 1
mod_cut1[mod_cut1 == 5] <- 4
mod_cut1[mod_cut1 == 11] <- 4

# rename
tmp <- names(mod_cut1)
mod_cut1 <- paste0('3.', as.character(mod_cut1))
names(mod_cut1) <- tmp


tmp <- c('L6b-M19', 'L345-M27', 'WM-M25', 'L23-M6')


##########################################
# take this and sub-cluster cuts 2
##########################################

D_sub <- D[mod_cut_full == 1, mod_cut_full == 1]
dist_mat_sub <- dist(D_sub, method = 'euclidean')
hclust_sub <- hclust(dist_mat_sub, method = 'average')
dend_sub <- as.dendrogram(hclust_sub)

pdf(paste0(fig_dir, 'test_module_hierarchy_12.pdf'), width=10, height=5)
for(cut_height in seq(1.1,1.35,0.01)){
    dend1 <- color_branches(dend_sub, h=cut_height)
    dend1 <- color_labels(dend1, h=cut_height)
    print(plot(dend1, main = paste0('cut height: ', cut_height), cex=0.3))
}
dev.off()

cut_height=1.3
dend1 <- color_branches(dend_sub, h=cut_height)
dend1 <- color_labels(dend1, h=cut_height)

mod_cut <- cutree(dend1, h=cut_height)
mod_cut2 <- mod_cut
table(mod_cut2)
mod_cut2

# need to manually merge the singletons to their nearest branch
mod_cut2[mod_cut2 == 5] <- 4

# rename
tmp <- names(mod_cut2)
mod_cut2 <- paste0('1.', as.character(mod_cut2))
names(mod_cut2) <- tmp


##########################################
# combine the three different cuts
##########################################

# rename
tmp <- names(mod_cut_full)
mod_cut_full <- paste0('0.', as.character(mod_cut_full))
names(mod_cut_full) <- tmp

mod_cut_full[names(mod_cut1)] <- mod_cut1
mod_cut_full[names(mod_cut2)] <- mod_cut2


##########################################
# Plot a circular dendrogram
##########################################

library(circlize)
library(dendextend)

dend <- as.dendrogram(hclust_avg)
# labels(dend) <- x[order.dendrogram(dend)]

dend_colors <- WGCNA::labels2colors(mod_cut_full)
names(dend_colors) <- names(mod_cut_full)
dend_colors <- dend_colors[labels(dend)]

dend <- dend %>% 
    set('leaves_pch', 19) %>%
    set('leaves_col', dend_colors)



pdf(paste0(fig_dir, 'test_module_hierarchy_circle2.pdf'), width=10, height=10)
circlize_dendrogram(
    dend,
    labels_track_height=NA,
    dend_track_height=0.4
)
dev.off()

'L6b-M19' %in% names(mod_cut1)

# save all the dendrogram stuff:
save(dend, hclust_avg, mod_cut_full, D, J, E, file=paste0(data_dir, 'dendro.rda'))
load(paste0(data_dir, 'dendro.rda'))

##########################################
# Make a new module table
##########################################

new_module_df <- data.frame()

for(mod in unique(mod_cut_full)){
    
    cur_mods <- names(mod_cut_full[mod_cut_full == mod])
    cur <- subset(all_mods, module %in% cur_mods) %>% 
        mutate(module = droplevels(module))
    cur_genes <- unique(cur$gene_name)

    cur_new_df <- data.frame(
        gene_name = cur_genes,
        module = mod
    )
    new_module_df <- rbind(new_module_df, cur_new_df)

}
new_module_df$color <- WGCNA::labels2colors(new_module_df$module)

# get kME values for each gene 
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
all_kmes <- Reduce(rbind, lapply(groups, function(x){
    GetHubGenes(seurat_human, Inf, wgcna_name=x)
}))

ix <- match(all_kmes$module, names(mod_cut_full))
all_kmes$new_mod <- as.character(mod_cut_full)[ix]

all_mods_reassigned <- all_kmes %>% group_by(gene_name) %>% slice_max(kME, n=1)

all_mods_reassigned <- all_mods_reassigned %>% 
    dplyr::select(c(gene_name, new_mod)) %>%
    dplyr::rename(module = new_mod) %>% 
    mutate(color = WGCNA::labels2colors(module)) %>%
    mutate(module = as.factor(module))

table(all_mods_reassigned$module)

seurat_human <- SetActiveWGCNA(seurat_human, 'consensus')
seurat_human <- SetModules(seurat_human, all_mods_reassigned)

###################################################################################
# Compute MEs and Connectivity for these modules
##################################################################################

seurat_human <- ModuleEigengenes(
    seurat_human,
    group.by.vars="Sample",
)

tmp <- as.data.frame(all_mods_reassigned)
rownames(tmp) <- tmp$gene_name
seurat_human <- SetModules(seurat_human, tmp)

seurat_human <- ModuleConnectivity(
    seurat_human,
    reassign_modules=FALSE
)

# run RenameModules
seurat_human <- ResetModuleNames(
  seurat_human,
  new_name = "M"
)

# module colors:
seurat_human <- SetActiveWGCNA(seurat_human, 'consensus')
modules <- GetModules(seurat_human)
modules$color <- WGCNA::labels2colors(as.character(modules$module))
seurat_human <- SetModules(seurat_human, modules)

library(MetBrewer)

modules <- GetModules(seurat_human)
mods <- levels(modules$module)
mod_colors <- dplyr::select(modules, c(module, color)) %>%
  distinct %>% arrange(module) %>% .$color
n_colors <- length(mod_colors) 

new_colors <- paste0(met.brewer("Signac", n=n_colors, type='continuous'))
new_colors <- sample(new_colors, n_colors)

seurat_human <- ResetModuleColors(seurat_human, new_colors)


unique(all_mods_reassigned$module)
table(all_mods_reassigned$module)
levels(all_mods_reassigned$module)

# re-color the dendrogram:
modules <- GetModules(seurat_human)
mods <- levels(modules$module)
mod_colors <- dplyr::select(modules, c(module, color)) %>%
  distinct %>% arrange(module) %>% .$color
names(mod_colors) <- levels(all_mods_reassigned$module)

# dend_colors <- WGCNA::labels2colors(mod_cut_full)
dend_colors <- mod_colors[mod_cut_full]
names(dend_colors) <- names(mod_cut_full)
dend_colors <- dend_colors[labels(dend)]

dend <- dend %>% 
   # set('leaves_pch', 19) %>%
    set('leaves_col', dend_colors)


saveRDS(seurat_human, file=paste0(data_dir, "ADDS_seurat_processed_annotated_hdWGCNA_all.rds"))


###################################################################################
# Run enrichr with these modules
##################################################################################


library(enrichR)

dbs <-c('GO_Biological_Process_2021','GO_Cellular_Component_2021','GO_Molecular_Function_2021','WikiPathway_2021_Human', 'KEGG_2021_Human')

# compute GO terms:
seurat_human <- RunEnrichr(seurat_human, dbs=dbs, max_genes=Inf)

# inspect the enrichr table:
enrichr_df <- GetEnrichrTable(seurat_human) %>% subset(P.value < 0.05)
enrichr_df$ngenes <- unlist(lapply(strsplit(enrichr_df$Genes, ';'), function(x){length(x)}))
enrichr_df <- subset(enrichr_df, ngenes >= 3)

write.table(enrichr_df, quote=FALSE, row.names=FALSE, sep='\t', file=paste0(data_dir, 'merged_human_vis_enrichr.tsv'))



################################################################################
# Plot selected GO terms:
#################################################################################


modules <- GetModules(seurat_human)
color_df <- modules %>% subset(module!='grey') %>%
  dplyr::select(c(module, color)) %>% distinct %>%
  mutate(module=droplevels(module)) %>%
  dplyr::rename(c(group=module, colour=color))
mods <- levels(modules$module); mods <- mods[mods != 'grey']

color_df$group <- factor(as.character(color_df$group), levels=mods)

# helper function to wrap text
wrapText <- function(x, len) {
    sapply(x, function(y) paste(strwrap(y, len), collapse = "\n"), USE.NAMES = FALSE)
}

combined_output <- GetEnrichrTable(seurat_human)
selected_terms <- read.delim('data/human_metamodule_selected_GO.txt', sep='\t', header=1)

# subset selected terms
combined_output$ngenes <- unlist(lapply(strsplit(combined_output$Genes, ';'), function(x){length(x)}))
selected_terms <- subset(combined_output, Term %in% selected_terms$Term & P.value < 0.05 & ngenes > 2)


selected_terms$group <- factor(
  as.character(selected_terms$module),
  levels = mods
)

# set max pval

quantile(-log(selected_terms$P.value), 0.95)
max_p <- 10

selected_terms$logp <- -log(selected_terms$P.value)
selected_terms$logp <- ifelse(selected_terms$logp > max_p, max_p, selected_terms$logp)

# remove GO Term ID
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

selected_terms <- selected_terms %>%
  arrange(group)


selected_terms$wrap <- wrapText(selected_terms$Term, 35)

selected_terms$Term <- factor(
  as.character(selected_terms$Term),
  levels = rev(unique(as.character(selected_terms$Term)))
)

# GO Term dot plot

p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color =logp, size=log(Combined.Score))) +
  geom_point() +
  scale_color_stepsn(colors=rev(magma(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(size=0.25, color='lightgrey')
  )+ labs(
    color = bquote("-log"[10]~"(P)"),
    size= bquote("log"[10]~"(Enrich)")
  )


p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color =logp, size=log(Combined.Score))) +
  geom_point() +
  scale_color_stepsn(colors=rev(magma(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  coord_equal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(size=0.25, color='lightgrey')
  )+ labs(
    color = bquote("-log"[10]~"(P)"),
    size= bquote("log"[10]~"(Enrich)")
  )



# # make the colorbar as its own heatmap
color_df$var <- 1
cp <- color_df$colour; names(cp) <- color_df$group
colorbar <- color_df %>%
  subset(group %in% unique(selected_terms$group)) %>%
  ggplot(aes(x=group, y=var, fill=group)) +
  geom_tile() +
  scale_fill_manual(values=cp) +
  coord_equal() +
  NoLegend() + RotatedAxis() +
  theme(
    plot.title=element_blank(),
    axis.line=element_blank(),
    axis.ticks.y =element_blank(),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.margin=margin(0,0,0,0),
  )



pdf(paste0(fig_dir, 'human_vis_selected_GO_terms_metamod_new.pdf'), width=7, height=7)
p / colorbar
dev.off()


pdf(paste0(fig_dir, 'human_vis_selected_GO_terms_metamod.pdf'), width=8, height=7)
p / colorbar
dev.off()


```

test module umap? 

```{r eval=FALSE}


RunModuleUMAP <- function(
  seurat_obj,
  n_hubs = 10,
  exclude_grey = TRUE,
  genes_use = NULL,
  wgcna_name = NULL,
  n_neighbors= 25,
  metric = "cosine",
  spread=1,
  min_dist=0.4,
  supervised = FALSE,
  ...
){

  if(is.null(wgcna_name)){wgcna_name <- seurat_obj@misc$active_wgcna}

  # get modules,
  modules <- GetModules(seurat_obj, wgcna_name)
  mods <- levels(modules$module)

  # check if we have eigengene-based connectivities:
  if(!all(paste0('kME_', as.character(mods)) %in% colnames(modules))){
    stop('Eigengene-based connectivity (kME) not found. Did you run ModuleEigengenes and ModuleConnectivity?')
  }

  # handle excluding grey module
  if(exclude_grey){
    mods <- mods[mods != 'grey']
    modules <- subset(modules, module != 'grey')
  }

  # get hub genes:
  hub_list <- lapply(mods, function(cur_mod){
    cur <- subset(modules, module == cur_mod)
    cur[,c('gene_name', paste0('kME_', cur_mod))] %>%
      top_n(n_hubs) %>% .$gene_name
  })
  names(hub_list) <- mods

  # get all genes that aren't in gray mod
  selected_genes <- modules[modules$module %in% mods,'gene_name']

  if(!is.null(genes_use)){
    selected_genes <- selected_genes[selected_genes %in% genes_use]
  } 

  # get the TOM
  TOM <- GetTOM(seurat_obj, wgcna_name)

  # subset the TOM for umap
  # keep all genes as rows, and keep only hubs as cols (features)
  feature_mat <- TOM[selected_genes,unlist(hub_list)]

  # run UMAP
  if(supervised){
    print('running supervised UMAP:')
    hub_umap <-  uwot::umap(
      X = feature_mat,
      min_dist = min_dist,
      n_neighbors= n_neighbors,
      metric = metric,
      spread=spread,
      y = modules$module, # for supervised UMAP
      ...
    )
  } else {
    hub_umap <-  uwot::umap(
      X = feature_mat,
      min_dist = min_dist,
      n_neighbors= n_neighbors,
      metric = metric,
      spread=spread,
      ...
    )
  }


  # set up plotting df
  plot_df <- as.data.frame(hub_umap)
  colnames(plot_df) <- c("UMAP1", "UMAP2")
  plot_df$gene <- rownames(feature_mat)

  # add module color, and hub gene status to the plotting df:
  ix <- match(plot_df$gene, modules$gene_name)
  plot_df$module <- modules$module[ix]
  plot_df$color <- modules$color[ix]
  plot_df$hub <- ifelse(
    plot_df$gene %in% as.character(unlist(hub_list)), 'hub', 'other'
  )

  # get kME values for each gene
  kMEs <- do.call(rbind, lapply(mods, function(cur_mod){
    cur <- subset(modules, module == cur_mod)
    cur <- cur[,c('gene_name', paste0('kME_', cur_mod))]
    colnames(cur) <- c('gene_name', 'kME')

    # scale kMEs between 0 & 1:
    cur$kME <- scale01(cur$kME)
    cur
  }))
  ix <- kMEs$gene_name[match(plot_df$gene, kMEs$gene_name)]
  plot_df$kME <- kMEs[ix, 'kME']

  # add the UMAP to the Seurat object:
  seurat_obj <- SetModuleUMAP(seurat_obj, plot_df, wgcna_name)

  seurat_obj
}




seurat_human <- RunModuleUMAP(
  seurat_human,
  n_hubs = 5,
  genes_use = VariableFeatures(seurat_human),
  n_neighbors=10,
  min_dist=0.1,
  spread=1,
    target_weight=0.5,
  supervised=TRUE
)

# get the hub gene UMAP table from the seurat object
umap_df <- GetModuleUMAP(
  seurat_human
)

# plot with ggplot
p <- ggplot(umap_df, aes(x=UMAP1, y=UMAP2)) +
  geom_point(
   color=umap_df$color,
   size=umap_df$kME*2
  ) 

pdf(paste0(fig_dir, 'hubgene_umap_ggplot_metamodule_selected.pdf'), width=5, height=5)
p
dev.off()

# individual module networks
ModuleNetworkPlot(
  seurat_human,
  outdir = paste0(fig_dir, 'metamod_hubNetworks/')
)


library(reshape2)
library(igraph)
pdf(paste0(fig_dir, 'hubgene_umap_igraph.pdf'), width=9, height=9)
ModuleUMAPPlot(
  seurat_human,
  edge.alpha=0.5,
  sample_edges=TRUE,
  vertex.label.cex = 0.1, #font size
  keep_grey_edges=FALSE,
  edge_prop=0.1, # taking the top 20% strongest edges in each module
  label_hubs=5 # how many hub genes to plot per module?
)
dev.off()

pdf(paste0(fig_dir, 'hubgene_umap_igraph_unlabeled.pdf'), width=7, height=7)
ModuleUMAPPlot(
  seurat_human,
  edge.alpha=0.5,
  sample_edges=TRUE,
  vertex.label.cex = 0.1, #font size
  keep_grey_edges=FALSE,
  edge_prop=0.15, # taking the top 20% strongest edges in each module
  label_hubs=0 # how many hub genes to plot per module?
)
dev.off()



```


Project these modules into the single-cell dataset 

```{r eval=FALSE}

modules <- GetModules(seurat_human, 'consensus')

seurat_obj <- ScaleData(seurat_obj, features=rownames(seurat_obj)[1:100])
seurat_obj <- ProjectModules(
  seurat_obj,
  modules = modules,
  group.by.vars = "Batch",
  seurat_ref = NULL,
  wgcna_name = "None",
  wgcna_name_proj = 'consensus'
)

saveRDS(seurat_obj, file=paste0(data_dir, 'ADDS_AD_integrated_hdWGCNA.rds'))
seurat_obj <- readRDS(file=paste0(data_dir, 'ADDS_AD_integrated_hdWGCNA.rds'))

MEs <- GetMEs(seurat_obj)
meta <- seurat_obj@meta.data
seurat_obj@meta.data <- cbind(meta, MEs)
mods <- colnames(MEs)

# make dotplot
p <- DotPlot(
    seurat_obj,
    group.by='cell_type',
    features = mods
    #scale=FALSE
    )  + RotatedAxis() +
    scale_color_gradient2(high='red', mid='grey95', low='blue') + xlab('') + ylab('') +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1)
    ) + ggtitle('meta-modules')

pdf(paste0(fig_dir, 'projected_MEs_dotplot_metamodules.pdf'), width=7, height=4)
p
dev.off()

seurat_obj@meta.data <- meta


################################
# DotPlot for the visium data 
###############################


MEs <- GetMEs(seurat_human)
meta <- seurat_human@meta.data
seurat_human@meta.data <- cbind(meta, MEs)
mods <- colnames(MEs)

# make dotplot
p <- DotPlot(
    seurat_human,
    group.by='annotation',
    features = mods,
    #cluster.idents=TRUE,
   # scale=FALSE
    )  + RotatedAxis() +
    scale_color_gradient2(high='red', mid='grey95', low='blue') + xlab('') + ylab('') +
    theme(
        plot.title = element_text(hjust = 0.5),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1)
    ) + ggtitle('meta-modules')

pdf(paste0(fig_dir, 'MEs_dotplot_metamodules.pdf'), width=7, height=4)
p
dev.off()

seurat_human@meta.data <- meta

```

DME analysis 

Compare condition vs control for each of the hdWGCNA modules in each of the annotations 

```{r eval=FALSE}

seurat_human$anno_DMEs <- ifelse(
    grepl('WM', as.character(seurat_human$annotation)), 
    'WM', as.character(seurat_human$annotation)
)

groups <- unique(seurat_human$anno_DMEs)
wgcna_groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM', 'consensus')
dx_list <- c('earlyAD', 'AD', 'AD_DS')
combos <- expand.grid(groups, wgcna_groups, dx_list)

DMEs <- Reduce(rbind, lapply(1:nrow(combos), function(i){
    print(i)
    cur_anno <- as.character(combos[i, 'Var1'])
    cur_wgcna <- as.character(combos[i,'Var2'])
    cur_dx <- as.character(combos[i, 'Var3'])

    # get list of cells with these annotations / diagnosis
    g1 <- seurat_human@meta.data %>% subset(anno_DMEs == cur_anno & Diagnosis == cur_dx) %>% rownames
    g2 <- seurat_human@meta.data %>% subset(anno_DMEs == cur_anno & Diagnosis == 'Control') %>% rownames

    # reset active WGCNA
    seurat_human <- SetActiveWGCNA(seurat_human, cur_wgcna)

    # run the DME comparison with a wilcox test
    DMEs <- FindDMEs(
        seurat_human,
        barcodes1 = g1,
        barcodes2 = g2,
        test.use='wilcox'
    )
    DMEs$ident.1 <- cur_dx
    DMEs$ident.2 <- 'Control'
    DMEs$annotation <- cur_anno
    DMEs$wgcna <- cur_wgcna
    DMEs

}))

# save the DME results as a table:
write.csv(DMEs, file=paste0(data_dir, 'differential_MEs_humanST.csv'), quote=FALSE, row.names=FALSE)
DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))

#DMEs %>% write.csv(file=paste0(data_dir, 'differential_MEs_humanST_match.csv'), quote=FALSE, row.names=FALSE)


# set up DME data for circular dendro plotting:
tmp <- subset(DMEs, annotation == wgcna) %>% 
    dplyr::select(c(module, ident.1, avg_log2FC)) %>%
    pivot_wider(names_from = ident.1, values_from = avg_log2FC) %>% 
    as.data.frame()

rownames(tmp) <- tmp$module
tmp <- tmp[,2:ncol(tmp)]


###################################################################
# Correlate DMEs across conditions
###################################################################

# compute the number of genes in each module:
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]

    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') 
mod_size <- table(all_mods$module)

# get a map of module and color
all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    #cur_mods <- cur_mods[genes_use,]
    #cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(module, color)) %>% 
    distinct()

    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') %>% distinct
mod_colors <- all_mods$color; names(mod_colors) <- all_mods$module

# load DME table
DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))

# AD info
plot_df1 <- subset(DMEs, ident.1 %in% c('AD')) %>% 
    dplyr::select(c(p_val_adj, avg_log2FC, module, annotation, wgcna)) %>% 
    dplyr::arrange(module, annotation, wgcna)

# ADDS info
plot_df2 <- subset(DMEs, ident.1 %in% c('AD_DS')) %>% 
    dplyr::arrange(module, annotation, wgcna) %>%
    dplyr::select(c(p_val_adj, avg_log2FC))
colnames(plot_df2) <- paste0(colnames(plot_df2), '2')

# combine
plot_df <- cbind(plot_df1, plot_df2) %>% 
    subset(annotation == wgcna)

# label the top 10 and bottom 10
nlabel <- 5
top_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_max(x, n=nlabel) %>% .$module %>% as.character
bottom_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_min(x, n=nlabel) %>% .$module %>% as.character
label_mods <- c(top_mods, bottom_mods)
plot_df$anno <- ifelse(plot_df$module %in% label_mods, plot_df$module, NA)

# set module factor level:
plot_df$module <- factor(as.character(plot_df$module), levels=names(mod_colors))

# add module sizes:
plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

plot_range <- max(max(abs(plot_df$avg_log2FC)), max(abs(plot_df$avg_log2FC2)))
p <- plot_df %>% 
    ggplot(aes(x = avg_log2FC, y = avg_log2FC2, color=module, size=log2(mod_size))) +
    geom_hline(yintercept = 0, linetype='dashed', color='grey30') +
    geom_vline(xintercept = 0, linetype='dashed', color='grey30',) +
    ggrastr::rasterise(geom_point(alpha=0.7), dpi=500) +
    geom_point(
       inherit.aes=FALSE,
       data=subset(plot_df, !is.na(anno)),
       aes(avg_log2FC, avg_log2FC2, fill=module, size=log2(mod_size)),
       shape=21, color='black'
     ) +
    geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC, y = avg_log2FC2), method='lm', color='black') +
    stat_cor(aes(label='..r.label..'), method='pearson') +
    geom_text_repel(aes(label=anno), color='black',  min.segment.length=0, max.overlaps=Inf, size=5) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    xlim(c(-plot_range, plot_range)) +
    ylim(c(-plot_range, plot_range)) +
    coord_equal() +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5),
        panel.grid = element_line(size=0.25, color='white'),
        panel.background = element_rect(fill='grey92'),
    ) +
    coord_fixed(ratio=1) + NoLegend() +
    xlab(bquote("LAD Avg. log"[2]~"(FC)")) +
    ylab(bquote("ADDS Avg. log"[2]~"(FC)")) 

pdf(paste0(fig_dir, 'DME_cor_compare_AD_vs_ADDS.pdf'), width=3, height=3)
p
dev.off()

##########################################################################
# LAD vs EAD
##########################################################################

# AD info
plot_df1 <- subset(DMEs, ident.1 %in% c('AD')) %>% 
    dplyr::select(c(p_val_adj, avg_log2FC, module, annotation, wgcna)) %>% 
    dplyr::arrange(module, annotation, wgcna)

# ADDS info
plot_df2 <- subset(DMEs, ident.1 %in% c('earlyAD')) %>% 
    dplyr::arrange(module, annotation, wgcna) %>%
    dplyr::select(c(p_val_adj, avg_log2FC))
colnames(plot_df2) <- paste0(colnames(plot_df2), '2')

# combine
plot_df <- cbind(plot_df1, plot_df2) %>% 
    subset(annotation == wgcna)

# label the top 10 and bottom 10
nlabel <- 5
top_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_max(x, n=nlabel) %>% .$module %>% as.character
bottom_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_min(x, n=nlabel) %>% .$module %>% as.character
label_mods <- c(top_mods, bottom_mods)
plot_df$anno <- ifelse(plot_df$module %in% label_mods, plot_df$module, NA)

# set module factor level:
plot_df$module <- factor(as.character(plot_df$module), levels=names(mod_colors))

# add module sizes:
plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

plot_range <- max(max(abs(plot_df$avg_log2FC)), max(abs(plot_df$avg_log2FC2)))
p <- plot_df %>% 
    ggplot(aes(x = avg_log2FC, y = avg_log2FC2, color=module, size=log2(mod_size))) +
    geom_hline(yintercept = 0, linetype='dashed', color='grey30') +
    geom_vline(xintercept = 0, linetype='dashed', color='grey30',) +
    ggrastr::rasterise(geom_point(alpha=0.7), dpi=500) +
    geom_point(
       inherit.aes=FALSE,
       data=subset(plot_df, !is.na(anno)),
       aes(avg_log2FC, avg_log2FC2, fill=module, size=log2(mod_size)),
       shape=21, color='black'
     ) +
    geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC, y = avg_log2FC2), method='lm', color='black') +
    stat_cor(aes(label='..r.label..'), method='pearson') +
    geom_text_repel(aes(label=anno), color='black',  min.segment.length=0, max.overlaps=Inf, size=5) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    xlim(c(-plot_range, plot_range)) +
    ylim(c(-plot_range, plot_range)) +
    coord_equal() +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5),
        panel.grid = element_line(size=0.25, color='white'),
        panel.background = element_rect(fill='grey92'),
    ) +
    coord_fixed(ratio=1) + NoLegend() +
    xlab(bquote("LAD Avg. log"[2]~"(FC)")) +
    ylab(bquote("EAD Avg. log"[2]~"(FC)")) 

pdf(paste0(fig_dir, 'DME_cor_compare_AD_vs_EAD.pdf'), width=3, height=3)
p
dev.off()


##########################################################################
# ADDS vs EAD
##########################################################################

# AD info
plot_df1 <- subset(DMEs, ident.1 %in% c('earlyAD')) %>% 
    dplyr::select(c(p_val_adj, avg_log2FC, module, annotation, wgcna)) %>% 
    dplyr::arrange(module, annotation, wgcna)

# ADDS info
plot_df2 <- subset(DMEs, ident.1 %in% c('AD_DS')) %>% 
    dplyr::arrange(module, annotation, wgcna) %>%
    dplyr::select(c(p_val_adj, avg_log2FC))
colnames(plot_df2) <- paste0(colnames(plot_df2), '2')

# combine
plot_df <- cbind(plot_df1, plot_df2) %>% 
    subset(annotation == wgcna)

# label the top 10 and bottom 10
nlabel <- 5
top_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_max(x, n=nlabel) %>% .$module %>% as.character
bottom_mods <- plot_df %>% mutate(x  = avg_log2FC + avg_log2FC2) %>%
    slice_min(x, n=nlabel) %>% .$module %>% as.character
label_mods <- c(top_mods, bottom_mods)
plot_df$anno <- ifelse(plot_df$module %in% label_mods, plot_df$module, NA)

# set module factor level:
plot_df$module <- factor(as.character(plot_df$module), levels=names(mod_colors))

# add module sizes:
plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

plot_range <- max(max(abs(plot_df$avg_log2FC)), max(abs(plot_df$avg_log2FC2)))
p <- plot_df %>% 
    ggplot(aes(x = avg_log2FC, y = avg_log2FC2, color=module, size=log2(mod_size))) +
    geom_hline(yintercept = 0, linetype='dashed', color='grey30') +
    geom_vline(xintercept = 0, linetype='dashed', color='grey30',) +
    ggrastr::rasterise(geom_point(alpha=0.7), dpi=500) +
    geom_point(
       inherit.aes=FALSE,
       data=subset(plot_df, !is.na(anno)),
       aes(avg_log2FC, avg_log2FC2, fill=module, size=log2(mod_size)),
       shape=21, color='black'
     ) +
    geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC, y = avg_log2FC2), method='lm', color='black') +
    stat_cor(aes(label='..r.label..'), method='pearson') +
    geom_text_repel(aes(label=anno), color='black',  min.segment.length=0, max.overlaps=Inf, size=5) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    xlim(c(-plot_range, plot_range)) +
    ylim(c(-plot_range, plot_range)) +
    coord_equal() +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5),
        panel.grid = element_line(size=0.25, color='white'),
        panel.background = element_rect(fill='grey92'),
    ) +
    coord_fixed(ratio=1) + NoLegend() +
    xlab(bquote("EAD Avg. log"[2]~"(FC)")) +
    ylab(bquote("ADDS Avg. log"[2]~"(FC)")) 

pdf(paste0(fig_dir, 'DME_cor_compare_EAD_vs_ADDS.pdf'), width=3, height=3)
p
dev.off()

##########################################################################
# TODO: upset plot comparing up- and down- regulated modules
##########################################################################

st_diag_cp <- c("Control" = "#B8DBC5", "earlyAD" = "#E7BDE1" , "AD" = "#CF8BA3", "AD_DS" = "#9E6D7F")


library(UpSetR)

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv')) %>% 
    subset(annotation == wgcna)


# fold-change cutoff:
fc_cutoff <- 0.25
groups <- unique(DMEs$ident.1)
DMEs$module <- factor(as.character(DMEs$module), levels=labels(dend))

upset_list <- list()
names <- c()
for(cur_group in groups){
  cur_df_up <- DMEs %>% subset(ident.1 == cur_group & p_val_adj < 0.05 & avg_log2FC > fc_cutoff)
  cur_df_down <- DMEs %>% subset(ident.1 == cur_group & p_val_adj < 0.05 & avg_log2FC < -fc_cutoff)
  upset_list[[paste0(cur_group, '_down')]] <- table(cur_df_down$module)
  upset_list[[paste0(cur_group, '_up')]] <- table(cur_df_up$module)
}

sum((upset_list$AD_DS_up + upset_list$AD_up) == 2)
sum((upset_df$AD_DS_up + upset_df$AD_up) == 2)
sum((upset_df$AD_DS_up + upset_df$AD_up + upset_df$earlyAD_up) ==  3)
sum((upset_df$AD_DS_down + upset_df$AD_down + upset_df$earlyAD_down) ==  3)


# tmp <- c(paste0(groups, '_up'), paste0(groups, '_down'))
# names <- tmp[order(tmp)]

# combine into one df
upset_df <- as.data.frame(Reduce(cbind, upset_list))
colnames(upset_df) <- names(upset_list)

# remove entries with all 0
upset_df <- upset_df[rowSums(upset_df) > 0,]

cur_colors <- st_diag_cp[groups]
names(cur_colors) <- paste0(names(cur_colors), '_up')
#cur_colors_down <- unlist(lapply(celltype_colors[groups], scales::muted))
cur_colors_down <- st_diag_cp[groups]
names(cur_colors_down) <- paste0(names(cur_colors_down), '_down')
cur_colors <- c(cur_colors, cur_colors_down)
cur_colors <- cur_colors[colnames(upset_df)]

tmp 

p <- UpSetR::upset(
  upset_df,
  sets=colnames(upset_df),
  group.by='sets',
  sets.bar.color=cur_colors,
)




pdf(paste0(fig_dir, 'DME_compare_upset.pdf'), width=6, height=5)
p
dev.off()


##########################################################################
# TODO: Spatial featureplot for module eigengenes
##########################################################################

```

Metamodule DME heatmaps 

```{r eval=FALSE}

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv')) %>% 
    subset(wgcna == 'consensus')

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))

test <- DMEs %>% subset(annotation == wgcna | wgcna == 'consensus') 

 write.csv(test, file=paste0(data_dir, 'differential_MEs_humanST_match.csv'), quote=FALSE, row.names=FALSE)



maxval <- 1; minval <- -1
plot_df <- DMEs
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC > maxval, maxval, plot_df$avg_log2FC)
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC < minval, minval, plot_df$avg_log2FC)

plot_df$textcolor <- ifelse(plot_df$avg_log2FC > 0.5, 'black', 'white')
plot_df$Significance <- ifelse(plot_df$p_val_adj > 0.05, 'x', NA)

# set factor levels:
plot_df$module <- factor(as.character(plot_df$module), levels=mods)
plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))

p <- plot_df %>% 
  ggplot(aes(x=module, y=annotation, fill=avg_log2FC)) +
  geom_tile() +
  geom_text(label=plot_df$Significance, color=plot_df$textcolor) +
  scale_fill_gradient2(low='purple', mid='black', high='yellow') +
  RotatedAxis() +
 # labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    #axis.title.x = element_blank(),
    #axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
   # axis.ticks.x = element_blank(),
   # axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() # + xlab(group1) + ylab(group2)


# Plot the result
pdf(paste0(fig_dir, 'DME_metamodule_heatmap.pdf'),height=7, width=5)
p + facet_wrap(~ident.1, ncol=1)
dev.off()


```

DMEs between Female & Male in AD/DS 

```{r eval=FALSE}

# re-load the sex downsamples object 

name <- 'ADDS'
name <- 'AD'
name <- 'earlyAD'

seurat_subset <- readRDS(file=paste0("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/data/", name, "_seurat_sex_downsampled.rds"))

seurat_human$anno_DMEs <- ifelse(
    grepl('WM', as.character(seurat_human$annotation)), 
    'WM', as.character(seurat_human$annotation)
)

groups <- unique(seurat_human$anno_DMEs)
wgcna_groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM', 'consensus')
combos <- expand.grid(groups, wgcna_groups)

DMEs <- Reduce(rbind, lapply(1:nrow(combos), function(i){
    print(i)
    cur_anno <- as.character(combos[i, 'Var1'])
    cur_wgcna <- as.character(combos[i,'Var2'])

    # get list of cells with these annotations / diagnosis
    g1 <- seurat_human@meta.data %>% subset(anno_DMEs == cur_anno & Sex == 'F') %>% rownames
    g2 <- seurat_human@meta.data %>% subset(anno_DMEs == cur_anno & Sex == 'M') %>% rownames

    # make sure the list of barcodes are in the downsampled dataset 
    g1 <- g1[g1 %in% colnames(seurat_subset)]
    g2 <- g2[g2 %in% colnames(seurat_subset)]

    # reset active WGCNA
    seurat_human <- SetActiveWGCNA(seurat_human, cur_wgcna)

    # run the DME comparison with a wilcox test
    DMEs <- FindDMEs(
        seurat_human,
        barcodes1 = g1,
        barcodes2 = g2,
        test.use='wilcox'
    )
    DMEs$ident.1 <- 'F'
    DMEs$ident.2 <- 'M'
    DMEs$annotation <- cur_anno
    DMEs$wgcna <- cur_wgcna
    DMEs

}))

# save the DME results as a table:
write.csv(DMEs, file=paste0(data_dir, 'differential_MEs_sex_', name, '_humanST.csv'), quote=FALSE, row.names=FALSE)



###############################################################
# Volcano Plot
###############################################################



name <- 'ADDS'
name <- 'AD'
name <- 'earlyAD'

# re-load
DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_sex_', name,'_humanST.csv'))


# compute the number of genes in each module:
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')
all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]

    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') 
mod_size <- table(all_mods$module)

# get a map of module and color
all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    #cur_mods <- cur_mods[genes_use,]
    #cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(module, color)) %>% 
    distinct()

    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') %>% distinct
mod_colors <- all_mods$color; names(mod_colors) <- all_mods$module


#DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_sex_humanST.csv')) 

# p <- DMEs %>% 
#   ggplot(aes(x=avg_log2FC, y=-log10(p_val_adj))) + 
#   geom_hline(yintercept=-log10(0.05), linetype='dashed') +
#   geom_point()

plot_df <- DMEs %>% subset(annotation==wgcna)


# label the top 10 and bottom 10
nlabel <- 10
top_mods <- plot_df %>% subset(p_val_adj < 0.05) %>%
    slice_max(avg_log2FC, n=nlabel) %>% .$module %>% as.character
bottom_mods <- plot_df %>% subset(p_val_adj < 0.05) %>%
    slice_min(avg_log2FC, n=nlabel) %>% .$module %>% as.character
label_mods <- c(top_mods, bottom_mods)
plot_df$anno <- ifelse(plot_df$module %in% label_mods, plot_df$module, NA)

print(top_mods)
print(bottom_mods)

# set module factor level:
plot_df$module <- factor(as.character(plot_df$module), levels=names(mod_colors))

# add module sizes:
plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

# add module color
ix <- match(as.character(plot_df$module), names(mod_colors))
plot_df$color <- as.character(mod_colors[ix])
plot_df$color <- ifelse(plot_df$p_val_adj > 0.05, 'grey', plot_df$color)

plot_range <- max(abs(range(plot_df$avg_log2FC)))

p <- plot_df %>% 
    ggplot(aes(x = avg_log2FC, y = -log10(p_val_adj), size=log2(mod_size)/2)) +
      geom_hline(yintercept=-log10(0.05), linetype='dashed', color='black', linewidth=0.5) +
    ggrastr::rasterise(geom_point(color=plot_df$color), dpi=500) +
    geom_point(
       inherit.aes=FALSE,
       data=subset(plot_df, !is.na(anno)),
       aes(avg_log2FC,-log10(p_val_adj),  size=log2(mod_size)/2),
       fill = subset(plot_df, !is.na(anno)) %>% .$color,
       shape=21, color='black'
     ) +
    geom_text_repel(aes(label=anno), color='black',  min.segment.length=0, max.overlaps=Inf, size=2) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
   # geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC, y = avg_log2FC2), method='lm', color='black') +
   # stat_cor(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC, y = avg_log2FC2), method='pearson') +
    xlim(c(-plot_range, plot_range)) +
   # ylim(c(-plot_range, plot_range)) +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5),
      panel.grid = element_line(size=0.25, color='white'),
      panel.background = element_rect(fill='grey92'),
    ) +
   # coord_fixed(ratio=1) + 
    NoLegend() +
    xlab(bquote("F vs M Average log"[2]~"(FC)")) +
     ylab(bquote("-log"[10]~"(Adj. P-value)")) 

pdf(paste0(fig_dir, 'DME_sex_volcano_', name, '.pdf'), width=4, height=4)
p
dev.off()


###############################################################
# metamodule heatmap
###############################################################


name <- 'ADDS'
name <- 'AD'
name <- 'earlyAD'

modules <- GetModules(seurat_human, 'consensus')
mods <- levels(modules$module)

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_sex_', name,'_humanST.csv')) %>%
  subset(wgcna == 'consensus')


maxval <- 1.5; minval <- -1.5
plot_df <- DMEs
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC > maxval, maxval, plot_df$avg_log2FC)
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC < minval, minval, plot_df$avg_log2FC)

plot_df$textcolor <- ifelse(plot_df$avg_log2FC > 0.5, 'black', 'white')
plot_df$Significance <- ifelse(plot_df$p_val_adj > 0.05, 'x', NA)
#plot_df$Significance <- gtools::stars.pval(plot_df$p_val_adj)
plot_df$alpha <- ifelse(plot_df$p_val_adj > 0.05, 0.25, 1)

# set factor levels:
plot_df$module <- factor(as.character(plot_df$module), levels=mods)
plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))

#plot_df <- na.omit(plot_df)

p <- plot_df %>% 
  ggplot(aes(x=module, y=annotation, fill=avg_log2FC)) +
  geom_tile(alpha=plot_df$alpha) +
  geom_text(label=plot_df$Significance, color='black') +
  scale_fill_gradient2(low='dodgerblue', mid='white', high='hotpink', limits=c(minval, maxval)) +
  RotatedAxis() +
 # labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() + xlab('') + ylab('')


# Plot the result
pdf(paste0(fig_dir, 'DME_sex_metamodule_heatmap_', name, '.pdf'),height=3, width=5)
p 
dev.off()



###############################################################
# Scatter plot comparison of sex DMEs and diagnosis DMEs 
###############################################################

#name <- 'AD_DS'

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_sex_', name,'_humanST.csv'))
DMEs_dx <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv')) %>% subset(ident.1 == name)

DMEs_dx$id <- paste0(DMEs_dx$module, '_', DMEs_dx$annotation, '_', DMEs_dx$wgcna)
DMEs$id <- paste0(DMEs$module, '_', DMEs$annotation, '_', DMEs$wgcna)

plot_df <- dplyr::inner_join(
  DMEs, 
  DMEs_dx %>% dplyr::select(c(p_val_adj, avg_log2FC, id)), 
  by = 'id'
  ) %>% 
  subset(annotation == wgcna)


# set module factor level:
plot_df$module <- factor(as.character(plot_df$module), levels=names(mod_colors))

# add module sizes:
plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

# add module color
ix <- match(as.character(plot_df$module), names(mod_colors))
plot_df$color <- as.character(mod_colors[ix])
#plot_df$color <- ifelse(plot_df$p_val_adj > 0.05, 'grey', plot_df$color)

plot_range <- max(c(abs(range(plot_df$avg_log2FC.y)), abs(range(plot_df$avg_log2FC.x))))

logfc_thresh <- 0.25

up_right <- plot_df %>% subset(avg_log2FC.x >= logfc_thresh & avg_log2FC.y >= logfc_thresh & (p_val_adj.y <= 0.05 | p_val_adj.x <= 0.05)) %>% nrow
down_right <- plot_df %>% subset(avg_log2FC.x >= logfc_thresh & avg_log2FC.y <= -logfc_thresh & (p_val_adj.y <= 0.05 | p_val_adj.x <= 0.05)) %>% nrow
up_left <- plot_df %>% subset(avg_log2FC.x <= -logfc_thresh & avg_log2FC.y >= logfc_thresh & (p_val_adj.y <= 0.05 | p_val_adj.x <= 0.05)) %>% nrow
down_left <- plot_df %>% subset(avg_log2FC.x <= -logfc_thresh & avg_log2FC.y <= -logfc_thresh & (p_val_adj.y <= 0.05 | p_val_adj.x <= 0.05)) %>% nrow

annotations <- data.frame(
        xpos = c(-Inf,-Inf,Inf,Inf),
        ypos =  c(-Inf, Inf,-Inf,Inf),
        annotateText = c(as.character(down_left),as.character(up_left), as.character(down_right),as.character(up_right)),
        hjustvar = c(-1,-1,2,2),
        vjustvar = c(-1,2,-1,2)) #<- adjust


p <- plot_df %>% 
  ggplot(aes(x=avg_log2FC.x, y=avg_log2FC.y, color=module, size=log2(mod_size)/2)) + 
 geom_hline(yintercept = 0, linetype='dashed', color='black', linewidth=0.25) +
  geom_vline(xintercept = 0, linetype='dashed', color='black', linewidth=0.25) +
  ggrastr::rasterise(geom_point(), dpi=500) +
  geom_smooth(inherit.aes=FALSE, data=plot_df, mapping = aes(x = avg_log2FC.x, y = avg_log2FC.y), method='lm', color='black') +
    stat_cor(
      inherit.aes=FALSE, 
      data=plot_df, 
      mapping = aes(x = avg_log2FC.x, y = avg_log2FC.y, label=..r.label..), 
      label.y = plot_range * 0.75,
      method='pearson') +
    xlim(c(-plot_range, plot_range)) +
    ylim(c(-plot_range, plot_range)) +
    scale_color_manual(values=mod_colors) +
    scale_size_continuous(range = c(0.5,4)) +
    theme(
      axis.line.x = element_blank(),
      axis.line.y = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      panel.border = element_rect(colour = "black", fill=NA, size=1),
      plot.title = element_text(hjust=0.5),
      panel.grid = element_line(size=0.25, color='white'),
      panel.background = element_rect(fill='grey92'),
    ) +
    coord_fixed(ratio=1) + NoLegend() + 
    geom_text(inherit.aes=FALSE, data=annotations,aes(x=xpos,y=ypos,hjust=hjustvar,vjust=vjustvar,label=annotateText))



# Plot the result
pdf(paste0(fig_dir, 'DME_sex_dx_compare_', name, '.pdf'), height=2.5, width=2.5)
print(p)
dev.off()




```




Module-trait correlation 

```{r eval=FALSE}

wgcna_groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM', 'consensus')
traits <- c('Age', 'Sex')

seurat_human$anno_DMEs <- ifelse(
    grepl('WM', as.character(seurat_human$annotation)), 
    'WM', as.character(seurat_human$annotation)
)

seurat_human$Sex <- factor(as.character(seurat_human$Sex), levels=c('F', 'M'))

cor_df <- Reduce(rbind, lapply(wgcna_groups, function(cur_wgcna){
    print(cur_wgcna)
    seurat_human <- SetActiveWGCNA(seurat_human, cur_wgcna)

    seurat_human <- ModuleTraitCorrelation(
        seurat_human,
        traits = traits,
        group.by = 'anno_DMEs',
        subset_groups = TRUE
    )

    mt_cor <- GetModuleTraitCorrelation(seurat_human)
    mt_cor_df <- reshape2::melt(mt_cor$cor)
    colnames(mt_cor_df) <- c('trait', 'module', 'cor', 'group')
    mt_cor_df$wgcna <- cur_wgcna 
    mt_cor_df

}))

#################################################
# amyloid mt cor
#################################################

# load seurat object with amyloid info:
seurat_amylo <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/voyager/data/ADDS_seurat_processed_geospatial.rds")

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_amylo@meta.data %>%
  dplyr::group_by(Sample) %>%
  dplyr::summarise(Mean = mean(Amyloglo_Score_log)) %>%
  dplyr::filter(Mean != 0) %>% .$Sample %>% as.character

seurat_human$amyloid_sample <- seurat_human$Sample %in% amyloid_samples
traits <- c('Amyloglo_Score_log', 'Amyloglo_Score_log_Gi', 'OC_Score_log', 'OC_Score_log_Gi')
amylo_meta <- seurat_amylo@meta.data
meta <- seurat_human@meta.data 
seurat_human@meta.data <- cbind(meta, amylo_meta[,traits])

conditions <- levels(seurat_human$Diagnosis)

plot_df <- data.frame()
for(cur_condition in conditions){
  for(cur_wgcna in wgcna_groups){

    print(cur_wgcna)
    seurat_human <- SetActiveWGCNA(seurat_human, cur_wgcna)
    seurat_human$amyloid_sample <- (seurat_human$Sample %in% amyloid_samples) & (seurat_human$Diagnosis == cur_condition)

    seurat_human <- ModuleTraitCorrelation(
        seurat_human,
        traits = traits,
        group.by = 'anno_DMEs',
        subset_by = 'amyloid_sample',
        subset_groups = TRUE
    )

    mt_cor <- GetModuleTraitCorrelation(seurat_human)

    cor_df <- do.call(rbind, lapply(names(mt_cor$cor), function(x){
        cur_cor <- mt_cor$cor[[x]]
        cur_fdr <- mt_cor$fdr[[x]]
        cur_cor_df <- reshape2::melt(cur_cor)
        cur_fdr_df <- reshape2::melt(cur_fdr)
        cur_cor_df$fdr <- cur_fdr_df$value
        cur_cor_df$group <- x
        cur_cor_df
    }))
    cor_df %<>% dplyr::rename(c(trait = Var1, module = Var2, cor=value))
    cor_df$Diagnosis <- cur_condition 
    cor_df$wgcna <- cur_wgcna

    plot_df <- rbind(plot_df, cor_df)

    }
}

write.csv(plot_df, file=paste0(data_dir, 'amylo_mtcor_humanST.csv'), quote=FALSE, row.names=FALSE)


tmp <- rbind(cor_df, amylo_cor_df)
write.csv(tmp, file=paste0(data_dir, 'mtcor_humanST.csv'), quote=FALSE, row.names=FALSE)

tmp <- read.csv(file=paste0(data_dir, 'mtcor_humanST.csv'))


#################################################
# TODO: combine the amylo cor df and cor ddf into a matrix to plot with the dendro
#################################################
traits <- c('Amyloglo_Score_log', 'Amyloglo_Score_log_Gi', 'OC_Score_log', 'OC_Score_log_Gi')

plot_df <- read.csv(file=paste0(data_dir, 'amylo_mtcor_humanST.csv'))

plot_df <- subset(plot_df, trait %in% traits & wgcna == 'consensus' & group != 'all_cells')


plot_df %>% subset(fdr < 0.05 & module == 'M11') %>% arrange(-cor) %>% head(25)
plot_df %>% subset(fdr < 0.05) %>% arrange(-cor) %>% head(25)



#---------------------------------------------------------------------
# code from Seurat DotPlot
#---------------------------------------------------------------------

# get wgcna info 
data.features <- GetMEs(seurat_human, wgcna_name='consensus') %>% as.data.frame()
data.features$id <- seurat_human$anno_DMEs
plot_df$id <- paste(plot_df$group, plot_df$Diagnosis, sep='_')



split.by <- 'Diagnosis'
scale <- TRUE
col.min <- -2.5
col.max <- 2.5
dot.min <- 0

data.features$id <- paste(seurat_human$anno_DMEs, seurat_human$Diagnosis, sep='_')

data.plot <- lapply(
    X = unique(x = data.features$id),
    FUN = function(ident) {
      data.use <- data.features[data.features$id == ident, 1:(ncol(x = data.features) - 1), drop = FALSE]
      avg.exp <- apply(
        X = data.use,
        MARGIN = 2,
        FUN = function(x) {
          return(mean(x = expm1(x = x)))
        }
      )
      pct.exp <- apply(X = data.use, MARGIN = 2, FUN = PercentAbove, threshold = 0)
      return(list(avg.exp = avg.exp, pct.exp = pct.exp))
    }
  )
names(x = data.plot) <- unique(x = data.features$id)
id.levels <- levels(seurat_obj$cell_type)
data.plot <- lapply(
    X = names(x = data.plot),
    FUN = function(x) {
      data.use <- as.data.frame(x = data.plot[[x]])
      data.use$features.plot <- rownames(x = data.use)
      data.use$id <- x
      return(data.use)
    }
  )
  data.plot <- do.call(what = 'rbind', args = data.plot)
  if (!is.null(x = id.levels)) {
    data.plot$id <- factor(x = data.plot$id, levels = id.levels)
  }
  ngroup <- length(x = levels(x = data.plot$id))

  avg.exp.scaled <- sapply(
    X = unique(x = data.plot$features.plot),
    FUN = function(x) {
      data.use <- data.plot[data.plot$features.plot == x, 'avg.exp']
      if (scale) {
        data.use <- scale(x = log1p(data.use))
        data.use <- MinMax(data = data.use, min = col.min, max = col.max)
      } else {
        data.use <- log1p(x = data.use)
      }
      return(data.use)
    }
  )
  avg.exp.scaled <- as.vector(x = t(x = avg.exp.scaled))
  data.plot$avg.exp.scaled <- avg.exp.scaled
  
  data.plot$pct.exp[data.plot$pct.exp < dot.min] <- NA
  data.plot$pct.exp <- data.plot$pct.exp * 100

# match with the other table
data.plot <- subset(data.plot, id %in% plot_df$id)

# merge the DFs
data.plot$ix <- paste0(data.plot$features.plot, '_', data.plot$id)
plot_df$ix <- paste0(plot_df$module, '_', plot_df$id)

ix <- match(plot_df$ix, data.plot$ix, )
all.equal(plot_df$ix, data.plot$ix[ix])

plot_df$avg.exp <- data.plot$avg.exp[ix]
plot_df$avg.exp.scaled <- data.plot$avg.exp.scaled[ix]
plot_df$pct.exp <- data.plot$pct.exp[ix]



plot_df$Significance <- ifelse(plot_df$fdr > 0.05, 'x', NA)

modules <- GetModules(seurat_human, wgcna_name = 'consensus')
mods <- levels(modules$module)
plot_df$module <- factor(as.character(plot_df$module), levels=mods)


plot_range <- plot_df %>% .$cor %>% range
size_range <- plot_df %>% .$pct.exp %>% range

# loop through Studies & diagnosis
diagnoses <- unique(as.character(seurat_human$Diagnosis))
plot_list <- list()
for(cur_diag in diagnoses){
 
    cur_df <- plot_df %>% subset(Diagnosis == cur_diag & trait %in% c('Amyloglo_Score_log_Gi', 'OC_Score_log_Gi'))

    p <-cur_df %>%
        ggplot(aes(x = module, y = group, color=cor, size=pct.exp)) + 
        geom_point() + 
        scale_color_gradient2(high='red', mid='white', low='blue', limits=plot_range) + 
        scale_size_continuous(limits=size_range) +
        geom_text(aes(label=Significance), color='black', size=3) +
        coord_equal() +
        RotatedAxis() +
        ylab(paste0(cur_diag)) + 
        theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            axis.title.x = element_blank(),
            axis.title.y = element_text(angle=0, vjust=0.5),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='lightgrey'),
            plot.margin = margin(c(0,0,0,0)),
            
        ) 

    if(length(plot_list) > 0){
        p <- p + theme(
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
        )
    }

    patch <- p + facet_wrap(~trait, ncol=4)

    plot_list[[cur_diag]] <- patch

}


pdf(paste0(fig_dir, 'MT_cor_heatmap_consensus_visium_dotplot_amylo.pdf'), width=12, height=8)
wrap_plots(rev(plot_list), ncol=1) + plot_layout(guides='collect')
dev.off()









```


Test plot circular dendrogram with heatmap using the circlized package

```{r eval=FALSE}

library(circlize)
library(dendextend)

load(paste0(data_dir, 'dendro.rda'))

##########################################
# Plot a circular dendrogram
###

dend <- as.dendrogram(hclust_avg)
# labels(dend) <- x[order.dendrogram(dend)]

dend_colors <- WGCNA::labels2colors(mod_cut_full)
names(dend_colors) <- names(mod_cut_full)
dend_colors <- dend_colors[labels(dend)]

dend <- dend %>% 
    set('leaves_pch', 19) %>%
    set('leaves_col', dend_colors) %>%
    dendextend::rotate(names(sort(mod_cut_full)))


# reorder labels:
# mod_cut_full[sort(mod_cut_full)]

pdf(paste0(fig_dir, 'test_module_hierarchy_circle3_today.pdf'), width=10, height=10)
circlize_dendrogram(
    dend,
    labels_track_height=NA,
    dend_track_height=0.4
)
dev.off()


# try making it with the circlize package 
labels <- labels(dend)
ct <- mod_cut_full
n <- length(labels)
pdf(paste0(fig_dir, 'test_module_hierarchy_circlize.pdf'), width=10, height=10)

###circos.par(cell.padding = c(0,0,0,0))
circos.initialize("a", xlim=c(0,n))

circos.track(ylim = c(0, 1), bg.border = NA, track.height = 0.3, 
    panel.fun = function(x, y) {
        for(i in seq_len(n)) {
            circos.text(i-0.5, 0, labels[i], adj = c(0, 0.5), 
                facing = "clockwise", niceFacing = TRUE,
                col = 'black', cex = 0.5)
        }
})

dend_height = attr(dend, "height")
circos.track(ylim = c(0, dend_height), bg.border = NA, 
    track.height = 0.4, panel.fun = function(x, y) {
        circos.dendrogram(dend)
})
circos.clear()
dev.off()

#####################################################################3
# make the circlize dendro with DME heamtap
#####################################################################

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))

# set up DME data for circular dendro plotting:
tmp <- subset(DMEs, annotation == wgcna) %>% 
    dplyr::select(c(module, ident.1, avg_log2FC)) %>%
    pivot_wider(names_from = ident.1, values_from = avg_log2FC) %>% 
    as.data.frame()

rownames(tmp) <- tmp$module
tmp <- tmp[,2:ncol(tmp)]


mat <- as.matrix(tmp)[labels(dend),]

# clip values
min_val <- -1 
max_val <- 1

mat[mat > max_val] <- max_val
mat[mat < min_val] <- min_val

#mat = matrix(rnorm(n*3), nrow = n, ncol = 3)
#mat2 = matrix(rnorm(n*3), nrow = n, ncol = 3)

col_fun = colorRamp2(c(min_val, 0, max_val), c("purple", "black", "yellow"))
#col_fun2 = colorRamp2(c(-2, 0, 2), c("green", "black", "red"))

# sectors = rep(letters[1:2], times = c(30, 70))
# mat_list = list(a = mat[sectors == "a", ],
#                 b = mat[sectors == "b", ])
# dend_list = list(a = as.dendrogram(hclust(dist(mat_list[["a"]]))),
#                  b = as.dendrogram(hclust(dist(mat_list[["b"]]))))


pdf(paste0(fig_dir, 'test_module_hierarchy_circlize_heat3.pdf'), width=9, height=9)

circos.par(cell.padding = c(0, 0, 0, 0), gap.degree = 5)
# circos.initialize(sectors, xlim = cbind(c(0, 0), table(sectors)))
circos.initialize("a", xlim=c(0,n))

# make labels
circos.track(ylim = c(0, 1), bg.border = NA, track.height = 0.2, 
    panel.fun = function(x, y) {
        for(i in seq_len(n)) {
            circos.text(i-0.5, 0, labels[i], adj = c(0, 0.5), 
                facing = "clockwise", niceFacing = TRUE,
                col = 'black', cex = 0.75)
        }
})

# make heatmap1
circos.track(ylim = c(0, 3), bg.border = NA, track.height=0.1, panel.fun = function(x, y) {
    sector.index = CELL_META$sector.index
    m = mat
    dend = dend

    m2 = m[order.dendrogram(dend), ]
    col_mat = col_fun(m2)
    nr = nrow(m2)
    nc = ncol(m2)
    cn = colnames(m)
    for(i in 1:nc) {
        circos.rect(1:nr - 1, rep(nc - i, nr), 
            1:nr, rep(nc - i + 1, nr), 
            border = col_mat[, i], col = col_mat[, i])
    }
    circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
            1:n - 0.5, cn, 
            cex = 0.5, adj = c(0, 0.5), facing = "inside")
})

# make dendrogram
max_height = dend_height
circos.track(ylim = c(0, max_height), bg.border = NA, track.height = 0.2, 
    panel.fun = function(x, y) {
        sector.index = get.cell.meta.data("sector.index")
        dend = dend
        circos.dendrogram(dend, max_height = max_height)
})
circos.clear()

dev.off()






# making this smaller for the paper revision
pdf(paste0(fig_dir, 'test_module_hierarchy_circlize_heat3_today.pdf'), width=12, height=12)

circos.par(cell.padding = c(0, 0, 0, 0), gap.degree = 5)
# circos.initialize(sectors, xlim = cbind(c(0, 0), table(sectors)))
circos.initialize("a", xlim=c(0,n))

# make labels
circos.track(ylim = c(0, 2), bg.border = NA, track.height = 0.45, 
    panel.fun = function(x, y) {
        for(i in seq_len(n)) {
            circos.text(i-0.5, 0, labels[i], adj = c(0, 0.5), 
                facing = "clockwise", niceFacing = TRUE,
                col = 'black', cex = 0.8)
        }
})

# make heatmap1
circos.track(ylim = c(0, 3), bg.border = NA, track.height=0.07, panel.fun = function(x, y) {
    sector.index = CELL_META$sector.index
    m = mat
    dend = dend

    m2 = m[order.dendrogram(dend), ]
    col_mat = col_fun(m2)
    nr = nrow(m2)
    nc = ncol(m2)
    cn = colnames(m)
    for(i in 1:nc) {
        circos.rect(1:nr - 1, rep(nc - i, nr), 
            1:nr, rep(nc - i + 1, nr), 
            border = col_mat[, i], col = col_mat[, i])
    }
    circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
            1:n - 0.5, cn, 
            cex = 0.5, adj = c(0, 0.5), facing = "inside")
})

# make dendrogram
max_height = dend_height
circos.track(ylim = c(0, max_height), bg.border = NA, track.height = 0.1, 
    panel.fun = function(x, y) {
        sector.index = get.cell.meta.data("sector.index")
        dend = dend
        circos.dendrogram(dend, max_height = max_height)
})
circos.clear()

dev.off()


#####################################################################3
# make the circlize dendro with DME heamtap + MT cor heatmap
#####################################################################

# # DMEs
# DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))

# # set up DME data for circular dendro plotting:
# tmp <- subset(DMEs, annotation == wgcna) %>% 
#     dplyr::select(c(module, ident.1, avg_log2FC)) %>%
#     pivot_wider(names_from = ident.1, values_from = avg_log2FC) %>% 
#     as.data.frame()

# rownames(tmp) <- tmp$module
# tmp <- tmp[,2:ncol(tmp)]

# mat <- as.matrix(tmp)[labels(dend),]

# # clip values
# min_val <- -1 
# max_val <- 1

# mat[mat > max_val] <- max_val
# mat[mat < min_val] <- min_val
# col_fun = colorRamp2(c(min_val, 0, max_val), c("purple", "black", "yellow"))



# # mt cor 
# tmp <- read.csv(file=paste0(data_dir, 'mtcor_humanST.csv'))

# traits <-  c('Age', 'Sex', 'Amyloglo_Score_log_Gi', 'OC_Score_log_Gi')

# tmp <- tmp %>% 
#     subset(group == wgcna) %>% 
#     subset(trait %in% traits) %>%
#     mutate(trait = factor(as.character(trait), levels=traits)) %>% 
#     dplyr::select(c(module, trait, cor)) %>%
#     pivot_wider(names_from = trait, values_from = cor) %>% 
#     as.data.frame()


# rownames(tmp) <- tmp$module
# tmp <- tmp[,2:ncol(tmp)]


# mat2 <- as.matrix(tmp)[labels(dend),]

# # clip values
# min_val <- -0.2
# max_val <- 0.2

# mat2[mat2 > max_val] <- max_val
# mat2[mat2 < min_val] <- min_val



# col_fun2 = colorRamp2(c(min_val, 0, max_val), c("orange", "white", "blue"))



# pdf(paste0(fig_dir, 'test_module_hierarchy_circlize_heat4.pdf'), width=12, height=12)

# circos.par(cell.padding = c(0, 0, 0, 0), gap.degree = 5)
# # circos.initialize(sectors, xlim = cbind(c(0, 0), table(sectors)))
# circos.initialize("a", xlim=c(0,n))

# # make labels
# circos.track(ylim = c(0, 1), bg.border = NA, track.height = 0.2, 
#     panel.fun = function(x, y) {
#         for(i in seq_len(n)) {
#             circos.text(i-0.5, 0, labels[i], adj = c(0, 0.5), 
#                 facing = "clockwise", niceFacing = TRUE,
#                 col = 'black', cex = 0.5)
#         }
# })

# # make heatmap1
# circos.track(ylim = c(0, ncol(mat)), bg.border = NA, track.height=0.1, panel.fun = function(x, y) {
#     sector.index = CELL_META$sector.index
#     m = mat
#     dend = dend

#     m2 = m[order.dendrogram(dend), ]
#     col_mat = col_fun(m2)
#     nr = nrow(m2)
#     nc = ncol(m2)
#     cn = colnames(m)
#     for(i in 1:nc) {
#         circos.rect(1:nr - 1, rep(nc - i, nr), 
#             1:nr, rep(nc - i + 1, nr), 
#             border = col_mat[, i], col = col_mat[, i])
#     }
#     circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
#             1:n - 0.5, cn, 
#             cex = 0.5, adj = c(0, 0.5), facing = "inside")
# })

# # # make heatmap2
# circos.track(ylim = c(0, ncol(mat2)), bg.border = NA, track.height=0.1, panel.fun = function(x, y) {
#     sector.index = CELL_META$sector.index
#     m = mat2
#     dend = dend

#     m2 = m[order.dendrogram(dend), ]
#     col_mat = col_fun2(m2)
#     nr = nrow(m2)
#     nc = ncol(m2)
#     cn = colnames(m)
#     for(i in 1:nc) {
#         circos.rect(1:nr - 1, rep(nc - i, nr), 
#             1:nr, rep(nc - i + 1, nr), 
#             border = col_mat[, i], col = col_mat[, i])
#     }
#     circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"), 
#         1:n - 0.5, cn, 
#         cex = 0.5, adj = c(0, 0.5), facing = "inside")
# })

# # make dendrogram
# max_height = dend_height
# circos.track(ylim = c(0, max_height), bg.border = NA, track.height = 0.3, 
#     panel.fun = function(x, y) {
#         sector.index = get.cell.meta.data("sector.index")
#         dend = dend
#         circos.dendrogram(dend, max_height = max_height)
# })
# circos.clear()

# dev.off()

```

Module FeaturePlots for the single-cell dataset:
```{r eval=FALSE}

head(GetModules(seurat_obj, 'consensus'))


MEs <- GetMEs(seurat_obj)
meta <- seurat_obj@meta.data
seurat_obj@meta.data <- cbind(meta, MEs)
mods <- colnames(MEs)





p <- FeatureEmbedding(
  seurat_obj,
  features=mods,
  same_range=FALSE,
  ncol=4,
  plot_max='q99',
  plot_min=0,
  dpi=200,
  point_size=1,
  dpi_scale=0.5,
  order_points=TRUE,
  colfunc=viridis::magma,
  combine=FALSE
)
class(p)

seurat_obj@meta.data <- seurat_obj@meta.data[,!(colnames(seurat_obj@meta.data) %in% colnames(MEs))]



pdf(paste0(fig_dir, 'projected_MEs_featureplot.pdf'), width=6, height=21)
wrap_plots(p, ncol=2)
dev.off()


plot_list <- ModuleFeaturePlot(
    seurat_obj, 
    order=TRUE, 
    raster=TRUE, 
    restrict_range=FALSE, 
    raster_dpi=200, 
    raster_scale=0.5, 
    point_size=1
)

pdf(paste0(fig_dir, 'projected_MEs_featureplot.pdf'), width=6, height=21)
wrap_plots(plot_list, ncol=2)
dev.off()

# plot the clustering
sc_color_df <- read.csv(file=paste0('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/cell_identity.csv'))

celltype_colors <- c(
   'EX' = sc_color_df %>% subset(group == 'EX L5') %>% .$colour,
   'INH' = sc_color_df %>% subset(group == 'INH VIP+') %>% .$colour,
   'ASC' = sc_color_df %>% subset(group == 'ASC2') %>% .$colour,
    'MG' = sc_color_df %>% subset(group == 'MG2') %>% .$colour,
    'ODC' = sc_color_df %>% subset(group == 'ODC3') %>% .$colour,
    'OPC' = sc_color_df %>% subset(group == 'OPC3') %>% .$colour,
    'END' = sc_color_df %>% subset(group == 'END Arterial') %>% .$colour,
    'PER' = sc_color_df %>% subset(group == 'T-Pericyte') %>% .$colour,
    'SMC' = sc_color_df %>% subset(group == 'SMC') %>% .$colour,
    'FBR' = sc_color_df %>% subset(group == 'Meningeal Fibroblast') %>% .$colour
)
cp_df <- data.frame(
    colour = as.character(celltype_colors),
    group = names(celltype_colors)
)


p <- PlotEmbedding(
  seurat_obj,
  group.by = 'cell_type',
  raster_dpi = 400,
  point_size=0.5,
  raster_scale = 0.5,
  plot_theme = umap_theme(),
  plot_under=TRUE,
  color_df = cp_df
)

pdf(paste0(fig_dir, 'umap_singlecell_celltypes.pdf'), width=6, height=6)
p
dev.off()


```

Plot MEs in snRNA-seq dataset in the spatial coordinates 

```{r eval=FALSE}

# load celltrek output coordinates:
anno_df <- read.csv(file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/spatial_mapping/celltrek_coords_combined.csv')

# add a clear coarse-grain celltype grouping:
seurat_obj$coarse_groups <- ifelse(
  seurat_obj$cell_type %in% c('FBR', 'PER', 'END', 'SMC'), 'VASC', seurat_obj$cell_type
)
seurat_obj$coarse_groups <- factor(
  as.character(seurat_obj$coarse_groups, levels=c('EX', 'INH', 'ODC', 'OPC', 'MG', 'ASC', 'VASC'))
)



# add MEs to seurat object
modules <- GetModules(seurat_obj)
MEs <- GetMEs(seurat_obj)
meta <- seurat_obj@meta.data
seurat_obj@meta.data <- cbind(meta, MEs)
mods <- colnames(MEs)

# select sample
cur_vis_sample <- 'Dec_13_2021_Human8'
cur_vis_sample <- 'Dec_13_2021_Human5'

# is this sample cntrl or AD/DS?
cur_vis_sample_dx <- subset(seurat_human@meta.data, Sample == cur_vis_sample) %>% .$Diagnosis %>% unique %>% as.character
if(cur_vis_sample_dx == 'AD_DS'){
  cur_vis_sample_dx <- 'DSAD'
}

# get ST coords for this sample, and split seurat object by cell-type
cur_coords <- subset(anno_df, vis_sample == cur_vis_sample)
cur_seurat <- seurat_obj[,seurat_obj$bc %in% cur_coords$sc_bc]
cur_seurat <- subset(cur_seurat, Diagnosis == cur_vis_sample_dx)

rownames(cur_coords) <- cur_coords$sc_bc
cur_coords <- cur_coords[as.character(cur_seurat$bc),]
reduc <- as.matrix(cur_coords[,c('celltrek_1', 'celltrek_2')])
rownames(reduc) <- colnames(cur_seurat)

cur_seurat@reductions$spatial <- CreateDimReducObject(
  embeddings = reduc
)

for(cur_mod in mods){
  cur_color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique

  p <- FeatureEmbedding(
      cur_seurat,
      features = cur_mod,
      facet_by = 'coarse_groups',
      reduction = 'spatial',
      point_size=0.5,
      plot_min = 0,
      plot_max = 'q99',
      colfunc=colorRampPalette(c('grey90', cur_color), bias=2),
      rev_colors=FALSE,
      dpi=400
    ) 


  pdf(paste0(fig_dir, 'snRNA_spatial_featureplots/', cur_mod, '_', cur_vis_sample, '_Featureplot.pdf'), width=8, height=4)
  print(p)
  dev.off()
}



seurat_obj@meta.data <- seurat_obj@meta.data[,!(colnames(seurat_obj@meta.data) %in% colnames(MEs))]


```


Human ST spatial featureplots 

```{r eval=FALSE}

MEs <- GetMEs(seurat_human)
meta <- seurat_human@meta.data
seurat_human@meta.data <- cbind(meta, MEs)
mods <- colnames(MEs)
mods <- paste0('M', 1:ncol(MEs))

# representative samples
human_rep_samples <- c(
  'Dec_13_2021_Human5', 'Dec_20_2021_Human1',
  'Dec_13_2021_Human6', 'Oct_2021_6',
  'Dec_13_2021_Human3', 'Dec_13_2021_Human7',
  'Nov_24_2021_VisiumHuman_12', 'Dec_13_2021_Human8'
)


plot_list <- list()
for(cur_mod in mods){
  print(cur_mod)
  cur_color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique

  p <- SampleFeaturePlot(
    seurat_human,
    feature=cur_mod,
    sample_labels = c("Sex", 'Diagnosis'),
    sample_col = "Sample",
    samples_to_plot = human_rep_samples,
    ncol = 8,
    raster=TRUE,
    plot_max = 'q99',
    plot_min = 0,
    #colfunc = inferno,
    #rev_colors=TRUE,
    colfunc=colorRampPalette(c('grey90', cur_color), bias=2),
    rev_colors=FALSE,
    dpi=600,
    combine=FALSE
  )

  p <- lapply(p, function(x){
    x <- x + theme(
      plot.title = element_text(face='bold', size=20, vjust=-1),
      plot.margin = margin(0,0,0,0)
      # plot.margin = margin(-1,-1,-1,-1)
    )
    if(cur_mod != mods[1]){
      x <- x + theme(
        plot.title=element_blank()
      )
    }
    x
  })

  p[[1]] <- p[[1]] + ylab(cur_mod) + theme(
    axis.title.y = element_text(face='bold', size=20, color = cur_color)
  )

  plot_list <- c(plot_list, p)

}

seurat_human@meta.data <- seurat_human@meta.data[,!(colnames(seurat_human@meta.data) %in% colnames(MEs))]


patch <- wrap_plots(plot_list, ncol=8, widths=1, heights=1) + plot_layout(guides='collect')

pdf(paste0(fig_dir, 'test_human_metamodule_ST_Featureplot.pdf'), width=14, height=18)
#  print(p)
print(patch)
dev.off()

```


Project the meta-modules into the mouse ST dataset 

```{r eval=FALSE}

# load the 5x dataset
seurat_5x <- readRDS(file="~/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/hdWGCNA/data/5xFAD_seurat_hdWGCNA.rds")
seurat_5x$annotation <- droplevels(seurat_5x$annotation)

# subset to just contain the cortex + WM
seurat_5x <- subset(seurat_5x, annotation %in% c('ctx-deep-layers', 'ctx-upper-layers', 'WM1', 'WM2'))


# load mouse <-> human gene name table:
hg38_mm10_genes <- read.table(
  "/dfs7/swaruplab/smorabit/resources/hg38_mm10_orthologs_2021.txt",
  sep='\t',
  header=TRUE
)
colnames(hg38_mm10_genes) <-c('hg38_id', 'mm10_id', 'mm10_name', 'hg38_name')
hg38_mm10_genes <- dplyr::select(hg38_mm10_genes, c(hg38_name, mm10_name, hg38_id, mm10_id))

hg38_mm10_genes <- subset(hg38_mm10_genes, mm10_name != '' & hg38_name != '')

# need to make sure that there's only one entry for each gene in hg38_mm10_genes
mm10_genes <- unique(hg38_mm10_genes$mm10_name)
hg38_genes <- unique(hg38_mm10_genes$hg38_name)
hg38_mm10_genes <- hg38_mm10_genes[match(mm10_genes, hg38_mm10_genes$mm10_name),]


seurat_5x <- ProjectModules(
  seurat_5x,
  seurat_ref = seurat_human,
  gene_mapping=hg38_mm10_genes,
  genome1_col="hg38_name", # genome of reference data
  genome2_col="mm10_name", # genome of query data
  group.by.vars="seqbatch",
  wgcna_name_proj="consensus_projected",
  wgcna_name = "consensus",
  verbose=FALSE
)




################################################################################
# ME violin plots:
################################################################################


library(ggpubr)

MEs <- GetMEs(seurat_5x)
modules <- GetModules(seurat_5x)
mods <- levels(modules$module)
mods <- mods[mods!='grey']
MEs <- MEs[,mods]

meta <- seurat_5x@meta.data
seurat_5x@meta.data <- cbind(meta, MEs)


cp_5x <- c("4 months" = "#E8C0FC", "6 months" = "#B57DD1",
        "8 months"= "#8C64A1", "12 months" = "#643D78")

cp_WT <- c("4 months" = "#BDFFDC", "6 months" = "#84DBB5",
          "8 months"= "#45BA7C", "12 months" = "#14703F")


color_df <- data.frame(
  colour = as.character(mouse_cp),
  group = names(mouse_cp)
) %>% subset(group != 'unknown')

p1 <- custom_vln(
    subset(seurat_5x, Condition == '5X'),
    features = mods,
    group.by = 'annotation',
    add_boxplot=FALSE,
    split.by = 'Age',
    split_colors= as.character(cp_5x),
    add_colorbar=TRUE,
    plot_ymin = NA,
    comparisons=NA,
    group_color_df = color_df
  )
  


p2 <- custom_vln(
    subset(seurat_5x, Condition == 'WT'),
    features = mods,
    group.by = 'annotation',
    add_boxplot=FALSE,
    split.by = 'Age',
    split_colors= as.character(cp_WT),
    add_colorbar=TRUE,
    plot_ymin = NA,
    comparisons=NA,
    group_color_df = color_df
  )

pdf(paste0(fig_dir, 'projected_ME_vln_stack_5x.pdf'), width=7, height=7)
p1 | p2 + plot_layout(guides='collect')
dev.off()


seurat_5x@meta.data <- meta


################################################################################
# selected ME violin plots:
################################################################################


library(ggpubr)

MEs <- GetMEs(seurat_5x)
modules <- GetModules(seurat_5x)
mods <- levels(modules$module)
mods <- mods[mods!='grey']
MEs <- MEs[,mods]

meta <- seurat_5x@meta.data
seurat_5x@meta.data <- cbind(meta, MEs)


cp_5x <- c("4 months" = "#E8C0FC", "6 months" = "#B57DD1",
        "8 months"= "#8C64A1", "12 months" = "#643D78")

cp_WT <- c("4 months" = "#BDFFDC", "6 months" = "#84DBB5",
          "8 months"= "#45BA7C", "12 months" = "#14703F")


color_df <- data.frame(
  colour = as.character(mouse_cp),
  group = names(mouse_cp)
) %>% subset(group != 'unknown')

select_mods <- c('M1', 'M6', 'M11')

p1 <- custom_vln(
    subset(seurat_5x, Condition == '5X'),
    features = select_mods,
    group.by = 'annotation',
    add_boxplot=FALSE,
    split.by = 'Age',
    split_colors= as.character(cp_5x),
    add_colorbar=TRUE,
    plot_ymin = NA,
    comparisons=NA,
    group_color_df = color_df
  )


p2 <- custom_vln(
    subset(seurat_5x, Condition == 'WT'),
    features = select_mods,
    group.by = 'annotation',
    add_boxplot=FALSE,
    split.by = 'Age',
    split_colors= as.character(cp_WT),
    add_colorbar=TRUE,
    plot_ymin = NA,
    comparisons=NA,
    group_color_df = color_df
  )

pdf(paste0(fig_dir, 'projected_ME_vln_stack_5x_selected.pdf'), width=8, height=3)
p1 | p2 + plot_layout(guides='collect')
dev.off()

seurat_5x@meta.data <- meta


################################################################################
# ME featureplot
################################################################################



#############################################
# plot the MEs just for representative samples
##############################################


# representative samples ordered by age
rep_samples_5x <- c('675', '421', '784', '314')
rep_samples_WT <- c('242', '721', '343', '313')
rep_samples <- c(rep_samples_5x, rep_samples_WT)


MEs <- GetMEs(seurat_5x)
modules <- GetModules(seurat_5x)
mods <- levels(modules$module)
mods <- mods[mods!='grey']
MEs <- MEs[,mods]

meta <- seurat_5x@meta.data
seurat_5x@meta.data <- cbind(meta, MEs)

plot_list <- list()
for(cur_mod in mods){
  print(cur_mod)
  cur_color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique

  p <- SampleFeaturePlot(
    seurat_5x,
    feature=cur_mod,
    sample_labels = c("Condition", 'Age'),
    sample_col = "SAMPLE",
    samples_to_plot = rep_samples,
    ncol = 10,
    raster=TRUE,
    plot_max = 'q99',
    plot_min = 0,
    #colfunc = inferno,
    #rev_colors=TRUE,
    colfunc=colorRampPalette(c('grey90', cur_color), bias=2),
    rev_colors=FALSE,
    dpi=600,
    combine=FALSE
  )

  p <- lapply(p, function(x){
    x <- x + theme(
      plot.title = element_text(face='bold', size=20, vjust=-1),
      plot.margin = margin(0,0,0,0)
      # plot.margin = margin(-1,-1,-1,-1)
    )
    if(cur_mod != mods[1]){
      x <- x + theme(
        plot.title=element_blank()
      )
    }
    x
  })

  p[[1]] <- p[[1]] + ylab(cur_mod) + theme(
    axis.title.y = element_text(face='bold', size=20, color = cur_color)
  )

  plot_list <- c(plot_list, p)

}

patch <- wrap_plots(plot_list, ncol=8, widths=1, heights=1) + plot_layout(guides='collect')

pdf(paste0(fig_dir, 'projected_ME_featureplots_representative.pdf'), width=14, height=18)
#  print(p)
print(patch)
dev.off()

seurat_5x@meta.data <- meta

#################################################################################
# MT correlation
################################################################################

# selected traits for MT correlation:
seurat_5x$Sex <- factor(as.character(seurat_5x$Sex), levels=c('M', 'F'))
traits <- c('Age', 'Sex', 'DAA_UCell', 'DAM_UCell', 'DOL_UCell', 'PIGs_UCell')

seurat_5x$annotation <- droplevels(seurat_5x$annotation)

seurat_5x <- ModuleTraitCorrelation(
  seurat_5x,
  traits = traits,
  group.by = 'annotation',
  subset_by = 'Condition',
  subset_groups = '5X'
)

p <- PlotModuleTraitCorrelation(
  seurat_5x,
  label = 'fdr',
  label_symbol = 'stars',
  text_size = 2,
  text_digits = 2,
  #trait_text_size = 5,
  text_color = 'black',
  high_color = 'red',
  mid_color = 'white',
  low_color = 'blue',
  plot_max = 0.5,
  #group_order = levels(seurat_5x$annotation),
  combine=TRUE
)

pdf(paste0(fig_dir, 'MT_correlation_5x.pdf'), width=5, height=10)
p
dev.off()



# correlation in WT mice:
seurat_5x <- ModuleTraitCorrelation(
  seurat_5x,
  traits = traits,
  group.by = 'annotation',
  subset_by = 'Condition',
  subset_groups = 'WT'
)

p <- PlotModuleTraitCorrelation(
  seurat_5x,
  label = 'fdr',
  label_symbol = 'stars',
  text_size = 2,
  text_digits = 2,
  #trait_text_size = 5,
  text_color = 'black',
  high_color = 'darkorchid3',
  mid_color = 'white',
  low_color = 'seagreen',
  plot_max = 0.5,
  #group_order = levels(seurat_5x$annotation),
  combine=TRUE
)

pdf(paste0(fig_dir, 'MT_correlation_WT.pdf'), width=5, height=10)
p
dev.off()

################################################################################
# Module Preservation
################################################################################


# need to add a grey module 
modules <- GetModules(seurat_human, 'consensus')
modules_orig <- GetModules(seurat_human, 'consensus')

# add a grey module using genes that were not included for clustering:
wgcna_genes <- GetWGCNAGenes(seurat_human, 'consensus')
not_wgcna_genes <- rownames(seurat_human)[!(rownames(seurat_human) %in% wgcna_genes)]
grey_genes <- sample(not_wgcna_genes, 1000)
grey_genes_mouse <- subset(hg38_mm10_genes, hg38_name %in% grey_genes) %>% .$mm10_name
grey_genes_mouse <- grey_genes_mouse[grey_genes_mouse %in% rownames(seurat_5x)]
grey_mod <- data.frame(
  gene_name = grey_genes,
  module = 'grey',
  color = 'grey'
)
grey_mod[,colnames(modules[4:ncol(modules)])] <- 0
modules <- rbind(modules, grey_mod)
seurat_human <- SetModules(seurat_human, modules, 'consensus') 


# 


genes_use <- subset(hg38_mm10_genes, mm10_name %in% c(GetWGCNAGenes(seurat_5x), grey_genes_mouse)) %>% .$hg38_name %>% unique
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')

datExpr <- hdWGCNA:::to_pseudobulk(
    seurat_human,
    replicate_col = 'Sample',
    cell_type_col = 'annotation',
    label_col = 'Diagnosis'
)
datExpr <- Reduce(cbind, lapply(names(datExpr), function(x){
    cur <- datExpr[[x]]
    colnames(cur) <- paste0(x, '_', colnames(cur))
    cur
}))

datExpr <- t(datExpr[genes_use,])

# compute CPM:
cpm <- t(apply(datExpr, 1, function(x){
    y <- (x) / sum(x) * 1000000
    log2(y + 1)
}))
cpm_human <- cpm


# mouse pseudobulk
genes_use <- c(GetWGCNAGenes(seurat_5x), grey_genes_mouse) %>% unique
datExpr <- hdWGCNA:::to_pseudobulk(
    seurat_5x,
    replicate_col = 'SAMPLE',
    cell_type_col = 'annotation',
    label_col = 'Condition'
)
datExpr <- Reduce(cbind, lapply(names(datExpr), function(x){
    cur <- datExpr[[x]]
    colnames(cur) <- paste0(x, '_', colnames(cur))
    cur
}))

datExpr <- t(datExpr[genes_use,])

# compute CPM:
cpm <- t(apply(datExpr, 1, function(x){
    y <- (x) / sum(x) * 1000000
    log2(y + 1)
}))
cpm_mouse <- cpm


# check samples / genes 
check_mouse <- WGCNA::goodSamplesGenes(cpm_mouse)
check_human <- WGCNA::goodSamplesGenes(cpm_human)
genes_use <- intersect(
  colnames(cpm_mouse)[check_mouse$goodGenes], 
  subset(hg38_mm10_genes, hg38_name %in% colnames(cpm_human)[check_human$goodGenes]) %>% .$mm10_name
)
genes_use_human <-   subset(hg38_mm10_genes, mm10_name %in% genes_use) %>% .$hg38_name

modules <- modules[genes_use_human,]
seurat_human <- SetModules(seurat_human, modules, 'consensus') 

cpm_human <- cpm_human[,genes_use_human]
cpm_mouse <- cpm_mouse[,genes_use]
seurat_human@misc[['consensus']]$datExpr <- cpm_human
seurat_5x@misc[['consensus_projected']]$datExpr <- cpm_mouse


# run module preservation function
seurat_5x <- ModulePreservation(
  seurat_5x,
  seurat_ref = seurat_human,
  name="consensus_projected",
  wgcna_name = 'consensus_projected',
  wgcna_name_ref = 'consensus',
  verbose=3,
  gene_mapping=hg38_mm10_genes,
  genome1_col="hg38_name", # genome of reference data
  genome2_col="mm10_name", # genome of query data
  n_permutations=100
)

saveRDS(seurat_5x, file=paste0(data_dir, "5xFAD_seurat_hdWGCNA_projected.rds"))
seurat_5x <- readRDS(file=paste0(data_dir, "5xFAD_seurat_hdWGCNA_projected.rds"))

# getthe module preservation table
mod_pres <- GetModulePreservation(seurat_5x, "consensus_projected")$Z
obs_df <- GetModulePreservation(seurat_5x, "consensus_projected")$obs


plot_list <- PlotModulePreservation(
  seurat_5x,
  name="consensus_projected",
  statistics = "summary"
)


pdf(paste0(fig_dir, 'module_preservation_5x.pdf'), width=10, height=5)
wrap_plots(plot_list, ncol=2)
dev.off()


################################################################################
# Plot as a lollipop
################################################################################

plot_df <- mod_pres[,c('moduleSize', 'Zsummary.pres')]
plot_df$module <- rownames(plot_df)
plot_df <- plot_df %>%subset(module != 'gold')
plot_df <-plot_df %>% dplyr::arrange(Zsummary.pres, descending=TRUE)
plot_df$module <- factor(as.character(plot_df$module), levels=as.character(plot_df$module))


mod_colors <- dplyr::select(modules, c(module, color)) %>%
  distinct
mod_cp <- mod_colors$color; names(mod_cp) <- mod_colors$module

p <- plot_df %>%
  ggplot(aes(y=module, x=Zsummary.pres, size= moduleSize, color=module, fill=module)) + 
  geom_vline(xintercept=2, linewidth=0.25) + 
  geom_vline(xintercept=10, linewidth=0.25) +
  #  geom_rect(
  #         data = plot_df[1,],
  #         aes(xmin=-Inf, ymax=Inf, ymin=-Inf, xmax=2), fill='grey75', alpha=0.8, color=NA) +
  #   geom_rect(
  #     data=plot_df[1,],
  #     aes(ymin=-Inf, ymax=Inf, xmin=2, xmax=10), fill='grey92', alpha=0.8, color=NA) + 
  geom_segment(aes(y=module, yend=module, x=0, xend=Zsummary.pres), size=0.5, alpha=0.5) +
  geom_point(shape=21, color='black') +
  scale_color_manual(values=mod_cp, guide='none') +
  scale_fill_manual(values=mod_cp, guide='none') +
  ylab('') + xlab("Z-summary preservation") +
  # scale_x_continuous(breaks = c(0, 1, 2,3)) +
 #labs(size='Size\nintersection') 

  theme(
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.title = element_text(hjust=0.5, face='plain')
  ) + RotatedAxis()

#p <- p + ggforestplot::geom_stripes(aes(y=module), inherit.aes=FALSE, data=overlap_df) 

pdf(paste0(fig_dir, 'module_preservation_lollipop.pdf'), width=4, height=5)
p
dev.off()


pdf(paste0(fig_dir, 'module_preservation_lollipop_smaller.pdf'), width=3, height=4)
p
dev.off()



###################################################################
# preservation in a WT
###################################################################

modules_full <- modules
cur_cpm <-cpm_mouse[grepl('WT', rownames(cpm_mouse)),]


# check samples / genes 
check_mouse <- WGCNA::goodSamplesGenes(cur_cpm)
genes_use <- intersect(
  colnames(cur_cpm)[check_mouse$goodGenes], 
  subset(hg38_mm10_genes, hg38_name %in% colnames(cpm_human)[check_human$goodGenes]) %>% .$mm10_name
)
genes_use_human <-   subset(hg38_mm10_genes, mm10_name %in% genes_use) %>% .$hg38_name

modules <- modules_full[genes_use_human,]
seurat_human <- SetModules(seurat_human, modules, 'consensus') 

cur_cpm_human <- cpm_human[,genes_use_human]
cur_cpm_mouse <- cur_cpm[,genes_use]
seurat_human@misc[['consensus']]$datExpr <- cur_cpm_human
seurat_5x@misc[['consensus_projected']]$datExpr <- cur_cpm_mouse


# run module preservation function
seurat_5x <- ModulePreservation(
  seurat_5x,
  seurat_ref = seurat_human,
  name="consensus_projected",
  wgcna_name = 'consensus_projected',
  wgcna_name_ref = 'consensus',
  verbose=3,
  gene_mapping=hg38_mm10_genes,
  genome1_col="hg38_name", # genome of reference data
  genome2_col="mm10_name", # genome of query data
  n_permutations=100
)


# getthe module preservation table
mod_pres_wt <- GetModulePreservation(seurat_5x, "consensus_projected")$Z
obs_df_wt <- GetModulePreservation(seurat_5x, "consensus_projected")$obs


plot_list <- PlotModulePreservation(
  seurat_5x,
  name="consensus_projected",
  statistics = "summary"
)


pdf(paste0(fig_dir, 'module_preservation_wt.pdf'), width=10, height=5)
wrap_plots(plot_list, ncol=2)
dev.off()




```


Correlation of Mouse amyloid scores with the human hdWGCNA modules 

```{r eval=FALSE}

################################################################################
# MT correlation
################################################################################

seurat_5x <- SetActiveWGCNA(seurat_5x, 'consensus_projected')

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amyloglo_Score_log)) %>%
  dplyr::filter(Mean != 0) %>% .$SAMPLE %>% as.character

seurat_5x$amyloid_sample <- seurat_5x$SAMPLE %in% amyloid_samples


# correlation of amyloid scores:
traits <- c('Amyloglo_Score_log', 'Amyloglo_Score_log_Gi', 'OC_Score_log', 'OC_Score_log_Gi')



seurat_5x$annotation <- droplevels(seurat_5x$annotation)


seurat_5x$Condition_Age <- paste0(
    as.character(seurat_5x$Condition), '_',
    as.character(seurat_5x$Age)
)

groups <- unique(seurat_5x$Condition_Age)
plot_df <- data.frame()
for(cur_group in groups){

    print(cur_group)

    seurat_5x <- ModuleTraitCorrelation(
        seurat_5x,
        traits = traits,
        group.by = 'annotation',
        subset_by = 'Condition_Age',
        subset_groups = cur_group
    )

    mt_cor <- GetModuleTraitCorrelation(seurat_5x)
    names(mt_cor)

    cor_df <- do.call(rbind, lapply(names(mt_cor$cor), function(x){
        cur_cor <- mt_cor$cor[[x]]
        cur_fdr <- mt_cor$fdr[[x]]
        cur_cor_df <- reshape2::melt(cur_cor)
        cur_fdr_df <- reshape2::melt(cur_fdr)
        cur_cor_df$fdr <- cur_fdr_df$value
        cur_cor_df$group <- x
        cur_cor_df
    }))
    cor_df %<>% dplyr::rename(c(trait = Var1, module = Var2, cor=value))

    # add age and genotype info
    tmp<- strsplit(cur_group, '_')[[1]]
    cor_df$Condition <- tmp[1]
    cor_df$Age <- tmp[2]

    plot_df <- rbind(plot_df, cor_df)

}

write.csv(plot_df, file=paste0(data_dir, 'amyloid_module_correlation_5x_metamodules.csv'), quote=FALSE)




plot_df <- read.csv(file=paste0(data_dir, 'amyloid_module_correlation_5x_metamodules.csv'))

plot_df <- subset(plot_df, 
  group != 'all_cells' & 
  trait %in% c('Amyloglo_Score_log_Gi', 'OC_Score_log_Gi') &
  !is.na(cor)
)


plot_df %>% subset(fdr < 0.05 & module == 'M11') %>% arrange(-cor) %>% head(25)
plot_df %>% subset(fdr < 0.05) %>% arrange(-cor) %>% head(25)



modules <- GetModules(seurat_5x)
mod_levels <- levels(modules$module)

plot_df$module <- factor(as.character(plot_df$module), levels=mod_levels)

# grouping variables:
plot_df$id <- paste(plot_df$group, plot_df$Condition, plot_df$Age, sep='_')

# get wgcna info 
data.features <- GetMEs(seurat_5x) %>% as.data.frame()
data.features$id <- paste(seurat_5x$annotation, seurat_5x$Condition, seurat_5x$Age, sep='_')


#---------------------------------------------------------------------
# code from Seurat DotPlot
#---------------------------------------------------------------------

scale <- TRUE
col.min <- -2.5
col.max <- 2.5
dot.min <- 0

data.plot <- lapply(
    X = unique(x = data.features$id),
    FUN = function(ident) {
      data.use <- data.features[data.features$id == ident, 1:(ncol(x = data.features) - 1), drop = FALSE]
      avg.exp <- apply(
        X = data.use,
        MARGIN = 2,
        FUN = function(x) {
          return(mean(x = expm1(x = x)))
        }
      )
      pct.exp <- apply(X = data.use, MARGIN = 2, FUN = PercentAbove, threshold = 0)
      return(list(avg.exp = avg.exp, pct.exp = pct.exp))
    }
  )
names(x = data.plot) <- unique(x = data.features$id)
id.levels <- levels(seurat_obj$cell_type)
data.plot <- lapply(
    X = names(x = data.plot),
    FUN = function(x) {
      data.use <- as.data.frame(x = data.plot[[x]])
      data.use$features.plot <- rownames(x = data.use)
      data.use$id <- x
      return(data.use)
    }
  )
  data.plot <- do.call(what = 'rbind', args = data.plot)
  if (!is.null(x = id.levels)) {
    data.plot$id <- factor(x = data.plot$id, levels = id.levels)
  }
  ngroup <- length(x = levels(x = data.plot$id))

  avg.exp.scaled <- sapply(
    X = unique(x = data.plot$features.plot),
    FUN = function(x) {
      data.use <- data.plot[data.plot$features.plot == x, 'avg.exp']
      if (scale) {
        data.use <- scale(x = log1p(data.use))
        data.use <- MinMax(data = data.use, min = col.min, max = col.max)
      } else {
        data.use <- log1p(x = data.use)
      }
      return(data.use)
    }
  )
  avg.exp.scaled <- as.vector(x = t(x = avg.exp.scaled))
  data.plot$avg.exp.scaled <- avg.exp.scaled
  
  data.plot$pct.exp[data.plot$pct.exp < dot.min] <- NA
  data.plot$pct.exp <- data.plot$pct.exp * 100

# match with the other table
data.plot <- subset(data.plot, id %in% plot_df$id)

# merge the DFs
data.plot$ix <- paste0(data.plot$features.plot, '_', data.plot$id)
plot_df$ix <- paste0(plot_df$module, '_', plot_df$id)

ix <- match(plot_df$ix, data.plot$ix, )
all.equal(plot_df$ix, data.plot$ix[ix])

plot_df$avg.exp <- data.plot$avg.exp[ix]
plot_df$avg.exp.scaled <- data.plot$avg.exp.scaled[ix]
plot_df$pct.exp <- data.plot$pct.exp[ix]



plot_df$Significance <- ifelse(plot_df$fdr > 0.05, 'x', NA)

plot_range <- plot_df %>% .$cor %>% range
size_range <- plot_df %>% .$pct.exp %>% range


# loop through Studies & diagnosis
ages <- c('12mo', '8mo', '6mo', '4mo')
genotypes <- unique(as.character(seurat_5x$Condition))
plot_list <- list()
for(cur_age in ages){
    for(cur_geno in genotypes){
        

    cur_df <- plot_df %>% subset(Condition == cur_geno & Age == cur_age)
    if(nrow(cur_df) == 0){next}    

        cur_df <- plot_df %>% subset(Condition == cur_geno & Age == cur_age)
        p <-cur_df %>%
            ggplot(aes(x = module, y = group, color=cor, size=pct.exp)) + 
            geom_point() + 
            scale_color_gradient2(high='red', mid='white', low='blue', limits=plot_range) + 
            scale_size_continuous(limits=size_range) +
            geom_text(aes(label=Significance), color='black', size=3) +
            coord_equal() +
            RotatedAxis() +
            ylab(paste0(cur_age, ', ', cur_geno)) + 
            theme(
                axis.line.x = element_blank(),
                axis.line.y = element_blank(),
                panel.border = element_rect(linewidth=1, color='black', fill=NA),
                axis.title.x = element_blank(),
                axis.title.y = element_text(angle=0, vjust=0.5),
                plot.title = element_text(hjust=0.5),
                panel.grid = element_line(size=0.25, color='lightgrey'),
                plot.margin = margin(c(0,0,0,0)),
                
            ) 

        if(cur_age != '12mo'){
            p <- p + theme(
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
            )
        }       

     patch <- p + facet_wrap(~trait, ncol=2)

     plot_list[[paste0(cur_age, ', ', cur_geno)]] <- patch
    }
}


pdf(paste0(fig_dir, 'MT_cor_heatmap_consensus_visium_dotplot_amylo_5x.pdf'), width=12, height=6)
 wrap_plots(rev(plot_list), ncol=1) + plot_layout(guides='collect')
#wrap_plots(p)
dev.off()


```










Below: have not run yet for the meta-modules







Compute the module-module distances like we did for the ST data but using the SC MEs.

```{r eval=FALSE}


########################################################################################
# Correlation of module eigengenes 
########################################################################################

# get the modules
MEs <- GetMEs(seurat_obj)

# correlate MEs in each of the brain regions
groups <- levels(seurat_obj$cell_identity)
mm_cor <- Reduce(rbind, lapply(groups, function(x){
    cur_spots <- seurat_obj@meta.data %>% subset(cell_identity == x) %>% rownames
    z <- qlcMatrix::corSparse(
        X = as.matrix(MEs[cur_spots,]),
        Y = as.matrix(MEs[cur_spots,])
    ) 
    colnames(z) <- colnames(MEs)
    rownames(z) <- colnames(MEs)
    z <-  reshape2::melt(z)
    z$group <- x
    z$pair <- paste0(z$Var1, '_', z$Var2)
    z
}))
head(mm_cor)

# take the component-wise maximum correlation across each of the regions
test <- mm_cor %>% 
    group_by(pair) %>% 
    mutate(val = max(value)) %>% 
    ungroup %>% 
    select(c(Var1, Var2, val)) %>% 
    distinct %>% 
    spread(Var1, val) 
test <- as.matrix(test[,-1])
rownames(test) <- colnames(test)

# re-order:
test <- test[levels(modules$module), levels(modules$module)]
E <- test

# set up Jaccard matrix:
J <- overlap_df %>% select(c(m1,m2,Jaccard)) %>% spread(m1,Jaccard)
J <- as.matrix(J[,-1])
rownames(J) <- colnames(J)
J <- J[rownames(E), colnames(E)]


# compute the dissimilarity matrix:
D = 1 - ((E + (3*J))/4)
D_sc <- D

################################
# Plot D_st vs D_sc
################################

plot_df <- reshape2::melt(D_sc)
plot_df2 <- reshape2::melt(D_st)
plot_df$val_st <- plot_df2$value

p <- plot_df %>%
    subset(Var1 != Var2) %>%
     ggplot(aes(x=value, y=val_st)) +
     geom_hex() +
     scale_fill_viridis(trans='log', breaks=c(7,50,400)) +
    # geom_point() + 
     # geom_smooth(method='lm') + 
     stat_cor() + 
     xlab('snRNA-seq module dissimilarity') +
     ylab('ST module dissimilarity') + 
     labs(fill = 'log(Density)')  +
     coord_equal() +
    theme(
        axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        panel.border = element_rect(linewidth=1, fill=NA, color='black')
    )



pdf(paste0(fig_dir, 'sc_st_module_distances.pdf'), width=6, height=4)
p
dev.off()




```

 consensus network analysis? 

```{r eval=FALSE}


library(magrittr)

genes_use <- GetWGCNAGenes(seurat_human)
groups <- c('L1', 'L2-3', 'L3-4', 'L3-4-5', 'L5-6', 'L6b', 'WM')

datExpr <- hdWGCNA:::to_pseudobulk(
    seurat_human,
    replicate_col = 'Sample',
    cell_type_col = 'annotation',
    label_col = 'Diagnosis'
)
datExpr <- Reduce(cbind, lapply(names(datExpr), function(x){
    cur <- datExpr[[x]]
    colnames(cur) <- paste0(x, '_', colnames(cur))
    cur
}))

datExpr <- t(datExpr[GetWGCNAGenes(seurat_human),])

# compute CPM:
cpm <- t(apply(datExpr, 1, function(x){
    y <- (x) / sum(x) * 1000000
    log2(y + 1)
}))

datExpr_list <- lapply(groups, function(x){
    cpm[grepl(x, rownames(cpm)),]
})
names(datExpr_list) <- groups

gene_names <- GetWGCNAGenes(seurat_human)
seurat_human <- SetupForWGCNA(
    seurat_human,
    features = gene_names,
    wgcna_name = 'consensus'
)


multiExpr <- WGCNA::list2multiData(datExpr_list)

genes_use <- WGCNA::goodGenesMS(multiExpr)
gene_names <- gene_names[genes_use]

# subset the multiExpr by the good genes::
datExpr_list <- lapply(1:length(multiExpr), function(i){
multiExpr[[i]]$data[,genes_use]
})
multiExpr <- WGCNA::list2multiData(datExpr_list)
names(multiExpr) <- groups

# update the WGCNA gene list:
seurat_human <- SetWGCNAGenes(seurat_human, gene_names, 'consensus')

# set the multiExpr in the Seurat object
seurat_human@misc[['consensus']]$multiExpr <- multiExpr


seurat_human <- TestSoftPowersConsensus(seurat_human)


seurat_human <- ConstructNetwork(
    seurat_human,
    minModuleSize=50,
    consensus=TRUE,
    #detectCutHeight=0.995,
    mergeCutHeight=0.3,
    tom_name='consensus',
    overwrite_tom=TRUE
)

length(unique(GetModules(seurat_human)$module))

pdf(paste0(fig_dir, "pb_consensus_dendro_cut3.pdf"),height=3, width=6)
print(PlotDendrogram(seurat_human, main='hdWGCNA Dendrogram'))
dev.off()

# run WGCNA separately for each group
for(cur_group in groups){
    print(cur_group)
    wgcna_name <- cur_group
    seurat_human <- SetupForWGCNA(
        seurat_human,
        features = genes_use,
        wgcna_name = cur_group
    )

    # just take from one brain layer 
    cur_cpm <- cpm[grepl(cur_group, rownames(cpm)),]
    seurat_human@misc[[wgcna_name]]$datExpr <- cur_cpm

    seurat_human <- TestSoftPowers(seurat_human)
    plot_list <- PlotSoftPowers(seurat_human)

    # assemble with patchwork
    pdf(paste0(fig_dir, cur_group, '_softpower.pdf'), width=12, height=8)
    print(wrap_plots(plot_list, ncol=2))
    dev.off()

    # construct wgcna network:
    seurat_human <- ConstructNetwork(
        seurat_human,
        #PamStage = T,
            minModuleSize=50,
            #detectCutHeight=0.995,
            mergeCutHeight=0.1,
        tom_name=cur_group,
        overwrite_tom=TRUE
    )

    # plot the dendrogram
    pdf(paste0(fig_dir, cur_group, "_pb_dendro.pdf"),height=3, width=6)
    print(PlotDendrogram(seurat_human, main='hdWGCNA Dendrogram'))
    dev.off()

    seurat_human <- ModuleEigengenes(
        seurat_human,
        group.by.vars="Sample",
    )

    seurat_human <- ModuleConnectivity(
        seurat_human,
        reassign_modules=FALSE
    )

}

saveRDS(seurat_human, file=paste0(data_dir, "ADDS_seurat_processed_annotated_hdWGCNA_all.rds"))


```



For peer review response:

Prioritize / sort the 166 co-expression modules and 15 meta-modules by their
patterns 

Direct quote:

"Regarding the multi-scale co-expression analysis, there are so many modules differentially 
regulated between AD and control. Is it possible to rank order these observations according to 
their strength and or temporal observations? Can we maybe infer putative causality?"


Maybe I could just do a bunch of different overlap / set operations to find the groups 
that I am interested in ...?

Pattern 1: Not DE in any group 
Pattern 2: DE in one condition 
Pattern 3: Opposite directions -- up in early but down in late etc

```{r eval=FALSE}

genes_use <- GetWGCNAGenes(seurat_human)

all_mods <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(module, color)) %>% 
    distinct()

    
    cur_mods 

}))
all_mods <- all_mods %>% subset(module != 'grey') %>% distinct
mod_colors <- all_mods$color; names(mod_colors) <- all_mods$module



# set a threshold for log2FC 
logfc_thresh <- 0.25

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv')) %>% 
    subset(wgcna == 'consensus')

DMEs <- read.csv(file=paste0(data_dir, 'differential_MEs_humanST.csv'))
DMEs <- DMEs %>% subset(annotation == wgcna ) 


#-------------------------------------------------------------------------#
# Which modules are not DE (below FC thresh) in any group
#-------------------------------------------------------------------------#

not_de <- subset(DMEs, p_val_adj > 0.05 | abs(avg_log2FC) < logfc_thresh) 
not_de <- names(which(table(not_de$module) == 3)) 

#-------------------------------------------------------------------------#
# Get a list of modules that are significantly up / down in each condition
#-------------------------------------------------------------------------#

# early 
up_eAD <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'earlyAD' & avg_log2FC > logfc_thresh) %>% .$module
down_eAD <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'earlyAD' & avg_log2FC < -logfc_thresh) %>% .$module

# AD
up_AD <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'AD' & avg_log2FC > logfc_thresh) %>% .$module
down_AD <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'AD' & avg_log2FC < -logfc_thresh) %>% .$module

# AD/DS
up_ADDS <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'AD_DS' & avg_log2FC > logfc_thresh) %>% .$module
down_ADDS <- subset(DMEs, p_val_adj <= 0.05 & ident.1 == 'AD_DS' & avg_log2FC < -logfc_thresh) %>% .$module

# check that all the modules are accounted for 
length(unique(c(
  not_de, 
  up_eAD, down_eAD,
  up_AD, down_AD, 
  up_ADDS, down_ADDS
)))
# [1] 166

#-------------------------------------------------------------------------#
# Set up overlapping sets of modules
#-------------------------------------------------------------------------#

up_all <- Reduce(intersect, list(up_eAD, up_AD, up_ADDS))
up_any <- Reduce(union, list(up_eAD, up_AD, up_ADDS))

# modules that are up only in one of the groups
up_eAD_only <- setdiff(up_eAD, Reduce(union, c(up_AD, up_ADDS, down_AD, down_ADDS)))
up_AD_only <- setdiff(up_AD, Reduce(union, c(up_eAD, up_ADDS, down_eAD, down_ADDS)))
up_ADDS_only <- setdiff(up_ADDS, Reduce(union, c(up_eAD, up_AD, down_eAD, down_AD)))

down_all <- Reduce(intersect, list(down_eAD, down_AD, down_ADDS))
down_any <- Reduce(union, list(down_eAD, down_AD, down_ADDS))

# modules that are down only in one of the groups
down_eAD_only <- setdiff(down_eAD, Reduce(union, c(up_AD, up_ADDS, down_AD, down_ADDS)))
down_AD_only <- setdiff(down_AD, Reduce(union, c(up_eAD, up_ADDS, down_eAD, down_ADDS)))
down_ADDS_only <- setdiff(down_ADDS, Reduce(union, c(up_eAD, up_AD, down_eAD, down_AD)))

#-------------------------------------------------------------------------#
# Identify conflicting sets of modules
#-------------------------------------------------------------------------#
up_eAD_down_other <- intersect(up_eAD, union(down_AD, down_ADDS))
up_AD_down_other <- intersect(up_AD, union(down_eAD, down_ADDS))
up_ADDS_down_other <- intersect(up_ADDS, union(down_eAD, down_AD))

down_eAD_up_other <- intersect(down_eAD, union(up_AD, up_ADDS))
down_AD_up_other <- intersect(down_AD, union(up_eAD, up_ADDS))
down_ADDS_up_other <- intersect(down_ADDS, union(up_eAD, up_AD))


```

Plot the modules that are not DE 

```{r eval=FALSE}


#-------------------------------------------------------------------------#
# Plot the modules that are up in all
#-------------------------------------------------------------------------#

mod_lists <- list(
  'not DE' = not_de
)

# make a plot -----------------------------------------------------------------
# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range
if(plot_range[1] > 0){plot_range[1] <- 0}

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'AD_DS')
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'earlyAD')

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.5) + 
    geom_point(size=2, alpha=0.5) + 
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
        size=3
      # size=2
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
        size=3
      # size=2
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group)

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_not_de.pdf'), width=3, height=3)
wrap_plots(plot_list, ncol=1)
dev.off()


```

plot modules that are up / down in all 

```{r eval=FALSE}

#-------------------------------------------------------------------------#
# Plot the modules that are up in all
#-------------------------------------------------------------------------#

# which of these are highest in eAD
temp <- subset(DMEs, module %in% up_all) %>% 
  group_by(module) %>% 
  slice_max(order_by=avg_log2FC, n=1)

highest_eAD <- subset(temp, ident.1 == 'earlyAD') %>% .$module
highest_AD <- subset(temp, ident.1 == 'AD') %>% .$module
highest_ADDS <- subset(temp, ident.1 == 'AD_DS') %>% .$module

mod_lists <- list(
  'earlyAD' = highest_eAD,
  'AD' = highest_AD,
  'AD_DS' = highest_ADDS
)

# make a plot -----------------------------------------------------------------
# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range
if(plot_range[1] > 0){plot_range[1] <- 0}



plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # add module sizes:
  plot_df$mod_size <- as.numeric(mod_size[as.character(plot_df$module)])

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_max(order_by=avg_log2FC, n=10) %>% .$module
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75, size=1) + 
    geom_point(alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC >= logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
        size=5
      # size=2
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
        size=5
      # size=2
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_up_all.pdf'), width=9, height=3.5)
wrap_plots(plot_list, ncol=3)
dev.off()


#-------------------------------------------------------------------------#
# Plot the modules that are down in all
#-------------------------------------------------------------------------#

temp <- subset(DMEs, module %in% down_all) %>% 
  group_by(module) %>% 
  slice_min(order_by=avg_log2FC, n=1)

highest_eAD <- subset(temp, ident.1 == 'earlyAD') %>% .$module
highest_AD <- subset(temp, ident.1 == 'AD') %>% .$module
highest_ADDS <- subset(temp, ident.1 == 'AD_DS') %>% .$module

mod_lists <- list(
  'earlyAD' = highest_eAD,
  'AD' = highest_AD,
  'AD_DS' = highest_ADDS
)

# make a plot -----------------------------------------------------------------

# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range
if(plot_range[2] < 0){plot_range[2] <- 0}

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  if(nrow(plot_df) == 0){
    next
  }
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_min(order_by=avg_log2FC, n=10) %>% .$module
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75) + 
    geom_point(size=2, alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC <= logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
        size=5
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
        size=5
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_down_all.pdf'), width=9, height=3.5)
wrap_plots(list(plot_list[[1]], plot_list[[2]], plot_list[[2]]), ncol=3)
dev.off()

```

Plot modules that are up/down in just one group 

```{r eval=FALSE}

#-------------------------------------------------------------------------#
# Plot the trajectories for modules that are up in just one group
#-------------------------------------------------------------------------#

mod_lists <- list(
  'earlyAD' = up_eAD_only,
  'AD' = up_AD_only,
  'AD_DS' = up_ADDS_only
)

# make a plot -----------------------------------------------------------------

# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_max(order_by=avg_log2FC, n=10) %>% .$module
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75) + 
    geom_point(size=2, alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC >= logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_up_exclusive.pdf'), width=9, height=3.5)
wrap_plots(plot_list, ncol=3)
dev.off()



#-------------------------------------------------------------------------#
# Plot the trajectories for modules that are down 
#-------------------------------------------------------------------------#

mod_lists <- list(
  'earlyAD' = down_eAD_only,
  'AD' = down_AD_only ,
  'AD_DS' = down_ADDS_only
)


# make a plot -----------------------------------------------------------------

# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_min(order_by=avg_log2FC, n=10) %>% .$module
 # label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'AD_DS' & module %in% top_mods)
 # label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'earlyAD' & module %in% top_mods)
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75) + 
    geom_point(size=2, alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC <= -logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_down_exclusive.pdf'), width=9, height=3.5)
wrap_plots(plot_list, ncol=3)
dev.off()

```

Plot modules that are conflicting across groups

```{r eval=FALSE}

#-------------------------------------------------------------------------#
# Plot the trajectories for modules that are up in one
#-------------------------------------------------------------------------#

mod_lists <- list(
  'earlyAD' = up_eAD_down_other,
  'AD' = up_AD_down_other,
  'AD_DS' = up_ADDS_down_other
)

# make a plot -----------------------------------------------------------------

# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_max(order_by=avg_log2FC, n=10) %>% .$module
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75) + 
    geom_point(size=2, alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC >= logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_up_conflicting.pdf'), width=9, height=3.5)
wrap_plots(plot_list, ncol=3)
dev.off()



#-------------------------------------------------------------------------#
# Plot the trajectories for modules that are down 
#-------------------------------------------------------------------------#

mod_lists <- list(
  'earlyAD' = down_eAD_up_other,
  'AD' = down_AD_up_other,
  'AD_DS' = down_ADDS_up_other
)


# make a plot -----------------------------------------------------------------

# get the min/max values
plot_range <- subset(DMEs, module %in% unlist(mod_lists)) %>% .$avg_log2FC %>% range

plot_list <- list()
for(cur_group in names(mod_lists)){
  
  cur_mod_list <- mod_lists[[cur_group]]

  plot_df <- subset(DMEs, module %in% cur_mod_list)
  if(nrow(plot_df) == 0){next}
  plot_df$ident.1 <- factor(as.character(plot_df$ident.1), levels=c('earlyAD', 'AD', 'AD_DS'))
  rownames(plot_df) <- 1:nrow(plot_df)

  # get the top mods for labeling
  top_mods <- plot_df %>% subset(ident.1 == cur_group) %>% slice_min(order_by=avg_log2FC, n=10) %>% .$module
 # label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'AD_DS' & module %in% top_mods)
 # label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == 'earlyAD' & module %in% top_mods)
  label_df_right <- plot_df[as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)
  label_df_left <- plot_df[!as.logical(as.numeric(rownames(plot_df)) %% 2),] %>% subset(ident.1 == cur_group & module %in% top_mods)

  p <- ggplot(plot_df, aes(x=ident.1, y = avg_log2FC, color=module, group=module)) + 
    geom_hline(yintercept=0, linetype='dashed', color='black') +
    geom_line(alpha=0.75) + 
    geom_point(size=2, alpha=0.75) + 
    geom_point(
      inherit.aes=FALSE,
      data = subset(plot_df, p_val_adj <= 0.05 & avg_log2FC <= -logfc_thresh),
      aes(x=ident.1, y = avg_log2FC, fill=module, group=module),
      color='black',
      shape=21,
      size=3
    ) +
    scale_color_manual(values=mod_colors) +
    scale_fill_manual(values=mod_colors) +
    ylim(plot_range[1] - 0.1, plot_range[2] + 0.1) +
    geom_text_repel(
        data = label_df_right,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = 0.5,
        direction = "y",
        hjust = "left",
        min.segment.length=0,
        max.overlaps=Inf,
      size=5
      ) +
      geom_text_repel(
        data = label_df_left,
        aes(x=ident.1, y = avg_log2FC , label=module, color=module),
        color='black',
        nudge_x = -0.5,
        direction = "y",
        hjust = "right",
        min.segment.length=0,
        max.overlaps=Inf,
       size=5
      ) +
    theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='white'),
            panel.background = element_rect(fill='grey92'),
            #plot.margin = margin(c(0,0,0,0))
    ) + NoLegend() + 
    ylab(bquote("Avg. log"[2]~"(FC)")) + 
    xlab('') + ggtitle(cur_group) + 
    RotatedAxis()

  plot_list[[cur_group]] <- p

}

pdf(paste0(fig_dir, 'module_lineplot_down_conflicting.pdf'), width=9, height=3.5)
wrap_plots(list(plot_list[[1]], plot_list[[2]], plot_list[[2]]), ncol=3)
dev.off()

```

Try running EnrichR on these sets of modules? 

```{r eval=FALSE}

library(enrichR)
dbs <-c('GO_Biological_Process_2021','GO_Cellular_Component_2021','GO_Molecular_Function_2021', 'WikiPathway_2021_Human', 'KEGG_2021_Human')

genes_use <- GetWGCNAGenes(seurat_human)

all_modules <- Reduce(rbind, lapply(groups, function(x){

    cur_mods <- GetModules(seurat_human, wgcna_name = x)[,1:3]
    cur_mods <- cur_mods[genes_use,]
    cur_mods$gene_name <- genes_use
    cur_mods$module <- ifelse(is.na(cur_mods$module), 'grey' , as.character(cur_mods$module))
    cur_mods$color <- ifelse(is.na(cur_mods$color), 'grey' , as.character(cur_mods$color))
    
    cur_mods <- cur_mods %>% 
    dplyr::select(c(gene_name, module, color))

    cur_mods 

}))
all_modules <- all_modules %>% subset(module != 'grey') 


n_hubs <- 25
all_hubs <- Reduce(rbind, lapply(groups, function(x){
  GetHubGenes(seurat_human, n_hubs, wgcna_name = x)[,1:3]
}))





# list of inputs to enrichr
input_list <- list(
  up_all = subset(all_hubs, module %in% up_all) %>% .$gene_name %>% unique,
  up_eAD_only = subset(all_hubs, module %in% up_eAD_only) %>% .$gene_name %>% unique,
  up_AD_only = subset(all_hubs, module %in% up_AD_only) %>% .$gene_name %>% unique,
  up_ADDS_only = subset(all_hubs, module %in% up_ADDS_only) %>% .$gene_name %>% unique,
  down_all = subset(all_hubs, module %in% down_all) %>% .$gene_name %>% unique,
  down_eAD_only = subset(all_hubs, module %in% down_eAD_only) %>% .$gene_name %>% unique,
  down_AD_only = subset(all_hubs, module %in% down_AD_only) %>% .$gene_name %>% unique,
  down_ADDS_only = subset(all_hubs, module %in% down_ADDS_only) %>% .$gene_name %>% unique
)

# size of lists
lapply(input_list, function(x){
  print(length(x))
})

# run enrichr and combine outputs
enrichr_df <- do.call(rbind, lapply(names(input_list), function(x){
  if(length(input_list[[x]]) > 0){
    cur_enrich <- enrichr(input_list[[x]], dbs)
    Sys.sleep(10)
  } else{return(data.frame())}
  cur_df <- do.call(rbind, lapply(dbs, function(cur_db){
    df <- cur_enrich[[cur_db]]
    if(nrow(df) > 1){df$group <- x; df$db <- cur_db}
    else{df <- data.frame()}
    df
  }))
}))

# number of genes
enrichr_df$ngenes <- unlist(lapply(strsplit(enrichr_df$Genes, ';'), function(x){length(x)}))

# write the significant results to a file
enrichr_df %>%
  subset(P.value < 0.05 & ngenes > 2) %>%
  write.table(
    file=paste0(data_dir, 'spatial_module_enrichR_results2.tsv'),
    quote=FALSE, row.names=FALSE, sep='\t'
  )

enrichr_df <- read.delim(paste0(data_dir, 'spatial_module_enrichR_results2.tsv'), sep='\t')

#---------------------------------------------------------------------------#
# plot the selected terms

selected_terms <- read.delim(paste0(data_dir, 'spatial_module_enrichR_selected.txt'), sep='\t')
selected_terms <- subset(enrichr_df, Term %in% selected_terms$Term & P.value < 0.005)

table(selected_terms$group)


cur_direction <- 'down'; cmap <- mako
cur_direction <- 'up'; cmap <- magma
selected_terms <- selected_terms[grepl(cur_direction, selected_terms$group),]

selected_terms$group <- factor(
  as.character(selected_terms$group),
  levels = paste0(cur_direction, '_', c('eAD_only', 'AD_only', 'ADDS_only', 'all'))
)

# set max pval
quantile(-log(selected_terms$P.value), 0.95)
max_p <- 10

selected_terms$logp <- -log(selected_terms$P.value)
selected_terms$logp <- ifelse(selected_terms$logp > max_p, max_p, selected_terms$logp)



# remove GO Term ID
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

selected_terms <- selected_terms %>%
  arrange(group)


selected_terms$Term <- factor(
  as.character(selected_terms$Term),
  levels = rev(unique(as.character(selected_terms$Term)))
)

p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color = logp, size=log(Odds.Ratio))) +
  geom_point() +
  scale_color_stepsn(colors=rev(cmap(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  coord_equal() + 
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(size=0.25, color='lightgrey')
  )

pdf(paste0(fig_dir, 'selected_GO_terms_modules_', cur_direction, '.pdf'), width=7, height=5)
p 
dev.off()

```