
```{r eval=FALSE}

library(Seurat)
library(tidyverse)
library(cowplot)
library(Matrix)
library(viridis)
library(harmony)
library(ggpubr)
library(patchwork)
library(RColorBrewer)
library(ggrepel)
library(CellTrek)

source("/pub/smorabit/hdWGCNA/bin/spatial_functions.R")
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')


# 2,000 MB limit:
options(future.globals.maxSize= 2000*1024^2)


setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/spatial_mapping/")

fig_dir <- "figures/"
data_dir <- "data/"

umap_theme() <- theme(
  axis.line=element_blank(),
  axis.text.x=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks=element_blank(),
  axis.title.x=element_blank(),
  axis.title.y=element_blank(),
  panel.background=element_blank(),
  panel.border=element_blank(),
  panel.grid.major=element_blank(),
  panel.grid.minor=element_blank(),
  plot.background=element_blank()
)

# load visium & split-seq seurat objects:
seurat_vis <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/data/ADDS_seurat_processed.rds')

# seurat_obj <- readRDS("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/data/ADDS_integrated_scvi.rds")

# load ADDS dataset:
seurat_obj <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_integrated.rds" )

# only get the FCX:
seurat_obj <- subset(seurat_obj, Tissue != 'PCC')

# saveRDS(seurat_obj, file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_integrated_FCX.rds")

# add barcode:
seurat_obj$bc <- colnames(seurat_obj)
seurat_vis$bc <- colnames(seurat_vis)

var_df <- read.delim(file = "/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_subclusters_var.tsv", sep='\t')

VariableFeatures(seurat_obj) <- var_df$X



human_cp <- c("WM1" = "#64BCDB", "WM2" = "#62A7D7", "WM3" = "#99C8D7",
      "L1" = "#8B3D5A", "L2/3" = "#E7BDE1", "L3/4" = "#E6A4CD",
      "L3/4/5" = "#CF8fBA3", "L5/6" = "#9E6D7F", "L6b" = "#CDAEB9")



```

it seems like we get fewer cells mapped the more samples we include togetgher so it
might be best to just map one sample at a time.

```{r eval=FALSE}

# variables from argparse:
index <- 1873
cell_id_col <- 'cell_identity'
n_iterations <- 3 # how many times to try mapping with celltrek

# set up variables for visium:
vis_sample_col <- 'Sample'
sc_sample_col <- 'Sample'
vis_samples <- unique(as.character(seurat_vis@meta.data[[vis_sample_col]]))
sc_samples <- unique(as.character(seurat_obj@meta.data[[sc_sample_col]]))

# get index of the visium & sc samples
vis_sample_ind <- ceiling(index / length(sc_samples))
sc_sample_ind <- 1 + (index %% length(sc_samples))
vis_sample_ind
sc_sample_ind

# get the name of the visium & sc samples
cur_vis_sample <- vis_samples[vis_sample_ind]
cur_sc_sample <- sc_samples[sc_sample_ind]

# output name is the concatenation of the sample names:
outname <- paste0('vis_', cur_vis_sample, '-sc_', cur_sc_sample)

# get the current visium sample and make sure to set up the image
cur_vis <- seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == cur_vis_sample]
cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])

# get the current sc seurat object
cur_seurat <- seurat_obj[,seurat_obj@meta.data[[sc_sample_col]] == cur_sc_sample]
cur_seurat$orig.ident <- cur_seurat@meta.data[[sc_sample_col]]

# add barcode column to vis & sc:
cur_vis$bc <- colnames(cur_vis)
cur_seurat$bc <- colnames(cur_seurat)

# this code gets a range of samples if we are mapping more than one at the same time
# not used here:
# index <- 12
# n_samples_total <- length(unique(seurat_obj@meta.data[[sc_sample_col]]))
# n_sc_samples <- 1
#
# sc_ind_start <- ((index - 1) * n_sc_samples) + 1
# sc_ind_end <- sc_ind_start + n_sc_samples - 1
# if(sc_ind_end > n_samples_total){
#   sc_ind_end <- n_samples_total
# }
# sc_ind_start
# sc_ind_end
# cur_sc_slice <- sc_ind_start:sc_ind_end


# co-embedding
ct_train <- CellTrek::traint(
  st_data=cur_vis,
  sc_data=cur_seurat,
  sc_assay='RNA',
  cell_names=cell_id_col
)

# add a label
# ct_train$label <- ifelse(
#   ct_train$type == 'sc', as.character(ct_train$cell_identity),
#   as.character(ct_train$annotation)
# )

saveRDS(ct_train, file=paste0(data_dir, outname, '_coembed.rds'))
#ct_train <- readRDS(file=paste0(data_dir, 'test_celltrek_sample.rds'))

#
#
#
# p <- DimPlot(ct_train, split.by = 'type', group.by = 'label', label=TRUE, raster=FALSE) + umap_theme()
#
#
# pdf(paste0(fig_dir, 'test_coembedding_sample.pdf'), width=16, height=8)
# p
# dev.off()



obj_list <- list()
mapped_bcs <- c(); keep <- colnames(ct_train)
for(i in 1:n_iterations){
  print(i)
  tic()
  ct_obj <- CellTrek::celltrek(
    st_sc_int=ct_train[,keep],
    int_assay='traint',
    sc_assay = 'RNA',
    reduction='pca',
    intp=T, intp_pnt=10000,
    intp_lin=F, nPCs=30,
    ntree=1000,
    dist_thresh=0.75,
    top_spot=1, spot_n=5,
    repel_r=20,
    repel_iter=20,
    keep_model=T,
    verbose=TRUE
  )$celltrek
  time_elapsed <- toc()

  # which cells were not mapped??
  mapped_bcs <- c(mapped_bcs, unique(ct_obj$bc))
  print(length(mapped_bcs) / ncol(cur_seurat))

  keep <- ifelse(ct_train$type == 'st', TRUE, !(as.character(ct_train$bc) %in% mapped_bcs))
  #table(keep)

  obj_list[[i]] <- ct_obj

}

ct_coords <- do.call(rbind, lapply(obj_list, function(x){
  x@reductions$celltrek@cell.embeddings
}))

ix <- match(rownames(ct_coords), colnames(ct_train))
ct_meta <- ct_train@meta.data[ix,colnames(cur_seurat@meta.data)]
ct_meta <- cbind(ct_meta, ct_coords)
write.csv(ct_meta, quote=FALSE, file=paste0(outname, '_mapped_coords.csv'))





#
#
#
#
#
#
#
#
#
#
# # https://github.com/navinlabcode/CellTrek/issues/13
#
# library(tictoc)
# tic()
# ct_obj <- CellTrek::celltrek(
#   st_sc_int=ct_train,
#   int_assay='traint',
#   # sc_data=cur_seurat,
#   #sc_data = seurat_obj,
#   sc_assay = 'RNA',
#   reduction='pca',
#   intp=T, intp_pnt=10000,
#   intp_lin=F, nPCs=30,
#   ntree=1000,
#   dist_thresh=0.75,
#   top_spot=1, spot_n=5,
#   repel_r=20,
#   repel_iter=20,
#   keep_model=T,
#   verbose=TRUE
# )$celltrek
# time_elapsed <- toc()
#
# #saveRDS(ct_obj, file=paste0(data_dir, 'test_celltrek_obj.rds'))
# #ct_obj <- readRDS(file=paste0(data_dir, 'test_celltrek_obj.rds'))
#
# # only 38% of cells were mapped!!!
# ncol(ct_obj) / ncol(cur_seurat)
#
# dim(ct_obj)
#
# # which cells were not mapped??
# mapped_bcs <- unique(ct_obj$bc); length(mapped_bcs)
# length(mapped_bcs) / ncol(cur_seurat)
#
# keep <- ifelse(ct_train$type == 'st', TRUE, !(ct_train$bc %in% mapped_bcs))
# table(keep)
#
#
# # mapping round 2:
#
# tic()
# ct_obj2 <- CellTrek::celltrek(
#   st_sc_int=ct_train[,keep],
#   int_assay='traint',
#   # sc_data=cur_seurat,
#   #sc_data = seurat_obj,
#   sc_assay = 'RNA',
#   reduction='pca',
#   intp=T, intp_pnt=10000,
#   intp_lin=F, nPCs=30,
#   ntree=1000,
#   dist_thresh=0.75,
#   top_spot=1, spot_n=5,
#   repel_r=20,
#   repel_iter=20,
#   keep_model=T,
#   verbose=TRUE
# )$celltrek
# time_elapsed <- toc()
#
# dim(ct_obj)
# dim(ct_obj2)
#
#
#
# mapped_bcs <- c(mapped_bcs, unique(ct_obj2$bc))
#
#
# keep <- ifelse(ct_train$type == 'st', TRUE, !(ct_train$bc %in% mapped_bcs))
# table(keep)
#
#
# # mapping round 2:
#
# tic()
# ct_obj3 <- CellTrek::celltrek(
#   st_sc_int=ct_train[,keep],
#   int_assay='traint',
#   # sc_data=cur_seurat,
#   #sc_data = seurat_obj,
#   sc_assay = 'RNA',
#   reduction='pca',
#   intp=T, intp_pnt=10000,
#   intp_lin=F, nPCs=30,
#   ntree=1000,
#   dist_thresh=0.75,
#   top_spot=1, spot_n=5,
#   repel_r=20,
#   repel_iter=20,
#   keep_model=T,
#   verbose=TRUE
# )$celltrek
# time_elapsed <- toc()
#
# dim(ct_obj)
# dim(ct_obj2)
# dim(ct_obj3)
#
#
#
# (ncol(ct_obj) + ncol(ct_obj2) + ncol(ct_obj3)) / ncol(cur_seurat)
#
# subset(ct_obj@meta.data, type == 'sc') %>% .$bc
#
#
#
# sc_data <- cur_seurat
# sc_assay <- 'RNA'
#
#
#
#

















sum(cur_seurat$bc %in% ct_obj$bc) / ncol(cur_seurat)

p1 <- cur_vis@meta.data %>%
  ggplot(aes(x=col, y=row, color=annotation)) +
  geom_point(size=1) +
  scale_color_manual(values=rev(magma(9))) +
  ggtitle('Visium') +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

p2 <- DimPlot(
    ct_obj,
    group.by = 'subtype',
    reduction = 'celltrek') +
  ggtitle('Celltrek') +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

pdf(paste0(fig_dir, 'test_cell_charting_sample.pdf'), width=8, height=3)
p1 + p2
dev.off()



p <- DimPlot(
    ct_obj,
    group.by = 'cell_names',
    split.by = 'cell_names',
    reduction = 'celltrek',
    ncol=7
  ) +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

pdf(paste0(fig_dir, 'test_celltrek_sample_split.pdf'), width=12, height=8)
p
dev.off()


p <- DimPlot(
    ct_obj,
    group.by = 'subtype',
    split.by = 'subtype',
    reduction = 'celltrek',
    ncol=8
  ) +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5)) +
  NoLegend()

pdf(paste0(fig_dir, 'test_celltrek_subtype_split.pdf'), width=12, height=4)
p
dev.off()


```

celltrek with full single-cell dataset:

```{r eval=FALSE}

# select current visium sample
cur_vis <- subset(seurat_vis, Sample == "Dec_20_2021_Human7")
cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])

###############################################################################
# Co-embed the single-cell and spatial transcriptomic data
###############################################################################

# needed something in orig.ident for some reason?
seurat_obj$orig.ident <- 'RNA'
Idents(seurat_obj) <- seurat_obj$cell_identity
Idents(cur_vis) <- cur_vis$annotation

# co-embedding
ct_train <- CellTrek::traint(
  st_data=cur_vis,
  sc_data=seurat_obj,
  sc_assay='RNA',
  cell_names='cell_identity',
  gene_kept = VariableFeatures(seurat_obj)
)
saveRDS(ct_train, file=paste0(data_dir, 'test_celltrek_full_snRNA.rds'))

# plot the integrated UMAP
p2 <- PlotEmbedding(
  ct_train,
  group.by = 'annotation',
  split.by = 'type',
  raster_dpi=200,
  plot_under=TRUE,
  plot_theme = umap_theme() + NoLegend()
)[[2]]

p1 <- PlotEmbedding(
  ct_train,
  group.by = 'cell_identity',
  split.by = 'type',
  raster_dpi=200,
  plot_under=TRUE,
  plot_theme = umap_theme() + NoLegend()
)[[1]]

pdf(paste0(fig_dir, 'test_coembedding_full2.pdf'), width=10, height=5)
wrap_plots(list(p1,p2), ncol=2) + plot_layout(guides='collect')
dev.off()

###############################################################################
# Map single-cell data to spatial coordinates
###############################################################################

# re-load ct train object
ct_train <- readRDS(file=paste0(data_dir, 'test_celltrek_full_snRNA.rds'))


# trying to figure out how many cells I can map at one time
#
# n_samples:
# 15: works
# 25: too big
# 20: seems fine

# maybe should just do 13 samples each time so it's exactly 3 bins?

n_samples <- 1

# subset ct_train with only a handful of the
all_samples <- unique(seurat_obj$Sample)
samples <- all_samples[1:n_samples]
ct_subset <- subset(ct_train, type == 'st' | Sample %in% samples)

dim(ct_subset)

ct_out <- CellTrek::celltrek(
  st_sc_int=ct_subset,
  int_assay='traint',
  # sc_data=cur_seurat,
  sc_assay = 'RNA',
  reduction='pca',
  intp=T, intp_pnt=10000,
  intp_lin=F, nPCs=30,
  ntree=1000,
  dist_thresh=0.55,
  top_spot=1, spot_n=100,
  repel_r=20, repel_iter=20,
  keep_model=T
)
ct_obj <- ct_out$celltrek

dim(ct_subset)
dim(ct_obj)

table(ct_subset$type)

saveRDS(ct_obj, file=paste0(data_dir, 'test_celltrek_obj_full.rds'))


p1 <- cur_vis@meta.data %>%
  ggplot(aes(x=col, y=row, color=annotation)) +
  geom_point(size=1) +
  scale_color_manual(values=rev(magma(9))) +
  ggtitle('Visium') +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

p2 <- DimPlot(
    ct_obj,
    group.by = 'cell_identity',
    reduction = 'celltrek') +
  ggtitle('Celltrek') +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

pdf(paste0(fig_dir, 'test_cell_charting2.pdf'), width=8, height=3)
p1 + p2
dev.off()



p <- DimPlot(
    ct_obj,
    group.by = 'cell_names',
    split.by = 'cell_names',
    reduction = 'celltrek',
    ncol=7
  ) +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5))

pdf(paste0(fig_dir, 'test_celltrek_split2.pdf'), width=12, height=8)
p
dev.off()


p <- DimPlot(
    ct_obj,
    group.by = 'cell_identity',
    split.by = 'cell_identity',
    reduction = 'celltrek',
    ncol=8
  ) +
  umap_theme() +
  theme(plot.title=element_text(hjust=0.5)) +
  NoLegend()

pdf(paste0(fig_dir, 'test_celltrek_subtype_split2.pdf'), width=12, height=4)
p
dev.off()

```

why did some of my jobs not work?

```{r eval=FALSE}

celltrek_dir <- '~/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/spatial_mapping/data/coords/'

vis_samples <- do.call(rbind, strsplit(dir(celltrek_dir), '-'))[,1]
vis_samples <- as.character(sapply(vis_samples, function(x){substr(x, 5, 100)}))

sc_samples <- as.character(sapply(dir(celltrek_dir), function(x){
  substr(x, 1, nchar(x) - 18)
}))
sc_samples <- do.call(rbind, strsplit(sc_samples, '-'))
sc_samples[,3] <- ifelse(sc_samples[,1] == sc_samples[,3], '', sc_samples[,3])
sc_samples <- sc_samples[,2:3]
sc_samples[,1] <- as.character(sapply(sc_samples[,1], function(x){substr(x,4, 100)}))

sc_samples <- apply(sc_samples, 1, function(x){
  if(x[2] != ""){
    out <- paste(c(x[1], x[2]), collapse='-')
  } else{
    out <- x[1]
  }
  out
})

vis_samples %in% seurat_vis$Sample %>% sum
sc_samples %in% seurat_obj$Sample %>% sum

sc_samples[!(sc_samples %in% seurat_obj$Sample)]
unique(seurat_obj$Sample)



# which combinations are missing?

combinations <- c()
all_vis_samples <- as.character(unique(seurat_vis$Sample))
all_sc_samples <- as.character(unique(seurat_obj$Sample))
for(i in all_vis_samples){
  for(j in all_sc_samples){
    combo <- paste(c(i, j), collapse='#')
    combinations <- c(combinations, combo)
  }
}
head(combinations)

run_combinations <- as.character(sapply(1:length(vis_samples), function(i){
  paste(c(vis_samples[i], sc_samples[i]), collapse='#')
}))

run_combinations %in% combinations %>% sum

# these are the combinations that failed:
failed_combos <- combinations[!(combinations %in% run_combinations)]

sapply(all_vis_samples, function(x){any(grepl(x, failed_combos))})
all_sc_samples[!sapply(all_sc_samples, function(x){any(grepl(x, failed_combos))})]

```

Get the indices of the failed combos

```{r eval=FALSE}

outnames <- c()

# list of vis & sc samples
vis_samples <- unique(as.character(seurat_vis@meta.data[[vis_sample_col]]))
sc_samples <- unique(as.character(seurat_obj@meta.data[[sc_sample_col]]))

for(index in 1:1872){

  # get index of the visium & sc samples
  vis_sample_ind <- ceiling(index / length(sc_samples))
  sc_sample_ind <- 1 + (index %% length(sc_samples))

  # get the name of the visium & sc samples
  cur_vis_sample <- vis_samples[vis_sample_ind]
  cur_sc_sample <- sc_samples[sc_sample_ind]

  # output name is the concatenation of the sample names:
  outname <- paste0('vis_', cur_vis_sample, '-sc_', cur_sc_sample)

  if(outname %in% outnames){
    print(vis_sample_ind)
    print(sc_sample_ind)
    print(index)
  }

  outnames <- c(outnames, outname)

}
outnames
max(table(outnames))

which(table(outnames) == 2)
outnames[outnames == "vis_Oct_2021_1-sc_ADDS_10"]

```


Run celltrek on one sample and make sure to flip the rows first!!!!!




```{r eval=FALSE}

# variables from argparse:
cell_id_col <- 'cell_identity'
n_iterations <- 3 # how many times to try mapping with celltrek
i <- 1

# set up variables for visium:
vis_sample_col <- 'Sample'
sc_sample_col <- 'Sample'
vis_samples <- unique(as.character(seurat_vis@meta.data[[vis_sample_col]]))
sc_samples <- unique(as.character(seurat_obj@meta.data[[sc_sample_col]]))


# get the name of the visium & sc samples
cur_vis_sample <- "Dec_13_2021_Human1"
cur_sc_sample <- "ADDS_118"

# output name is the concatenation of the sample names:
outname <- paste0('vis_', cur_vis_sample, '-sc_', cur_sc_sample)

# get the current visium sample and make sure to set up the image
# cur_vis <- seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == cur_vis_sample]
# cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
# cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])



# subset the seurat object for this visium sample
cur_vis <- seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == cur_vis_sample]
cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])

# add coords to seurat metadata

# need to fix the image row:
cur_row <- cur_vis@images$cur_image@coordinates$row
row_min <- min(cur_row);
cur_row <- cur_row - row_min
row_max <- max(cur_row)
cur_row <- cur_row - row_max
cur_row <- cur_row * -1
cur_row <- cur_row + row_min

cur_vis$row <- cur_row
cur_vis@images$cur_image@coordinates$row <- cur_row
cur_vis$col <- cur_vis@images$cur_image@coordinates$col
cur_vis$imagerow <- cur_vis@images$cur_image@coordinates$imagerow
cur_vis$imagecol <- cur_vis@images$cur_image@coordinates$imagecol



# get the current sc seurat object
cur_seurat <- seurat_obj[,seurat_obj@meta.data[[sc_sample_col]] == cur_sc_sample]
cur_seurat$orig.ident <- cur_seurat@meta.data[[sc_sample_col]]

# add barcode column to vis & sc:
cur_vis$bc <- colnames(cur_vis)
cur_seurat$bc <- colnames(cur_seurat)


# co-embedding
ct_train <- CellTrek::traint(
  st_data=cur_vis,
  sc_data=cur_seurat,
  sc_assay='RNA',
  cell_names=cell_id_col
)


library(tictoc)

obj_list <- list()
mapped_bcs <- c(); keep <- colnames(ct_train)
for(i in 1:n_iterations){
  print(i)
  tic()
  ct_obj <- CellTrek::celltrek(
    st_sc_int=ct_train[,keep],
    int_assay='traint',
    sc_assay = 'RNA',
    reduction='pca',
    intp=T, intp_pnt=10000,
    intp_lin=F, nPCs=30,
    ntree=1000,
    dist_thresh=0.75,
    top_spot=1, spot_n=5,
    repel_r=20,
    repel_iter=20,
    keep_model=T,
    verbose=TRUE
  )$celltrek
  time_elapsed <- toc()

  # which cells were not mapped??
  mapped_bcs <- c(mapped_bcs, unique(ct_obj$bc))
  print(length(mapped_bcs) / ncol(cur_seurat))

  keep <- ifelse(ct_train$type == 'st', TRUE, !(as.character(ct_train$bc) %in% mapped_bcs))
  #table(keep)

  obj_list[[i]] <- ct_obj

}

ct_coords <- do.call(rbind, lapply(obj_list, function(x){
  x@reductions$celltrek@cell.embeddings
}))

ix <- match(rownames(ct_coords), colnames(ct_train))
ct_meta <- ct_train@meta.data[ix,colnames(cur_seurat@meta.data)]
ct_meta <- cbind(ct_meta, ct_coords)

# only take the barcode and the coordinates to save:
ct_meta <- dplyr::select(ct_meta, c(bc, celltrek_1, celltrek_2))

write.csv(ct_meta, quote=FALSE, file=paste0(outname, '_mapped_coords.csv'))

write.table(ct_meta, quote=FALSE, file=paste0(outname, '_mapped_coords.tsv'), sep='\t')

test <- read.csv(file=paste0(outname, '_mapped_coords.csv'))
test <- read.table(paste0(outname, '_mapped_coords.tsv'), sep='\t')

```

Load one of the results:

```{r eval=FALSE}

library(magrittr)


# get a list of the celltrek output files
celltrek_dir <- '~/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/spatial_mapping/data/coords/'
coords_files <- dir(celltrek_dir)

# get a list of the visium / sc samples
vis_sample_col <- 'Sample'
sc_sample_col <- 'Sample'

vis_samples <- do.call(rbind, strsplit(dir(celltrek_dir), '-'))[,1]
vis_samples <- as.character(sapply(vis_samples, function(x){substr(x, 5, 100)}))

# extract the visium / sc samples from the list of output files
sc_samples <- as.character(sapply(dir(celltrek_dir), function(x){
  substr(x, 1, nchar(x) - 18)
}))
sc_samples <- do.call(rbind, strsplit(sc_samples, '-'))
sc_samples[,3] <- ifelse(sc_samples[,1] == sc_samples[,3], '', sc_samples[,3])
sc_samples <- sc_samples[,2:3]
sc_samples[,1] <- as.character(sapply(sc_samples[,1], function(x){substr(x,4, 100)}))
sc_samples <- apply(sc_samples, 1, function(x){
  if(x[2] != ""){
    out <- paste(c(x[1], x[2]), collapse='-')
  } else{
    out <- x[1]
  }
  out
})


# make list of seurat objects for each sample:
seurat_list <- lapply(unique(seurat_obj@meta.data[[sc_sample_col]]), function(x){
  seurat_obj[,seurat_obj@meta.data[[sc_sample_col]] == x]
})
names(seurat_list) <- unique(seurat_obj@meta.data[[sc_sample_col]])

# make list of seurat visium objects for each sample
vis_list <- lapply(unique(seurat_vis@meta.data[[vis_sample_col]]), function(y){
  cur_vis <- seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == y]
  cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
  cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])
  cur_vis
})
names(vis_list) <- unique(seurat_vis@meta.data[[vis_sample_col]])

# number of closest spots:
n_closest <- 10

# initialize progress bar
pb <- utils::txtProgressBar(min = 0, max = length(coords_files),style = 3, width = 50, char = "=")

anno_list <- list()
for(i in 1:length(coords_files)){
#for(i in 1:50){
  print(i)
  setTxtProgressBar(pb, i)

  # load the celltrek coordinates
  cur_file <- coords_files[i]
  cur_coords <- read.csv(paste0(celltrek_dir, cur_file))

  # remove any cells that were multi-mapped
  cur_coords <- cur_coords[match(unique(cur_coords$bc), cur_coords$bc),]

  # get current sample
  cur_sc_sample <- sc_samples[i]
  cur_vis_sample <- vis_samples[i]

  # add sample info to coords
  cur_coords$vis_sample <- cur_vis_sample
  cur_coords$sc_sample <- cur_sc_sample

  # flip celltrek2
  cur_row <- cur_coords$celltrek_2
  row_min <- min(cur_row);
  cur_row <- cur_row - row_min
  row_max <- max(cur_row)
  cur_row <- cur_row - row_max
  cur_row <- cur_row * -1
  cur_row <- cur_row + row_min
  cur_coords$celltrek_2 <- cur_row

  # subset the seurat object for this visium sample
  cur_vis <- vis_list[[cur_vis_sample]]
  # cur_vis <- seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == cur_vis_sample]
  # cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
  # cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])

  # subset the seurat object for this sc sample
  # cur_seurat <- seurat_obj[,seurat_obj@meta.data[[sc_sample_col]] == cur_sc_sample]
  cur_seurat <- seurat_list[[cur_sc_sample]]

  # setup the celltrek coords dataframe to compute the distance
  sc_coords <- cur_coords %>% dplyr::select(c(celltrek_1, celltrek_2)) %>%
    dplyr::rename(c(x=celltrek_1, y=celltrek_2)) %>%
    as.matrix() %>% t()
  colnames(sc_coords) <- as.character(cur_coords$bc)

  # setup the visium coords dataframe to compute the distance
  spots <- cur_vis@meta.data[,c('imagecol', 'imagerow')] %>% dplyr::rename(c(x=imagecol, y=imagerow))
  spots <- spots %>% as.matrix() %>% t()

  # compute distance between the projected single-cells and the spots
  dists <- apply(sc_coords, 2, function(z){
      sqrt(colSums((spots - z)^2))
  })

  # flatten into a dataframe
  dist_df <- reshape2::melt(dists)
  dist_df %<>% dplyr::rename(c(vis_bc=Var1, sc_bc=Var2, dist=value))

  # get the n closest spots to each mapped cell based on distance
  top_closest <- dist_df %>% group_by(sc_bc) %>% top_n(-n_closest, wt=dist)

  # add the spatial annotation of the closest spots
  ix <- match(top_closest$vis_bc, colnames(cur_vis))
  top_closest$annotation <- cur_vis$annotation[ix]

  # what's the most frequent annotation for each cell?
  anno_freq <- top_closest %>% group_by(sc_bc, annotation) %>%
    summarize(Freq=n())
  anno_freq <- anno_freq %>% group_by(sc_bc) %>% top_n(1, wt=Freq) %>%
    mutate(annotation = paste0(annotation, collapse=' ')) %>%
    distinct()

  # add the cluster identity from the singlecell data
  ix <- match(anno_freq$sc_bc, colnames(cur_seurat))
  anno_freq$cell_identity <- cur_seurat$cell_identity[ix]
  anno_freq$cell_type <- cur_seurat$cell_type[ix]

  # add visium sample and sc sample:
  anno_freq$vis_sample <- cur_vis_sample
  anno_freq$sc_sample <- cur_sc_sample

  # add the celltrek coords in
  anno_freq$celltrek_1 <- cur_coords$celltrek_1
  anno_freq$celltrek_2 <- cur_coords$celltrek_2

  anno_list[[i]] <- anno_freq

}

close(pb)

# combine into one dataframe
anno_df <- do.call(rbind, anno_list)

write.csv(anno_df, quote=FALSE, file='celltrek_coords_combined.csv')


# re-load the output:


```

Plot the cell mapping for each visium sample

```{r eval=FALSE}

anno_df <- read.csv('celltrek_coords_combined.csv')
color_df <- read.csv('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/cell_identity.csv')
cp  <- color_df$colour
names(cp) <- color_df$group

# ix <- match(anno_df$sc_bc, seurat_obj$bc)
# anno_df$cell_identity <- seurat_obj$cell_identity[ix]
# anno_df$cell_type <- seurat_obj$cell_type[ix]
#

# make a plot for each vis sample
vis_sample_col <- 'Sample'
visium_samples <- unique(seurat_vis@meta.data[[vis_sample_col]])
cur_vis_sample <- visium_samples[1]

#visium_samples <- c("Dec_13_2021_Human5", "Dec_13_2021_Human8")

for(cur_vis_sample in visium_samples){

  print(cur_vis_sample)

  plot_df <- subset(anno_df, vis_sample == cur_vis_sample)

  p_vis <- VisDimPlot(
    seurat_vis[,seurat_vis@meta.data[[vis_sample_col]] == cur_vis_sample],
    group.by = 'annotation',
    dpi=600,
    ncol = 1,
    color_df=human_cp,
    text_size=15
  )
  #
  # # plot colored by region
  # p <- plot_df %>%
  #   ggplot(aes(x=celltrek_1, y=celltrek_2, color=annotation)) +
  #   ggrastr::rasterise(geom_point(size=0.5), dpi=150, scale=0.5) +
  #   umap_theme() + coord_fixed()
  #
  # ct_plots <- p + facet_wrap(~sc_sample, ncol=12) + theme(strip.text.x = element_text(size = 5))
  #
  # pdf(paste0(fig_dir,'celltrek_mapped/', cur_vis_sample, '_anno.pdf'), width=18, height=6)
  # print(p_vis + ct_plots + plot_layout(widths=c(3, 9)))
  # dev.off()

  # plot colored by cell identity
  p <- plot_df %>%
    ggplot(aes(x=celltrek_1, y=celltrek_2, color=cell_identity)) +
    ggrastr::rasterise(geom_point(size=0.5), dpi=500, scale=0.5) +
    scale_color_manual(values=cp) +
    umap_theme() + coord_fixed() +
    guides(colour = guide_legend(override.aes = list(size=5)))

  ct_plots <- p + facet_wrap(~sc_sample, ncol=12) + theme(strip.text.x = element_text(size = 5))

  pdf(paste0(fig_dir,'celltrek_mapped/', cur_vis_sample, '_cell_ids.pdf'), width=18, height=6)
  print(p_vis + ct_plots + plot_layout(widths=c(3, 9)))
  dev.off()

  # plot colored by cell type
  # p <- plot_df %>%
  #   ggplot(aes(x=celltrek_1, y=celltrek_2, color=cell_type)) +
  #   ggrastr::rasterise(geom_point(size=0.5), dpi=150, scale=0.5) +
  #   umap_theme() + coord_fixed()
  #
  # ct_plots <- p + facet_wrap(~sc_sample, ncol=12) + theme(strip.text.x = element_text(size = 5))
  #
  # pdf(paste0(fig_dir,'celltrek_mapped/', cur_vis_sample, '_celltype.pdf'), width=18, height=6)
  # print(p_vis + ct_plots + plot_layout(widths=c(3, 9)))
  # dev.off()

}



annotation_names <- c('WM', as.character(unique(seurat_vis$annotation)))
table(anno_df$annotation)




# re-name white matter:
anno_df$anno_fix <- ifelse(
  grepl('WM', anno_df$annotation) & !(grepl('L', anno_df$annotation)),
  'WM',
  anno_df$annotation
)


sum(anno_df$annotation %in% annotation_names) / nrow(anno_df)
sum(anno_df$anno_fix %in% annotation_names) / nrow(anno_df)


length(table(anno_df$anno_fix))
length(table(anno_df$anno_fix)[table(anno_df$anno_fix) >= 500])
table(anno_df$anno_fix)[table(anno_df$anno_fix) >= 500]

sum(table(anno_df$anno_fix)[table(anno_df$anno_fix) > 500]) / nrow(anno_df)
sum(table(anno_df$anno_fix)[table(anno_df$anno_fix) <= 500]) / nrow(anno_df)

sum(colnames(seurat_obj) %in% anno_df$sc_bc)

```

Which region was each cell most frequently mapped to across all of the samples?



Seems odd that for some clusters like Microglia are 99% in white matter? I think
for MG it would broadly split into WM vs GM for microglia

```{r eval=FALSE}

# anno_df_full <- anno_df


anno_df <- subset(anno_df_full, anno_fix %in% annotation_names)
anno_df$sc_bc <- as.character(anno_df$sc_bc)
anno_df$anno_fix <- factor(
  as.character(anno_df$anno_fix),
  levels = c('L1', 'L2/3', 'L3/4', 'L3/4/5', 'L5/6', 'L6b', 'WM')
)

# count how many times each cell was mapped to each annotation
test <- anno_df %>% dplyr::group_by(sc_bc, anno_fix, .drop=FALSE) %>%
    dplyr::summarize(Freq=n())

bc_counts <- table(anno_df$sc_bc)
ix <- match(test$sc_bc, names(bc_counts))
test$Freq / as.numeric(bc_counts[ix])
test$prop <- test$Freq / as.numeric(bc_counts[ix])


props <- data.frame(bc = colnames(seurat_plot))
for(cur_group in as.character(unique(test$anno_fix))){

  test_prop <- subset(test, anno_fix == cur_group)
  ix <- match(colnames(seurat_plot), as.character(test_prop$sc_bc))
  cur_prop <- data.frame(
    cur_group = test_prop$prop[ix]
  )
  names(cur_prop) <- cur_group
  props <- cbind(props, cur_prop)

#  p <- VlnPlot(seurat_plot, features='prop', group.by='cell_identity', pt.size=0) + NoLegend()


}

seurat_plot@meta.data <- cbind(seurat_plot@meta.data, props[,-1])


p <- custom_vln(
    seurat_plot,
    features = names(props[,-1]),
    group.by = 'cell_identity',
    #groups = c('INH1 VIP+', 'INH4 LAMP5+', 'INH2 SST+',  'INH5 SST+',  'INH3 PVALB+', 'INH6 PVALB+'),
    add_boxplot=FALSE
  #  split.by = 'Diagnosis',
  #  selected_split = c('ASD', 'Control'),
  #  split_colors=c('darkorchid', 'grey'),
  #  add_colorbar=FALSE,
    #plot_ymin = NA
  )




pdf(paste0(fig_dir, 'test_prop_vln_big.pdf'), width=10, height=7)
p + NoLegend()
dev.off()

pdf(paste0(fig_dir, 'test_prop_vln.pdf'), width=10, height=4)
p
dev.off()





mg_bcs <- colnames(seurat_plot[, seurat_plot$cell_type == 'MG'])
test_mg <- test %>% subset(sc_bc %in% mg_bcs)

test_mg %>% subset(anno_fix == 'WM' & Freq > 20) %>% dim


test <- test %>% group_by(sc_bc) %>% top_n(1, wt=Freq) %>%
  mutate(anno_fix = paste0(anno_fix, collapse=' ')) %>%
  distinct()


sum(test$anno_fix %in% annotation_names) / nrow(test)

table(test$anno_fix)


test2 <- subset(test, anno_fix %in% annotation_names)



seurat_plot <- seurat_obj[,test2$sc_bc]
ix <- match(as.character(colnames(seurat_plot)), test2$sc_bc)
seurat_plot$spatial_anno <- as.character(test2$anno_fix)[ix]



p <- PlotEmbedding(
  seurat_plot,
  group.by = 'spatial_anno',
  raster_dpi = 600,
  raster_scale=0.5, point_size=1,
  plot_theme = umap_theme()
)


p1 <- DimPlot(seurat_plot, group.by = 'spatial_anno', raster=F, label=TRUE) + umap_theme()
p2 <- DimPlot(seurat_plot, group.by = 'cell_identity', raster=F, label=TRUE) + umap_theme()

pdf(paste0(fig_dir, 'umap_spatial_anno.pdf'), width=6, height=6)
p
dev.off()


pdf(paste0(fig_dir, 'umap_spatial_anno.pdf'), width=12, height=4)
p1 + p2
dev.off()

test3 <- subset(seurat_plot@meta.data, cell_type == 'MG')
table(test3$spatial_anno) / nrow(test3) * 100

```
