# Load libraries and visium samples

```{r eval=FALSE}

# conda activate cicero
library(Seurat)
library(harmony)
library(tidyverse)
library(cowplot)
library(patchwork)
library(RColorBrewer)
library(tictoc)
library(BayesSpace)
library(scater)

colfunc <- colorRampPalette(rev(brewer.pal(11, 'Spectral' )))
theme_set(theme_cowplot())

setwd("/dfs3b/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD")

fig_dir <- "figures/"
data_dir <- "data/"


umap_theme <- theme(
  axis.line=element_blank(),
  axis.text.x=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks=element_blank(),
  axis.title.x=element_blank(),
  axis.title.y=element_blank(),
  panel.background=element_blank(),
  panel.border=element_blank(),
  panel.grid.major=element_blank(),
  panel.grid.minor=element_blank(),
  plot.background=element_blank()
)

# re-load Seurat obj
seurat_obj <- readRDS(paste0(data_dir,'5XFAD_seurat_processed.rds'))

# re-load BayesSpace object:
sce.combined <- readRDS(file=paste0(data_dir, '5xFAD_bayesspace.rds'))

seurat_obj$bs.q15 <- sce.combined$spatial.cluster.q15

```

# Seurat clustering analysis

```{r eval=FALSE}

# re-load unprocessed data
seurat_obj <- readRDS(paste0(data_dir, '5xFAD_visium_unprocessed.rds'))
seurat_obj <- subset(seurat_obj, nCount_Spatial != 0)

# process data
seurat_obj <- NormalizeData(seurat_obj)
seurat_obj <- FindVariableFeatures(seurat_obj, nfeatures=3500)
seurat_obj <- ScaleData(seurat_obj, features=VariableFeatures(seurat_obj))

# dim reduction and clustering
seurat_obj <- RunPCA(seurat_obj, verbose = FALSE)
seurat_obj <- RunHarmony(
  seurat_obj,
  group.by.vars='seqbatch',
  assay='Spatial'
)

# UMAP + clustering
seurat_obj <- RunUMAP(
  seurat_obj,
  reduction = "harmony",
  dims = 1:30,
  min.dist=0.05,
  #spread=1.5,
  n.neighbors=5,
  return.model=TRUE
)
seurat_obj <- FindNeighbors(seurat_obj, reduction = "harmony", dims = 1:30)
seurat_obj <- FindClusters(seurat_obj, verbose = TRUE, res=1)

```

# Plotting stuff on UMAP

```{r eval=FALSE}

# plot umap colored by cluster
p <- DimPlot(seurat_obj, group.by='seurat_clusters', label=TRUE, raster=FALSE) + NoLegend() + umap_theme

pdf(paste0(fig_dir, 'umap_clusters.pdf'), width=7, height=7, useDingbats=FALSE)
p
dev.off()

# color by cluster, split by batch
p <- DimPlot(seurat_obj, group.by='seurat_clusters', label=TRUE, raster=FALSE, split.by='seqbatch', ncol=3) + umap_theme + NoLegend() + ggtitle('')

pdf(paste0(fig_dir, 'umap_batches.pdf'), width=9, height=6, useDingbats=FALSE)
p
dev.off()

# color by cluster, split by sample
p <- DimPlot(seurat_obj, group.by='seurat_clusters', label=TRUE, raster=FALSE, split.by='SAMPLE', ncol=10) + umap_theme + NoLegend() + ggtitle('')

pdf(paste0(fig_dir, 'umap_samples.pdf'), width=20, height=16, useDingbats=FALSE)
p
dev.off()

# color by nCountSpatial
seurat_obj$log_nCount_Spatial <- log(seurat_obj$nCount_Spatial)
p <- FeaturePlot(seurat_obj, features='nCount_Spatial', raster=FALSE, order=TRUE) + umap_theme
  # scale_color_gradientn(colors=colfunc(256), guide = guide_colorbar(barwidth=15, barheight=0.5, ticks=FALSE)) + theme(legend.position='bottom')

pdf(paste0(fig_dir, 'umap_nCountSpatial.pdf'), width=9, height=8, useDingbats=FALSE)
p
dev.off()


plot_list <- SpatialDimPlot(seurat_obj, label=TRUE, combine=FALSE)
for(i in 1:length(plot_list)){
  plot_list[[i]] <- plot_list[[i]] + NoLegend()
}

pdf(paste0(fig_dir, 'spatial_clusters.pdf'), width=15, height=15, useDingbats=FALSE)
(p[[1]] + p[[2]]) / (p[[3]] + p[[4]])
dev.off()



# marker genes:
# color by nCountSpatial
p <- FeaturePlot(seurat_obj, features=c('Mbp', 'Csf1r', 'Gfap', 'Gad1'), raster=TRUE, order=TRUE, ncol=2) + umap_theme
  # scale_color_gradientn(colors=colfunc(256), guide = guide_colorbar(barwidth=15, barheight=0.5, ticks=FALSE)) + theme(legend.position='bottom')

pdf(paste0(fig_dir, 'umap_markers.pdf'), width=10, height=10, useDingbats=FALSE)
p
dev.off()

saveRDS(seurat_obj, paste0(data_dir,'5XFAD_seurat_processed.rds'))

```




## Label transfer

```{r eval=FALSE}

# load processed data:
rosenberg <- readRDS('~/swaruplab/smorabit/collab/Harvard_visium/rosenberg_2018/data/rosenberg_brain_seurat_processed.rds')

# keep genes in rosenberg that are in seurat_obj:
rosenberg <- rosenberg[rownames(rosenberg)[rownames(rosenberg) %in% rownames(seurat_obj)],]

# transfer anchors between ros and seurat_obj:
anchors <- FindTransferAnchors(
  reference = rosenberg,
  query = seurat_obj
)
saveRDS(anchors, 'data/rosenberg_anchors.rds')

# make predictions using anchors:
predictions.assay <- TransferData(
  anchorset = anchors,
  refdata = rosenberg$cluster_assignment,
  prediction.assay = TRUE,
  dims=1:30,
  weight.reduction = seurat_obj[["harmony"]]
)

# add to seurat_obj seurat obj
seurat_obj[["predictions"]] <- predictions.assay
saveRDS(seurat_obj, paste0(data_dir,'5XFAD_seurat_processed.rds')

```


Plot prediction scores:

```{r eval=FALSE}

dir.create(paste0(fig_dir, 'rosenberg_label_transfer'))

# Plot prediction scores for some clusters:
DefaultAssay(seurat_obj) <- "predictions"

prediction_matrix <- GetAssayData(seurat_obj, assay='predictions')

for(label in rownames(seurat_obj)[rowSums(prediction_matrix) > 0]){

  name <- gsub(' ', '_', label)
  name <- gsub('/', '_', label)
  print(name)

  # umap feature plot
  p1 <- FeaturePlot(seurat_obj, features=label, order=TRUE) +
    scale_color_gradientn(colors=colfunc(256), guide = guide_colorbar(barwidth=15, barheight=0.5, ticks=FALSE)) +
    umap_theme + theme(legend.position='bottom')

  # spatial feature plot
  # plot_list <- list()
  # for(sample in unique(seurat_obj$SampleID)){
  #   cur <- subset(seurat_obj, SampleID == sample)
  #   cur_image <- names(cur@images)[sapply(names(cur@images), function(x){nrow(cur@images[[x]]@coordinates) > 0})]
  #   cur@images <- list(cur_image = cur@images[[cur_image]])
  #   plot_list[[sample]] <- SpatialFeaturePlot(cur, features=label) + ggtitle(sample)  +
  #   #  scale_color_gradientn(colors=colfunc(256), guide = guide_colorbar(barwidth=15, barheight=0.5, ticks=FALSE)) +
  #     theme(legend.position='bottom', legend.title=element_blank())
  # }

  # cluster violin plot:
  p3 <- VlnPlot(seurat_obj, features=label, pt.size=0) +
   NoLegend() + ggtitle('') +
   ylab(paste(label, 'score')) + xlab('clusters')

  # patchwork
  patch <- (p1 / p3)

  pdf(paste0(fig_dir, 'rosenberg_label_transfer/', name, '.pdf'), width=12, height=12, useDingbats=FALSE)
  print(patch + plot_layout(heights=c(4,1)))
  dev.off()

}

```


# get the image coordinates for BayesSpace:
```{r eval=FALSE}

# get all of the image coordinates for BayesSpace
image_df <- do.call(rbind, lapply(names(seurat_obj@images), function(cur_image){seurat_obj@images[[cur_image]]@coordinates}))

# re-order the rows of image_df to match the seurat_obj
image_df <- image_df[colnames(seurat_obj),]
all.equal(rownames(image_df), colnames(seurat_obj))


```



## Run BayesSpace

* How do I get the correct image coordinates?????
  - Need to loop over each sample I think...
  - Actually maybe just do this while I setup the data initially?

```{r eval=FALSE}


# convert from Seurat to SCE format:
sce.combined <- seurat_obj %>% as.SingleCellExperiment()

# add the row, col, imagerow, imagecol
sce.combined$row <- image_df$row
sce.combined$imagerow <- image_df$imagerow
sce.combined$col <- image_df$col
sce.combined$imagecol <- image_df$imagecol

# pre-process
sce.combined = spatialPreprocess(sce.combined, n.PCs = 50) #lognormalize, PCA

# correct with harmony
sce.combined = RunHarmony(sce.combined, "SAMPLE", verbose = T)

# run UMAP
sce.combined = runUMAP(
  sce.combined,
  dimred = "HARMONY",
  name = "UMAP.HARMONY",
  n_neighbors=10
)

# where are the row & col in sce.combined

colnames(reducedDim(sce.combined, "UMAP.HARMONY")) = c("UMAP1", "UMAP2")


p <- ggplot(data.frame(reducedDim(sce.combined, "UMAP.HARMONY")),
  aes(x = UMAP1, y = UMAP2, color = factor(sce.combined$SAMPLE))) +
  geom_point() +
  labs(color = "Sample") +
  theme_bw() + NoLegend()

pdf(paste0(fig_dir, 'baysespace_umap.pdf'), width=7, height=7)
p
dev.off()

saveRDS(sce.combined, file=paste0(data_dir, '5xFAD_bayesspace.rds'))


```

Come up with a way to offset the samples

10 x 8 grid


```{r eval=FALSE}

sce.combined <- readRDS(file=paste0(data_dir, '5xFAD_bayesspace.rds'))

# get all of the image coordinates for BayesSpace
image_df <- do.call(rbind, lapply(names(seurat_obj@images), function(cur_image){
  cur_coords <- seurat_obj@images[[cur_image]]@coordinates
  cur_coords$image <- cur_image
  cur_coords
}))

# re-order the rows of image_df to match the seurat_obj
image_df <- image_df[colnames(seurat_obj),]
all.equal(rownames(image_df), colnames(seurat_obj))


range(image_df$row)
range(image_df$col)

# based on the max of the row/col
row_offset <- 100
col_offset <- 150

images <- unique(image_df$image)
offset_df <- data.frame()
cur_ind <- 1
for(i in 1:10){
  cur_row_offset <- row_offset*i
  for(j in 1:8){

    print(cur_ind)
    cur_col_offset <- col_offset*j

    # get cur_coords:
    cur_img <- images[cur_ind]
    cur_coords <- image_df %>% subset(image == cur_img)

    # apply offset:
    cur_coords$row <- cur_coords$row + cur_row_offset
    cur_coords$col <- cur_coords$col + cur_col_offset

    offset_df <- rbind(offset_df, cur_coords)

    cur_ind <- cur_ind + 1

  }
}

offset_df <- offset_df[rownames(image_df),]

# add the row, col, imagerow, imagecol
sce.combined$row <- offset_df$row
sce.combined$imagerow <- offset_df$imagerow
sce.combined$col <- offset_df$col
sce.combined$imagecol <- offset_df$imagecol


# add it to the seurat object:
seurat_obj$row <- offset_df$row
seurat_obj$col <- offset_df$col
seurat_obj$imagerow <- offset_df$imagerow
seurat_obj$imagecol <- offset_df$imagecol

p <- clusterPlot(sce.combined, "SAMPLE", color=NA) + #make sure no overlap between samples
  labs(color = "Sample", title = "Offset check")

pdf(paste0(fig_dir, 'slide_offset_test.pdf'), width=10, height=10)
p
dev.off()


library(MetBrewer)

# test plotting with ggplot on image df
p <- offset_df %>% ggplot(aes(x=row, y=col, color=image)) +
  geom_point() + umap_theme + NoLegend() +
  scale_color_manual(values=met.brewer("Signac", length(unique(offset_df$image))))

pdf(paste0(fig_dir, 'test_offset.pdf'), width=20, height=16)
p
dev.off()

```

```{r eval=FALSE}


library(tictoc)



# run BayesSpace clustering with q=15:
tic()
sce.combined = spatialCluster(sce.combined, use.dimred = "HARMONY", platform='Visium', q = 15, nrep = 5000) #use HARMONY
x <- toc() # took about 3 hours

sce.combined$spatial.cluster.q15 <- sce.combined$spatial.cluster
seurat_obj$bs.q15 <- sce.combined$spatial.cluster




p <- clusterPlot(sce.combined, color = NA) + #plot clusters
labs(title = "BayesSpace joint clustering")

pdf(paste0(fig_dir, 'BayesSpace_test.pdf'), width=10, height=10)
p
dev.off()

p <- ggplot(data.frame(reducedDim(sce.combined, "UMAP.HARMONY")),
  aes(x = UMAP1, y = UMAP2, color = factor(sce.combined$spatial.cluster))) +
  geom_point(size=0.2, alpha=0.75) +
  labs(color = "BayesSpace Clusters") + ggtitle('BayesSpace Clusters') +
  umap_theme + NoLegend()

pdf(paste0(fig_dir, 'baysespace_umap_cluster.pdf'), width=7, height=7)
p
dev.off()

# run BayesSpace clustering with q=10:
tic()
sce.combined = spatialCluster(sce.combined, use.dimred = "HARMONY", platform='Visium', q = 10, nrep = 5000, burn.in=500) #use HARMONY
x1 <- toc() # took about 3 hours
sce.combined$spatial.cluster.q10 <- sce.combined$spatial.cluster
seurat_obj$bs.q10 <- sce.combined$spatial.cluster

# run BayesSpace clustering with q=20:
tic()
sce.combined = spatialCluster(sce.combined, use.dimred = "HARMONY", platform='Visium', q = 20, nrep = 5000, burn.in=500) #use HARMONY
x2 <- toc() # took about 3 hours
sce.combined$spatial.cluster.q20 <- sce.combined$spatial.cluster
seurat_obj$bs.q20 <- sce.combined$spatial.cluster

saveRDS(sce.combined, file=paste0(data_dir, '5xFAD_bayesspace.rds'))





p1 <- clusterPlot(sce.combined, label = "spatial.cluster.q10", color=NA) +
  ggtitle('BayesSpace clusters, q=10')

p2 <- clusterPlot(sce.combined, label = "spatial.cluster.q15", color=NA) +
  ggtitle('BayesSpace clusters, q=15')

p3 <- clusterPlot(sce.combined, label = "spatial.cluster.q20", color=NA) +
  ggtitle('BayesSpace clusters, q=20')

pdf(paste0(fig_dir, 'baysespace_clusters.pdf'), width=10, height=10)
p1
p2
p3
dev.off()








# run bayesspace enhanced clustering:
# this can't run it is asking for 6 TB of RAM lmaoooo
tic()
sce.enhanced <- spatialEnhance(
  sce.combined,
  q = 15,
  d = 15, # number of components
  use.dimred = "HARMONY",
  platform = "Visium",
  nrep = 5000, burn.in = 1000,
  gamma=3, verbose=TRUE,
  jitter_scale=5.5, jitter_prior=0.3,
  save.chain=TRUE,
  chain.fname = 'test_bayesspace_mcmc.hdf5'
)
y <- toc()






```


Annotate clusters

```{r eval=FALSE}

anno_df <- read.csv("data/5xFAD_bayesspace_cluster_annotations.csv")

ix <- match(seurat_obj$bs.q15, anno_df$cluster)
seurat_obj$annotation <- anno_df$annotation[ix]


# plot umap colored by cluster
p <- DimPlot(seurat_obj, group.by='annotation', label=TRUE, raster=FALSE) + umap_theme

pdf(paste0(fig_dir, 'umap_annotation.pdf'), width=9, height=7, useDingbats=FALSE)
p
dev.off()

saveRDS(seurat_obj, paste0(data_dir,'5XFAD_seurat_processed.rds'))


```
