
Load the required libraries and the Seurat dataset

```{r eval=FALSE}

library(Seurat)
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)
library(tidyverse)
library(cowplot)
library(patchwork)
library(ggrepel)
library(viridis)
library(magrittr)
theme_set(theme_cowplot())

setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/milo/")


fig_dir <- 'figures/'
data_dir <- 'data/'

# load ADDS + AD integrated single-cell dataset:
seurat_obj <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_AD_integrated.rds" )

# add new column to collapse VASC clusters 
seurat_obj$cell_group <- ifelse(
    seurat_obj$cell_type %in% c('FBR', 'PER', 'SMC', 'END'),
    'VASC', seurat_obj$cell_type
)
seurat_obj$cell_group <- factor(
    as.character(seurat_obj$cell_group),
    levels = c('EX', 'INH', 'ODC', 'OPC', 'MG', 'ASC', 'VASC')
)

dx_cp <- c("Control" = "#B8DBC5", "Early-AD" = "#E7BDE1" , "AD" = "#CF8BA3", "DSAD" = "#9E6D7F")


subset(seurat_obj@meta.data, cell_type == 'INH') %>% .$cell_identity %>% droplevels %>% table 
subset(seurat_obj@meta.data, cell_group == 'INH') %>% .$cell_identity %>% droplevels %>% table 

#--------------------------------------------------------------------#
# Important meta-data variables
#--------------------------------------------------------------------#

# this seurat object contains data from our new AD/DS project and 3 other 
# studies. 
table(seurat_obj$Study)

# variable denoting the individual of origin 
table(seurat_obj$Sample)

# our AD/DS dataset has two different brain regions (frontal cortex, FCX
# and posterior cingulate cortex, PCC). 
table(seurat_obj$Tissue)

# annotated clusters 
table(seurat_obj$annotation)

# major cell type 
table(seurat_obj$cell_type)

# disease status 
# our new dataset has control and AD/DS
# Mathys et al has control, early-ad, and late AD (AD)
# other two studies have control and late AD
table(seurat_obj$Diagnosis)

```

We are trying to use MiloR to perform differential cell type composition analysis 
in order to respond to Reviewer 2's comment:

"Figure 1e – There seems to be no difference in coarse cluster structure of 
the snRNA-Seq data between control and AD, does higher resolution clustering 
show further differences?"

The Reviewer is (incorrectly) trying to infer differences in cell type composition 
between control and AD based on the UMAP plot, which is not possible. I think they 
don't really care about the UMAP itself rather than are interested in understanding 
if there are any significant changes in cell type / cluster proportions in diseae. 
We can address this point using MiloR.

MiloR can perform diffferential cell type composition analysis comparing two 
conditions. Since our dataset has 4 conditions (control, AD/DS, early-AD, 
and late-AD) and four datasets, and two brain regions, we will need to run separate 
MiloR analyses for the following groups.

* Current dataset, control vs AD/DS in FCX region 
* Current dataset, control vs AD/DS in PCC region 
* Mathys et al, control vs early AD 
* Mathys et al, control vs late AD 
* Morabito & Miyoshi et al, control vs AD 
* Zhou et al, control vs AD 

For this analysis we should use the "annotaiton" meta-data column for the cluster groups.

I suggest following this notebook written by the MiloR authors:
https://rawcdn.githack.com/MarioniLab/miloR/7c7f906b94a73e62e36e095ddb3e3567b414144e/vignettes/milo_gastrulation.html#1_Load_data

I would ultimately like to have the "Beeswarm" plots that are shown in this notebook
as the main result of this analysis (Controls should correspond to negative log fold 
change). 

In the next block I will subset the Seurat object for the AD/DS and FCX analysis 
and convert it to the right format for MiloR.

```{r eval=FALSE}

# subset the seurat object 
cur_seurat <- subset(seurat_obj, Study == 'ADDS' & Tissue == 'FCX')

# not run yet: subset for the mathys et al early AD vs Control
#cur_seurat <- subset(seurat_obj, Study == 'Mathys_2019' & Diagnosis %in% c("Control", 'Early-AD'))

# convert the dataset to singlecellexperiment:
sce <- as.SingleCellExperiment(cur_seurat)

# convert from SCE to milo object
milo_obj <- Milo(sce)

# compute knn
# for all purposes downstream where applicable use SCVI as the reduced dim
milo_obj <- buildGraph(milo_obj, k = 30, d = 30, reduced.dim = "SCANVI")


# continue the rest of the MiloR pipeline...


```

Plotting with one of Sai's outputs 

```{r eval=FALSE}

milo_dir <- '/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/milo/miloR_result/'
milo_outputs <- dir(milo_dir)

cur_milo_file <- milo_outputs[1]


for(cur_milo_file in milo_outputs){
    print(cur_milo_file)

    milo_obj <- readRDS(paste0(milo_dir, cur_milo_file, '/milo_obj.rds'))

    milo_obj$cell_group <- ifelse(
        milo_obj$cell_type %in% c('FBR', 'PER', 'SMC', 'END'),
        'VASC', milo_obj$cell_type
    )
    milo_obj$cell_group <- factor(
        as.character(milo_obj$cell_group),
        levels = c('EX', 'INH', 'ODC', 'OPC', 'MG', 'ASC', 'VASC')
    )

    # get the conditions
    conditions <- unique(as.character(milo_obj$Diagnosis))
    dx_condition <- conditions[conditions != 'Control']

    # change factor levels:
    milo_obj$Diagnosis <- factor(
        as.character(milo_obj$Diagnosis),
        levels = c('Control', dx_condition)
    )

    # make the design table
    design_df <- data.frame(colData(milo_obj))[,c("Sample", "Diagnosis")] %>% distinct()
    rownames(design_df) <- design_df$Sample

    # run the differential abundance test
    da_results <- testNhoods(
        milo_obj, design = ~ Diagnosis, 
        design.df = design_df,
        reduced.dim = "SCANVI"
    )

    # add neighborhood annotations
    da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cell_identity")
    da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cell_group")
    head(da_results)

    da_results$cell_group <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cell_group)
    da_results$cell_identity <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cell_identity)

    # remove the NA results:
    da_results %<>% subset(!is.na(cell_identity))

    write.csv(da_results, file=paste0(data_dir, 'milo_da_results_', cur_milo_file,'.csv'))

    #-------------------------------------------------------#
    # Making the beeswarm plot
    #-------------------------------------------------------#

    library(ggbeeswarm)

    # order them by major cell type and then by average log2FC
    groups <- rev(levels(seurat_obj$cell_group))
    df <- data.frame()
    plot_df <- do.call(rbind, lapply(groups, function(cur_group){
        subset(da_results, cell_group == cur_group) %>%
            group_by(cell_identity) %>%
            summarise(mean=median(logFC)) %>%
            arrange(mean)
    }))
    plot_df$cell_identity <- factor(
        as.character(plot_df$cell_identity),
        levels =  as.character(plot_df$cell_identity)
    )

    da_results$cell_identity <- factor(
        as.character(da_results$cell_identity),
        levels =  as.character(plot_df$cell_identity)
    )

      # remove the NA results:
    da_results %<>% subset(!is.na(cell_identity))

    length(unique(da_results$cell_identity))

    da_signif <- subset(da_results, FDR < 0.05)

    # max range of the plot 
    plot_range <- max(abs(da_results$logFC)) + 0.1

    p <- da_results %>%
    ggplot(aes(x = logFC, y = cell_identity)) +
    # geom_vline(xintercept=0, linetype='dashed', color='black') + 
    ggrastr::rasterise(geom_quasirandom(
        method = "pseudorandom",
        color = 'lightgrey'
        ), dpi=200) + 
    ggrastr::rasterise(geom_quasirandom(
        data = da_signif,
        aes(color = logFC),
        method = "pseudorandom"
        ), dpi=200) + 
    geom_boxplot(fill=NA, outlier.shape=NA) + 
    scale_color_gradient2() +
    xlim(-plot_range, plot_range) + 
    scale_color_gradient2(high=dx_cp[dx_condition], mid='lightgrey', low=dx_cp['Control'], midpoint=0) +
    ylab('') + 
    theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.border = element_rect(size=1,color='black', fill=NA),
        panel.grid.major.x = element_line(size-0.5, color='grey'),
        plot.margin=margin(0,0,0,0)
    )

    # make a color bar
    color_df <- read.csv(file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/cell_identity.csv')
    color_df <- subset(color_df, group %in% da_results$cell_identity)
    color_df$group <- factor(
    as.character(color_df$group),
    levels = levels(da_results$cell_identity)
    )
    color_df %<>% arrange(group)

    colorbar <- color_df %>%
        ggplot(aes(y=group, x=1, fill=group)) +
        geom_tile() +
        scale_fill_manual(values=color_df$colour) +
        NoLegend() +
        theme(
            plot.title=element_blank(),
            axis.line=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x = element_blank(),
            axis.title = element_blank(),
            plot.margin=margin(0,0,0,0)
        ) 
    
    patch <- colorbar + p + plot_layout(widths=c(1, 10))

    pdf(paste0(fig_dir, 'beeswarm_', cur_milo_file, '.pdf'), width=5, height=9)
    print(patch)
    dev.off()
}


```

Milo + Celltrek

```{r eval=FALSE}


milo_dir <- '/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/milo/miloR_result/'
milo_outputs <- dir(milo_dir)

cur_milo_file <- milo_outputs[1]


milo_obj <- readRDS(paste0(milo_dir, cur_milo_file, '/milo_obj.rds'))

#milo_obj <- milo_obj[,milo_obj$bc %in% seurat_obj_celltrek$bc]

ix <- match(as.character(milo_obj$bc), as.character(seurat_obj_celltrek$bc))
length(ix)

all.equal(as.character(milo_obj$bc), as.character(seurat_obj_celltrek$bc[ix]))

milo_obj$cellchat_anno <- seurat_obj_celltrek$cellchat_anno[ix]

sum(is.na(milo_obj$cellchat_anno))
milo_obj$cellchat_anno <- ifelse(is.na(milo_obj$cellchat_anno), 'unmapped', milo_obj$cellchat_anno)

milo_obj$cell_group <- ifelse(
    milo_obj$cell_type %in% c('FBR', 'PER', 'SMC', 'END'),
    'VASC', milo_obj$cell_type
)
milo_obj$cell_group <- factor(
    as.character(milo_obj$cell_group),
    levels = c('EX', 'INH', 'ODC', 'OPC', 'MG', 'ASC', 'VASC')
)

subset(colData(milo_obj), cell_group == 'INH') %>% .$cellchat_anno %>% table

# get the conditions
conditions <- unique(as.character(milo_obj$Diagnosis))
dx_condition <- conditions[conditions != 'Control']

# change factor levels:
milo_obj$Diagnosis <- factor(
    as.character(milo_obj$Diagnosis),
    levels = c('Control', dx_condition)
)

# make the design table
design_df <- data.frame(colData(milo_obj))[,c("Sample", "Diagnosis")] %>% distinct()
rownames(design_df) <- design_df$Sample

# run the differential abundance test
da_results <- testNhoods(
    milo_obj, design = ~ Diagnosis, 
    design.df = design_df,
    reduced.dim = "SCANVI"
)


# add neighborhood annotations
da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cellchat_anno")
da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cell_group")
head(da_results)

length(unique(da_results$cellchat_anno))


da_results$cell_group <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cell_group)
da_results$cellchat_anno <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cellchat_anno)

subset(da_results, cell_group == 'INH') %>% .$cellchat_anno %>% table

# remove the NA results:
da_results %<>% subset(!is.na(cellchat_anno))

write.csv(da_results, file=paste0(data_dir, 'milo_da_results_celltrek_anno.csv'))

da_results <- read.csv(file=paste0(data_dir, 'milo_da_results_celltrek_anno.csv'))


#-------------------------------------------------------#
# Making the beeswarm plot
#-------------------------------------------------------#

# note: some of the very small groups are missing from this analysis

groups_exclude <- c(names(which(table(da_results$cellchat_anno) < 10)), 'Mixed', 'unmapped')
da_results <- subset(da_results, ! cellchat_anno %in% groups_exclude)

length(unique(da_results$cellchat_anno))
length(unique(milo_obj$cellchat_anno))
length(unique(seurat_obj$cellchat_anno))

subset(seurat_obj@meta.data, cell_type == 'ASC') %>% .$cellchat_anno %>% table
subset(seurat_obj@meta.data, cell_type == 'MG') %>% .$cellchat_anno %>% table

library(ggbeeswarm)

# order them by major cell type and then by average log2FC
groups <- rev(levels(seurat_obj$cell_group))
df <- data.frame()
plot_df <- do.call(rbind, lapply(groups, function(cur_group){
    good_groups <- subset(seurat_obj@meta.data, cell_group == cur_group) %>% .$cellchat_anno %>% unique
    subset(da_results, cellchat_anno %in% good_groups) %>%
        group_by(cellchat_anno) %>%
        summarise(mean=median(logFC)) %>%
        arrange(mean)
}))
plot_df$cellchat_anno <- factor(
    as.character(plot_df$cellchat_anno),
    levels =  unique(as.character(plot_df$cellchat_anno))
)

da_results$cellchat_anno <- factor(
    as.character(da_results$cellchat_anno),
    levels =  unique(as.character(plot_df$cellchat_anno))
)

    # remove the NA results:
da_results %<>% subset(!is.na(cellchat_anno))


da_signif <- subset(da_results, FDR < 0.05)

# max range of the plot 
plot_range <- max(abs(da_results$logFC)) + 0.1

p <- da_results %>%
ggplot(aes(x = logFC, y = cellchat_anno)) +
# geom_vline(xintercept=0, linetype='dashed', color='black') + 
ggrastr::rasterise(geom_quasirandom(
    method = "pseudorandom",
    color = 'lightgrey'
    ), dpi=800) + 
ggrastr::rasterise(geom_quasirandom(
    data = da_signif,
    aes(color = logFC),
    method = "pseudorandom"
    ), dpi=800) + 
geom_boxplot(fill=NA, outlier.shape=NA) + 
scale_color_gradient2() +
xlim(-plot_range, plot_range) + 
scale_color_gradient2(high=dx_cp[dx_condition], mid='lightgrey', low=dx_cp['Control'], midpoint=0) +
ylab('') + 
theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.border = element_rect(size=1,color='black', fill=NA),
    panel.grid.major.x = element_line(size-0.5, color='grey'),
    plot.margin=margin(0,0,0,0)
)

pdf(paste0(fig_dir, 'beeswarm_celltrek2.pdf'), width=4, height=7)
print(p)
dev.off()


subset(seurat_obj_celltrek@meta.data, cell_type == 'ASC') %>% .$cellchat_anno %>% table



```


Run MiloR again but on the spatial annotations 

```{r eval=FALSE}

seurat_obj_celltrek <- readRDS(file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_for_cellchat.rds')
seurat_obj <- seurat_obj_celltrek


# add new column to collapse VASC clusters 
seurat_obj$cell_group <- ifelse(
    seurat_obj$cell_type %in% c('FBR', 'PER', 'SMC', 'END'),
    'VASC', seurat_obj$cell_type
)
seurat_obj$cell_group <- factor(
    as.character(seurat_obj$cell_group),
    levels = c('EX', 'INH', 'ODC', 'OPC', 'MG', 'ASC', 'VASC')
)

head(milo_obj$bc)
head(seurat_obj_celltrek$bc)


# convert the dataset to singlecellexperiment:
sce <- as.SingleCellExperiment(seurat_obj)

# convert from SCE to milo object
milo_obj <- Milo(sce)

# compute knn
# for all purposes downstream where applicable use SCVI as the reduced dim
milo_obj <- buildGraph(milo_obj, k = 30, d = 30, reduced.dim = "SCVI")

# defining neighborhoods
milo_obj <- makeNhoods(milo_obj, prop = 0.1, k = 30, d=30, refined = TRUE, reduced_dims = "SCVI")

# count cells in each neighborhood 
milo_obj <- countCells(
    milo_obj, 
    meta.data = as.data.frame(colData(milo_obj)), 
    sample="Sample"
)



# from above, need to modify to work for this data


# get the conditions
conditions <- unique(as.character(milo_obj$Diagnosis))
dx_condition <- conditions[conditions != 'Control']

# change factor levels:
milo_obj$Diagnosis <- factor(
    as.character(milo_obj$Diagnosis),
    levels = c('Control', dx_condition)
)

# make the design table
design_df <- data.frame(colData(milo_obj))[,c("Sample", "Diagnosis")] %>% distinct()
rownames(design_df) <- design_df$Sample

# run the differential abundance test
da_results <- testNhoods(
    milo_obj, design = ~ Diagnosis, 
    design.df = design_df,
    reduced.dim = "SCVI"
)

# add neighborhood annotations
da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cellchat_anno")
da_results <- annotateNhoods(milo_obj, da_results, coldata_col = "cell_group")
head(da_results)

da_results$cell_group <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cell_group)
da_results$cell_identity <- ifelse(da_results$cell_group_fraction < 0.7, "Mixed", da_results$cell_identity)

# remove the NA results:
da_results %<>% subset(!is.na(cell_identity))

write.csv(da_results, file=paste0(data_dir, 'milo_da_results_', cur_milo_file,'.csv'))




```