
```{r eval=FALSE}

library(tidyverse)
library(cowplot)
library(viridis)
library(Seurat)
library(ggrastr)
library(ggpubr)

theme_set(theme_cowplot())

setwd('~/swaruplab/smorabit/analysis/ADDS_2021/hyperion/')

data_dir <- 'data/'
fig_dir <- 'figures/'

# re-load seurat object:
seurat_obj <- readRDS(paste0(data_dir, 'hyperion_seurat_processed.rds'))


# load snRNA-seq data just to plot some expressions
seurat_rna <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_AD_integrated.rds" )


tmp <- distinct(seurat_obj@meta.data[,c('Diagnosis', 'sample', 'TargetRegion')])

```

make a table with all of the samples 

```{r eval=FALSE}

imc_files <- dir(data_dir)
imc_files <- imc_files[grepl('.tsv', imc_files)]
imc_files <- imc_files[!grepl('TEST', imc_files)]


df <- Reduce(rbind, lapply(imc_files, function(cur_file){
    tmp <- strsplit(cur_file, '\\[')[[1]]
    tmp1 <- tmp[1]; tmp2 <- tmp[2]
    roi <- strsplit(tmp2, '\\]')[[1]][1]

    if(grepl('final', cur_file)){
        slide <- paste0('Slide_', strsplit(tmp1, '_')[[1]][4])
    } else{
        slide <- strsplit(tmp1, 'IMC_')[[1]][2]
    }

    df <- read.delim(paste0(data_dir, cur_file), sep='\t')
    df$roi <- roi 
    df$slide <- slide 
    df$sample <- paste0(slide, '-', roi)
    df
}))



# exclude list: 
 
# slide 1: 
# 1,2,3,15

# slide 2: 
# 1,17

# slide 3:
# 3,4,7,10,16,17

exclude <- c(
    'Slide_1-ROI_001',
    'Slide_1-ROI_002',
    'Slide_1-ROI_003',
    'Slide_1-ROI_015',
    'Slide_2-ROI_001',
    'Slide_2-ROI_013',
    'Slide_2-ROI_017',
    'Slide_3-ROI_003',
    'Slide_3-ROI_004',
    'Slide_3-ROI_007',
    'Slide_3-ROI_010',
    'Slide_3-ROI_016',
    'Slide_3-ROI_017'
)
df <- df %>% subset(!(sample %in% exclude))
meta_df <- df[,c('slide', 'roi', 'sample', 'nuclei_X', 'nuclei_y')]
meta_df <- meta_df %>% dplyr::rename(c(x = nuclei_X, y = nuclei_y))

length(unique(df$sample))

```

Selecting markers 

```{r eval=FALSE}

# just get the whole cell features
# pca_df <- df[,grepl('Entire.cell', colnames(df))] 
# pca_df <- pca_df[,!grepl('Type', colnames(pca_df))]
# pca_df <- na.omit(pca_df)


test <- colnames(df)
#test <- test[grepl('Entire', test)]
test <- test[grepl('body', test)]
test <- test[2:length(test)]
df <- df[,test]

pattern <- "\\b\\.{4}\\b"

probe_names <- unlist(lapply(test, function(x){
    tmp <- strsplit(x, pattern)[[1]]
    tmp <- tmp[length(tmp)]
    tmp
}))
names(probe_names) <- colnames(df)
as.character(probe_names)

probes_keep <- c(
    'CD44', 'rPTPRM', 'Moesin', 'Cystatin_C',
    'Beta_Amyloid', 'CD68', 'Map2', 'ERBIN',
    'BIN1', 'CD163', 'GFAP', 'Foxp2',
    'PDGFRa', 'NeuN', 'APOE', 'Midkine', 'Olig2', 'CL.P1',
    'COL25A1', 'CX3CR1', 'GCP5', 'CALB1', 'Phospho_Tau_AT8',
    'Iba1', 'Mac_2', 'YKL.40', 'S100b', 'Netrin1', 'Apolipo_J_CLU'
)

probe_names <- probe_names[probe_names %in% probes_keep]
df <- df[,names(probe_names)]
colnames(df) <- as.character(probe_names)


```

Process data using Seurat 

```{r eval=FALSE}

# create the seurat object
mat <- t(df)
mat[is.na(mat)] <- 0
colnames(mat) <- 1:ncol(mat)
rownames(meta_df) <- 1:nrow(meta_df)
seurat_obj <- CreateSeuratObject(counts = mat, meta=meta_df)

# get ROI number:
seurat_obj$roi_num <- as.numeric(gsub('ROI_0', '', seurat_obj$roi))
seurat_obj$slide_num <- as.numeric(gsub('Slide_', '', seurat_obj$slide))
seurat_obj$slide_roi <- paste0(seurat_obj$slide_num, '-', seurat_obj$roi_num)


# additional metadata
slide_meta <- read.table('data/slide_meta.txt', sep = '\t', header=1)

slide_meta$roi_num <- as.numeric(unlist(lapply(slide_meta$TIFF.Name, function(x){
    tmp <- strsplit(x, '_')[[1]]
    tmp[length(tmp)]
})))
slide_meta$slide_roi <- paste0(slide_meta$SlideNo, '-', slide_meta$roi_num)

all(seurat_obj$slide_roi %in% slide_meta$slide_roi)

ix <- match(seurat_obj$slide_roi, slide_meta$slide_roi)
meta <- slide_meta[ix, c('TIFF.Name', 'TargetRegion', 'post', 'SlideSample')]
rownames(meta) <- colnames(seurat_obj)

seurat_obj$TargetRegion <- meta$TargetRegion
seurat_obj$post <- meta$post
seurat_obj$SlideSample <- meta$SlideSample
seurat_obj$TIFF.Name <- meta$TIFF.Name

seurat_obj$Slide_SlideSample <- paste0(seurat_obj$slide_num, '-', seurat_obj$SlideSample)

# load the other metadata 
meta <- read.csv('data/ADDS_Hyperion_slides.csv')
meta$Slide_SlideSample <- paste0(meta$SlideNo, '-', meta$SlideSampleNo)

ix <- match(seurat_obj$Slide_SlideSample, meta$Slide_SlideSample)
seurat_obj@meta.data<- cbind(seurat_obj@meta.data, meta[ix,c('Age', 'Sex', 'PMI', 'APoE', 'Tangle.Stage', 'Plaque.Stage', 'Diagnosis')])


###########################################################################################
# run Seurat 
###########################################################################################

seurat_obj <- NormalizeData(seurat_obj)
seurat_obj <- ScaleData(seurat_obj)
seurat_obj <- RunPCA(seurat_obj, features=rownames(seurat_obj))

p <- ElbowPlot(seurat_obj)

pdf(paste0(fig_dir, 'pca_elbow.pdf'), width=6, height=4)
p
dev.off()


seurat_obj <- RunUMAP(seurat_obj, dims=1:10)
seurat_obj <- FindNeighbors(seurat_obj, dims=1:10)
seurat_obj <- FindClusters(seurat_obj, resolution=1)


p <- DimPlot(seurat_obj)

pdf(paste0(fig_dir, 'seurat_umap.pdf'), width=7, height=7)
p
dev.off()

p1 <- DimPlot(seurat_obj, group.by = 'sample')
p2 <- DimPlot(seurat_obj, group.by = 'slide')

pdf(paste0(fig_dir, 'seurat_umap_samples.pdf'), width=14, height=6)
p1 | p2
dev.off()


######################################################################
# With harmony 
######################################################################

library(harmony)


# seurat_obj <- RunHarmony(seurat_obj, dims=1:10, group.by.vars = 'slide')
# seurat_obj <- RunHarmony(seurat_obj, dims=1:10, group.by.vars = 'Slide_SlideSample')
seurat_obj <- RunHarmony(seurat_obj, dims=1:10, group.by.vars = 'TIFF.Name')

seurat_obj <- RunUMAP(seurat_obj, reduction='harmony', dims=1:10, n.neighbors=25L, min.dist=0.05, spread=0.75)

seurat_obj <- FindNeighbors(seurat_obj, reduction='harmony', dims=1:10)
seurat_obj <- FindClusters(seurat_obj, resolution=0.75)

# remove the smallest cluster 
seurat_obj <- subset(seurat_obj, seurat_clusters != '12')
seurat_obj$seurat_clusters <- droplevels(seurat_obj$seurat_clusters)

# cluster dendro
seurat_obj <- BuildClusterTree(seurat_obj, dims=1:5, reorder=TRUE)


###########################################################################################
# plotting
###########################################################################################


pdf(paste0(fig_dir, 'seurat_clustertree.pdf'), width=4, height=7)
PlotClusterTree(seurat_obj, direction='rightwards')
dev.off()


p <- DimPlot(seurat_obj, label=TRUE)

pdf(paste0(fig_dir, 'seurat_umap_harmony.pdf'), width=7, height=7)
p
dev.off()

# p <- DimPlot(seurat_obj, label=TRUE, reduction='tsne')
# pdf(paste0(fig_dir, 'seurat_tsne_harmony.pdf'), width=7, height=7)
# p
# dev.off()


p <- DimPlot(seurat_obj, label=FALSE, split.by = 'seurat_clusters', ncol=5)

pdf(paste0(fig_dir, 'seurat_umap_harmony_split.pdf'), width=10, height=10)
p
dev.off()

p1 <- DimPlot(seurat_obj, group.by = 'sample')
p2 <- DimPlot(seurat_obj, group.by = 'slide')

pdf(paste0(fig_dir, 'seurat_umap_harmony_samples.pdf'), width=14, height=6)
p1 | p2
dev.off()

p1 <- DimPlot(seurat_obj, group.by = 'seurat_clusters', split.by = 'TargetRegion', label=TRUE)

pdf(paste0(fig_dir, 'seurat_umap_harmony_regions.pdf'), width=14, height=7)
p1 
dev.off()


p1 <- DimPlot(seurat_obj, group.by = 'sample', split.by='slide')

pdf(paste0(fig_dir, 'seurat_umap_harmony_samples_split.pdf'), width=14, height=6)
p1
dev.off()


p <- DimPlot(seurat_obj, group.by = 'Slide_SlideSample', split.by = 'Diagnosis') 
pdf(paste0(fig_dir, 'seurat_umap_sample_dx.pdf'), width=15, height=5)
p
dev.off()


p <- DotPlot(seurat_obj, features=rownames(seurat_obj)) + 
coord_flip() + RotatedAxis()

pdf(paste0(fig_dir, 'seurat_dotplot_harmony.pdf'), width=14, height=7)
p
dev.off()

p <- FeaturePlot(seurat_obj, features=rownames(seurat_obj), ncol=6, raster=TRUE)

pdf(paste0(fig_dir, 'seurat_featureplot_harmony.pdf'), width=18, height=15)
p
dev.off()

p <- VlnPlot(seurat_obj, features=rownames(seurat_obj), pt.size=0, stack=TRUE, flip=TRUE) + 
RotatedAxis() + NoLegend()

pdf(paste0(fig_dir, 'seurat_vlnplot_harmony.pdf'), width=14, height=14)
p
dev.off()



p <- VlnPlot(
  seurat_obj %>% subset(seurat_clusters %in% c(13, 15)), 
  features = 'Beta-Amyloid', group.by = 'Diagnosis', pt.size=0)

pdf(paste0(fig_dir, 'test_vln.pdf'), width=7, height=4)
p
dev.off()

subset(seurat_obj@meta.data, seurat_clusters == 15) %>% .$Diagnosis %>% table


p <- DotPlot(seurat_obj, features=rownames(seurat_obj), cluster.idents=TRUE) + 
coord_flip() + RotatedAxis()

pdf(paste0(fig_dir, 'seurat_dotplot.pdf'), width=14, height=7)
p
dev.off()

p <- FeaturePlot(seurat_obj, features=rownames(seurat_obj), ncol=6, raster=TRUE)

pdf(paste0(fig_dir, 'seurat_featureplot.pdf'), width=18, height=15)
p
dev.off()


```

Plotting the X/Y coords 

```{r eval=FALSE}

dir.create(paste0(fig_dir, 'coords/'))

images <- unique(seurat_obj$TIFF.Name)
cur_image <- images[1]
plot_df <- subset(seurat_obj@meta.data, TIFF.Name == cur_image)

for(cur_image in images){
  print(cur_image)
  p <- ggplot(plot_df, aes(x=x, y=y, color=seurat_clusters)) + 
    geom_point() + theme_void() + NoLegend()

  file_name <- paste0(cur_image, '_clusters.pdf')

  pdf(paste0(fig_dir, 'coords/', file_name), width=7, height=7)
  print(p)
  dev.off()
}


```

cluster Markers 

```{r eval-FALSE}

markers <- FindAllMarkers(
    seurat_obj,
    logfc.threshold = 0,
    min.pct = 0,
    min.cells.feature = 1,
    min.cells.group = 1,
    features = rownames(seurat_obj)
)
table(markers$cluster)

subset(markers, cluster == 11 & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(-avg_log2FC)
subset(markers, cluster == 5 & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(-avg_log2FC)

subset(markers, gene == 'Mac-2') %>% arrange(-avg_log2FC) %>%  head



# plot the marker gene heatmap:
markers$cluster <- factor(as.character(markers$cluster), levels = levels(Idents(seurat_obj)))


library(dendextend)

mat <- markers %>% 
    dplyr::select(c(cluster, gene, avg_log2FC)) %>% 
   spread(cluster, avg_log2FC) 
mat[is.na(mat)] <- 0
labels <-mat[,1]
mat <- as.matrix(mat[,-1])
rownames(mat) <- labels

dist_mat <- dist(mat, method = 'euclidean')
hclust_avg <- hclust(dist_mat, method = 'average')
dend <- as.dendrogram(hclust_avg)


pdf(paste0(fig_dir, 'test_hclust.pdf'), width=5, height=5)
plot(hclust_avg, cex=0.5, hang=-1)
dev.off()

labels(dend)

markers$gene <- factor(as.character(markers$gene), levels = labels(dend))

maxval <- 1; minval <- -1
plot_df <- markers
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC > maxval, maxval, plot_df$avg_log2FC)
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC < minval, minval, plot_df$avg_log2FC)

plot_df$textcolor <- ifelse(plot_df$avg_log2FC > 0.5, 'black', 'white')
plot_df$Significance <- ifelse(plot_df$p_val_adj > 0.05, 'x', NA)


p <- plot_df %>% 
  ggplot(aes(x=gene, y=cluster, fill=avg_log2FC)) +
  geom_tile() +
  #geom_text(label=plot_df$Significance, color=plot_df$textcolor) +
  scale_fill_gradient2(low='purple', mid='black', high='yellow') +
  RotatedAxis() +
 # labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.background = element_rect(fill='black', color='black'),
    #axis.title.x = element_blank(),
    #axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
   # axis.ticks.x = element_blank(),
   # axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() # + xlab(group1) + ylab(group2)


# Plot the result
pdf(paste0(fig_dir, 'marker_heatmap.pdf'),height=7, width=8)
p
dev.off()


```

Annotate clusters 

```{r eval=FALSE}

anno <- read.table('data/hyperion_clusters2.txt', sep='\t', header=1)

# anno$cluster <- factor(
#   as.character(anno$cluster),
#   levels = levels(Idents(seurat_obj))
# )

color_df <- dplyr::select(anno, c(annotation, color)) %>% dplyr::rename(group = annotation, colour = color)

anno <- anno %>% arrange(desc(cluster))
anno$annotation <- factor(anno$annotation, levels=unique(anno$annotation))

cp <- anno$color; names(cp) <- anno$annotation; 

ix <- match(as.numeric(as.character(seurat_obj$seurat_clusters)), anno$cluster)
seurat_obj$cell_type <- anno[ix, 'cell_type']
seurat_obj$annotation <- anno[ix, 'annotation']





p <- DimPlot(seurat_obj, group.by = 'cell_type', label=TRUE)
pdf(paste0(fig_dir, 'seurat_umap_celltype.pdf'), width=7, height=7)
p
dev.off()

p <- DimPlot(seurat_obj, group.by = 'annotation', label=TRUE, repel=TRUE) +
  scale_color_manual(values=cp)
pdf(paste0(fig_dir, 'seurat_umap_annotation.pdf'), width=10, height=7)
p
dev.off()


p <- DimPlot(seurat_obj, group.by = 'annotation', split.by = 'Diagnosis', label=TRUE, repel=TRUE) + NoLegend()
pdf(paste0(fig_dir, 'seurat_umap_annotation_dx.pdf'), width=10, height=7)
p
dev.off()



p <- DimPlot(seurat_obj, group.by = 'TIFF.Name', split.by = 'TIFF.Name', label=FALSE, repel=TRUE,) + NoLegend()
pdf(paste0(fig_dir, 'seurat_umap_tiff_slide.pdf'), width=18, height=18)
p
dev.off()


saveRDS(seurat_obj, paste0(data_dir, 'hyperion_seurat_processed.rds'))

```



Plotting for paper:

```{r eval=FALSE}

source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

library(hdWGCNA)



Idents(seurat_obj) <- seurat_obj$annotation

# cluster dendro

seurat_obj <- BuildClusterTree(seurat_obj, dims=1:5, reorder=TRUE)

seurat_obj$annotation <- factor(
  as.character(seurat_obj$annotation),
  levels = levels(Idents(seurat_obj))
)

###########################################################################################
# plotting
###########################################################################################


pdf(paste0(fig_dir, 'seurat_clustertree_annotation.pdf'), width=7, height=7)
PlotClusterTree(seurat_obj, direction='rightwards')
dev.off()






#################################################################################
# Plot the UMAP colored by cluster
#################################################################################

p <- PlotEmbedding(
  seurat_obj,
  group.by = 'annotation',
  raster_dpi = 500,
  raster_scale=0.5, point_size=0.5,
  plot_theme = umap_theme(),
  color_df = color_df
)

pdf(paste0(fig_dir, 'seurat_umap_annotation.pdf'), width=10, height=10)
p
dev.off()

#################################################################################
# Plot the UMAP colored by cluster split by DX
#################################################################################

p <- PlotEmbedding(
  seurat_obj,
  group.by = 'annotation',
  split.by = 'Diagnosis',
  raster_dpi = 500,
  raster_scale=0.5, point_size=0.5,
  plot_theme = umap_theme(),
  plot_under = TRUE,
  color_df = color_df
)

pdf(paste0(fig_dir, 'seurat_umap_annotation_dx.pdf'), width=15, height=5)
wrap_plots(p, ncol=3) + plot_layout(guides='collect')
dev.off()

#################################################################################
# Plot the UMAP colored by cluster split by Region
#################################################################################

p <- PlotEmbedding(
  seurat_obj,
  group.by = 'annotation',
  split.by = 'TargetRegion',
  raster_dpi = 500,
  raster_scale=0.5, point_size=0.5,
  plot_theme = umap_theme(),
  plot_under = TRUE,
  color_df = color_df
)

pdf(paste0(fig_dir, 'seurat_umap_annotation_region.pdf'), width=10, height=5)
wrap_plots(p, ncol=2) + plot_layout(guides='collect')
dev.off()












# plot_list <- FeatureEmbedding(
#   seurat_obj, 
#   features=rownames(seurat_obj), 
#   combine=FALSE,
#   raster=TRUE, 
#   slot="scale.data"
# )

# plot_list <- lapply(plot_list, function(x){
#   x <- x + scale_color_gradient2(high='red', mid='white', low='blue')
# })

# pdf(paste0(fig_dir, 'seurat_featureplot_harmony.pdf'), width=18, height=15)
# wrap_plots(plot_list, ncol=6)
# dev.off()



```

Make violin plots for each feature in each cluster 

```{r eval=FALSE}

library(patchwork)


seurat_obj$Diagnosis <- factor(
  as.character(seurat_obj$Diagnosis),
  levels = c("Control", 'AD', 'AD_DS')
)


st_diag_cp <- c("Control" = "#B8DBC5",  "AD" = "#CF8BA3", "AD_DS" = "#9E6D7F")
features <- rownames(seurat_obj)
features <- labels(dend) # run the marker gene hclust above to use this
clusters <- levels(seurat_obj$annotation)


for(cur_cluster in clusters){
  print(cur_cluster)
  cur_seurat <- subset(seurat_obj, annotation == cur_cluster)

  # if there's fewer than 25 cells, remove the groups:
  tmp <- table(cur_seurat$Diagnosis)
  remove <- names(which(tmp < 25))

  if(length(remove) > 0){
    cur_seurat <- cur_seurat[,!(cur_seurat$Diagnosis %in% remove)]
    cur_seurat$Diagnosis <- droplevels(cur_seurat$Diagnosis)
  }

  if(length(levels(cur_seurat$Diagnosis)) == 2){
    my_comparisons <- list(levels(cur_seurat$Diagnosis))
  } else{
    my_comparisons <- list( c('Control', 'AD'), c('AD', 'AD_DS'), c('Control', 'AD_DS') )
  }

  exp <- GetAssayData(cur_seurat, slot='data')
  plot_list <- lapply(features, function(cur_feature){
    plot_range <- range(exp[cur_feature,])
    VlnPlot(
        cur_seurat, 
        features = cur_feature, group.by = 'Diagnosis', pt.size=0) + 
        scale_fill_manual(values=st_diag_cp
        ) + 
        geom_boxplot(width=0.25, outlier.shape=NA, fill='white') + 
        stat_compare_means(
          comparisons=my_comparisons,
         # label = 'p.signif',
          test = 'wilcox'
          ) + 
        scale_y_continuous(limits=c(plot_range[1],plot_range[2]*1.33)) + 
        xlab('') + theme(
          axis.line.y = element_blank(),
          axis.line.x = element_blank(),
          panel.border = element_rect(linewidth=1, fill=NA, color='black'),
          panel.grid.major.y = element_line(color='lightgrey'),
          plot.margin=margin(0,0,0,0),
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          plot.title = element_text(vjust=-0.2)

        ) + ylab('') + NoLegend()
  })

  patch <-   wrap_plots(plot_list, ncol=6) + plot_annotation(
    title = cur_cluster
  ) & theme(plot.title=element_text(hjust=0.5))

  pdf(paste0(fig_dir, 'VlnPlots/', cur_cluster,'_vln_combine.pdf'), width=15, height=15)
  print(patch)
  dev.off()

}


```


Markers with anotations 


cluster Markers 

```{r eval-FALSE}

markers <- FindAllMarkers(
    seurat_obj,
    logfc.threshold = 0,
    min.pct = 0,
    min.cells.feature = 1,
    min.cells.group = 1,
    features = rownames(seurat_obj)
)
table(markers$cluster)

write.table(markers, file='data/imc_cluster_markers.tsv', quote=FALSE, row.names=FALSE, sep='\t')

markers <- read.table(file='data/imc_cluster_markers.tsv', header=1, sep='\t')

subset(markers, cluster == 'Neuron (Abeta+, APOE+, Map2+)' & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(desc(avg_log2FC)) %>% select(c(gene, avg_log2FC))
subset(markers, cluster == 'Neuron (Map2+)' & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(desc(avg_log2FC)) %>% select(c(gene, avg_log2FC))
subset(markers, cluster == 'ASC (GFAP+ , Tau+)' & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(desc(avg_log2FC)) %>% select(c(gene, avg_log2FC))

subset(markers, cluster == 'ASC (GFAP+)' & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(desc(avg_log2FC)) %>% select(c(gene, avg_log2FC))
subset(markers, cluster == 'MG (CD44+)' & p_val_adj < 0.05 & avg_log2FC > 0) %>% arrange(desc(avg_log2FC)) %>% select(c(gene, avg_log2FC))


test <- DotPlot(
  seurat_rna, 
  features=c(
    'CD44', 'CD163', 'CD68', 'CX3CR1', 'CLU',
    'ERBIN', 'BIN1', 'FOXP2', 'APP', 'MAPT', 'APOE', 'GFAP', 'S100B'),
  group.by = 'annotation'
  ) + coord_flip() + RotatedAxis()


pdf(paste0(fig_dir, 'snRNA_dotplot.pdf'), width=12, height=8)
test
dev.off()

# plot the marker gene heatmap:

library(dendextend)

mat <- markers %>% 
    dplyr::select(c(cluster, gene, avg_log2FC)) %>% 
   spread(cluster, avg_log2FC) 
mat[is.na(mat)] <- 0
labels <-mat[,1]
mat <- as.matrix(mat[,-1])
rownames(mat) <- labels

dist_mat <- dist(mat, method = 'euclidean')
hclust_avg <- hclust(dist_mat, method = 'average')
dend <- as.dendrogram(hclust_avg)


pdf(paste0(fig_dir, 'test_hclust_anno.pdf'), width=5, height=5)
plot(hclust_avg, cex=0.5, hang=-1)
dev.off()

labels(dend)

markers$gene <- factor(as.character(markers$gene), levels = labels(dend))

maxval <- 1; minval <- -1
plot_df <- markers
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC > maxval, maxval, plot_df$avg_log2FC)
plot_df$avg_log2FC <- ifelse(plot_df$avg_log2FC < minval, minval, plot_df$avg_log2FC)

plot_df$textcolor <- ifelse(plot_df$avg_log2FC > 0.5, 'black', 'white')
plot_df$Significance <- ifelse(plot_df$p_val_adj > 0.05, 'x', NA)


p <- plot_df %>% 
  ggplot(aes(x=gene, y=cluster, fill=avg_log2FC)) +
  geom_tile() +
  #geom_text(label=plot_df$Significance, color=plot_df$textcolor) +
  scale_fill_gradient2(low='purple', mid='black', high='yellow') +
  RotatedAxis() +
 # labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', size=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.background = element_rect(fill='black', color='black'),
    #axis.title.x = element_blank(),
    #axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
   # axis.ticks.x = element_blank(),
   # axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal() # + xlab(group1) + ylab(group2)


# Plot the result
pdf(paste0(fig_dir, 'marker_heatmap_anno.pdf'),height=10, width=10)
p
dev.off()


```




Stacked bar plots 

```{r eval=FALSE}



################################################################################
# snRNA-seq Proportion of Condition
################################################################################


seurat_meta <- seurat_obj@meta.data
batches <- seurat_meta$Diagnosis %>% unique %>% as.character
cur_clusters  <- seurat_meta$annotation %>% unique %>% as.character

scale_vec <- table(seurat_meta$Diagnosis)

proportion_df <- data.frame()
for(i in 1:length(cur_clusters)){
  cluster_meta <- subset(seurat_meta, annotation == cur_clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$Diagnosis)) %>% dplyr::rename(Count = Freq)

  # compute the non-normalized % of cells in each group
  cur_df$Freq <- cur_df$Count / sum(cur_df$Count)

  # scale frequency to the total number of clusters in each Diagnosis
  cur_df$Scaled <- cur_df$Count / scale_vec
  cur_df$Scaled <- cur_df$Scaled / sum(cur_df$Scaled)

  # add to ongoing proportion df
  cur_df$cluster <- cur_clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, Diagnosis = Var1)
proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_meta$annotation))

p <- ggplot(proportion_df, aes(y=Scaled, x=cluster, fill=Diagnosis)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=c('gray65', 'gray45')) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y =  element_blank(),
    axis.line.x = element_blank(),
    panel.border = element_rect(size=1, fill=NA, color='black')
  ) +
  scale_fill_manual(values=st_diag_cp)



pdf(paste0(fig_dir, 'cluster_proportion_Diagnosis_scaled.pdf'), height=4, width=10)
print(p)
dev.off()

p <- ggplot(proportion_df, aes(y=Freq, x=cluster, fill=Diagnosis)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=c('gray65', 'gray45')) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y =  element_blank(),
    axis.line.x = element_blank(),
    panel.border = element_rect(size=1, fill=NA, color='black')
  ) +
  scale_fill_manual(values=st_diag_cp)


pdf(paste0(fig_dir, 'cluster_proportion_Diagnosis_raw.pdf'), height=4, width=10)
print(p)
dev.off()





################################################################################
# snRNA-seq Proportion of Sample
################################################################################

seurat_obj$sample <- as.factor(seurat_obj$sample)

seurat_meta <- seurat_obj@meta.data
batches <- seurat_meta$sample %>% unique %>% as.character
cur_clusters  <- seurat_meta$annotation %>% unique %>% as.character

scale_vec <- table(seurat_meta$sample)

proportion_df <- data.frame()
for(i in 1:length(cur_clusters)){
  cluster_meta <- subset(seurat_meta, annotation == cur_clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$sample)) %>% dplyr::rename(Count = Freq)

  # compute the non-normalized % of cells in each group
  cur_df$Freq <- cur_df$Count / sum(cur_df$Count)

  # scale frequency to the total number of clusters in each sample
  cur_df$Scaled <- cur_df$Count / scale_vec
  cur_df$Scaled <- cur_df$Scaled / sum(cur_df$Scaled)

  # add to ongoing proportion df
  cur_df$cluster <- cur_clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, sample = Var1)
proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_meta$annotation))

p <- ggplot(proportion_df, aes(y=Scaled, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  geom_hline(yintercept = 0.5, linetype='dashed') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=c('gray65', 'gray45')) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_line(colour = "black"),
    axis.line.x = element_blank()
  )

pdf(paste0(fig_dir, 'cluster_proportion_sample_scaled.pdf'), height=4, width=10)
print(p)
dev.off()

p <- ggplot(proportion_df, aes(y=Freq, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  geom_hline(yintercept = 0.5, linetype='dashed') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=c('gray65', 'gray45')) +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_line(colour = "black"),
    axis.line.x = element_blank()
  )

pdf(paste0(fig_dir, 'cluster_proportion_sample_raw.pdf'), height=4, width=10)
print(p)
dev.off()





################################################################################
# Proportion of Sample in each cluster
################################################################################

seurat_meta <- seurat_obj@meta.data
samples <- seurat_obj$sample %>% unique %>% as.character
clusters <- seurat_obj$annotation %>% unique


proportion_df <- data.frame()
for(i in 1:length(clusters)){
  cluster_meta <- subset(seurat_meta, annotation == clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$sample)/dim(cluster_meta)[1])
  print(sum(cur_df$Freq))
  cur_df$cluster <- clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, sample = Var1)
proportion_df <- dplyr::rename(proportion_df, fraction = Freq)
#proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_obj$cluster_name))

p <- ggplot(proportion_df, aes(y=fraction, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=color.scheme) +
  #theme_minimal() +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank()
  ) 
pdf(paste0(fig_dir, "/sample_clusters_barplot.pdf"), height=4, width=10)
print(p)
dev.off()


################################################################################
# Proportion of Sample in each cluster
################################################################################

seurat_meta <- seurat_obj@meta.data
samples <- seurat_obj$Diagnosis %>% unique %>% as.character
clusters <- seurat_obj$annotation %>% unique


proportion_df <- data.frame()
for(i in 1:length(clusters)){
  cluster_meta <- subset(seurat_meta, annotation == clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$Diagnosis)/dim(cluster_meta)[1])
  print(sum(cur_df$Freq))
  cur_df$cluster <- clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, sample = Var1)
proportion_df <- dplyr::rename(proportion_df, fraction = Freq)
#proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_obj$cluster_name))

p <- ggplot(proportion_df, aes(y=fraction, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=color.scheme) +
  #theme_minimal() +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank()
  ) 
pdf(paste0(fig_dir, "/condition_clusters_barplot.pdf"), height=4, width=10)
print(p)
dev.off()


################################################################################
# Proportion of GM/WM in each cluster
################################################################################

seurat_meta <- seurat_obj@meta.data
samples <- seurat_obj$TargetRegion %>% unique %>% as.character
clusters <- seurat_obj$annotation %>% unique


proportion_df <- data.frame()
for(i in 1:length(clusters)){
  cluster_meta <- subset(seurat_meta, annotation == clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$TargetRegion)/dim(cluster_meta)[1])
  print(sum(cur_df$Freq))
  cur_df$cluster <- clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, sample = Var1)
proportion_df <- dplyr::rename(proportion_df, fraction = Freq)
#proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_obj$cluster_name))

p <- ggplot(proportion_df, aes(y=fraction, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=color.scheme) +
  #theme_minimal() +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank()
  ) 
pdf(paste0(fig_dir, "/region_clusters_barplot.pdf"), height=4, width=10)
print(p)
dev.off()

################################################################################
# Proportion of slidein each cluster
################################################################################

seurat_meta <- seurat_obj@meta.data
samples <- seurat_obj$slide %>% unique %>% as.character
clusters <- seurat_obj$annotation %>% unique


proportion_df <- data.frame()
for(i in 1:length(clusters)){
  cluster_meta <- subset(seurat_meta, annotation == clusters[i])
  cur_df <- as.data.frame(table(cluster_meta$slide)/dim(cluster_meta)[1])
  print(sum(cur_df$Freq))
  cur_df$cluster <- clusters[i]
  proportion_df <- rbind(proportion_df, cur_df)
}

proportion_df <- dplyr::rename(proportion_df, sample = Var1)
proportion_df <- dplyr::rename(proportion_df, fraction = Freq)
#proportion_df$cluster <- factor(proportion_df$cluster, levels=levels(seurat_obj$cluster_name))



p <- ggplot(proportion_df, aes(y=fraction, x=cluster, fill=sample)) +
  geom_bar(stat='identity') +
  scale_y_continuous(expand = c(0,0)) +
  #scale_fill_manual(values=color.scheme) +
  #theme_minimal() +
  theme(
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    axis.line.y = element_blank(),
    axis.line.x = element_blank()
  ) 
pdf(paste0(fig_dir, "/slide_clusters_barplot.pdf"), height=4, width=10)
print(p)
dev.off()



```







What did the pixie downstream processing notebook do? 

Should we just use the cell-type markers for PCA + UMAP?

Should I try to include all of the images together for this?

```{r eval=FALSE}

colnames(df)

colnames(df)[grepl('Map2', colnames(df))]
colnames(df)[grepl('S100b', colnames(df))]
colnames(df)[grepl('Iba', colnames(df))]
colnames(df)[grepl('GFAP', colnames(df))]

quantile(na.omit(df$Biomarker.intensity..Entire.cell....Yb.173....S100b))
sum(na.omit(df$Biomarker.intensity..Entire.cell....Yb.173....S100b) > 0)
min(na.omit(df$Biomarker.intensity..Entire.cell....Yb.173....S100b))



quantile(na.omit(df$Biomarker.intensity..Cell.body....Gd.156....GFAP))
sum(na.omit(df$Biomarker.intensity..Entire.cell....Yb.173....S100b) > 0)
min(na.omit(df$Biomarker.intensity..Entire.cell....Yb.173....S100b))


p <- df %>%
    ggplot(aes(
        x=Biomarker.intensity..Entire.cell....Yb.173....S100b, 
        y=Biomarker.intensity..Entire.cell....Gd.156....GFAP, 
        color=sample)) + 
    rasterise(geom_point(), dpi=500) +
    xlab('S100b') + 
    ylab('GFAP') +
    RotatedAxis()
    #scale_color_viridis() +
    #theme_void() 



pdf(paste0(fig_dir, 'test_imc_scatter.pdf'), width=15, height=15)
p + facet_wrap(~sample, ncol=8)
dev.off()


#############################################################
# Try to run PCA on the cells
############################################################

library(harmony)

# just get the whole cell features
pca_df <- df[,grepl('Entire.cell', colnames(df))] 
pca_df <- pca_df[,!grepl('Type', colnames(pca_df))]
pca_df <- na.omit(pca_df)

plot_list <- list()
samples <- unique(df$sample)
for(cur_sample in samples){
    print(cur_sample)
    cur_df <- subset(df, sample == cur_sample)

    tmp <- pca_df[rownames(pca_df) %in% rownames(cur_df),]
    plot_df <- reshape2::melt(tmp)


plot_df$variable <- gsub('[BiomarkerintensityEntirecell]', '', as.character(plot_df$variable))

    p <- ggplot(
        plot_df,
        aes(x = variable, y=value, fill=variable)
    ) + 
    geom_violin() + 
    RotatedAxis() + 
    NoLegend() + 
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text.x = element_text(size = 6)
    )

    plot_list[[cur_sample]] <- p

}



pdf(paste0(fig_dir, 'test_imc_vln.pdf'), width=16, height=8)
for(i in 1:length(plot_list)){
    print(plot_list[[i]] + facet_wrap(~ variable, scales='free', ncol=10) + ggtitle(samples[i]))
}
dev.off()


scale_df <- scale(pca_df)
pcs <- svd(scale_df)
pc_mat <- pcs$u[,1:15]

# Harmony?
# harm <- HarmonyMatrix(
#     pc_mat,
#     meta_data = df[rownames(pca_df),],
#     vars_use = 'sample',
#     do_pca=FALSE
# )

umap_coords <- uwot::umap(
    X = pc_mat,
      #scale_df,
      #X = harm,
      min_dist = 0.2,
      n_neighbors= 25,
      metric = 'cosine',
      spread=0.5
    )

plot_df <- umap_coords %>% as.data.frame()
plot_df$feat <- df[rownames(pca_df),'Biomarker.intensity..Entire.cell....Yb.173....S100b']

p <- plot_df %>%
    ggplot(aes(x=V1, y=V2, color=feat)) + 
    rasterise(geom_point(), dpi=500) + 
    scale_color_viridis() +
    theme_void() 



pdf(paste0(fig_dir, 'test_imc_umap2.pdf'), width=5, height=5)
p
dev.off()



plot_df <- umap_coords %>% as.data.frame()
plot_df$sample <- df[rownames(pca_df),'sample']

p <- plot_df %>%
    ggplot(aes(x=V1, y=V2, color=sample)) + 
    rasterise(geom_point(), dpi=500) + 
    #scale_color_viridis() +
    theme_void() 



pdf(paste0(fig_dir, 'test_imc_umap_sample.pdf'), width=10, height=5)
p
dev.off()


```