
# Load libraries and visium samples

```{r eval=FALSE}
# conda activate cicero
library(Seurat)
library(harmony)
library(tidyverse)
library(cowplot)
library(patchwork)
library(RColorBrewer)
colfunc <- colorRampPalette(rev(brewer.pal(11, 'Spectral' )))
theme_set(theme_cowplot())

setwd("/dfs3b/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/July_2021")

fig_dir <- "figures/"
data_dir <- "data/"

umap_theme <- theme(
  axis.line=element_blank(),
  axis.text.x=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks=element_blank(),
  axis.title.x=element_blank(),
  axis.title.y=element_blank(),
  panel.background=element_blank(),
  panel.border=element_blank(),
  panel.grid.major=element_blank(),
  panel.grid.minor=element_blank(),
  plot.background=element_blank()
)

# re-load data:
seurat_obj <- readRDS(file=paste0(data_dir, '5XFAD_seurat_processed.rds'))


################################################################################
# Load batch 1 data into Seurat:
################################################################################

spaceranger_outdir <- '~/swaruplab/smorabit/data/Visium_Mouse_2021/spaceranger_count/5XFAD_samples/'
spaceranger_samples <- dir(spaceranger_outdir)

# load metadata:
meta <- read.csv('~/swaruplab/smorabit/analysis/5XFAD_visium_2021/pilot/data/pilot_samples.csv', stringsAsFactors=FALSE)
rownames(meta) <- meta$SAMPLE

seurat_list <- list()
for(sample in spaceranger_samples){

  print(sample)
  cur <- Seurat::Load10X_Spatial(paste0(spaceranger_outdir, sample, '/outs/'))
  cur$SAMPLE <- sample

  # Add metadata:
  cur$Condition <- meta[sample,'Condition']
  cur$Age <- meta[sample,'Age']
  cur$Slide <- 'V10M17-063'

  # add basic metadata cols:
  seurat_list[[sample]] <- cur
}


################################################################################
# Load July 2021 data into Seurat:
################################################################################

spaceranger_outdir <- '~/swaruplab/smorabit/data/Visium_Mouse_2021/July_2021/spaceranger_count/'
spaceranger_samples <- dir(spaceranger_outdir)

# load metadata:
meta <- read.csv('~/swaruplab/smorabit/data/Visium_Mouse_2021/July_2021/5XFAD_8samples.csv', stringsAsFactors=FALSE)
meta <- meta[1:8,]
meta$visium_sample <- paste0('Visium', meta$X)
rownames(meta) <- meta$visium_sample

# only keep important rows:
meta <- meta %>% select(c(SAMPLE, Condition, Sex, Age, CaptureArea, Slide, visium_sample ))

for(sample in spaceranger_samples){

  print(sample)
  cur <- Seurat::Load10X_Spatial(paste0(spaceranger_outdir, sample, '/outs/'))

  # Add metadata:
  for(meta_col in names(meta)){
    cur@meta.data[[meta_col]] <- meta[sample, meta_col]
  }

  # add basic metadata cols:
  seurat_list[[sample]] <- cur
}

# merge into one seurat obj:
seurat_obj <- merge(seurat_list[[1]], seurat_list[2:length(seurat_list)])
rm(seurat_list); gc();


```



# Processing

```{r eval=FALSE}

# QC plots
plot_list <- list()
for(sample in unique(seurat_obj$SAMPLE)){
  cur <- subset(seurat_obj, SAMPLE == sample)
  cur_condition <- as.character(unique(cur$Condition))
  cur_image <- names(cur@images)[sapply(names(cur@images), function(x){nrow(cur@images[[x]]@coordinates) > 0})]
  cur@images <- list(cur_image = cur@images[[cur_image]])
  plot_list[[sample]] <- SpatialFeaturePlot(
    cur, features='nCount_Spatial'
  ) +
    ggtitle(paste0('Sample ', sample, ' (',cur_condition, ')')) +
    scale_color_gradientn(colors=colfunc(256), guide = guide_colorbar(barwidth=15, barheight=0.5, ticks=FALSE, angle=30)) +
    theme(legend.position='bottom', legend.title = element_blank())
}

pdf(paste0(fig_dir, 'qc.pdf'), width=12, height=9, useDingbats=FALSE)
# patch1 <- wrap_plots(plot_list[1:2], ncol=2)
# patch2 <- wrap_plots(plot_list[3:4], ncol=2)
# print(patch1 / patch2 + plot_annotation(title='n_count_spatial'))
wrap_plots(plot_list, ncol=4)
dev.off()

# normalization:
seurat_obj <- NormalizeData(seurat_obj)
seurat_obj <- FindVariableFeatures(seurat_obj, nfeatures=3000)
seurat_obj <- ScaleData(seurat_obj, features=VariableFeatures(seurat_obj))

# dim reduction and clustering
seurat_obj <- RunPCA(seurat_obj, verbose = FALSE)
seurat_obj <- RunHarmony(seurat_obj, group.by.vars='Slide', assay='Spatial')
seurat_obj <- RunUMAP(seurat_obj, reduction = "harmony", dims = 1:30)
seurat_obj <- FindNeighbors(seurat_obj, reduction = "harmony", dims = 1:30)
seurat_obj <- FindClusters(seurat_obj, verbose = FALSE, res=0.5)

# add column for Time + Condition
seurat_obj$Condition.Age <- paste0(as.character(seurat_obj$Condition), '_', as.character(seurat_obj$Age))

saveRDS(seurat_obj, file=paste0(data_dir, '5XFAD_seurat_processed.rds'))

# plot UMAP
p1 <- DimPlot(seurat_obj, group.by='seurat_clusters', reduction = "umap", label = TRUE) + umap_theme
p2 <- DimPlot(seurat_obj, group.by='SAMPLE', reduction = "umap", label = FALSE) + umap_theme

pdf(paste0(fig_dir, 'umap_clusters.pdf'), width=12, height=5, useDingbats=FALSE)
p1 | p2
dev.off()


# plot clusters on spatial image:
p1 <- DimPlot(seurat_obj, group.by='seurat_clusters', reduction = "umap", label = TRUE) + umap_theme
g <- ggplot_build(p1)
colors <- g$data[[1]]["colour"]
groups <- g$data[[1]]['group']
color_df <- unique(data.frame(colors, groups)) %>% arrange(group)
color_df$seurat_clusters <- levels(seurat_obj$seurat_clusters)
colors.use <- color_df$colour
names(colors.use) <- color_df$seurat_clusters

# plot clusters as spatial dim plot:
plot_list <- list()
plot_list_tissue <- list()
for(sample in unique(seurat_obj$SAMPLE)){

  print(sample)

  # subset by this sample
  cur <- subset(seurat_obj, SAMPLE == sample)
  cur_condition <- as.character(unique(cur$Condition))

  # only keep current image:
  cur_image <- names(cur@images)[sapply(names(cur@images), function(x){nrow(cur@images[[x]]@coordinates) > 0})]
  cur@images <- list(cur_image = cur@images[[cur_image]])

  # plot clusters
  plot_list[[sample]] <- SpatialDimPlot(
    cur,
    group.by='seurat_clusters',
    label = TRUE, label.size = 3
  ) + NoLegend() + ggtitle(paste0(sample, ' (',cur_condition, ')')) + scale_fill_manual(values=colors.use)

  plot_list_tissue[[sample]] <- SpatialFeaturePlot(cur,features='Mobp', alpha=0) + NoLegend() + ggtitle(paste0(sample, ' (',cur_condition, ')'))

}


pdf(paste0(fig_dir, 'spatial_clusters_seurat_obj.pdf'), width=12, height=12, useDingbats=FALSE)
wrap_plots(plot_list, ncol=2)
dev.off()


# plot UMAP + spatial
p1 <- DimPlot(seurat_obj, group.by='seurat_clusters', reduction = "umap", label = TRUE) +
theme(
  legend.position='bottom',
  axis.ticks=element_blank(),
  axis.text.x=element_blank(),
  axis.text.y=element_blank(),
) + ylab('UMAP 2') + xlab("UMAP 1")

pdf(paste0(fig_dir, 'clusters_spatial.pdf'), width=18, height=8, useDingbats=FALSE)
p1 + wrap_plots(plot_list, ncol=2) + wrap_plots(plot_list_tissue, ncol=2) + plot_layout(widths=c(1,1,1))
dev.off()


```

Spatial expression plot of ERBIN:

```{r eval=FALSE}

plot_list <- list()
pdf(paste0(fig_dir, 'Erbin_spatial.pdf'), width=6, height=7, useDingbats=FALSE)
for(sample in unique(seurat_obj$SAMPLE)){
  cur <- subset(seurat_obj, SAMPLE == sample)
  cur_condition <- cur$Condition %>% unique
  cur_age <- cur$Age %>% unique
  cur_image <- names(cur@images)[sapply(names(cur@images), function(x){nrow(cur@images[[x]]@coordinates) > 0})]
  cur@images <- list(cur_image = cur@images[[cur_image]])
  p <- SpatialFeaturePlot(cur, features='Erbin') +
    ggtitle(paste0('Sample ', sample, ', ', cur_condition, ', ', cur_age))
  print(p)
  plot_list[[sample]] <- p
}
dev.off()

pdf(paste0(fig_dir, 'Erbin_spatial_combined.pdf'), width=12, height=12, useDingbats=FALSE)
wrap_plots(plot_list, ncol=4)
dev.off()

```

Differential expression analysis

* ran the cluster_markers.sub script to find marker DEGs for clusters
* ran 5xFAD 8mo vs 5xFAD 4 mo

```{r eval=FALSE}

###########################################################################
# marker DEGs:
###########################################################################

DEG_dir <- "/dfs3b/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/July_2021/DEGs/data/cluster_markers/"
DEG_tests <- dir(DEG_dir)

# combine  all tests into one table:
combined <- Reduce(rbind, lapply(dir(DEG_dir), function(file){
  read.csv(paste0(DEG_dir, file))
}))
write.csv(combined, file=paste0('DEGs/data/cluster_markers.csv'), quote=FALSE, row.names=FALSE)

###########################################################################
# 5xFAD 8mo vs 4mo
###########################################################################

DEG_dir <- "/dfs3b/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/July_2021/DEGs/data/5XFAD_8mo_vs_4mo/"
DEG_tests <- dir(DEG_dir)

# combine  all tests into one table:
combined <- Reduce(rbind, lapply(dir(DEG_dir), function(file){
  read.csv(paste0(DEG_dir, file))
})) %>% subset(p_val_adj < 0.05)
write.csv(combined, file=paste0('DEGs/data/5XFAD_8mo_vs_4mo.csv'), quote=FALSE, row.names=FALSE)


```

DEGs that are in the cortex that are different between 5x 8mo & 4mo:

```{r eval=FALSE}

marker_degs <- read.csv(file='DEGs/data/cluster_markers.csv')
age_degs <- read.csv(file='DEGs/data/5XFAD_8mo_vs_4mo.csv')

cur_cluster <- 11

cur_markers <- marker_degs %>% subset(group == cur_cluster)
cur_degs <- age_degs %>% subset(group == cur_cluster)

top_markers <- cur_markers %>% top_n(250, wt=abs(avg_log2FC)) %>% .$ gene %>% unique

cur_degs %>% subset(gene %in% top_markers) %>% top_n(25, wt=abs(avg_log2FC)) %>% arrange(avg_log2FC)


cur_gene <- 'Bst2'
cur_gene <- 'Ifitm3'
cur_gene <- 'Tmem107'

plot_list <- list()
pdf(paste0(fig_dir, cur_gene,'_spatial_genes.pdf'), width=6, height=7, useDingbats=FALSE)
for(sample in unique(seurat_obj$SAMPLE)){
  cur <- subset(seurat_obj, SAMPLE == sample)
  cur_condition <- cur$Condition %>% unique
  cur_age <- cur$Age %>% unique
  cur_image <- names(cur@images)[sapply(names(cur@images), function(x){nrow(cur@images[[x]]@coordinates) > 0})]
  cur@images <- list(cur_image = cur@images[[cur_image]])
  p <- SpatialFeaturePlot(cur, features=cur_gene, alpha=c(0.1,1)) +
    ggtitle(paste0('Sample ', sample, ', ', cur_condition, ', ', cur_age))
  print(p)
  plot_list[[sample]] <- p
}
dev.off()


pdf(paste0(fig_dir, cur_gene,'_spatial_combined.pdf'), width=12, height=12, useDingbats=FALSE)
wrap_plots(plot_list, ncol=4)
dev.off()




```



5xFAD 8 months vs 4 months for each cluster
Overlap with cluster marker genes
send vivek 2 up and 2 down plotted on the spatial slide

```{r eval=FALSE}



```
