```{r eval=FALSE}

library(Seurat)
#library(tidyverse)
library(dplyr)
library(ggplot2)
library(cowplot)
library(Matrix)
library(viridis)
library(harmony)
library(ggpubr)
library(patchwork)
library(RColorBrewer)
library(ggrepel)
library(CellTrek)
library(hdWGCNA)

#source("/pub/smorabit/hdWGCNA/bin/spatial_functions.R")
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')
source('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/amyloid/bin/Visium_amyloid_function.R')


setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/amyloid/")
fig_dir <- "figures/"
data_dir <- "data/"


# load visium & split-seq seurat objects:
seurat_vis <- readRDS("/dfs7/swaruplab/emiyoshi/Visium_ADDS/ADDS_seurat_processed_annotated.rds")
seurat_vis$Diagnosis <- factor(
  as.character(seurat_vis$Diagnosis),
  levels = c("Control", "earlyAD", "AD", "AD_DS")
)
seurat_vis$SlideCap <- paste0(as.character(seurat_vis$Slide), '_', as.character(seurat_vis$CaptureArea))

dplyr::select(seurat_vis@meta.data, c(Sample, Slide, CaptureArea, seqbatch)) %>% distinct()



# amyloid visium seurat object
seurat_amyloid <- readRDS("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/voyager/data/ADDS_seurat_processed_geospatial.rds")

# seurat_obj <- readRDS("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/data/ADDS_integrated_scvi.rds")

# load ADDS dataset:
seurat_obj <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_integrated.rds" )

# only get the FCX:
seurat_obj <- subset(seurat_obj, Tissue != 'PCC')

# saveRDS(seurat_obj, file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_integrated_FCX.rds")

# add barcode:
seurat_obj$bc <- colnames(seurat_obj)
seurat_vis$bc <- colnames(seurat_vis)

var_df <- read.delim(file = "/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/ADDS_subclusters_var.tsv", sep='\t')

VariableFeatures(seurat_obj) <- var_df$X


# representative samples
human_rep_samples <- c(
  'Dec_13_2021_Human5', 'Dec_20_2021_Human1',
  'Dec_13_2021_Human6', 'Oct_2021_6',
  'Dec_13_2021_Human3', 'Dec_13_2021_Human7',
  'Nov_24_2021_VisiumHuman_12', 'Dec_13_2021_Human8'
)

#################################################################
# Load color schemes
#################################################################


human_cp <- c(
      "L1" = "#8B3D5A", "L2-3" = "#E7BDE1", "L3-4" = "#E6A4CD",
      "L3-4-5" = "#CF8BA3", "L5-6" = "#9E6D7F", "L6b" = "#CDAEB9", "WM1" = "#64BCDB", "WM2" = "#62A7D7", "WM3" = "#99C8D7")


# color scheme:
color_df <- read.csv(file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/data/cell_identity.csv')
color_df <- color_df[,2:3]

group_levels <- c(
  'EX L2', 'EX L2-3', 'EX L3-5', 'EX L5', 'EX L5-6', 'EX L6',
  'INH VIP+', 'INH', 'INH LAMP5+', 'INH PVALB+', 'INH SST+',
  'ASC1', 'ASC2', 'ASC3', 'ASC4',
  'MG1', 'MG2',
  'ODC1', 'ODC2', 'ODC3',
  'OPC1', 'OPC2', 'OPC3',
  'END Arterial', 'END Capillary',
  'T-Pericyte', 'M-Pericyte', 'SMC',
  'Perivascular Fibroblast', 'Meningeal Fibroblast'
)
color_df$group <- factor(as.character(color_df$group), levels=group_levels)
color_df <- arrange(color_df, group)
cp <- as.character(color_df$colour)
names(cp) <- as.character(color_df$group)



st_color_df <- data.frame(
  colour = as.character(human_cp),
  group = names(human_cp)
)
st_color_df$group <- factor(as.character(st_color_df$group), levels=names(human_cp))


seurat_obj$cell_identity <- factor(
  as.character(seurat_obj$cell_identity),
  levels = group_levels
)

#################################################################
# Load celltrek coordinates:
#################################################################

anno_df <- read.csv('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/spatial_mapping/celltrek_coords_combined.csv')


# load the emily amyloid table:
emily_amyloid_results <- read.csv("/dfs7/swaruplab/emiyoshi/Visium_ADDS/amyloid_meta.csv")


```

Write my own functions to do the amyloid analysis!! 

```{r eval=FALSE}

CalculateSpatialDistances <- function(
    seurat_obj,
    scale_factor,
    img_analysis_files,
    sample_col = 'Sample',
    slide_col = 'Slide',
    capture_col = 'CaptureArea',
    spot_diameter = 55,
    subset_samples = TRUE
){

    # list of columns to check
    meta_check <- c(sample_col, slide_col, capture_col, 'imagerow', 'imagecol')

    # check that each of those columns are in the seurat object metadata
    for(col in meta_check){
        if(!(col %in% colnames(seurat_obj@meta.data))){
            stop(paste0('Column ', col, ' not found in seurat_obj@meta.data'))
        }
    }

    # check that the image analysis files exist
    if(!all(file.exists(img_analysis_files))){
        bad_files <- img_analysis_files[!file.exists(img_analysis_files)]
        stop(paste0('Some img_analysis_files do not exist: ', paste0(bad_files, collapse=', ')))
    }

    # get a list of the capture areas and the slides in the seurat obj 
    valid_slide_caps <- unique(paste0(seurat_obj@meta.data[[slide_col]], '_', seurat_obj@meta.data[[capture_col]]))

    # load the quant files:
    quant_df <- data.frame()
    for(img_analysis_file in img_analysis_files){
        
        print(img_analysis_file)

        # get the slide num and cap area from the file 
        tmp <- strsplit(img_analysis_file, '/')[[1]]
        tmp <- tmp[length(tmp)]
        tmp <- strsplit(tmp, '_')[[1]]
        cur_slide <- tmp[1]
        cur_capture <- substr(tmp[2], start=1, stop=2)
        cur_slide_cap <- paste0(cur_slide, '_', cur_capture)

        # check that cur_slide_cap is valid:
        if(!(cur_slide_cap %in% valid_slide_caps)){
            stop(paste0(cur_slide_cap, ' is not a valid combination of Visium Slide number and capture area in this seurat object. Filepath: ', img_analysis_file))
        }

         # load the amyloid file:
        quant <- read.table(img_analysis_file, sep = "\t", skip = 1, skipNul = TRUE)
        quant <- quant[,1:7]

        # give the columns names 
        colnames(quant) <- c('row', 'Source', 'BinaryID', 'Area', 'Diameter', 'X', 'Y')

        # re-name the binaryID for legibility:
        quant$BinaryID <- ifelse(
            grepl('OC', quant$BinaryID), 'OC', 'Amylo'
        )

        # convert to pixels based on the scale facrtor
        quant$PixelX <- quant$X / scale_factor
        quant$PixelY <- quant$Y / scale_factor
        quant$PixelDiameter <- quant$Diameter / scale_factor
        quant$PixelArea <- quant$Area / scale_factor

        # add the slide info 
        quant$Slide <- cur_slide 
        quant$CaptureArea <- cur_capture 
        quant$SlideCap <- cur_slide_cap 
        
        quant_df <- rbind(quant_df, quant)
    }

    # loop through each slide / cap and compute the distances
    img_df <- data.frame()
    for(cur_slide_cap in unique(quant_df$SlideCap)){
        
        print(cur_slide_cap)

        quant <- subset(quant_df, SlideCap == cur_slide_cap)
        cur_slide <- unique(quant$Slide); cur_capture <- unique(quant$CaptureArea)

        # which visium sample corresponds to this slide/cap? 
        if(subset_samples){
            bcs_keep <- seurat_obj@meta.data[[slide_col]] == cur_slide & seurat_obj@meta.data[[capture_col]] == cur_capture
            cur_meta <- seurat_obj@meta.data[bcs_keep,]

            # check that this only corresponds to one sample:
            cur_sample <- unique(cur_meta[[sample_col]])
            if(length(cur_sample) != 1){
                stop('Current Slide number (', cur_slide, ') and capture area (', cur_capture, ') must correspond to exactly one sample in the seurat_obj@meta.data[[sample_col]]. Filepath: ', img_analysis_file)
            }

            # subset the seurat object for this sample 
            cur_seurat <- seurat_obj[,bcs_keep]
        } else{
            cur_seurat <- seurat_obj
        }
        
        # split by amylo & OC 
        quant_list <- quant %>% dplyr::group_by(BinaryID) %>% dplyr::group_split()

        # get the snRNA seurat meta-data
        cur_meta <- cur_seurat@meta.data

        # compute the distances between plaques and spots
        out_df <- do.call(cbind, lapply(quant_list, function(binary){

            cur_bid <- unique(binary$BinaryID)

            # make a matrix for the quants 
            X <- binary %>% 
                dplyr::select(c(PixelX,PixelY)) %>%
                as.matrix() %>% t()

            # make matrix for the visium data:
            spots <- cur_meta[,c('imagecol', 'imagerow')] %>% 
                dplyr::rename(c(X=imagecol, Y=imagerow)) %>% 
                as.matrix() %>% t()

            dists <- apply(X, 2, function(x){
                sqrt(colSums((spots - x)^2))
            })
            colnames(dists) <- binary$row

            # flatten into a dataframe
            dist_df <- reshape2::melt(dists) %>% 
            dplyr::rename(distance=value)

            # transfer the plaque size info to the new df
            ix <- match(dist_df$Var2, binary$row)
            dist_df$PixelArea <- binary$PixelArea[ix]
            dist_df$PixelDiameter <- binary$PixelDiameter[ix]

            # Just count how many are within a certain distance 
            dist_df$close <- dist_df$distance <= ((dist_df$PixelDiameter/2) + spot_diameter)
            bc_totals <- dist_df %>% subset(close) %>% .$Var1 %>% table 

            # make the area zero if the spot isn't close enough 
            dist_df$PixelArea <- ifelse(dist_df$close, dist_df$PixelArea, 0)

            # sum up the areas and add the total number of plaques to the output df
            out_df <- dist_df %>% 
                group_by(Var1) %>% 
                summarise(Area = sum(PixelArea)) %>% 
                as.data.frame()
            out_df$Number <- as.numeric(bc_totals)
            rownames(out_df) <- as.character(out_df$Var1)
            out_df <- out_df[,-1]
            colnames(out_df) <- paste0(cur_bid, '_', colnames(out_df))
            out_df
        }))

        out_df$Sample <- cur_sample 
        out_df$Slide <- cur_slide 
        out_df$CaptureArea <- cur_capture
        img_df <- rbind(img_df, out_df)

    }

    # re-name columns to match inputs:
    colnames(img_df)[5:7] <- c(sample_col, slide_col, capture_col)

    # return the dataframe
    img_df

}


```

Male vs Female sample numbers?

```{r eval=FALSE}

tmp <- seurat_obj@meta.data %>% dplyr::select(c(Sample, Sex, Diagnosis)) %>% distinct
table(tmp$Sex, tmp$Diagnosis)

tmp <- seurat_vis@meta.data %>% dplyr::select(c(Sample, Sex, Diagnosis)) %>% distinct
table(tmp$Sex, tmp$Diagnosis)

```


Run CalculateSpatialDistances on the Human Visium dataset 

```{r eval=FALSE}

img_analysis_dir <- '/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/amyloid/data/VisiumImageAnalysis/human/'
img_analysis_files <- dir(img_analysis_dir)
img_analysis_files <- img_analysis_files[grepl('txt', img_analysis_files)]
img_analysis_files <- paste0(img_analysis_dir, img_analysis_files)

# run on the human visium dataset
img_df <- CalculateSpatialDistances(
    seurat_vis,
    scale_factor = 0.42,
    img_analysis_files = img_analysis_files
)
table(img_df$Sample)


# subset the seurat object 
seurat_vis_amyloid <- seurat_vis[,rownames(img_df)]

# add the amyloid info:L
seurat_vis_amyloid@meta.data <- cbind(seurat_vis_amyloid@meta.data, img_df[,1:4])

saveRDS(seurat_vis_amyloid, file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/amyloid/data/ADDS_seurat_processed_annotated_amyloid.rds')
write.csv(img_df, file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/amyloid/data/ADDS_seurat_processed_annotated_amyloid.csv', quote=FALSE)

img_df <- read.csv(file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/amyloid/data/ADDS_seurat_processed_annotated_amyloid.csv')



##############################################################
# Plot the side-by-side for each sample:
##############################################################

# load the quant files:
quant_df <- data.frame()
for(img_analysis_file in img_analysis_files){
    
    print(img_analysis_file)

    # get the slide num and cap area from the file 
    tmp <- strsplit(img_analysis_file, '/')[[1]]
    tmp <- tmp[length(tmp)]
    tmp <- strsplit(tmp, '_')[[1]]
    cur_slide <- tmp[1]
    cur_capture <- substr(tmp[2], start=1, stop=2)
    cur_slide_cap <- paste0(cur_slide, '_', cur_capture)

    # load the amyloid file:
    quant <- read.table(img_analysis_file, sep = "\t", skip = 1, skipNul = TRUE)
    quant <- quant[,1:7]

    # give the columns names 
    colnames(quant) <- c('row', 'Source', 'BinaryID', 'Area', 'Diameter', 'X', 'Y')

    # re-name the binaryID for legibility:
    quant$BinaryID <- ifelse(
        grepl('OC', quant$BinaryID), 'OC', 'Amylo'
    )

    # convert to pixels based on the scale facrtor
    quant$PixelX <- quant$X / scale_factor
    quant$PixelY <- quant$Y / scale_factor
    quant$PixelDiameter <- quant$Diameter / scale_factor
    quant$PixelArea <- quant$Area / scale_factor

    # add the slide info 
    quant$Slide <- cur_slide 
    quant$CaptureArea <- cur_capture 
    quant$SlideCap <- cur_slide_cap 
    
    quant_df <- rbind(quant_df, quant)
}

samples <- unique(img_df$Sample)

for(cur_sample in samples){

    print(cur_sample)
    out_df <- subset(img_df, Sample == cur_sample)

    # subset visium 
    cur_vis <- subset(seurat_vis, Sample == cur_sample)
    cur_slide <- unique(cur_vis$Slide)
    cur_capture <- unique(cur_vis$CaptureArea)

    # subset the quant 
    quant <- subset(quant_df, Slide == cur_slide & CaptureArea == cur_capture)

    # subset old visium:
    test <- subset(seurat_amyloid_human, Sample == cur_sample)



    all.equal(rownames(out_df), colnames(cur_vis))
    for(x in colnames(out_df)){
        cur_vis@meta.data[,x] <- out_df[,x]
    }

        p1 <- SampleFeaturePlot(
            cur_vis,
            feature='Amylo_Area',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        )

        p2 <- SampleFeaturePlot(
            cur_vis,
            feature='Amylo_Number',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        )


        p3 <- SampleFeaturePlot(
            cur_vis,
            feature='OC_Area',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        )

        p4 <- SampleFeaturePlot(
            cur_vis,
            feature='OC_Number',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        )

        xrange <- range(cur_vis@meta.data$imagecol); 
        xrange[1] <- xrange[1] - 100
        xrange[2] <- xrange[2] + 100
        yrange <- range(cur_vis@meta.data$imagerow)
        yrange[1] <- yrange[1] - 100
        yrange[2] <- yrange[2] + 100

        p5 <- quant %>% subset(BinaryID == 'Amylo') %>% 
            ggplot(aes(x=PixelX, y=PixelY, color=PixelArea)) + 
            geom_point(inherit.aes=FALSE, data=cur_vis@meta.data, aes(x=imagecol,y=imagerow), color='grey') +
            geom_point() +
            scale_color_viridis() + coord_fixed() + 
            scale_x_continuous(limits=xrange) + 
            scale_y_continuous(limits=yrange)+ hdWGCNA::umap_theme()


        p6 <- quant %>% subset(BinaryID == 'OC') %>% 
            ggplot(aes(x=PixelX, y=PixelY, color=PixelArea)) + 
            geom_point(inherit.aes=FALSE, data=cur_vis@meta.data, aes(x=imagecol,y=imagerow), color='grey') +
            geom_point() +
            scale_color_viridis() + coord_fixed()+ 
            scale_x_continuous(limits=xrange) + 
            scale_y_continuous(limits=yrange) + hdWGCNA::umap_theme()



        # patch <- (p5 | p1 | p2 ) / (p6 | p3 | p4 )
        # pdf(paste0(fig_dir, 'test_image_analysis_', cur_vis_sample, '.pdf'), width=12, height=6)
        # print(patch)
        # dev.off()


        p7 <- SampleFeaturePlot(
            test,
            feature='Amyloglo_Score',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        ) + ggtitle('Old Amylo Score') 
        p8 <- SampleFeaturePlot(
            test,
            feature='OC_Score',
            raster=TRUE,
            colfunc = inferno,
            rev_colors=TRUE,
            dpi=300,
            combine=FALSE
        )+ ggtitle('Old OC Score')


        patch <- (p5 | p1 | p2 | p7) / (p6 | p3 | p4 | p8)
        pdf(paste0(fig_dir, 'human_image_analysis_', cur_sample, '.pdf'), width=16, height=6)
        print(patch)
        dev.off()

}





##############################################################
# Plot the full results:
##############################################################

cur_vis <- seurat_vis[,rownames(img_df)]

all.equal(rownames(img_df), colnames(cur_vis))
for(x in colnames(img_df)){
    cur_vis@meta.data[,x] <- img_df[,x]
}

p1 <- SampleFeaturePlot(
    cur_vis,
    feature='Amylo_Area',
    sample_labels = c("Diagnosis", "Sex", 'Age'),
    ncol = 10,
    plot_max = 'q99',
    raster=TRUE,
    colfunc = inferno,
    rev_colors=TRUE,
    dpi=300,
    combine=FALSE
)

p1 <- lapply(p1, function(x){
    x + theme(
    plot.title = element_text(face='bold', size=12, vjust=-1),
    plot.margin = margin(0,0,0,0)
    ) + labs(fill = paste0('Amylo Area'))
})


patch <- wrap_plots(p1, ncol=8, widths=1, heights=1) + plot_layout(guides='collect')
pdf(paste0(fig_dir, 'human_image_analysis_amylo_area.pdf'), width=18, height=9)
print(patch)
dev.off()


```

Run CalculateSpatialDistances on the Mouse Visium dataset 

```{r eval=FALSE}

img_analysis_dir <- '/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/amyloid/data/VisiumImageAnalysis/mouse/'
img_analysis_files <- dir(img_analysis_dir)
img_analysis_files <- img_analysis_files[grepl('txt', img_analysis_files)]
img_analysis_files <- paste0(img_analysis_dir, img_analysis_files)

# load Seurat object
seurat_5x <- readRDS('/dfs7/swaruplab/emiyoshi/Visium_5X/5XFAD_seurat_processed_annotated.rds')

# TODO 
mouse_img_df <- CalculateSpatialDistances(
    seurat_5x,
    scale_factor = 0.42,
    img_analysis_files = img_analysis_files,
    sample_col = 'SAMPLE',
    slide_col = 'SlideSerial'
)

# subset the seurat object 
seurat_5x_amyloid <- seurat_5x[,rownames(mouse_img_df)]

# add the amyloid info:L
seurat_5x_amyloid@meta.data <- cbind(seurat_5x_amyloid@meta.data, mouse_img_df[,1:4])

saveRDS(seurat_5x_amyloid, file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/amyloid/data/5XFAD_seurat_processed_annotated_amyloid.rds')

write.csv(mouse_img_df, file='/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/amyloid/data/5XFAD_seurat_processed_annotated_amyloid.csv', quote=FALSE)

##############################################################
# Plot the full results:
##############################################################

cur_vis <- seurat_vis[,rownames(img_df)]

all.equal(rownames(img_df), colnames(cur_vis))
for(x in colnames(img_df)){
    cur_vis@meta.data[,x] <- img_df[,x]
}

p1 <- SampleFeaturePlot(
    seurat_5x_amyloid,
    feature='Amylo_Area',
    sample_labels = c('Age', "Condition", 'Sex'),
    sample_col = "SAMPLE",
    ncol = 10,
    plot_max = 'q99',
    raster=TRUE,
    colfunc = inferno,
    rev_colors=TRUE,
    dpi=300,
    combine=FALSE
)

p1 <- lapply(p1, function(x){
    x + theme(
    plot.title = element_text(face='bold', size=12, vjust=-1),
    plot.margin = margin(0,0,0,0)
    ) + labs(fill = paste0('Amylo Area'))
})


patch <- wrap_plots(p1, ncol=8, widths=1, heights=1) + plot_layout(guides='collect')
pdf(paste0(fig_dir, 'mouse_image_analysis_amylo_area.pdf'), width=12, height=4)
print(patch)
dev.off()


```


Run CalculateSpatialDistances on the snRNA-seq dataset 

```{r eval=FALSE}

img_analysis_dir <- '/dfs7/swaruplab/smorabit/analysis/ADDS_2021/splitseq/integration/amyloid/data/VisiumImageAnalysis/human/'
img_analysis_files <- dir(img_analysis_dir)
img_analysis_files <- img_analysis_files[grepl('txt', img_analysis_files)]
img_analysis_files <- paste0(img_analysis_dir, img_analysis_files)

# list of samples:
visium_samples <- unique(anno_df$vis_sample)

snRNA_img_df <- data.frame()
for(cur_vis_sample in visium_samples){

    print(cur_vis_sample)

    cur_vis_meta <- subset(seurat_vis@meta.data, Sample == cur_vis_sample)
    cur_slide <- unique(cur_vis_meta$Slide)
    cur_capture <- unique(cur_vis_meta$CaptureArea)
    cur_diagnosis <- as.character(unique(cur_vis_meta$Diagnosis))

    # get the current image analysis file:
    cur_img_file <- img_analysis_files[grepl(paste0(cur_slide, '_', cur_capture), img_analysis_files)]

    if(length(cur_img_file) == 0){
        print('skipping')
        next
    }

    # get the celltrek coords and subset the snRNA seurat object
    cur_coords <- subset(anno_df, vis_sample == cur_vis_sample)
    cur_seurat <- seurat_obj[,cur_coords$sc_bc]

   # cur_seurat <- subset(cur_seurat, Diagnosis == cur_diagnosis)
   # rownames(cur_coords) <- cur_coords$sc_bc
    ix <- match(colnames(cur_seurat), cur_coords$sc_bc)
    cur_coords <- cur_coords[ix,]
    all.equal(cur_coords$sc_bc, colnames(cur_seurat))
   
    # add the image info
    cur_seurat$Slide <- cur_slide 
    cur_seurat$CaptureArea <- cur_capture
    cur_seurat$imagecol <- cur_coords$celltrek_1 
    cur_seurat$imagerow <- cur_coords$celltrek_2

    # add visium sample info:
    cur_seurat$VisiumSample <- cur_vis_sample

    # add spatial reduction
    cur_seurat@reductions$spatial <- CreateDimReducObject(
        embeddings = as.matrix(cur_coords[,c('celltrek_1', 'celltrek_2')])
    )

    # run on the human visium dataset
    cur_img_df <- CalculateSpatialDistances(
        cur_seurat,
        scale_factor = 0.42,
        img_analysis_files = cur_img_file,
        subset_samples = FALSE, 
        sample_col = 'VisiumSample'
    )

    # combine with the coords df:
    all.equal(rownames(cur_img_df), cur_coords$sc_bc)
    cur_img_df <- cbind(cur_img_df, cur_coords)

    snRNA_img_df <- rbind(snRNA_img_df, cur_img_df)

}

write.csv(snRNA_img_df, file=paste0(data_dir, 'snRNA_Human_amyloid.csv'), quote=FALSE)


```




Run monocle3 GLM for DEGs based on the amyloid scores 

```{r eval=FALSE}

# re-load 
snRNA_img_df <- read.csv(file=paste0(data_dir, 'snRNA_Human_amyloid.csv'))

# only plot samples that have amyloid scores:
amyloid_samples <- snRNA_img_df %>%
  group_by(vis_sample) %>%
  summarise(Mean = mean(Amylo_Area)) %>%
  filter(Mean != 0) %>% .$vis_sample %>% as.character

ds_samples <- subset(seurat_vis@meta.data, Diagnosis == 'AD_DS') %>% .$Sample %>% unique
ds_samples <- ds_samples[ds_samples %in% amyloid_samples]



cur_vis_sample <- ds_samples[1]

# get the celltrek coords and subset the snRNA seurat object
cur_coords <- subset(snRNA_img_df, vis_sample == cur_vis_sample)
cur_seurat <- seurat_obj[,cur_coords$sc_bc]
cur_seurat <- subset(cur_seurat, Diagnosis == 'DSAD')

# cur_seurat <- subset(cur_seurat, Diagnosis == cur_diagnosis)
# rownames(cur_coords) <- cur_coords$sc_bc
ix <- match(colnames(cur_seurat), cur_coords$sc_bc)
cur_coords <- cur_coords[ix,]
all.equal(cur_coords$sc_bc, colnames(cur_seurat))

# add the image info
cur_seurat$imagecol <- cur_coords$celltrek_1 
cur_seurat$imagerow <- cur_coords$celltrek_2
cur_seurat$Amylo_Area <- cur_coords$Amylo_Area
cur_seurat$OC_Area <- cur_coords$OC_Area

# add visium sample info:
cur_seurat$VisiumSample <- cur_vis_sample

# add spatial reduction
cur_seurat@reductions$spatial <- CreateDimReducObject(
    embeddings = as.matrix(cur_coords[,c('celltrek_1', 'celltrek_2')])
)



library(monocle3)
library(tictoc)


clusters <- levels(cur_seurat$cell_identity)
clusters <- clusters[clusters != 'unknown']

cur_cluster <- 'EX L3-5'
expr_proportion <- 0.05


# make a new CDS with just the RNA-seq data in order to do t-DEGs:
deg_list <- list()
#for(cur_cluster in clusters){

  print(cur_cluster)

  expr_matrix  <- GetAssayData(
    subset(cur_seurat, cell_identity== cur_cluster),
    slot='data',
  )

  tmp <- expr_matrix > 0
  tmp <- Matrix::rowSums(tmp) / ncol(expr_matrix)
  test_genes <- names(which(tmp > expr_proportion))

  # create celldataset object
  cds <- new_cell_data_set(
    expr_matrix[test_genes,],
    cell_metadata=cur_seurat@meta.data[colnames(expr_matrix),]
  )
  print(dim(cds))

  # identify amylo hotspot DEGs
  gene_fits <- monocle3::fit_models(
    cds,
    model_formula_str = "~Amylo_Area + Sample + nCount_RNA",
    verbose=FALSE,
    cores=1
  )

  fit_coefs <- coefficient_table(gene_fits)

  degs <- fit_coefs %>% filter(term == "Amylo_Area") %>% select(gene_id, term, p_value, q_value, test_val, std_err, estimate, normalized_effect)

  degs$group <- cur_cluster
  deg_list[[cur_cluster]] <- degs

#}

degs <- do.call(rbind, deg_list)


```

Try to run Voyager on one of the mapped samples? 

```{r eval=FALSE}

library(SpatialFeatureExperiment)
library(SpatialExperiment)
library(Voyager)
library(sf)





# re-load 
snRNA_img_df <- read.csv(file=paste0(data_dir, 'snRNA_Human_amyloid.csv'))

# only plot samples that have amyloid scores:
amyloid_samples <- snRNA_img_df %>%
  group_by(vis_sample) %>%
  summarise(Mean = mean(Amylo_Area)) %>%
  filter(Mean != 0) %>% .$vis_sample %>% as.character

ds_samples <- subset(seurat_vis@meta.data, Diagnosis == 'AD_DS') %>% .$Sample %>% unique
ds_samples <- ds_samples[ds_samples %in% amyloid_samples]

cur_vis_sample <- ds_samples[1]

# get the celltrek coords and subset the snRNA seurat object
cur_coords <- subset(snRNA_img_df, vis_sample == cur_vis_sample)
cur_seurat <- seurat_obj[,cur_coords$sc_bc]
cur_seurat <- subset(cur_seurat, Diagnosis == 'AD_DS')

# cur_seurat <- subset(cur_seurat, Diagnosis == cur_diagnosis)
# rownames(cur_coords) <- cur_coords$sc_bc
ix <- match(colnames(cur_seurat), cur_coords$sc_bc)
cur_coords <- cur_coords[ix,]
all.equal(cur_coords$sc_bc, colnames(cur_seurat))

# add the image info
cur_seurat$imagecol <- cur_coords$celltrek_1 
cur_seurat$imagerow <- cur_coords$celltrek_2
cur_seurat$Amylo_Area <- cur_coords$Amylo_Area
cur_seurat$OC_Area <- cur_coords$OC_Area

# add visium sample info:
cur_seurat$VisiumSample <- cur_vis_sample

# add spatial reduction
cur_seurat@reductions$spatial <- CreateDimReducObject(
    embeddings = as.matrix(cur_coords[,c('celltrek_1', 'celltrek_2')])
)



cur_coords <- dplyr::select(cur_coords, c(sc_bc, celltrek_1, celltrek_2)) %>%
    dplyr::rename(barcode=sc_bc, col=celltrek_1, row=celltrek_2)

meta <- cbind(cur_coords,cur_seurat@meta.data)

X <- GetAssayData(cur_seurat, slot='counts')

sfe <- SpatialFeatureExperiment(
    list(counts=X),
    colData = meta,
    spatialCoordsNames = c('col', 'row')
)






data("visium_row_col")
coords1 <- visium_row_col[visium_row_col$col < 6 & visium_row_col$row < 6,]
coords1$row <- coords1$row * sqrt(3)

# Random toy sparse matrix
set.seed(29)
col_inds <- sample(1:13, 13)
row_inds <- sample(1:5, 13, replace = TRUE)
values <- sample(1:5, 13, replace = TRUE)
mat <- sparseMatrix(i = row_inds, j = col_inds, x = values)
colnames(mat) <- coords1$barcode
rownames(mat) <- sample(LETTERS, 5)

sfe3 <- SpatialFeatureExperiment(list(counts = mat), colData = coords1,
                                spatialCoordsNames = c("col", "row"),
                                spotDiameter = 0.7)

```


























Plot all mapped cells in representative samples

```{r eval=FALSE}

ex_neurons <- c(
  'EX L2', 'EX L2-3', 'EX L3-5', 'EX L5', 'EX L5-6', 'EX L6')

inh_neurons <- c(
  'INH VIP+', 'INH', 'INH LAMP5+', 'INH PVALB+', 'INH SST+')

glia <- c(
  'ASC1', 'ASC2', 'ASC3', 'ASC4',
  'MG1', 'MG2',
  'ODC1', 'ODC2', 'ODC3',
  'OPC1', 'OPC2', 'OPC3'
)
vasc <- c(
  'END Arterial', 'END Capillary',
  'T-Pericyte', 'M-Pericyte', 'SMC',
  'Perivascular Fibroblast', 'Meningeal Fibroblast'
)

group_list <- list(
  EX = ex_neurons,
  INH = inh_neurons,
  glia = glia,
  VASC = vasc
)

# get sc coordinates for the current visium sample
plot_list <- list()
for(i in 1:length(human_rep_samples)){

  cur_vis_sample <- human_rep_samples[i]
  print(cur_vis_sample)
  plot_list <- list()
  for(cur_group in names(group_list)){
    cur_coords <- subset(anno_df, vis_sample == cur_vis_sample & cell_identity %in% group_list[[cur_group]])

    # plot the main dataset
    p <- cur_coords %>%
      ggplot(aes(x=celltrek_1, y=celltrek_2, color=cell_identity)) +
      ggrastr::rasterise(geom_point(size=1, alpha=1), dpi=300, scale=0.5) +
      scale_color_manual(values=cp) +
      umap_theme() + coord_fixed() +
      guides(colour = guide_legend(override.aes = list(size=5))) +
      NoLegend()

    plot_list[[cur_group]] <- p
  }

  pdf(paste0(fig_dir, 'rep_samples_mapped/', human_rep_samples[i],'.pdf'), width=12, height=3)
  print(wrap_plots(plot_list, ncol=4))
  dev.off()

}

pdf(paste0(fig_dir, 'rep_samples_mapped_cell_ids.pdf'), width=18, height=3)
print(wrap_plots(plot_list, ncol=8))
dev.off()


```

