
In this notebook, we will compute gene signature scores for known relevant  gene
signatures (DAMs, DAAs, etc), and we will also use the Voyager R package to try to
perform geospatial statistical analysis based on these features.

Load the Seurat data

```{r eval=FALSE}

# conda activate voyager
library(Seurat)
library(tidyverse)
library(cowplot)
library(patchwork)
library(RColorBrewer)
library(Voyager)
library(UCell)
library(ggrastr)
library(ggrepel)
library(viridis)
library(Voyager)
library(SpatialFeatureExperiment)
library(scater)
library(scran)
library(SFEData)
library(sf)
library(ggplot2)
library(scales)
library(patchwork)
library(BiocParallel)
library(bluster)
library(tidyverse)
library(ggpubr)
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

#detach("package:hdWGCNA", unload=TRUE)
#devtools::install_github('smorabit/hdWGCNA', ref='dev')
library(hdWGCNA)


colfunc <- colorRampPalette(rev(brewer.pal(11, 'Spectral' )))
theme_set(theme_cowplot())

setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/amyloid/")

fig_dir <- "figures/"
data_dir <- "data/"


# re-load
sfe <- readRDS(paste0(data_dir, '5xFAD_SFE.rds'))
seurat_5x <- readRDS(paste0(data_dir, '5xFAD_seurat_processed_amyloid_voyager.rds'))


# 5x data:
seurat_5x <- readRDS('/dfs7/swaruplab/emiyoshi/Visium_5X/5XFAD_seurat_processed_annotated.rds')

seurat_5x$Age <- factor(
  as.character(seurat_5x$Age),
  levels = c("4mo", "6mo", "8mo", "12mo")
)

mouse_cp <- c(
      "ctx-deep-layers" = "#5b4468", "ctx-upper-layers" = "#9581a4", "ctx-olfactory" = "#8073ab",
      "hippocampus" = "#d6a5b2", "hippocampus-pyramidal" = "#e193aa",
      "lateral-ventricle" = "#d2c3b2",
      "striatum" = "#F0DE8C",
      "thalamus1" = "#00CBA7", "thalamus2" = "#abcc94",
      "hypothalamus-amygdala" = "#72c9b1",
      "WM1" = "#325ea8", "WM2" = "#64bcdb", "WM-cerebral-peduncle" = "#62a7d7",
      "erythrocytes-neurons" = "#d07fc4", "unknown" = "#c47c4d"
    )


# set factor levels for mouse clusters
seurat_5x$annotation <- factor(
  as.character(seurat_5x$annotation),
  levels = names(mouse_cp)
)

cp_5x <- c("4mo" = "#E8C0FC", "6mo" = "#B57DD1",
        "8mo"= "#8C64A1", "12mo" = "#643D78")



# seurat objets with amyloid data:
seurat_amyloid_5x <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/amyloid/data/5XFAD_seurat_processed_annotated_amyloid.rds')



```

Compute module scores in the mouse dataset

```{r eval=FALSE}


# Chen 2021 (compilation of multiple studies)
homeostatic <- c("Tmem119", "P2ry12", "P2ry13", "Csf1r",
				 "Hexb", "Cst3", "Cx3cr1", "Siglech",
				 "Cd33", "Tgfbr1", "Sall1", "Selplg",
				 "Mef2a", "Jun", "Ms4a6d", "Bin1", "Serinc3")

DAM <- c("Cd9", "Apoe", "Trem2", "Tyrobp",
		"Cd63", "Lgals3", "Axl", #"Clec7a",
		"Spp1", "Ctsb", "Lpl", # "Cstz",
		"Ctsl", "Ctsd", "Itgax", "B2m",
		"Cst7", "Csf1", "Gpnmb", "Igf1",
		"Irf8", "Fth1", "Lyz1", # "Lilrb4",
		"Ccl3", "Ccl6", "Timp2")

IFN <- c("Irf7", "Ifitm3", "Ifit2", "Ifit3",
		 "Cxcl10", "Oasl2", "Cd69", "Isg15",
		 "Usp18")

MHC <- c("H2-D1", "H2-K1", "H2-Eb1", "H2-Aa",
		 "H2-Ab1", "H2-DMa", "Cd74")

Cyc_M <- c("Top2a", "Mki67", "Cenpe", "Mcm5",
		   "Birc5", "H2afz", "H2afv")

# Habib et al 2020
DAA_markers <- list()
DAA_markers$GFAP_low <- c("Luzp2", "Slc7a10", "Mfge8")
DAA_markers$GFAP_high <- c("Gfap", "Id3", "Aqp4", "Myoc", "Id1", "Fabp7")
DAA_markers$DAA <- c("Gfap", "Ctsb", "Vim", "Osmr", "Serpina3n", "Gsn", "Ggta1")

# Kenigsbuch et al 2022
DOL <- read.csv("/dfs7/swaruplab/emiyoshi/Visium_5X/genescores/oligodendrocytes/DOL_genes.csv")

# from our old paper check the citation again
DOL_markers <- list("DOL" = DOL[,1], 'NFOL'=c('Tcf7l2', 'Casr', 'Cemip2', 'Itpr2'),
								  'MFOL'=c('Mal', 'Mog', 'Plp1', 'Opalin', 'Serinc5', 'Ctps1'),
								  'MOL'=c('Klk6', 'Apod', 'Slc5a11', 'Pde1a'))

# Load PIGs from Chen et al 2020
pig_table <- read.csv('~/swaruplab/smorabit/analysis/ADDS_2021/data/chen_modules.csv')
pigs <- subset(pig_table, module == 'PIG') %>% .$gene
oligs <- subset(pig_table, module == 'OLIG') %>% .$gene



gene_lists <- list("homeostatic" = homeostatic, "DAM" = DAM, "IFN" = IFN, "MHC" = MHC, "Cyc_M" = Cyc_M, 'PIGs' = pigs, 'OLIG' = oligs)
gene_lists <- c(gene_lists, DAA_markers, DOL_markers)

seurat_5x <- AddModuleScore_UCell(seurat_5x , features = gene_lists)


```

Create an SFE object holding all mouse samples

```{r eval=FALSE}

# spaceranger output dir
spaceranger_dir <- '/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/'
batches <- c('5xFAD_pilot_2021', 'July_2021', 'Nov_15_2021', 'Nov_24_2021', 'Dec_13_2021', 'Dec_20_2021')

# spaceranger output folder for each sample in each batch
files <- unlist(lapply(batches, function(x){
  cur_dir <- paste0(spaceranger_dir, x, '/spaceranger_count/')
  paste0(cur_dir, dir(cur_dir), '/outs/')
}))
files

# concatenate batch name and sample folder name
samples <- unlist(lapply(batches, function(x){
  cur_dir <- paste0(spaceranger_dir, x, '/spaceranger_count/')
  paste0(x,'#', dir(cur_dir))
}))
samples

# create SFE from visium data
sfe <- read10xVisiumSFE(files, sample_id=samples, type='sparse', data='raw', load=FALSE)

# only keep spots that are in the tissue
sfe <- sfe[, colData(sfe)$in_tissue]

##############################################################
# add the sample-level meta-data:
##############################################################

table(seurat_5x$SAMPLE)
table(sfe$sample_id)
tmp <- do.call(rbind, strsplit(sfe$sample_id, '#'))
sfe$seqbatch <- tmp[,1]
sfe$SAMPLE <- tmp[,2]
table(sfe$seqbatch)
table(seurat_5x$seqbatch)

colnames(seurat_5x@meta.data)

pilot_meta <- read.csv("/dfs7/swaruplab/smorabit/analysis/5XFAD_visium_2021/pilot/data/pilot_samples.csv", stringsAsFactors=FALSE)
pilot_meta$LibraryID <- pilot_meta$SAMPLE
pilot_meta <- pilot_meta %>% select(c(SAMPLE, LibraryID))
pilot_meta$seqbatch <- '5xFAD_pilot_2021'

july_meta <- read.csv("/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/July_2021/5XFAD_8samples.csv", stringsAsFactors=FALSE)
july_meta <- na.omit(july_meta)
july_meta$LibraryID <- paste0('Visium', july_meta$X)
july_meta <- july_meta %>% select(c(SAMPLE, LibraryID))
july_meta$seqbatch <- 'July_2021'

nov15_meta <- read.csv("/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/Nov_15_2021/visium_5xFAD_11-15-21_sample_meta.csv", stringsAsFactors=FALSE)
nov15_meta$LibraryID <- paste0("Visium5xFAD_", nov15_meta$SAMPLE_num)
nov15_meta <- nov15_meta %>% select(c(SAMPLE, LibraryID))
nov15_meta$seqbatch <- 'Nov_15_2021'

nov24_meta <- read.csv("/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/Nov_24_2021/visium_5xFAD_11-24-21_sample_meta.csv", stringsAsFactors=FALSE)
nov24_meta <- nov24_meta %>% dplyr::rename(LibraryID = LibraryName) %>% select(c(SAMPLE, LibraryID))
nov24_meta$seqbatch <- 'Nov_24_2021'

dec13_meta <- read.csv("/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/Dec_13_2021/visium_5xFAD_12-13-21_sample_meta.csv", stringsAsFactors=FALSE)
dec13_meta$LibraryID <- paste0('5XFAD', dec13_meta$LibraryName)
dec13_meta <- dec13_meta %>% select(c(SAMPLE, LibraryID))
dec13_meta$seqbatch <- 'Dec_13_2021'

dec20_meta <- read.csv("/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/Dec_20_2021/visium_5xFAD_12-20-21_sample_meta.csv", stringsAsFactors=FALSE)
dec20_meta <- dec20_meta %>% dplyr::rename(LibraryID = LibraryName) %>% select(c(SAMPLE, LibraryID))
dec20_meta$seqbatch <- 'Dec_20_2021'

sample_meta <- rbind(
  pilot_meta,
  july_meta,
  nov15_meta,
  nov24_meta,
  dec13_meta,
  dec20_meta
)
sample_meta$sample_id <- paste0(sample_meta$seqbatch, '#', sample_meta$LibraryID)
all(sample_meta$sample_id %in% sfe$sample_id)

# transfer the sample ID to the SFE object
ix <- match(as.character(sfe$sample_id), as.character(sample_meta$sample_id))
sfe$SAMPLE <- sample_meta$SAMPLE[ix]

##############################################################
# transfer the Seurat meta-data to the SFE object
##############################################################

# match barcodes between the seurat object and the SFE object
bc <- do.call(rbind, strsplit(colnames(seurat_5x), '_'))[,1]
bc <- do.call(rbind, strsplit(bc, '-'))[,1]
seurat_5x$bc <- paste0(bc, '-', as.character(seurat_5x$SAMPLE))
bc <- do.call(rbind, strsplit(colnames(sfe), '-'))[,1]
sfe$bc <- paste0(bc, '-', as.character(sfe$SAMPLE))

# make sure there's only spots that are in the Seurat object!!!
sfe <- sfe[,sfe$bc %in% seurat_5x$bc]

# match seurat metadata
meta <- seurat_5x@meta.data
ix <- match(as.character(seurat_5x$bc), as.character(sfe$bc))
meta <- meta[ix,] # %>% dplyr::select(-bc)
all.equal(as.character(sfe$bc), as.character(meta$bc))

# keep Seurat meta-data columns that aren't already in the SFE object
meta <- meta[,!(colnames(meta) %in% colnames(colData(sfe)))]

# add Seurat metadata to SFE
colData(sfe) <- cbind(colData(sfe), meta)

##############################################################
# transfer the amyloid data to the sfe object
##############################################################

# match barcodes between the seurat object and the SFE object
bc <- do.call(rbind, strsplit(colnames(seurat_amyloid_5x), '_'))[,1]
bc <- do.call(rbind, strsplit(bc, '-'))[,1]
seurat_amyloid_5x$bc <- paste0(bc, '-', as.character(seurat_amyloid_5x$SAMPLE))

# match seurat metadata
ix <- match(sfe$bc, seurat_amyloid_5x$bc)

# add amyloig + OC scores to the object:
sfe$Amylo_Area <- seurat_amyloid_5x$Amylo_Area[ix]
sfe$OC_Area <- seurat_amyloid_5x$OC_Area[ix]

# compute log scale of OC and Amyloglo scores
colData(sfe)$Amylo_Area_log <- log(sfe$Amylo_Area + 1)
colData(sfe)$OC_Area_log <- log(sfe$OC_Area + 1)


##############################################################
# change gene names from ensembl ID to gene symbols
##############################################################

gene_file <- '/dfs7/swaruplab/smorabit/data/Visium_Mouse_2021/Dec_20_2021/spaceranger_count/5XFAD1/outs/filtered_feature_bc_matrix/features.tsv.gz'
gene_table <- read.table(gene_file, sep='\t')

# set gene names
rownames(sfe) <- gene_table$V2

##############################################################
# normalize data and compute spatial graphs
##############################################################

clusters <- sfe$annotation
sfe <- computeSumFactors(sfe, clusters=clusters)
sfe <- logNormCounts(sfe)

# compute spatial neighborhood graph
gs <- findVisiumGraph(sfe, sample_id = 'all')
for(sample in names(gs)){
  colGraph(sfe, sample_id = sample, 'visium') <- gs[[sample]]
}

# compute a binary-weight graph
gs <- findVisiumGraph(sfe, sample_id= 'all', style = "B")
for(sample in names(gs)){
  colGraph(sfe, sample_id = sample, 'visium_B') <- gs[[sample]]
}

saveRDS(sfe, paste0(data_dir, '5xFAD_SFE.rds'))
saveRDS(seurat_5x, paste0(data_dir, '5xFAD_seurat_processed_amyloid.rds'))

################################################################################
# geospatial stats for amyloid scores
################################################################################

feats <- c('Amylo_Area_log', 'OC_Area_log', 'Amylo_Area', 'OC_Area')

# get a list of samples that we have amyloid scores for:
amyloid_samples <- colData(sfe) %>%
  as.data.frame() %>%
  group_by(sample_id) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$sample_id %>% as.character


# compute Moran's I for amyloglo and OC scores
for(cur_sample in amyloid_samples){
  print(cur_sample)
  sfe <- colDataUnivariate(
    sfe, type = "localmoran", features = feats,
    colGraphName = "visium", sample_id = cur_sample
  )
}

# compute Getis-Ord Gi* or amyloglo and OC scores
for(cur_sample in amyloid_samples){
  print(cur_sample)
  sfe <- colDataUnivariate(
    sfe, type = "localG_perm", features = feats,
    colGraphName = "visium_B", include_self=TRUE,
    sample_id = cur_sample
  )
}



################################################################################
# geospatial stats for gene signatures
################################################################################

feats <- c(paste0(names(gene_lists), '_UCell'))
#feats <- c('PIGs_UCell', 'OLIG_UCell')

# compute Moran's I
for(cur_sample in unique(sfe$sample_id)){
  print(cur_sample)
  sfe <- colDataUnivariate(
    sfe, type = "localmoran", features = feats,
    colGraphName = "visium", sample_id = cur_sample
  )
}

# compute Getis-Ord Gi*
for(cur_sample in unique(sfe$sample_id)){
  print(cur_sample)
  sfe <- colDataUnivariate(
    sfe, type = "localG_perm", features = feats,
    colGraphName = "visium_B", include_self=TRUE,
    sample_id = cur_sample
  )
}

###############################################################
# save the SFE object
###############################################################

saveRDS(sfe, paste0(data_dir, '5xFAD_SFE.rds'))

###############################################################
# Plot the results in Seurat
###############################################################

# move the results to the seurat object:
ix <- match(as.character(sfe$bc), as.character(seurat_5x$bc))
all.equal(as.character(sfe$bc[ix]), as.character(seurat_5x$bc))

feats <- c(paste0(names(gene_lists), '_UCell'), 'Amylo_Area_log', 'OC_Area_log', 'Amylo_Area', 'OC_Area')



# move features to seurat obj
for(feat in feats){

  # Getis-Ord Gi*
  feat_new <- paste0(feat, '_Gi')
  test <- localResult(sfe, 'localG_perm', feat, sample_id='all')[,'localG']
  seurat_5x@meta.data[[feat_new]] <- test[ix]
  seurat_5x@meta.data[[feat_new]]  <- ifelse(is.na(seurat_5x@meta.data[[feat_new]]), 0, seurat_5x@meta.data[[feat_new]])
  seurat_5x@meta.data[[feat_new]]  <- ifelse(seurat_5x@meta.data[[feat_new]]  == Inf, 0, seurat_5x@meta.data[[feat_new]])

  # moran's I
  feat_new <- paste0(feat, '_lmi')
  test <- localResult(sfe, 'localmoran', feat, sample_id='all')[,'Ii']
  seurat_5x@meta.data[[feat_new]] <- test[ix]
  seurat_5x@meta.data[[feat_new]]  <- ifelse(is.na(seurat_5x@meta.data[[feat_new]]), 0, seurat_5x@meta.data[[feat_new]])
  seurat_5x@meta.data[[feat_new]]  <- ifelse(seurat_5x@meta.data[[feat_new]]  == Inf, 0, seurat_5x@meta.data[[feat_new]])
}

seurat_5x$Amylo_Area_log <- sfe$Amylo_Area_log[ix]
seurat_5x@meta.data[['Amylo_Area_log']]  <- ifelse(is.na(seurat_5x@meta.data[['Amylo_Area_log']]), 0, seurat_5x@meta.data[['Amylo_Area_log']])
seurat_5x@meta.data[['Amylo_Area_log']]  <- ifelse(seurat_5x@meta.data[['Amylo_Area_log']]  == Inf, 0, seurat_5x@meta.data[['Amylo_Area_log']])

seurat_5x$OC_Area_log <- sfe$OC_Area_log[ix]
seurat_5x@meta.data[['OC_Area_log']]  <- ifelse(is.na(seurat_5x@meta.data[['OC_Area_log']]), 0, seurat_5x@meta.data[['OC_Area_log']])
seurat_5x@meta.data[['OC_Area_log']]  <- ifelse(seurat_5x@meta.data[['OC_Area_log']]  == Inf, 0, seurat_5x@meta.data[['OC_Area_log']])

seurat_5x$Amylo_Area <- sfe$Amylo_Area[ix]
seurat_5x@meta.data[['Amylo_Area']]  <- ifelse(is.na(seurat_5x@meta.data[['Amylo_Area']]), 0, seurat_5x@meta.data[['Amylo_Area']])
seurat_5x@meta.data[['Amylo_Area']]  <- ifelse(seurat_5x@meta.data[['Amylo_Area']]  == Inf, 0, seurat_5x@meta.data[['Amylo_Area']])

seurat_5x$OC_Area <- sfe$OC_Area[ix]
seurat_5x@meta.data[['OC_Area']]  <- ifelse(is.na(seurat_5x@meta.data[['OC_Area']]), 0, seurat_5x@meta.data[['OC_Area']])
seurat_5x@meta.data[['OC_Area']]  <- ifelse(seurat_5x@meta.data[['OC_Area']]  == Inf, 0, seurat_5x@meta.data[['OC_Area']])


saveRDS(seurat_5x, paste0(data_dir, '5xFAD_seurat_processed_amyloid_voyager.rds'))


################################################################################
# plot amyloid geospatial stats
################################################################################

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character


feats <- c('Amylo_Area_log', 'OC_Area_log', 'Amylo_Area', 'OC_Area')

for(feat in feats){
  # feat_name <- strsplit(feat, '_')[[1]][1]
  feat_name <- feat
  for(modifier in c('', '_Gi', '_lmi')){

    print(feat_name)
    print(modifier)

    if(modifier == '_lmi'){
      plot_min <- 0.5
    } else{
      plot_min = 0
    }

    p <- SampleFeaturePlot(
      seurat_5x,
      feature=paste0(feat, modifier),
      sample_labels = c("Condition", 'Age', 'Sex'),
      samples_to_plot = amyloid_samples,
      sample_col = "SAMPLE",
      ncol = 10,
      raster=TRUE,
      plot_max = 'q99',
      plot_min = plot_min,
      colfunc = inferno,
      rev_colors=TRUE,
      dpi=800,
      combine=FALSE
    )

    p <- lapply(p, function(x){
      x + theme(
        plot.title = element_text(face='bold', size=12, vjust=-1),
        plot.margin = margin(0,0,0,0)
      ) + labs(fill = paste0(feat_name, ' ', modifier))

    })

    patch <- wrap_plots(p, ncol=6, widths=1, heights=1) + plot_layout(guides='collect')

    pdf(paste0(fig_dir, 'featureplots/featureplot_', feat_name, modifier, '.pdf'), width=12, height=6)
    print(patch)
    dev.off()

  }
}


################################################################################
# plot the genescore geospatial stats
################################################################################


feats <- c(paste0(names(gene_lists), '_UCell'))
feats <- c('PIGs_UCell', 'OLIG_UCell')

for(feat in feats){
  feat_name <- strsplit(feat, '_')[[1]][1]
  for(modifier in c('', '_Gi', '_lmi')){

    print(feat_name)
    print(modifier)

    if(modifier == '_lmi'){
      plot_min <- 0.5
    } else{
      plot_min = 0
    }

    if(grepl('UCell', feat)){
      plot_min <- 0.05
    }

    p <- SampleFeaturePlot(
      seurat_5x,
      feature=paste0(feat, modifier),
      sample_labels = c('Age', "Condition", 'Sex'),
      sample_col = "SAMPLE",
      ncol = 10,
      raster=TRUE,
      plot_max = 'q99',
      plot_min = plot_min,
      colfunc = inferno,
      rev_colors=TRUE,
      dpi=600,
      combine=FALSE
    )

    p <- lapply(p, function(x){
      x + theme(
        plot.title = element_text(face='bold', size=15, vjust=-1),
        plot.margin = margin(0,0,0,0)
      ) + labs(fill = paste0(feat_name, ' ', modifier))

    })

    patch <- wrap_plots(p, ncol=8, widths=1, heights=1) + plot_layout(guides='collect')

    pdf(paste0(fig_dir, 'featureplots/featureplot_', feat_name, modifier, '.pdf'), width=14, height=16)
    print(patch)
    dev.off()

  }
}



```

Plot the distribution of amylo & OC scores in the mouse data 

```{r eval=FALSE}

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character



plot_df <- seurat_5x@meta.data %>% 
  subset(SAMPLE %in% amyloid_samples & annotation != 'unknown') %>%
  group_by(SAMPLE, annotation, Age) %>%
  summarise(
    oc_sum = sum(OC_Area) / n(), 
    amylo_sum = sum(Amylo_Area) / n(),
  ) %>% 
  ungroup 

my_comparisons <- list(c('Control', 'earlyAD'), c('Control', 'AD'), c('Control', 'AD_DS'))

p1 <- ggplot(plot_df, aes(x=Age, y = log(1+oc_sum), fill = Age)) + 
  geom_boxplot(outlier.shape=NA, width=0.75) + 
  geom_jitter(size=0.5, width=0.25) +
  scale_fill_manual(values=cp_5x) +
  RotatedAxis() +
  # stat_compare_means(
  #   comparisons = my_comparisons, 
  #   label='p.signif', 
  #   method='wilcox'
  # ) + 
   theme(
    axis.text.x= element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    plot.margin = margin(c(0,0,0,0))
  ) +
  ylab('Avg. OC Area') + 
  facet_wrap(~annotation, ncol=7) 
 
p2 <- ggplot(plot_df, aes(x=Age, y = log(1+amylo_sum), fill = Age)) + 
  geom_boxplot(outlier.shape=NA, width=0.75) + 
  geom_jitter(size=0.5, width=0.25) +
  scale_fill_manual(values=cp_5x) +
  RotatedAxis() +
  # stat_compare_means(
  #   comparisons = my_comparisons, 
  #   label='p.signif', 
  #   method='wilcox'
  # ) +
  theme(
    axis.text.x= element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    plot.margin = margin(c(0,0,0,0))
  ) +
  ylab('Avg. Amyloglo Area') + 
  facet_wrap(~annotation, ncol=7) 

pdf(paste0(fig_dir, 'amyloid_box_5x.pdf'), width=8, height=6)
p1 / p2 + plot_layout(guides='collect')
dev.off()



```


Amyloid Differential Gene expression analysis with Monocle3 GLM

Running this analysis with Grey + White matter together has issues, since there's
more plaques in the grey matter, it looked like the amyloid DEGs and associated
GO terms were biased towards neuronal genes up near amyloid, and glial genes
up away from the amyloid. However there were still some interesting GO results
in this more "global" analysis, with APP metabolism down and response to amyloid up.
Now going to modify to try this out in just the white matter to see what happens.


```{r eval=FALSE}

library(monocle3)
library(tictoc)

# get a list of samples that we have amyloid scores for:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character


# plot violin plots of Amyloglo scores
p <- VlnPlot(
  subset(seurat_5x, SAMPLE %in% amyloid_samples),
  features = 'Amylo_Area_log',
  group.by = 'annotation',
  split.by = 'Age',
  pt.size=0
) + RotatedAxis() +
  xlab('') + ylab('')

pdf(paste0(fig_dir, 'test_vln_Amylo_Area.pdf'), width=9, height=3.5)
print(p)
dev.off()


################################################################################
# Run DEGs for the OC scores
################################################################################

clusters <- levels(seurat_5x$annotation)
clusters <- clusters[clusters != 'unknown']

# make a new CDS with just the RNA-seq data in order to do t-DEGs:
deg_list <- list()
for(cur_cluster in clusters){

  print(cur_cluster)

  expr_matrix  <- GetAssayData(
    subset(seurat_5x, Condition != 'WT' & SAMPLE %in% amyloid_samples & annotation == cur_cluster),
    slot='data',
  )
  test_genes <- rownames(expr_matrix )[rowSums(expr_matrix )!=0]

  # create celldataset object
  cds <- new_cell_data_set(
    expr_matrix[test_genes,],
    cell_metadata=seurat_5x@meta.data[colnames(expr_matrix),]
  )
  print(dim(cds))

  # identify amylo hotspot DEGs
  gene_fits <- monocle3::fit_models(
    cds,
    model_formula_str = "~OC_Area_log_Gi + SAMPLE + nCount_Spatial",
    verbose=FALSE,
    cores=1
  )

  fit_coefs <- coefficient_table(gene_fits)

  degs <- fit_coefs %>% filter(term == "OC_Area_log_Gi") %>% select(gene_id, term, p_value, q_value, test_val, std_err, estimate, normalized_effect)

  degs$group <- cur_cluster
  deg_list[[cur_cluster]] <- degs

}

degs <- do.call(rbind, deg_list)

# write amyloid-degs table as a csv:
write.csv(degs, file=paste0(data_dir, '5x_OC_Area_log_Gi_DEGs.csv'), quote=FALSE, row.names=FALSE)

subset(degs, q_value < 0.05) %>% dim


################################################################################
# Run DEGs for the Amylo scores
################################################################################

clusters <- levels(seurat_5x$annotation)
clusters <- clusters[clusters != 'unknown']

# make a new CDS with just the RNA-seq data in order to do t-DEGs:
deg_list <- list()
for(cur_cluster in clusters){

  print(cur_cluster)

  expr_matrix  <- GetAssayData(
    subset(seurat_5x, Condition != 'WT' & SAMPLE %in% amyloid_samples & annotation == cur_cluster),
    slot='data',
  )
  test_genes <- rownames(expr_matrix )[rowSums(expr_matrix )!=0]

  # create celldataset object
  cds <- new_cell_data_set(
    expr_matrix[test_genes,],
    cell_metadata=seurat_5x@meta.data[colnames(expr_matrix),]
  )
  print(dim(cds))

  # identify amylo hotspot DEGs
  gene_fits <- monocle3::fit_models(
    cds,
    model_formula_str = "~Amylo_Area_log_Gi + SAMPLE + nCount_Spatial",
    verbose=FALSE,
    cores=1
  )

  fit_coefs <- coefficient_table(gene_fits)

  degs <- fit_coefs %>% filter(term == "Amylo_Area_log_Gi") %>% select(gene_id, term, p_value, q_value, test_val, std_err, estimate, normalized_effect)

  degs$group <- cur_cluster
  deg_list[[cur_cluster]] <- degs

}

degs <- do.call(rbind, deg_list)

# write amyloid-degs table as a csv:
write.csv(degs, file=paste0(data_dir, '5x_Amylo_Area_log_Gi_DEGs.csv'), quote=FALSE, row.names=FALSE)

# write amyloid-degs table as a csv:
write.csv(degs, file=paste0(data_dir, 'Amylo_Area_log_Gi_ctx-deep-layers_DEGs.csv'), quote=FALSE, row.names=FALSE)

subset(degs, q_value < 0.05) %>% dim

################################################################################
# Plot it as a Volcano plot?
################################################################################

library(ggrepel)

degs <- read.csv(file=paste0(data_dir, 'Amylo_Area_log_Gi_ctx-deep-layers_DEGs.csv'))

cp <- c("Control" = "#B8DBC5", "earlyAD" = "#E7BDE1" , "AD" = "#CF8BA3", "AD_DS" = "#9E6D7F")

# late-AD
color1 <- cp['AD']; color2 <- cp['Control']

# early-AD
color1 <- cp['earlyAD']; color2 <- cp['Control']

# AD/DS
color1 <- cp['AD_DS']; color2 <- cp['Control']

# exclude MT genes
degs <- degs[!grepl('MT-', degs$gene_id),]

# lowest non-zero value
lowest <- degs %>% subset(q_value != 0) %>% top_n(-1, wt=q_value) %>% .$q_value
degs$q_value <- ifelse(degs$q_value == 0, lowest, degs$q_value)

nlabel <- 10


cur_degs <- Reduce(rbind, lapply(unique(degs$term), function(x){
  cur <- subset(degs, term == x)

  top_thresh <- cur %>% subset(q_value <= 0.05 & normalized_effect > 0) %>% top_n(nlabel, wt=normalized_effect) %>% .$normalized_effect %>% min
  bottom_thresh <- cur %>% subset(q_value <= 0.05 & normalized_effect < 0) %>% top_n(-1*nlabel, wt=normalized_effect) %>% .$normalized_effect %>% max

  cur$anno <- ifelse(cur$q_value <= 0.05 & cur$normalized_effect >= top_thresh, cur$gene, NA)
  cur$anno <- ifelse(cur$q_value <= 0.05 & cur$normalized_effect <= bottom_thresh, cur$gene, cur$anno)
  cur$color <- ifelse(cur$q_value > 0.05, 'gray', ifelse(cur$normalized_effect > 0, color1, color2))
  cur

}))


groups <- unique(degs$term)
plot_list <- list()
for(cluster  in groups){


  print(cluster)
  plot_degs <- cur_degs %>% subset(term == cluster)

  p <- plot_degs  %>%
     ggplot(aes(x=normalized_effect, y=-log10(q_value))) +
     geom_hline(yintercept=-log10(0.05), linetype='dashed')

  # plot genes that are Nr4a2 targets
  p <- p + ggrastr::rasterise(geom_point(
    alpha=0.5,
    color=plot_degs %>% .$color
  ), dpi=500)

  p <- p +
     geom_point(
       inherit.aes=FALSE,
       data=subset(plot_degs, !is.na(anno)),
       aes(normalized_effect, -log10(q_value)),
       fill=subset(plot_degs, !is.na(anno)) %>% .$color,
       shape=21, size=3, color='black'
     ) +
     geom_text_repel(aes(label=anno), color='black', fontface='italic',  min.segment.length=0, max.overlaps=Inf) +
     xlim(-1*max(abs(plot_degs$normalized_effect))-0.1, max(abs(plot_degs$normalized_effect))+0.1) +
     ggtitle(paste0(cluster)) +
     xlab(bquote("Normalized effect size")) +
     ylab(bquote("-log"[10]~"(Adj. P-value)")) +
     theme(
       panel.border = element_rect(color='black', fill=NA, linewidth=1),
       panel.grid.major = element_blank(),
       axis.line = element_blank(),
       plot.title = element_text(hjust = 0.5, vjust=-1),
       legend.position='bottom'
     )

    plot_list[[cluster]] <- p

}

name <- 'Amylo_Area_log_Gi_GM'
out <- paste0(fig_dir, 'volcano_', name, '_visium.pdf')

pdf(out, width=5, height=5, useDingbats=FALSE)
plot_list[[1]]
dev.off()

# plot_list <- lapply(plot_list, function(x){
#   x + theme(
#     axis.title.x = element_blank(),
#     axis.title.y = element_blank(),
#     plot.margin = margin(0,0,0,0),
#     plot.title = element_text(vjust=-0.2)
#   )
# })
#
# pdf(out, width=12, height=12, useDingbats=FALSE)
# wrap_plots(plot_list, ncol=3)
# dev.off()

```

Correlation of gene expression and amyloid scores 

```{r eval=FALSE}


# load the amylo and the OC DEGs
degs_amylo <- read.csv(file=paste0(data_dir, '5x_Amylo_Area_log_Gi_DEGs.csv'))
degs_oc <- read.csv(file=paste0(data_dir, '5x_OC_Area_log_Gi_DEGs.csv'))


# get a list of samples that we have amyloid scores for:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character

groups <- unique(degs_amylo$group)

# get the expression matrix:
expr_matrix  <- GetAssayData(seurat_5x, slot='data')
cor_df <- data.frame()
amylo_signif <- data.frame()
oc_signif <- data.frame()

for(cur_group in groups){

  print(cur_group)

  # degs for this cluster
  cur_amylo_genes <- degs_amylo %>% subset(group == cur_group & q_value < 0.05) %>% .$gene
  cur_oc_genes <- degs_oc %>% subset(group == cur_group & q_value < 0.05) %>% .$gene

  # get relevant seurat metadata for this cluster
  cur_meta <- subset(seurat_5x@meta.data, Condition != 'WT' & SAMPLE %in% amyloid_samples & annotation == cur_group)

  # subset expression matrix by the amylo and oc gene
  cur_amylo_expr <- expr_matrix[cur_amylo_genes,rownames(cur_meta)]
  cur_oc_expr <- expr_matrix[cur_oc_genes,rownames(cur_meta)]

  # amylo correlation
  temp <- Hmisc::rcorr(t(as.matrix(cur_amylo_expr)), cur_meta$Amylo_Area_log_Gi)
  amylo_cor <- temp$r[cur_amylo_genes,ncol(temp$r)]

  # oc correlation
  temp <- Hmisc::rcorr(t(as.matrix(cur_oc_expr)), cur_meta$OC_Area_log_Gi)
  oc_cor <- temp$r[cur_oc_genes,ncol(temp$r)]

  cur_cor_df <- data.frame(
    gene = c(cur_amylo_genes, cur_oc_genes),
    cor = c(as.numeric(amylo_cor), as.numeric(oc_cor)),
    type = c(rep('amylo', length(cur_amylo_genes)), rep('oc', length(cur_oc_genes))),
    group = cur_group
  )
  cor_df <- rbind(cur_cor_df, cor_df)


  cur_amylo_genes <- subset(cur_cor_df, type == 'amylo' & cor > 0) %>% .$gene
  cur_oc_genes <- subset(cur_cor_df, type == 'oc' & cor > 0) %>% .$gene

  amylo_signif <- rbind(
    amylo_signif,
    degs_amylo %>% subset(group == cur_group & gene_id %in% cur_amylo_genes)
  )

 oc_signif <- rbind(
    oc_signif,
    degs_oc %>% subset(group == cur_group & gene_id %in% cur_oc_genes)
  )


}

write.csv(cor_df, file=paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))


# save the significant DEGs only:
write.csv(amylo_signif, file=paste0(data_dir, '5x_Amylo_Area_log_Gi_DEGs_signif.csv'))
write.csv(oc_signif, file=paste0(data_dir, '5x_OC_Area_log_Gi_DEGs_signif.csv'))


```

Plot the number of amyloid DEGs in each group 

```{r eval=FALSE}

amylo_cp <- c("#FF01F8", "grey", "#FCFD03")
names(amylo_cp) <- c('amylo', 'shared', 'oc')

# load the correlation df  
cor_df <- read.csv(paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))
cor_df$group <- factor(as.character(cor_df$group), levels = levels(seurat_5x$annotation)) 
cor_df <- arrange(cor_df, group)


groups <- unique(cor_df$group); groups <- groups[groups != 'unknoqn']

plot_df <- data.frame()
totals <- c()
for(cur_group in groups){
  
  cur_amylo <- cor_df %>% subset(type == 'amylo' & group == cur_group & cor > 0) %>% .$gene
  cur_oc <- cor_df %>% subset(type == 'oc' & group == cur_group & cor > 0) %>% .$gene


  shared <- intersect(cur_amylo, cur_oc)
  amylo_only <- cur_amylo[! cur_amylo %in% shared]
  oc_only <- cur_oc[! cur_oc %in% shared]


  cur_df <- data.frame(
    group = cur_group,
    n = c(length(amylo_only), length(shared), length(oc_only)),
    genes = c('amylo', 'shared', 'oc')
  )

  totals <- c(totals, sum(cur_df$n))

  plot_df <- rbind(plot_df, cur_df)

}
names(totals) <- groups
totals <- totals[order(totals)]
group_order <- names(totals)

plot_df$group <- factor(as.character(plot_df$group), levels=group_order)
plot_df$genes <- factor(as.character(plot_df$genes), levels=c('amylo', 'shared', 'oc'))

plot_df %<>% group_by(group) %>%   
  mutate(pos = cumsum(n) - (0.5 * n))


p <- plot_df %>%
  ggplot(aes(x = n, y=group, fill=genes)) + 
  geom_bar(position='stack', stat='identity') + 
  scale_fill_manual(values=amylo_cp) + 
  geom_text(aes(label=n), position = position_stack(vjust=0.5)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank()
  ) + 
  xlab(bquote("N"[genes]))
#      ylab(bquote("-log"[10]~"(Adj. P-value)")) +


pdf(paste0(fig_dir, 'amyloid_genes_barchart.pdf'), width=9, height=4, useDingbats=FALSE)
p
dev.off()


```

Feature plots for selected amyloid genes

```{r eval=FALSE}

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area_log)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character

features <- c('Cst3', 'Aldoc', 'Serpine2', 'Pvalb', 'Clu', 'Vim')


for(feat in features){


    p <- SampleFeaturePlot(
      seurat_5x,
      feature=feat,
      sample_labels = c("Condition", 'Age', 'Sex'),
      samples_to_plot = amyloid_samples,
      sample_col = "SAMPLE",
      ncol = 10,
      raster=TRUE,
      plot_max = 'q99',
      plot_min = 'q5',
      colfunc = inferno,
      rev_colors=TRUE,
      dpi=800,
      combine=FALSE
    )

    p <- lapply(p, function(x){
      x + theme(
        plot.title = element_text(face='bold', size=12, vjust=-1),
        plot.margin = margin(0,0,0,0)
      ) 

    })

    patch <- wrap_plots(p, ncol=6, widths=1, heights=1) + plot_layout(guides='collect')

    pdf(paste0(fig_dir, 'featureplots/featureplot_', feat, '_5xFAD.pdf'), width=12, height=6)
    print(patch)
    dev.off()

}


```

Heatmap of the amylo genes in the mouse dataset 

Need to narrow down the list of genes somehow!!!

```{r eval=FALSE}


# load the amylo and the OC DEGs
degs_amylo <- read.csv(file=paste0(data_dir, '5x_Amylo_Area_log_Gi_DEGs.csv'))
degs_oc <- read.csv(file=paste0(data_dir, '5x_OC_Area_log_Gi_DEGs.csv'))

# subset by significant results
degs_amylo <- subset(degs_amylo, q_value < 0.05)
degs_oc <- subset(degs_oc, q_value < 0.05)

# load the correlation df  
cor_df <- read.csv(paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))
cor_df$group <- factor(as.character(cor_df$group), levels = levels(seurat_5x$annotation)) 
cor_df <- arrange(cor_df, group)

# how many unique genes for amylo & oc across all the groups?
amylo_genes <- cor_df %>% subset(type == 'amylo' & cor > 0) %>% 
  group_by(group) %>% 
  arrange(desc(cor), .by_group=TRUE) %>% .$gene %>% unique

tmp <- cor_df %>% subset(type == 'amylo' & cor > 0)
table(tmp$group)
length(amylo_genes)

# how many unique genes for amylo & oc across all the groups?
oc_genes <- cor_df %>% subset(type == 'oc' & cor > 0) %>%
  group_by(group) %>% 
  arrange(desc(cor), .by_group=TRUE) %>% .$gene %>% unique

tmp <- cor_df %>% subset(type == 'oc' & cor > 0)
table(tmp$group)
length(oc_genes)

length(intersect(amylo_genes, oc_genes))

seurat_5x <- ScaleData(seurat_5x, features = amylo_genes)


seurat_5x$barcode <- colnames(seurat_5x)
temp <- table(seurat_5x@meta.data$annotation)
temp <- temp[names(temp) != 'unknown']
df <- data.frame()
for(i in 1:length(temp)){

  if(temp[[i]] < 500){
    cur_df <- seurat_5x@meta.data %>% subset(annotation == names(temp)[i])
  } else{
    cur_df <- seurat_5x@meta.data %>% subset(annotation == names(temp)[i]) %>% sample_n(500)
  }
  df <- rbind(df, cur_df)
}


p <- DoHeatmap(
  seurat_5x %>% subset(barcode %in% df$barcode),
  features=intersect(amylo_genes, oc_genes),
  group.by='annotation',
  #group.colors = human_cp,
  raster=TRUE, slot='scale.data'
) + theme(
  axis.text.y = element_text(face='italic', size=3)
)

pdf(paste0(fig_dir, 'amyloid_genes_heatmap_5x.pdf'), width=12, height=8, useDingbats=FALSE)
p
dev.off()


```

Compare the OC and Amylo Results 

```{r eval=FALSE}

# load the amylo and the OC DEGs
degs_amylo <- read.csv(file=paste0(data_dir, '5x_Amylo_Area_log_Gi_DEGs.csv'))
degs_oc <- read.csv(file=paste0(data_dir, '5x_OC_Area_log_Gi_DEGs.csv'))

# subset by significant results
degs_amylo <- subset(degs_amylo, q_value < 0.05)
degs_oc <- subset(degs_oc, q_value < 0.05)

# load the correlation df  
cor_df <- read.csv(paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))

# how many unique genes for amylo & oc across all the groups?
amylo_genes <- cor_df %>% subset(type == 'amylo' & cor > 0) %>% .$gene %>% unique
tmp <- cor_df %>% subset(type == 'amylo' & cor > 0)
table(tmp$group)
length(amylo_genes)

# how many unique genes for amylo & oc across all the groups?
oc_genes <- cor_df %>% subset(type == 'oc' & cor > 0) %>% .$gene %>% unique
tmp <- cor_df %>% subset(type == 'oc' & cor > 0)
table(tmp$group)
length(oc_genes)

library(GeneOverlap)

groups <- levels(seurat_5x$annotation); groups <- groups[groups != 'unknown']
genome.size <- nrow(seurat_5x)
overlap_df <- do.call(rbind, lapply(groups, function(cur_group){

  cur_amylo_up <- cor_df %>% subset(type == 'amylo' & group == cur_group & cor > 0) %>% .$gene
  cur_oc_up <- cor_df %>% subset(type == 'oc' & group == cur_group & cor > 0) %>% .$gene

  cur_overlap <- testGeneOverlap(newGeneOverlap(
      cur_amylo_up,
      cur_oc_up,
      genome.size=genome.size
  ))
  

  cur_overlap <- data.frame(
    'odds.ratio' = cur_overlap@odds.ratio,
    'pval' = cur_overlap@pval,
    'Jaccard' = cur_overlap@Jaccard,
    'size_intersection' = length(cur_overlap@intersection),
    'group' = cur_group,
    'size_set1' = length(cur_amylo_up),
    'size_set2' = length(cur_oc_up),
    'size_union' = length(cur_overlap@union)
  )

  cur_overlap
})) %>% as.data.frame
overlap_df$fdr <- p.adjust(overlap_df$pval)
overlap_df$group <- factor(
  as.character(overlap_df$group),
  levels = groups
)

overlap_df$shape <- ifelse(overlap_df$fdr < 0.05, 21, 4)
overlap_df <- overlap_df %>% arrange(odds.ratio, descending=TRUE) %>%
  mutate(group = factor(as.character(group), levels=as.character(group))) %>%
  mutate(percent_overlap = size_intersection / size_union)

p <- overlap_df %>%
  ggplot(aes(x=odds.ratio, y=group, size=percent_overlap, fill=group, color=group)) + 
  geom_segment(aes(y=group, yend=group, x=0, xend=odds.ratio), size=0.5, alpha=0.5) +
  geom_point() +
  geom_point(shape=overlap_df$shape, color='black', fill=NA) +
  scale_color_manual(values=mouse_cp, guide='none') +
  scale_fill_manual(values=mouse_cp, guide='none') +
  ylab('') + xlab("Odds ratio") +
  #scale_x_continuous(breaks = c(0, 1, 2,3)) +
  labs(size='Percent overlap') +
  ggtitle('Overlap of Amylo and OC DEGs') +
    theme(
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    plot.title = element_text(hjust=0.5, face='plain')
  )

p <- p + ggforestplot::geom_stripes(aes(y=group), inherit.aes=FALSE, data=overlap_df) 


pdf(paste0(fig_dir, 'amylo_oc_overlap_lolipop.pdf'), width=6, height=5)
p
dev.off()


```


EnrichR on DEGs

```{r eval=FALSE}


# load the correlation df  
cor_df <- read.csv(paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))


################################################################################
# Run enrichr
################################################################################

library(enrichR)

dbs <-c('GO_Biological_Process_2021','GO_Cellular_Component_2021','GO_Molecular_Function_2021', 'WikiPathway_2021_Mouse', 'KEGG_2021_Mouse')

# plot settings
groups <- unique(cor_df$group)
logfc_thresh <- 0
combined_output <- data.frame()

# run loop over each group
for(cur_group in groups){
  print(cur_group)

  # get genes that are up / down
  cur_amylo <- subset(cor_df, group == cur_group & cor > 0 & type == 'amylo') %>% .$gene
  cur_oc <- subset(cor_df, group == cur_group & cor > 0 & type == 'oc') %>% .$gene
  cur_shared <- intersect(cur_amylo, cur_oc)
  cur_amylo_only <- setdiff(cur_amylo, cur_oc)
  cur_oc_only <- setdiff(cur_oc, cur_amylo)

  # setup input list for enrichr loop
  input_list <- list(
    amylo = cur_amylo,
    oc = cur_oc,
    shared = cur_shared,
    amylo_only = cur_amylo_only,
    oc_only = cur_oc_only
  )

  # size of lists
  lapply(input_list, function(x){
    print(length(x))
  })

  # run enrichr and combine outputs
  enriched_df <- do.call(rbind, lapply(names(input_list), function(x){
    Sys.sleep(5)
    if(length(input_list[[x]]) > 0){
      cur_enrich <- enrichr(input_list[[x]], dbs)
    } else{return(data.frame())}
    cur_df <- do.call(rbind, lapply(dbs, function(cur_db){
      df <- cur_enrich[[cur_db]]
      if(nrow(df) > 1){df$degs <- x; df$group <- cur_group; df$db <- cur_db}
      else{df <- data.frame()}
      df
    }))
  }))

  combined_output <- rbind(combined_output, enriched_df)

}

enriched_df$ngenes <- unlist(lapply(strsplit(enriched_df$Genes, ';'), function(x){length(x)}))
subset(enriched_df, P.value < 0.05 & ngenes >= 3) %>% .$degs %>% table

# write the output to a tsv
write.table(combined_output, file=paste0(data_dir, '5x_Amyloglo_GO_terms.tsv'), quote=FALSE, row.names=FALSE, sep='\t')


combined_output$ngenes <- unlist(lapply(strsplit(combined_output$Genes, ';'), function(x){length(x)}))
combined_output <- subset(combined_output, ngenes >= 3 & P.value < 0.05)

subset(combined_output, group == 'ctx-deep-layers') %>% .$degs %>% table

combined_output %>%
  subset(P.value < 0.05) %>%
  write.table(
    file=paste0(data_dir, '5x_Amyloglo_GO_terms_signif.tsv'),
    quote=FALSE, row.names=FALSE, sep='\t'
  )


################################################################################
# Plot selected GO terms for upreg
################################################################################

# helper function to wrap text
wrapText <- function(x, len) {
    sapply(x, function(y) paste(strwrap(y, len), collapse = "\n"), USE.NAMES = FALSE)
}

combined_output <- read.delim(
  file=paste0(data_dir, '5x_Amyloglo_GO_terms.tsv'),
  header=1, sep='\t'
)

selected_terms <- read.delim('data/5x_Amyloglo_GO_terms_selected.txt', sep='\t', header=1)
table(selected_terms$group)

# subset selected terms for up-regulated
selected_terms <- subset(combined_output, Term %in% selected_terms$Term & P.value < 0.05 & degs == 'shared')
table(selected_terms$group)

groups <- levels(seurat_5x$annotation)
groups <- groups[groups != 'unknown']
selected_terms$group <- factor(
  as.character(selected_terms$group),
  levels = groups
)
table(selected_terms$group)

# set max pval

quantile(-log(selected_terms$P.value), 0.95)
max_p <- 10

selected_terms$logp <- -log(selected_terms$P.value)
selected_terms$logp <- ifelse(selected_terms$logp > max_p, max_p, selected_terms$logp)

# remove GO Term ID
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

selected_terms <- selected_terms %>%
  arrange(group)


selected_terms$wrap <- wrapText(selected_terms$Term, 35)

selected_terms$Term <- factor(
  as.character(selected_terms$Term),
  levels = rev(unique(as.character(selected_terms$Term)))
)

# GO Term dot plot

p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color =logp, size=log(Combined.Score))) +
  geom_point() +
  scale_color_stepsn(colors=rev(magma(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(linewidth=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(linewidth=0.25, color='lightgrey')
  ) + labs(
    color = bquote("-log"[10]~"(P)"),
    size= bquote("log"[10]~"(Enrich)")
  )




color_df <- data.frame(
  group = names(mouse_cp),
  colour = as.character(mouse_cp)
)
color_df <- color_df %>% subset(group %in% selected_terms$group)

# make the colorbar as its own heatmap
color_df$var <- 1
cp <- color_df$colour; names(cp) <- color_df$group
colorbar <- color_df %>%
  ggplot(aes(x=group, y=var, fill=group)) +
  geom_tile() +
  scale_fill_manual(values=cp) +
  coord_equal() +
  NoLegend() + RotatedAxis() +
  theme(
    plot.title=element_blank(),
    axis.line=element_blank(),
    axis.ticks.y =element_blank(),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.margin=margin(0,0,0,0),
  )


pdf(paste0(fig_dir, '5x_Amyloglo_selected_GO_terms.pdf'), width=7, height=5)
p / colorbar
dev.off()


################################################################################
# Plot selected GO terms for upreg
################################################################################


# helper function to wrap text
wrapText <- function(x, len) {
    sapply(x, function(y) paste(strwrap(y, len), collapse = "\n"), USE.NAMES = FALSE)
}

combined_output <- read.delim(
  file=paste0(data_dir, '5x_Amyloglo_GO_terms.tsv'),
  header=1, sep='\t'
)

selected_terms <- read.delim('data/5x_Amyloglo_GO_terms_selected.txt', sep='\t', header=1)
table(selected_terms$group)

# subset selected terms for up-regulated
selected_terms <- subset(combined_output, Term %in% selected_terms$Term & P.value < 0.05 & degs == 'down_amyloid')
table(selected_terms$group)



groups <- levels(seurat_5x$annotation)
groups <- groups[groups != 'unknown']
selected_terms$group <- factor(
  as.character(selected_terms$group),
  levels = groups
)
table(selected_terms$group)

# set max pval

quantile(-log(selected_terms$P.value), 0.95)
max_p <- 10

selected_terms$logp <- -log(selected_terms$P.value)
selected_terms$logp <- ifelse(selected_terms$logp > max_p, max_p, selected_terms$logp)

# remove GO Term ID
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

selected_terms <- selected_terms %>%
  arrange(group)


selected_terms$wrap <- wrapText(selected_terms$Term, 35)

selected_terms$Term <- factor(
  as.character(selected_terms$Term),
  levels = rev(unique(as.character(selected_terms$Term)))
)

# GO Term dot plot

p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color =logp, size=log(Combined.Score))) +
  geom_point() +
  scale_color_stepsn(colors=rev(magma(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(linewidth=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(linewidth=0.25, color='lightgrey')
  ) + labs(
    color = bquote("-log"[10]~"(P)"),
    size= bquote("log"[10]~"(Enrich)")
  )




color_df <- data.frame(
  group = names(mouse_cp),
  colour = as.character(mouse_cp)
)
color_df <- color_df %>% subset(group %in% selected_terms$group)

# make the colorbar as its own heatmap
color_df$var <- 1
cp <- color_df$colour; names(cp) <- color_df$group
colorbar <- color_df %>%
  ggplot(aes(x=group, y=var, fill=group)) +
  geom_tile() +
  scale_fill_manual(values=cp) +
  coord_equal() +
  NoLegend() + RotatedAxis() +
  theme(
    plot.title=element_blank(),
    axis.line=element_blank(),
    axis.ticks.y =element_blank(),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.margin=margin(0,0,0,0),
  )


pdf(paste0(fig_dir, '5x_Amyloglo_selected_GO_terms_down.pdf'), width=9, height=9)
p / colorbar
dev.off()


```

```{r eval=FALSE}


################################################################################
# Overlap with other gene lists:
#
# * ST marker genes
# * SC marker genes
# * AD vs Control DEGs ST & SC
# * ADDS vs Control DEGs ST & SC
# * Known gene signatures (DAM, DAA, DOL, PIGs, GWAS)
################################################################################


# load mouse <-> human gene name table:
hg38_mm10_genes <- read.table(
  "/dfs7/swaruplab/smorabit/resources/hg38_mm10_orthologs_2021.txt",
  sep='\t',
  header=TRUE
)
colnames(hg38_mm10_genes) <-c('hg38_id', 'mm10_id', 'mm10_name', 'hg38_name')
hg38_mm10_genes <- dplyr::select(hg38_mm10_genes, c(hg38_name, mm10_name, hg38_id, mm10_id))

hg38_mm10_genes <- subset(hg38_mm10_genes, mm10_name != '' & hg38_name != '')

# need to make sure that there's only one entry for each gene in hg38_mm10_genes
mm10_genes <- unique(hg38_mm10_genes$mm10_name)
hg38_genes <- unique(hg38_mm10_genes$hg38_name)
hg38_mm10_genes <- hg38_mm10_genes[match(mm10_genes, hg38_mm10_genes$mm10_name),]


# Load PIGs from Chen et al 2020
pig_table <- read.csv('~/swaruplab/smorabit/analysis/ADDS_2021/data/chen_modules.csv')

pigs <- subset(pig_table, module == 'PIG') %>% .$gene
oligs <- subset(pig_table, module == 'OLIG') %>% .$gene


intersect(pigs, cur_up)
intersect(pigs, cur_down)

intersect(oligs, cur_up)
intersect(oligs, cur_down)








# load the correlation df  
cor_df <- read.csv(paste0(data_dir, '5x_amylo_oc_correlation_DEGs.csv'))
cor_df$group <- factor(as.character(cor_df$group), levels = levels(seurat_5x$annotation)) 
cor_df <- arrange(cor_df, group)

groups <- as.character(levels(cor_df$group)); groups <- groups[groups != 'unknown']
list_names <- c('DAM', 'DAA', 'DOL', 'PIGs')



# run overlaps between module gene lists and DEG lists:
overlap_df <- do.call(rbind, lapply(list_names, function(cur_list){
  cur_genes <- gene_lists[[cur_list]]
  cur_overlap_df <- do.call(rbind, lapply(groups, function(cur_group){

  
  amylo_genes <- subset(cor_df, cor > 0 & group == cur_group & type == 'amylo') %>% .$gene
  oc_genes <- subset(cor_df, cor > 0 & group == cur_group & type == 'oc') %>% .$gene


    cur_overlap_amylo <- testGeneOverlap(newGeneOverlap(
        cur_genes,
        amylo_genes,
        genome.size=nrow(seurat_5x)
    ))

    # cur_overlap_oc <- testGeneOverlap(newGeneOverlap(
    #     cur_genes,
    #     oc_genes,
    #     genome.size=nrow(seurat_5x)
    # ))

    cur_overlap_oc<- testGeneOverlap(newGeneOverlap(
      cur_genes,
      intersect(amylo_genes, oc_genes),
      genome.size=nrow(seurat_5x)
    ))

    cur_overlap <- data.frame(
      'odds.ratio' = c(cur_overlap_amylo@odds.ratio, cur_overlap_oc@odds.ratio),
      'pval' = c(cur_overlap_amylo@pval, cur_overlap_oc@pval),
      'Jaccard' = c(cur_overlap_amylo@Jaccard, cur_overlap_oc@Jaccard),
      'size_set1' = length(cur_genes),
      'size_set2' = c(length(amylo_genes), length(oc_genes)),
      'size_intersection' = c(length(cur_overlap_amylo@intersection), length(cur_overlap_oc@intersection)),
      'size_union' = c(length(cur_overlap_amylo@union), length(cur_overlap_oc@union)),
      'list' = cur_list,
      'gene_type' = c('amylo', 'oc'),
      'group' = cur_group,
      'gene_list' = cur_list
    )


  })) %>% as.data.frame
  #colnames(cur_overlap_df) <- c('odds_ratio', 'pval', 'Jaccard', 'size_intersection')
  #cur_overlap_df$m1 <- cur_mod1
  #cur_overlap_df$m2 <- mods2

  # module color:
  #cur_overlap_df$color <- modules %>% subset(module == cur_mod) %>% .$color %>% unique
  cur_overlap_df
}))


overlap_df <- overlap_df %>% subset(gene_type == 'oc') %>%
  group_by(group) %>%
  mutate(fdr=p.adjust(pval, method='fdr')) %>%
  ungroup


# significance level:
overlap_df$Significance <- gtools::stars.pval(overlap_df$fdr)
overlap_df$Significance <- ifelse(
  overlap_df$Significance == '.', '',
  overlap_df$Significance
)



# plot the results as a heatmap:
maxval <- 100
plot_df <- overlap_df
plot_df$odds.ratio <- ifelse(plot_df$odds.ratio > maxval, maxval, plot_df$odds.ratio)
plot_df$textcolor <- ifelse(plot_df$odds.ratio > 0.7*maxval, 'white', 'black')

plot_df$group <- fct_rev(factor(as.character(plot_df$group)))

p <- plot_df %>%
  ggplot(aes(y=group, x=gene_list, fill=odds.ratio)) +
  geom_tile() +
  geom_text(label=plot_df$Significance, vjust = 0.72, color=plot_df$textcolor) +
  scale_fill_gradient(low='white', high='blue') +
  RotatedAxis() +
  labs(fill = 'Odds ratio') +
  theme(
    panel.border = element_rect(fill=NA, color='black', linewidth=1),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    #axis.text.x = element_blank(),
    #axis.ticks.x = element_blank(),
    #axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    plot.margin=margin(0,0,0,0)
  ) +
  coord_equal()


pdf(paste0(fig_dir, 'amyloid_geneset_overlap_heatmap.pdf'), width=5, height=4, useDingbats=FALSE)
p 
dev.off()







```


For each cluster and in each sample, compute the total amyloid area / number of ST spots 

```{r eval=FALSE}

# only plot samples that have amyloid scores:
amyloid_samples <- seurat_amyloid_5x@meta.data %>%
  group_by(SAMPLE) %>%
  summarise(Mean = mean(Amylo_Area)) %>%
  filter(Mean != 0) %>% .$SAMPLE %>% as.character


plot_df <- seurat_amyloid_5x@meta.data %>% 
  group_by(SAMPLE, annotation, Age) %>%
  summarise(
    oc_sum = sum(OC_Number) / n(), 
    amylo_sum = sum(Amylo_Number) / n(),
  ) %>% 
  subset(annotation != 'unknown') %>%
  ungroup 


plot_df$Age <- factor(
  as.character(plot_df$Age), 
  levels = names(cp_5x)
)


my_comparisons <- list(c('4mo', '6mo'), c('4mo', '8mo'), c('4mo', '12mo'))

p1 <- ggplot(plot_df, aes(x=as.numeric(Age), y = oc_sum)) + 
  geom_boxplot(aes(fill=Age), outlier.shape=NA, width=0.75) + 
  geom_jitter(size=0.5, width=0.25) +
  geom_smooth(method='lm', color='darkgrey') +
  scale_fill_manual(values=cp_5x) +
  RotatedAxis() +
  stat_cor(size=2) +  theme(
    axis.text.x= element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    plot.margin = margin(c(0,0,0,0)),
    strip.text = element_text(size=7)
  ) +
  ylab('Avg. OC Num') + 
  facet_wrap(~annotation, ncol=7) 
 
p2 <- ggplot(plot_df, aes(x=as.numeric(Age), y = amylo_sum)) + 
  geom_boxplot(aes(fill=Age), outlier.shape=NA, width=0.75) + 
  geom_jitter(size=0.5, width=0.25) +
  geom_smooth(method='lm', color='darkgrey') +
  scale_fill_manual(values=cp_5x) +
  RotatedAxis() +
  stat_cor(size=2)+
  theme(
    axis.text.x= element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    plot.margin = margin(c(0,0,0,0)),
    strip.text = element_text(size=7)
  ) +
  ylab('Avg. Amyloglo Num') + 
  facet_wrap(~annotation, ncol=7) 

pdf(paste0(fig_dir, 'amyloid_box2.pdf'), width=10, height=7)
p1 / p2 + plot_layout(guides='collect')
dev.off()



```























OLD ZONE BELOW

Test setting up a SpatialFeatureExperiment for one mouse sample

```{r eval=FALSE}

setwd('voyager/')

library(Voyager)
library(SpatialFeatureExperiment)
library(scater)
library(scran)
library(SFEData)
library(sf)
library(ggplot2)
library(scales)
library(patchwork)
library(BiocParallel)
library(bluster)
library(tidyverse)
library(ggpubr)


dir.create('figures/5x')
fig_dir <- 'figures/5x/'

# which sample to pick?
names(seurat_amyloid_5x@meta.data)
table(seurat_amyloid_5x$SAMPLE, seurat_amyloid_5x$Age)
table(seurat_amyloid_5x$SampleID)


gene_file <- '~/swaruplab/smorabit/data/Visium_Mouse_2021/Dec_20_2021/spaceranger_count/5XFAD28/outs/filtered_feature_bc_matrix/features.tsv.gz'
gene_table <- read.table(gene_file, sep='\t')


cur_vis_sample <- '5101'

# get the current visium sample and make sure to set up the image
cur_vis <- seurat_amyloid_5x[,seurat_amyloid_5x@meta.data[["SAMPLE"]] == cur_vis_sample]
cur_image <- names(cur_vis@images)[sapply(names(cur_vis@images), function(x){nrow(cur_vis@images[[x]]@coordinates) > 0})]
cur_vis@images <- list(cur_image = cur_vis@images[[cur_image]])

# get the barcode:
cur_vis$bc <- do.call(rbind, strsplit(colnames(cur_vis), '_'))[,1]

######################################################################
# make a sfe object from the spaceranger output:
#
# note: I tried to make it from scratch and it didn't work
######################################################################

file <- '~/swaruplab/smorabit/data/Visium_Mouse_2021/Dec_20_2021/spaceranger_count/5XFAD28/outs'
sfe <- read10xVisiumSFE(file, type='sparse', data='raw', load=FALSE)

# change to gene names
all.equal(gene_table$V1, rownames(sfe))
rownames(sfe) <- gene_table$V2

# only keep spots that are in the tissue
sfe <- sfe[, colData(sfe)$in_tissue]
#sfe <- sfe[rowSums(counts(sfe)) > 0,]

# same size as the Seurat object?
dim(sfe)
dim(cur_vis)
all.equal(as.character(colnames(sfe)), as.character(cur_vis$bc))

meta <- cur_vis@meta.data
rownames(meta) <- colnames(sfe)
colData(sfe) <- cbind(colData(sfe), meta)

# log normalize data
clusters <- sfe$annotation
sfe <- computeSumFactors(sfe, clusters=clusters)
sfe <- logNormCounts(sfe)

# compute spatial neighborhood graph
colGraph(sfe, "visium") <- findVisiumGraph(sfe)

# compute a binary-weight graph
colGraph(sfe, "visium_B") <- findVisiumGraph(sfe, style = "B")


################################################################################
# Compute Local Moran's I & Getis-Ord(*) for different features
################################################################################

# compute log scale of OC and Amyloglo scores
colData(sfe)$Amylo_Area_log <- log(sfe$Amylo_Area + 1)
colData(sfe)$OC_Area_log <- log(sfe$OC_Area + 1)

# select features
feats <- c(paste0(names(gene_lists), '_UCell'), 'GFAP_high_UCell', 'DAA_UCell', 'GFAP_low_UCell', 'Amylo_Area_log', 'OC_Area_log')

# compute Moran's I
sfe <- colDataUnivariate(sfe, type = "localmoran", features = feats,
                            colGraphName = "visium")

# compute Getis-Ord Gi*
sfe <- colDataUnivariate(sfe, type = "localG_perm", features = feats,
                            colGraphName = "visium_B", include_self=TRUE)


for(cur_feat in feats){
  print(cur_feat)
  p1 <- plotSpatialFeature(sfe, features = cur_feat, colGeometryName = "spotPoly") +
    RotatedAxis() + ggtitle(cur_feat) + labs(fill='') + umap_theme()
  p2 <- plotLocalResult(sfe, "localmoran", features = cur_feat,
                  colGeometryName = "spotPoly",divergent = TRUE,
                  diverge_center = 0) +
                    RotatedAxis() + labs(fill='') + umap_theme()
  p3 <- plotLocalResult(sfe, "localmoran", features = cur_feat,
                  colGeometryName = "spotPoly", attribute = "-log10p_adj", divergent = TRUE,
                  diverge_center = -log10(0.05)) +
                    RotatedAxis() + labs(fill='') + umap_theme()

  p4 <- plotLocalResult(sfe, "localG_perm", features = cur_feat,
                  colGeometryName = "spotPoly",divergent = TRUE,
                  diverge_center = 0) +
                  ggtitle('Getis-Ord Gi(*)') +
                  RotatedAxis() + labs(fill='') + umap_theme()

  # p4 <- plotLocalResult(sfe, "localmoran", features = cur_feat,
  #                 colGeometryName = "spotPoly", attribute = "pysal") +
  #                 RotatedAxis() + labs(fill='') + umap_theme()

  pdf(paste0(fig_dir, 'plot_', cur_feat,'_5x.pdf'), width=8, height=6)
  print((p1 + p2) / (p3 + p4))
  dev.off()
}


################################################################################
# Correlate Getis-Ord Gi(*) for DAM & OC
################################################################################


for(amyloid in c('Amylo_Area_log', 'OC_Area_log')){
  for(feat in c(paste0(names(gene_lists), '_UCell'), 'GFAP_high_UCell', 'DAA_UCell', 'GFAP_low_UCell')){

    print(paste0(feat, ' ', amyloid))
    meta <- as.data.frame(colData(sfe))
    meta$feat1 <- as.numeric( localResult(sfe, 'localG_perm', amyloid)[,'localG'])
    meta$feat2 <- as.numeric( localResult(sfe, 'localG_perm', feat)[,'localG'])

    meta <- subset(meta, annotation!= 'unknown')

    p <- meta %>%
      ggplot(aes(x = feat1, y = feat2, color = annotation)) +
      geom_hline(yintercept=0, linetype='dashed', color='grey90') +
      geom_vline(xintercept=0, linetype='dashed', color='grey90') +
      geom_point(alpha=0.2) +
      geom_smooth(inherit.aes=FALSE, data=meta, mapping = aes(x = feat1, y = feat2, color=annotation), method='lm') +
      stat_cor(inherit.aes=FALSE, data=meta, mapping = aes(x = feat1, y = feat2), method='pearson') +
      #xlim(c(-plot_range, plot_range)) +
      #ylim(c(-plot_range, plot_range)) +
      theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
        plot.title = element_text(hjust=0.5)
      ) + NoLegend() +
      scale_color_manual(values=mouse_cp)

    p <- p +
    xlab(paste0(amyloid, " Getis-Ord Gi(*)")) +
    ylab(paste0(feat, " Getis-Ord Gi(*)"))

    pdf(paste0(fig_dir, 'GI_corr/corr_', amyloid, '_', feat, '_Gi_corr.pdf'), width=8, height=8)
    print(p + facet_wrap(~ annotation, ncol=4))
    dev.off()

  }
}


################################################################################
# Correlate Moran's I for DAM & OC
################################################################################


for(amyloid in c('Amylo_Area_log', 'OC_Area_log')){
  for(feat in c(paste0(names(gene_lists), '_UCell'), 'GFAP_high_UCell', 'DAA_UCell', 'GFAP_low_UCell')){

    print(paste0(feat, ' ', amyloid))
    meta <- as.data.frame(colData(sfe))
    meta$feat1 <- as.numeric( localResult(sfe, 'localmoran', amyloid)[,'Ii'])
    meta$feat2 <- as.numeric( localResult(sfe, 'localmoran', feat)[,'Ii'])

    meta <- subset(meta, annotation!= 'unknown')

    p <- meta %>%
      ggplot(aes(x = feat1, y = feat2, color = annotation)) +
      geom_hline(yintercept=0, linetype='dashed', color='grey90') +
      geom_vline(xintercept=0, linetype='dashed', color='grey90') +
      geom_point(alpha=0.2) +
      geom_smooth(inherit.aes=FALSE, data=meta, mapping = aes(x = feat1, y = feat2, color=annotation), method='lm') +
      stat_cor(inherit.aes=FALSE, data=meta, mapping = aes(x = feat1, y = feat2), method='pearson') +
      #xlim(c(-plot_range, plot_range)) +
      #ylim(c(-plot_range, plot_range)) +
      theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
        plot.title = element_text(hjust=0.5)
      ) + NoLegend() +
      scale_color_manual(values=mouse_cp)

    p <- p +
    xlab(paste0(amyloid, " local Moran's I")) +
    ylab(paste0(feat, " local Moran's I"))

    pdf(paste0(fig_dir, 'LMI_corr/corr_', amyloid, '_', feat, '_LMI_corr.pdf'), width=8, height=8)
    print(p + facet_wrap(~ annotation, ncol=4))
    dev.off()

  }
}

```
