
```{r eval=FALSE}

library(Seurat)
library(tidyverse)
library(cowplot)
library(patchwork)
library(ggrepel)
library(viridis)
library(magrittr)
theme_set(theme_cowplot())

source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')


setwd("/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/scDRS/")

fig_dir <- 'figures/'
data_dir <- 'data/'

# load 5xFAD
seurat_5x <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/voyager/data/5xFAD_seurat_processed_amyloid.rds')

cp_5x <- c("4mo" = "#E8C0FC", "6mo" = "#B57DD1",
        "8mo"= "#8C64A1", "12mo" = "#643D78")


cp_5x <- c(
    "4mo-5X" = "#E8C0FC", "4mo-WT" = "#BDFFDC",
    "6mo-5X" = "#B57DD1", "6mo-WT" = "#84DBB5",
    "8mo-5X"= "#8C64A1",  "8mo-WT"= "#45BA7C", 
    "12mo-5X" = "#643D78", "12mo-WT" = "#14703F",
    'All' = 'grey20'
)



mouse_cp <- c(
      "ctx-deep-layers" = "#5b4468", "ctx-upper-layers" = "#9581a4", "ctx-olfactory" = "#8073ab",
      "hippocampus" = "#d6a5b2", "hippocampus-pyramidal" = "#e193aa",
      "lateral-ventricle" = "#d2c3b2",
      "striatum" = "#F0DE8C",
      "thalamus1" = "#00CBA7", "thalamus2" = "#abcc94",
      "hypothalamus-amygdala" = "#72c9b1",
      "WM1" = "#325ea8", "WM2" = "#64bcdb", "WM-cerebral-peduncle" = "#62a7d7",
      "erythrocytes-neurons" = "#d07fc4", "unknown" = "#c47c4d"
    )

color_df <- data.frame(
  colour = as.character(mouse_cp),
  group = names(mouse_cp)
)
color_df$group <- factor(
  as.character(color_df$group),
    levels=names(mouse_cp)
)
color_df <- arrange(color_df, group)



```


Function to compute cell group / disease associations 

```{r eval=FALSE}


TraitGroupAssociations <- function(
    seurat_obj,
    scdrs_file,
    group.by,
    percentile = 0.95,
    pval_thresh = c(0.01, 0.05, 0.1)
){

    # check if "full" is in the name of the scdrs file  
    if(!grepl("full_score", scdrs_file)){
        warning("Please check that this is the full_score output from scDRS.")
    }

    # check that the group.by is in the Seurat object:
    if(!(group.by %in% names(seurat_obj@meta.data))){
        stop(paste0(group.by, ' not found in seurat_obj@meta.data'))
    }

    # load the scDRS output file
    print('Loading scDRS output file ...')
    scdrs_df <- read_delim(scdrs_file,  delim='\t')
    names(scdrs_df)[1] <- 'barcode'

    # compute the FDR:
    scdrs_df$fdr <- p.adjust(scdrs_df$pval, method='fdr')

    # match the barcodes with the seurat object
   # dup_bcs <- names(which(table(seurat_obj@meta.data$bc) > 1))
   # seurat_meta <- subset(seurat_obj@meta.data, !(bc %in% dup_bcs))
    seurat_meta <- subset(seurat_obj@meta.data, bc %in% scdrs_df$barcode)
    scdrs_df <- subset(scdrs_df, barcode %in% seurat_meta$bc)
    ix <- match(seurat_meta$bc, scdrs_df$barcode)

    print(dim(seurat_meta))
    print(dim(scdrs_df))
    print(length(ix))

    all.equal(as.character(scdrs_df$barcode), as.character(seurat_meta$bc)[ix])

    # add the cell cluster info
    scdrs_df$cell_group <- as.character(seurat_meta[[group.by]][ix])
    group_list <- unique(scdrs_df$cell_group)

    # set up progress bar
    print('Computing associations between cell groups and current trait...')
    pb <- utils::txtProgressBar(min = 0, max = length(group_list),style = 3, width = 50, char = "=")
    i <- 1

    # loop through each cluster
    assoc_df <- data.frame()
    for(cur_group in group_list){
        
        # get the scores for this cluster
        group_scores <- subset(scdrs_df, cell_group == cur_group)

        # compute the percentile
        score_q <- as.numeric(quantile(group_scores$norm_score, percentile))

        # compute the percentile for each control score
        control_score_q <- group_scores %>% 
            summarise(across(starts_with("ctrl_norm_score"), .fns = ~ quantile(.x, percentile))) %>% 
            as.numeric

        # comptue p-val
        mc_p <- (sum(control_score_q >= score_q) + 1) / (length(control_score_q) + 1)
        mc_z <- (score_q - mean(control_score_q)) / sd(control_score_q)

        # compute the proportions of cells with signif p-vals:
        pval_props <- unlist(lapply(pval_thresh, function(thresh){
            as.numeric(table(group_scores$pval < thresh)[2] / nrow(group_scores))
        }))
        names(pval_props) <- paste0('prop_p_', pval_thresh)
        pval_props[is.na(pval_props)] <- 0
        fdr_props <- unlist(lapply(pval_thresh, function(thresh){
            as.numeric(table(group_scores$fdr < thresh)[2] / nrow(group_scores))
        }))
        names(fdr_props) <- paste0('prop_fdr_', pval_thresh)
        fdr_props[is.na(fdr_props)] <- 0

        # assemble output data for this cluster
        cur_assoc_df <- data.frame(group = cur_group, mc_p = mc_p, mc_z = mc_z)
        cur_assoc_df[,names(pval_props)] <- pval_props 
        cur_assoc_df[,names(fdr_props)] <- fdr_props
        assoc_df <- rbind(assoc_df, cur_assoc_df)

        # update progress bar:
        i <- i+ 1
        setTxtProgressBar(pb, i)
    }

    # close progress bar
    close(pb)

    # return the resutls
    assoc_df
}

```

Combine the scDRS output tables for each dataset 

```{r eval=FALSE}

# get the file paths for the output files
scdrs_output_dir <- paste0('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/scDRS/data/outputs/')
output_files <- dir(scdrs_output_dir)
output_files <- output_files[!grepl('full', output_files)]
traits <- gsub(".score.gz", '', output_files)
traits <- gsub("PASS_", '', traits)
output_files <- paste0(scdrs_output_dir, output_files)

# combine the individual tables
scdrs_df <- do.call(rbind, lapply(1:length(traits), function(i){
    cur_df <- read_delim(output_files[i],  delim='\t')
    names(cur_df)[1] <- 'barcode'
    cur_df$trait <- traits[i]
    cur_df
}))

# write the output file:
write.csv(scdrs_df, quote=FALSE, row.names=FALSE, file=paste0(data_dir,'5xFAD_visium_scDRS_results.csv'))




```


Compute the celltype / disease associations 

```{r eval=FALSE}

all(seurat_5x$bc %in% scdrs_df$barcode)
all(colnames(seurat_5x) %in% scdrs_df$barcode)

seurat_5x$bc <- colnames(seurat_5x)

scdrs_output_dir <- paste0('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/scDRS/data/outputs/')
output_files <- dir(scdrs_output_dir)
output_files <- output_files[grepl('full', output_files)]
traits <- gsub(".score.gz", '', output_files)
traits <- gsub("PASS_", '', traits)
output_files <- paste0(scdrs_output_dir, output_files)


assoc_df <- data.frame()
for(i in 1:length(traits)){
    tic()
    out <- TraitGroupAssociations(
        seurat_obj = seurat_5x,
        scdrs_file = output_files[i],
        group.by = 'annotation',
        percentile = 0.95
    )
    out$trait <- traits[i]
    time_elapsed <- toc()
    print(time_elapsed)
    assoc_df <- rbind(assoc_df, out)
}
write.csv(assoc_df, file=paste0(data_dir, '5xFAD_visium_scDRS_celltype_associations.csv'))





assoc_df$mc_fdr <- p.adjust(assoc_df$mc_p, method='fdr')

p <- ggplot(assoc_df, aes(x=group, y=trait, fill=prop_p_0.05)) + 
    geom_tile() + 
    geom_point(
        inherit.aes=FALSE,
        data = subset(assoc_df, mc_fdr < 0.05),
        aes(x=group, y=trait),
        color='black', 
        shape=0,
        size=4,
        fill=NA
    ) + 
    coord_equal() + 
    scale_fill_steps(low='white', high='red') + 
RotatedAxis()

pdf(paste0(fig_dir, '5xFAD_visium_scdrs_heatmap.pdf'), width=12, height=15)
p
dev.off()



```




Compute the celltype / disease associations separately for each condition 

```{r eval=FALSE}

all(seurat_5x$bc %in% scdrs_df$barcode)
all(colnames(seurat_5x) %in% scdrs_df$barcode)

seurat_5x$bc <- colnames(seurat_5x)


# get a list of scDRS output files:
scdrs_output_dir <- paste0('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/scDRS/data/outputs/')
output_files <- dir(scdrs_output_dir)
output_files <- output_files[grepl('full', output_files)]
traits <- gsub(".score.gz", '', output_files)
traits <- gsub("PASS_", '', traits)
output_files <- paste0(scdrs_output_dir, output_files)

# list of different conditions:
conditions <- unique(seurat_5x$Group)


assoc_df <- data.frame()
for(j in 1:length(conditions)){
    cur_condition <- conditions[j]
    print(cur_condition)
    cur_seurat <- subset(seurat_5x, Group == cur_condition)
    for(i in 1:length(traits)){
        out <- TraitGroupAssociations(
            seurat_obj = cur_seurat,
            scdrs_file = output_files[i],
            group.by = 'annotation',
            percentile = 0.95
        )
        out$trait <- traits[i]
        out$condition <- cur_condition
        assoc_df <- rbind(assoc_df, out) 
    }
}

# compute the FDR
assoc_df$mc_fdr <- p.adjust(assoc_df$mc_p, method='fdr')
write.csv(assoc_df, file=paste0(data_dir, '5xFAD_visium_diagnosis_scDRS_celltype_associations.csv'))

# assoc_df %<>% 
#     group_by(condition) %>% 
#     mutate(fdr = p.adjust(mc_p, method='fdr')) %>% 
#     ungroup

#-------------------------------------------------------------------------------------#
# quick and dirty of plot the results as a heatmap 
#-------------------------------------------------------------------------------------#

p <- ggplot(assoc_df, aes(x=group, y=trait, color=prop_p_0.05, size=-log10(fdr))) + 
    geom_point() + 
    geom_point(
        inherit.aes=FALSE,
        data = subset(assoc_df, fdr <= 0.05),
        aes(x=group, y=trait, size=-log10(fdr)),
        color='black', 
        shape=1,
        #size=-log10(fdr),
        fill=NA
    ) + 
    coord_equal() + 
    scale_color_steps(low='white', high='red') + 
    RotatedAxis()

pdf(paste0(fig_dir, 'ADDS_visium_scdrs_heatmap.pdf'), width=20, height=15)
p + facet_wrap(~condition, ncol=4)
dev.off()

assoc_df %>% subset(trait == 'UKB_460K.body_HEIGHTz.full')


p <- ggplot(assoc_df, aes(x=group, y=trait, color=prop_p_0.05, size=-log10(fdr))) + 
    geom_point() + 
    geom_point(
        inherit.aes=FALSE,
        data = subset(assoc_df, fdr <= 0.05),
        aes(x=group, y=trait, size=-log10(fdr)),
        color='black', 
        shape=1,
        #size=-log10(fdr),
        fill=NA
    ) + 
    coord_equal() + 
    scale_color_steps(low='white', high='red') + 
    RotatedAxis()

pdf(paste0(fig_dir, 'ADDS_visium_scdrs_heatmap_ratio.pdf'), width=20, height=15)
p + facet_wrap(~condition, ncol=4)
dev.off()


```



Ordering of the traits in the heatmap 

```{r eval=FALSE}


blood_immune_traits <- c(
    'CD_deLange2017', # chron's disease
    'Celiac',
    'IBD_deLange2017',
    'Lupus',
    'Multiple_sclerosis',
    'Rheumatoid_Arthritis',
    'Primary_biliary_cirrhosis',
    'UC_deLange2017',
    'Type_1_Diabetes',
    'EOSINOPHIL_COUNT',
    'LYMPHOCYTE_COUNT',
    'MEAN_CORPUSCULAR_HEMOGLOBIN',
    'MONOCYTE_COUNT',
    'PLATELET_COUNT',
    'RBC_DISTRIB_WIDTH',
    'RED_COUNT',
    'WHITE_COUNT',
    'AID_ALL', #AID = autoimmune disease
    'ALLERGY_ECZEMA_DIAGNOSED',
    'ASTHMA_DIAGNOSED',
    'HYPOTHYROIDISM_SELF_REP',
    'RESPIRATORY_ENT'
)

brain_traits <- c(
    'ADHD_Demontis2018',
    'Alzheimers_Jansen2019',
    'BIP_Mullins2021',
    'DrinksPerWeek_Liu2019',
    'GeneralRiskTolerance_KarlssonLinner2019',
    'Insomnia_Jansen2019',
    'Intelligence_SavageJansen2018',
    'MDD_Howard2019',
    'ReactionTime_Davies2018',
    'SWB',
    'Schizophrenia_Pardinas2018',
    'SleepDuration_Dashti2019',
    'VerbalNumericReasoning_Davies2018',
    'EDU_COLLEGE',
    'EDU_YEARS',
    'NEUROTICISM',
    'MORNINGPERSON',
    'BMIz',
    'SMOKING_STATUS',
    'NumberChildrenEverBorn_Pooled',
    'Worry_Nagel2018'
)

other_diseases <- c(
    'Type_2_Diabetes',
    'AtrialFibrillation_Nielsen2018',
    'Coronary_Artery_Disease',
    'SYSTOLICadjMEDz',
    'DIASTOLICadjMEDz',
    'CARDIOVASCULAR',
    'HYPERTENSION_DIAGNOSED',
    'FastingGlucose_Manning',
    'AlanineAminotransferase',
    'AlkalinePhosphatase',
    'Cholesterol',
    'Glucose',
    'HDLcholesterol',
    'HbA1c',
    'LDLdirect',
    'SHBG',
    'Testosterone_Male',
    'TotalBilirubin',
    'TotalProtein',
    'Triglycerides',
    'HEEL_TSCOREz',
    'BALDING1',
    'HEIGHTz',
    'WHRadjBMIz',
    'BREAST',
    'BASAL_METABOLIC_RATEz',
    'FEV1FVCzSMOKE',
    'FVCzSMOKE',
    'HAIR',
    'MENARCHE_AGE',
    'MENOPAUSE_AGE'
)



traits_order <- c(brain_traits, blood_immune_traits, other_diseases)

```

```{r eval=FALSE}

# re-load  file 
combined_assoc_df <- read.csv(file=paste0(data_dir, '5xFAD_visium_scDRS_celltype_associations.csv'))
combined_assoc_df$condition <- 'All'
combined_assoc_df$mc_fdr <- p.adjust(combined_assoc_df$mc_p, method='fdr')

# re-load  file 
assoc_df <- read.csv(file=paste0(data_dir, '5xFAD_visium_diagnosis_scDRS_celltype_associations.csv'))
assoc_df$mc_fdr <- p.adjust(assoc_df$mc_p, method='fdr')


# assoc_df %<>% 
#     group_by(condition) %>% 
#     mutate(mc_fdr = p.adjust(mc_p, method='fdr')) %>% 
#     ungroup


# combine
assoc_df <- rbind(assoc_df, combined_assoc_df)
assoc_df$group <- factor(as.character(assoc_df$group), levels=levels(seurat_5x$annotation))
assoc_df <- subset(assoc_df, group != 'unknown')
assoc_df$group <- droplevels(assoc_df$group)

# rename for convenience
assoc_df$trait <- gsub('UKB_460K.biochemistry_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.blood_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.bmd_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.body_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.bp_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.cancer_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.cov_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.disease_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.impedance_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.lung_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.mental_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.other_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.pigment_', '', assoc_df$trait)
assoc_df$trait <- gsub('UKB_460K.repro_', '', assoc_df$trait)
assoc_df$trait <- gsub('.full', '', assoc_df$trait)

traits <- unique(assoc_df$trait)
traits_order <- traits_order[traits_order %in% traits]

assoc_df <- subset(assoc_df, trait %in% traits_order)

# add trait type:
assoc_df$trait_type <- ifelse(assoc_df$trait %in% brain_traits, 'brain', 'other')
assoc_df$trait_type <- ifelse(assoc_df$trait %in% blood_immune_traits, 'blood/immune', assoc_df$trait)

#table(assoc_df$condition, assoc_df$trait)

#------------------------------------------------------------------------#
# different plots for each condition 
#------------------------------------------------------------------------#

#plot_groups <- unique(assoc_df$condition)
plot_groups <- c('4mo-WT', '4mo-5X', '6mo-WT', '6mo-5X', '8mo-WT', '8mo-5X', '12mo-WT', '12mo-5X', 'All')
plot_groups <- c('4mo-WT', '6mo-WT', '8mo-WT',  '12mo-WT', 'All')
plot_groups <- c('4mo-5X', '6mo-5X', '8mo-5X',  '12mo-5X', 'All')

plot_list <- list()
i <- 1
for(cur_group in plot_groups){

    plot_df <- subset(assoc_df, condition == cur_group)
    plot_df$trait <- factor(as.character(plot_df$trait), levels=rev(traits_order))

    p <- ggplot(plot_df, aes(x=group, y=trait, color=prop_p_0.05, size=-log10(mc_fdr))) + 
        geom_point() + 
        geom_point(
            inherit.aes=FALSE,
            data = subset(plot_df, mc_fdr <= 0.05),
            aes(x=group, y=trait, size=-log10(mc_fdr)),
            color='black', 
            shape=1,
            fill=NA
        ) + 
        coord_equal() + 
        scale_color_steps(
            low='white', high=as.character(cp_5x[cur_group]),
           # breaks=seq(0:10)/10
        ) + 
       # scale_size_continuous(range = c(1, 3 * max(-log10(assoc_df$mc_fdr)))) +
        xlab('') + ylab('') +
        ggtitle(cur_group) +
        theme(
            axis.line.x = element_blank(),
            axis.line.y = element_blank(),
            panel.border = element_rect(linewidth=1, color='black', fill=NA),
            axis.text.x = element_blank(),
            axis.title.x = element_blank(),
            axis.ticks.x = element_blank(),
            plot.title = element_text(hjust=0.5),
            panel.grid = element_line(size=0.25, color='lightgrey'),
            plot.margin = margin(c(0,0,0,0))
        ) 

    if(cur_group != plot_groups[1]){
        p <- p + theme( axis.text.y = element_blank())
    }

    # # make the colorbar as its own heatmap
    color_df$var <- 1
    cp <- color_df$colour; names(cp) <- color_df$group
    colorbar <- color_df %>% 
        subset(group %in% assoc_df$group) %>% 
        mutate(group = droplevels(group)) %>%
        ggplot(aes(x=group, y=var, fill=group)) +
        geom_tile() +
        scale_fill_manual(values=cp) +
        coord_equal() +
        NoLegend() + RotatedAxis() +
        theme(
            plot.title=element_blank(),
            axis.line=element_blank(),
            axis.ticks.y =element_blank(),
            axis.text.y = element_blank(),
            axis.title = element_blank(),
            plot.margin=margin(0,0,0,0),
        )


    patch <- p  / colorbar
    plot_list[[cur_group]] <- patch

}

pdf(paste0(fig_dir, '5xFAD_visium_scdrs_heatmap_ordered_WT.pdf'), width=20, height=12)
wrap_plots(plot_list, ncol=5) + plot_layout(guides='collect')
dev.off()


pdf(paste0(fig_dir, '5xFAD_visium_scdrs_heatmap_ordered_5X.pdf'), width=20, height=12)
wrap_plots(plot_list, ncol=5) + plot_layout(guides='collect')
dev.off()



cur_trait <- 'Alzheimers_Jansen2019'
plot_groups <- c('4mo-WT', '4mo-5X', '6mo-WT', '6mo-5X', '8mo-WT', '8mo-5X', '12mo-WT', '12mo-5X')

# get the info for this trait 
plot_df <- subset(assoc_df, trait == cur_trait & condition != 'All')
#plot_df$condition <- factor(as.character(plot_df$condition), levels=plot_groups)
plot_df$genotype <- ifelse(grepl('5X', plot_df$condition), '5X', 'WT')
plot_df$age <- do.call(rbind, strsplit(plot_df$condition, '-'))[,1]
plot_df$age <- factor(as.character(plot_df$age), levels=c('4mo', '6mo', '8mo', '12mo'))
plot_df$group <- factor(as.character(plot_df$group), levels=rev(names(mouse_cp)))


p <- ggplot(plot_df, aes(x=age, y=group, color=prop_p_0.05, size=-log10(mc_fdr))) + 
    geom_point() + 
    geom_point(
        inherit.aes=FALSE,
        data = subset(plot_df, mc_fdr <= 0.05),
        aes(x=age, y=group, size=-log10(mc_fdr)),
        color='black', 
        shape=1,
        fill=NA
    ) + 
    coord_equal() + 
    scale_color_steps(
        low='white', high='red',
        # breaks=seq(0:10)/10
    ) + 
    scale_size(limits = c(0, 2.5)) +
    xlab('') + ylab('') +
   # ggtitle(cur_group) +
    theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(linewidth=1, color='black', fill=NA),
       # axis.text.x = element_blank(),
        axis.title.x = element_blank(),
       # axis.ticks.x = element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.grid = element_line(size=0.25, color='lightgrey'),
        plot.margin = margin(c(0,0,0,0))
    ) + RotatedAxis()



pdf(paste0(fig_dir, '5x_scDRS_dotplot.pdf'), width=7, height=3.5)
p + facet_wrap(~genotype, ncol=2)
dev.off()


```

Correlation with Amyloid scores 

```{r eval=FALSE}

seurat_amyloid_5x <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/amyloid/data/5xFAD_seurat_processed_amyloid_voyager.rds')

# load the AD/DS results
scdrs_df <- read_csv(file=paste0(data_dir, '5xFAD_visium_scDRS_results.csv'))

# subset for one trait (AD)
cur_df <- subset(scdrs_df, trait == 'Alzheimers_Jansen2019')
cur_df <- subset(cur_df, barcode %in% colnames(seurat_amyloid_5x))

ix <- match(colnames(seurat_amyloid_5x), cur_df$barcode)
all.equal(as.character(cur_df$barcode[ix]), colnames(seurat_amyloid_5x))

seurat_amyloid_5x$scdrs_score <- cur_df[ix,]$norm_score
seurat_amyloid_5x$scdrs_zscore <- cur_df[ix,]$zscore
seurat_amyloid_5x$scdrs_pval <- cur_df[ix,]$pval
seurat_amyloid_5x$scdrs_score_signif <- ifelse(seurat_amyloid_5x$scdrs_pval < 0.05, seurat_amyloid_5x$scdrs_score, 0)



samples <- unique(seurat_amyloid_5x$SAMPLE)

amyloid_col <- 'DAM_UCell'
amyloid_col <- 'Amylo_Area_log_Gi'
amyloid_col <- 'OC_Area_log_Gi'
scdrs_col <- 'scdrs_zscore'
cor_df <- data.frame()

for(cur_sample in samples){

    print(cur_sample)
    cur_meta <- subset(seurat_amyloid_5x@meta.data, SAMPLE == cur_sample)
    cur_cor <- cor.test(
      cur_meta[[amyloid_col]],
      cur_meta[[scdrs_col]]
    )

    cur_df <- data.frame(
        sample = cur_sample,
        cor = as.numeric(cur_cor$estimate),
        pval = as.numeric(cur_cor$p.value)
    )
    cor_df <- rbind(cor_df, cur_df)

}

cor_df <- subset(cor_df, ! is.na(cor))
quantile(cor_df$cor)

subset(seurat_amyloid_5x@meta.data, scdrs_pval <= 0.05) %>% .$scdrs_score %>% min

# plot the scDRS score
p <- SampleFeaturePlot(
    seurat_amyloid_5x,
    feature='scdrs_zscore',
    sample_labels = c("Diagnosis", "Sex", 'Age'),
    ncol = 10,
    raster=TRUE,
    plot_max = 'q99',
    plot_min = 1.5,
    colfunc = inferno,
    rev_colors=TRUE,
    dpi=800,
    combine=FALSE
)

p <- lapply(p, function(x){
    x + theme(
        plot.title = element_text(face='bold', size=10, vjust=-1),
        plot.margin = margin(0,0,0,0)
    ) 
   # scale_fill_gradient2()
})


patch <- wrap_plots(p, ncol=10, widths=1, heights=1) + plot_layout(guides='collect')

pdf(paste0(fig_dir, 'featureplot_scDRS_AD.pdf'), width=12, height=6)
print(patch)
dev.off()


```


Correlation of scDRS score and module eigengenes 

```{r eval=FALSE}

library(hdWGCNA)

# load the 5xFAD data with hdWGCNA info:
seurat_5x <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/hdWGCNA/data/5xFAD_seurat_hdWGCNA.rds')
wgcna_name <- 'vis'

# load the 5xFAD data with human projected hdWGCNA info:
seurat_5x <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/hdWGCNA/data/5xFAD_seurat_hdWGCNA_projected.rds")
wgcna_name <- 'consensus_projected'

seurat_5x <- SetActiveWGCNA(seurat_5x, wgcna_name)

# load the AD/DS results
scdrs_df <- read_csv(file=paste0(data_dir, '5xFAD_visium_scDRS_results.csv'))

# subset for one trait (AD)
cur_df <- subset(scdrs_df, trait == 'Alzheimers_Jansen2019')
cur_df <- subset(cur_df, barcode %in% colnames(seurat_5x))

ix <- match(colnames(seurat_5x), cur_df$barcode)
all.equal(as.character(cur_df$barcode[ix]), colnames(seurat_5x))

seurat_5x$scdrs_score <- cur_df[ix,]$norm_score
seurat_5x$scdrs_zscore <- cur_df[ix,]$zscore
seurat_5x$scdrs_pval <- cur_df[ix,]$pval
seurat_5x$scdrs_score_signif <- ifelse(seurat_5x$scdrs_pval < 0.05, seurat_5x$scdrs_score, 0)


#################################################################################
# MT correlation
################################################################################


# selected traits for MT correlation:
seurat_5x$Sex <- factor(as.character(seurat_5x$Sex), levels=c('M', 'F'))
traits <- c('Age', 'Sex', 'DAA_UCell', 'DAM_UCell', 'DOL_UCell', 'PIGs_UCell', 'scdrs_score')

seurat_5x$annotation <- droplevels(seurat_5x$annotation)


seurat_5x$Condition_Age <- paste0(
    as.character(seurat_5x$Condition), '_',
    as.character(seurat_5x$Age)
)

groups <- unique(seurat_5x$Condition_Age)
plot_df <- data.frame()
for(cur_group in groups){

    print(cur_group)

    seurat_5x <- ModuleTraitCorrelation(
        seurat_5x,
        traits = traits,
        group.by = 'annotation',
        subset_by = 'Condition_Age',
        subset_groups = cur_group
    )

    mt_cor <- GetModuleTraitCorrelation(seurat_5x)
    names(mt_cor)

    cor_df <- do.call(rbind, lapply(names(mt_cor$cor), function(x){
        cur_cor <- mt_cor$cor[[x]]
        cur_fdr <- mt_cor$fdr[[x]]
        cur_cor_df <- reshape2::melt(cur_cor)
        cur_fdr_df <- reshape2::melt(cur_fdr)
        cur_cor_df$fdr <- cur_fdr_df$value
        cur_cor_df$group <- x
        cur_cor_df
    }))
    cor_df %<>% dplyr::rename(c(trait = Var1, module = Var2, cor=value))

    # add age and genotype info
    tmp<- strsplit(cur_group, '_')[[1]]
    cor_df$Condition <- tmp[1]
    cor_df$Age <- tmp[2]

    plot_df <- rbind(plot_df, cor_df)

}


write.csv(plot_df, file=paste0(data_dir, 'scDRS_module_correlation_5x_', wgcna_name, '.csv'), quote=FALSE)
plot_df <- read.csv(file=paste0(data_dir, 'scDRS_module_correlation_5x_', wgcna_name, '.csv'))



# which modules have a high correlation with scDRS score?
# subset(cor_df_5x, trait == 'scdrs_score' & abs(cor) > 0.25)

# subset(cor_df_wt, trait == 'scdrs_score' & abs(cor) > 0.25)
# subset(cor_df_wt, trait == 'scdrs_score' & module == 'SM6') %>% .$cor %>% quantile
# subset(cor_df_5x, trait == 'scdrs_score' & module == 'SM6') %>% .$cor %>% quantile

plot_df$group <- factor(as.character(plot_df$group), levels=c('all_cells', names(mouse_cp)))
plot_df$group <- fct_rev(plot_df$group)

plot_range <- subset(plot_df, trait == 'scdrs_score') %>% .$cor %>% range
plot_max <- max(abs(plot_range))

plot_df$Age <- factor(as.character(plot_df$Age), levels=c('4mo', '6mo', '8mo', '12mo'))

p1 <- plot_df %>% subset(trait == 'scdrs_score') %>%
    ggplot(aes(x = module, y = group, fill=cor)) + 
    geom_tile() + 
    geom_text(
        data = subset(plot_df %>% subset(trait == 'scdrs_score' & abs(cor) > 0.2)),
        aes(x = module, y = group, label=round(cor, 2)), size=1.5
    ) +
    scale_fill_gradient2(high='red', mid='white', low='blue', limits=c(-plot_max, plot_max)) + 
    coord_equal() + RotatedAxis() +
    theme(
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.border = element_rect(linewidth=1, color='black', fill=NA),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust=0.5),
        panel.grid = element_line(size=0.25, color='lightgrey'),
    ) 



pdf(paste0(fig_dir, 'MT_cor_heatmap_', wgcna_name,'_5x.pdf'), width=12, height=6)
p1 + facet_grid(Condition ~ Age)
dev.off()


```

Make a Dot Plot 

```{r eval=FALSE}

# load the 5xFAD data with hdWGCNA info:
seurat_5x <- readRDS('/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/5xFAD/hdWGCNA/data/5xFAD_seurat_hdWGCNA.rds')
wgcna_name <- 'vis'

# load the 5xFAD data with human projected hdWGCNA info:
seurat_5x <- readRDS(file="/dfs7/swaruplab/smorabit/analysis/ADDS_2021/visium/human/hdWGCNA/data/5xFAD_seurat_hdWGCNA_projected.rds")
wgcna_name <- 'consensus_projected'


# what group shad a significant scDRS?
signif_groups <- c('ctx-deep-layers', 'ctx-olfactory', 'hippocampus', 'lateral-ventricle', 'thalamus2', 'WM1', 'WM2', 'WM-cerebral-peduncle')


plot_df <- read.csv(file=paste0(data_dir, 'scDRS_module_correlation_5x_', wgcna_name, '.csv'))

plot_df <- subset(plot_df, group != 'all_cells' & trait == 'scdrs_score')

plot_df$group <- factor(as.character(plot_df$group), levels=c('all_cells', names(mouse_cp)))
plot_df$group <- fct_rev(plot_df$group)

plot_range <- subset(plot_df, trait == 'scdrs_score') %>% .$cor %>% range
plot_max <- max(abs(plot_range))

plot_df$Age <- factor(as.character(plot_df$Age), levels=c('4mo', '6mo', '8mo', '12mo'))


modules <- GetModules(seurat_5x, wgcna_name = wgcna_name)
mod_levels <- levels(modules$module)

plot_df$module <- factor(as.character(plot_df$module), levels=mod_levels)

# grouping variables:
plot_df$id <- paste(plot_df$group, plot_df$Condition, plot_df$Age, sep='_')

# get wgcna info 
data.features <- GetMEs(seurat_5x, wgcna_name=wgcna_name) %>% as.data.frame()
data.features$id <- paste(seurat_5x$annotation, seurat_5x$Condition, seurat_5x$Age, sep='_')

#---------------------------------------------------------------------
# code from Seurat DotPlot
#---------------------------------------------------------------------

scale <- TRUE
col.min <- -2.5
col.max <- 2.5
dot.min <- 0

data.plot <- lapply(
    X = unique(x = data.features$id),
    FUN = function(ident) {
      data.use <- data.features[data.features$id == ident, 1:(ncol(x = data.features) - 1), drop = FALSE]
      avg.exp <- apply(
        X = data.use,
        MARGIN = 2,
        FUN = function(x) {
          return(mean(x = expm1(x = x)))
        }
      )
      pct.exp <- apply(X = data.use, MARGIN = 2, FUN = PercentAbove, threshold = 0)
      return(list(avg.exp = avg.exp, pct.exp = pct.exp))
    }
  )
names(x = data.plot) <- unique(x = data.features$id)
id.levels <- levels(seurat_obj$cell_type)
data.plot <- lapply(
    X = names(x = data.plot),
    FUN = function(x) {
      data.use <- as.data.frame(x = data.plot[[x]])
      data.use$features.plot <- rownames(x = data.use)
      data.use$id <- x
      return(data.use)
    }
  )
  data.plot <- do.call(what = 'rbind', args = data.plot)
  if (!is.null(x = id.levels)) {
    data.plot$id <- factor(x = data.plot$id, levels = id.levels)
  }
  ngroup <- length(x = levels(x = data.plot$id))

  avg.exp.scaled <- sapply(
    X = unique(x = data.plot$features.plot),
    FUN = function(x) {
      data.use <- data.plot[data.plot$features.plot == x, 'avg.exp']
      if (scale) {
        data.use <- scale(x = log1p(data.use))
        data.use <- MinMax(data = data.use, min = col.min, max = col.max)
      } else {
        data.use <- log1p(x = data.use)
      }
      return(data.use)
    }
  )
  avg.exp.scaled <- as.vector(x = t(x = avg.exp.scaled))
  data.plot$avg.exp.scaled <- avg.exp.scaled
  
  data.plot$pct.exp[data.plot$pct.exp < dot.min] <- NA
  data.plot$pct.exp <- data.plot$pct.exp * 100

# match with the other table
data.plot <- subset(data.plot, id %in% plot_df$id)

# merge the DFs
data.plot$ix <- paste0(data.plot$features.plot, '_', data.plot$id)
plot_df$ix <- paste0(plot_df$module, '_', plot_df$id)

ix <- match(plot_df$ix, data.plot$ix, )
all.equal(plot_df$ix, data.plot$ix[ix])

plot_df$avg.exp <- data.plot$avg.exp[ix]
plot_df$avg.exp.scaled <- data.plot$avg.exp.scaled[ix]
plot_df$pct.exp <- data.plot$pct.exp[ix]



plot_range <- plot_df %>% subset(group %in% signif_groups) %>% .$cor %>% range
size_range <- plot_df %>% subset(group %in% signif_groups) %>% .$pct.exp %>% range

# loop through Studies & diagnosis
ages <- c('12mo', '8mo', '6mo', '4mo')
genotypes <- unique(as.character(seurat_5x$Condition))
plot_list <- list()
for(cur_age in ages){
    for(cur_geno in genotypes){
        
    
        cur_df <- plot_df %>% subset(group %in% signif_groups & Condition == cur_geno & Age == cur_age)
        p <-cur_df %>%
            ggplot(aes(x = module, y = group, color=cor, size=pct.exp)) + 
            geom_point() + 
            scale_color_gradient2(high='red', mid='white', low='blue', limits=plot_range) + 
            scale_size_continuous(limits=size_range) +
            coord_equal() +
            RotatedAxis() +
            ylab(paste0(cur_age, ', ', cur_geno)) + 
            theme(
                axis.line.x = element_blank(),
                axis.line.y = element_blank(),
                panel.border = element_rect(linewidth=1, color='black', fill=NA),
                axis.title.x = element_blank(),
                axis.title.y = element_text(angle=0, vjust=0.5),
                plot.title = element_text(hjust=0.5),
                panel.grid = element_line(size=0.25, color='lightgrey'),
                plot.margin = margin(c(0,0,0,0)),
                
            ) 

        if(cur_age != '12mo'){
            p <- p + theme(
                axis.text.x = element_blank(),
                axis.ticks.x = element_blank(),
            )
        }       

        plot_list[[paste0(cur_age, ', ', cur_geno)]] <- p

    }
}


pdf(paste0(fig_dir, 'MT_cor_heatmap_', wgcna_name, '_5x_visium_dotplot.pdf'), width=12, height=8)
wrap_plots(rev(plot_list), ncol=2) + plot_layout(guides='collect')
dev.off()

pdf(paste0(fig_dir, 'MT_cor_heatmap_', wgcna_name, '_5x_visium_dotplot.pdf'), width=6, height=10)
wrap_plots(rev(plot_list), ncol=1) + plot_layout(guides='collect')
dev.off()


```



Plot the scDRS score in representative samples 

```{r eval=FALSE}


# representative samples ordered by age
rep_samples_5x <- c('675', '421', '784', '314')
rep_samples_WT <- c('242', '721', '343', '313')
rep_samples <- c(rep_samples_5x, rep_samples_WT)


# load the AD/DS results
scdrs_df <- read_csv(file=paste0(data_dir, '5xFAD_visium_scDRS_results.csv'))

# subset for one trait (AD)
cur_df <- subset(scdrs_df, trait == 'Alzheimers_Jansen2019')
cur_df <- subset(cur_df, barcode %in% colnames(seurat_5x))

ix <- match(colnames(seurat_5x), cur_df$barcode)
all.equal(as.character(cur_df$barcode[ix]), colnames(seurat_5x))

seurat_5x$scdrs_score <- cur_df[ix,]$norm_score
seurat_5x$scdrs_zscore <- cur_df[ix,]$zscore
seurat_5x$scdrs_pval <- cur_df[ix,]$pval
seurat_5x$scdrs_score_signif <- ifelse(seurat_5x$scdrs_pval < 0.05, seurat_5x$scdrs_score, 0)



library(RColorBrewer)

p <- SampleFeaturePlot(
    seurat_5x,
    feature='scdrs_score',
    sample_labels = c("Condition", 'Age'),
    sample_col = "SAMPLE",
    samples_to_plot = rep_samples,
    ncol = 8,
    raster=TRUE,
    plot_max = 'q99',
    plot_min = 0,
    #colfunc = inferno,
    rev_colors=TRUE,
    colfunc=colorRampPalette(c('red', 'grey90')),
   #rev_colors=FALSE,
    dpi=500,
    combine=FALSE
)

# p <- lapply(p, function(x){
#     x + scale_color_gradient2()
# })

patch <- wrap_plots(p, ncol=4) + plot_layout(guides='collect')

pdf(paste0(fig_dir, '5x_visium_scDRS_AD_featureplot.pdf'), width=8, height=8)
patch
dev.off()






```




NOT RUN BELOW:


Try to get the Moran's I or the Geary's C to work from the Voyager package:

Following this tutorial from Voyager:
https://pachterlab.github.io/voyager/articles/nonspatial.html

```{r eval=FALSE}


library(Voyager)
library(SpatialFeatureExperiment)
library(SpatialExperiment)
library(DropletUtils)
library(BiocNeighbors)
library(scater)
library(scran)
library(bluster)
library(BiocParallel)
library(scuttle)
library(stringr)
library(BiocSingular)
library(spdep)
library(patchwork)
library(dplyr)
library(reticulate)

# seurat_morabito <- readRDS("/dfs7/swaruplab/smorabit/analysis/scWGCNA/data/Swarup_2021.rds")

# just get the Morabito / Miyoshi 2021 dataset 
cur_seurat <- subset(seurat_5x, Study == 'ADDS')

# load the scDRS results
scdrs_df <- read.csv(file=paste0(data_dir,'ADDS_scDRS_results.csv'))

# subset for one trait
cur_df <- subset(scdrs_df, trait == 'Alzheimers_Jansen2019')

# make sure to only use barcodes that match between the Seurat obj and the scDRS 
cur_seurat <- subset(seurat_5x, bc %in% cur_df$barcode)

ix <- match(cur_seurat$bc, cur_df$barcode)
all.equal(as.character(cur_df$barcode)[ix], as.character(cur_seurat$bc))

cur_seurat$scdrs_score <- cur_df[ix,]$norm_score
cur_seurat$scdrs_zscore <- cur_df[ix,]$zscore
cur_seurat$scdrs_pval <- cur_df[ix,]$pval


# harmony <- seurat_morabito@reductions$harmony@cell.embeddings
# harmony <- harmony[cur_seurat$bc,]
# rownames(harmony) <- colnames(cur_seurat)

# cur_seurat@reductions$harmony <- Seurat::CreateDimReducObject(
#   embeddings = harmony,
#   key="harmony",
#   assay="RNA"
# )




p1 <- VlnPlot(
    cur_seurat, 
    features = 'scdrs_score',
    group.by = 'cell_identity', 
    pt.size=0) + NoLegend()

p2 <- VlnPlot(
    cur_seurat, 
    features = 'scdrs_zscore',
    group.by = 'cell_identity', 
    pt.size=0) + NoLegend()


pdf(paste0(fig_dir, 'test_scdrs_vln_ADDS_AD.pdf'), width=10, height=6)
p1 / p2
dev.off()


#cur_seurat <- subset(seurat_5x, Study == 'Morabito_Miyoshi')

# convert it to SingleCellExperiment
sce <- as.SingleCellExperiment(cur_seurat)

#-------------------------------------------------------------#
# Data processing
#-------------------------------------------------------------#

sce <- logNormCounts(sce)

#-------------------------------------------------------------#
# Run KNN on the SCVANI representation
#-------------------------------------------------------------#

# run KNN
foo <- findKNN(reducedDim(sce, "SCANVI")[,1:30], k=10, BNPARAM=AnnoyParam())
#foo <- findKNN(reducedDim(sce, "HARMONY")[,1:30], k=10, BNPARAM=AnnoyParam())

# Split by row
foo_nb <- asplit(foo$index, 1)
dmat <- 1/foo$distance

# Row normalize the weights
dmat <- sweep(dmat, 1, rowSums(dmat), FUN = "/")
glist <- asplit(dmat, 1)

# Sort based on index
ord <- lapply(foo_nb, order)
foo_nb <- lapply(seq_along(foo_nb), function(i) foo_nb[[i]][ord[[i]]])
class(foo_nb) <- "nb"
glist <- lapply(seq_along(glist), function(i) glist[[i]][ord[[i]]])

listw <- list(style = "W",
              neighbours = foo_nb,
              weights = glist)
class(listw) <- "listw"
attr(listw, "region.id") <- colnames(sce)

#-------------------------------------------------------------#
# Convert formats from SCE to SFE
#-------------------------------------------------------------#

# convert to SFE format
(sfe <- toSpatialFeatureExperiment(sce, spatialCoords = reducedDim(sce, "UMAP")[,1:2],
                                  spatialCoordsNames = NULL))

# add the knn to the SFE
colGraph(sfe, "knn") <- listw

#-------------------------------------------------------------#
# Test running moran's I
#-------------------------------------------------------------#

# why nans?
sfe <- runMoransI(sfe,  
    features = c('CSF1R', 'SLC17A7', 'MOBP'))
rowData(sfe)[c('CSF1R', 'SLC17A7', 'MOBP'),]


# gloabl Moran's I
sfe <- colDataMoransI(sfe, c("nCount_RNA", "doublet_scores", 'scdrs_score'))
colFeatureData(sfe)[ c("nCount_RNA", "doublet_scores", 'scdrs_score'),]

sfe <- colDataUnivariate(sfe, "moran.plot",  c("nCount_RNA", "doublet_scores", 'scdrs_score'))

p <- moranPlot(sfe, "scdrs_score", color_by = "cell_identity")

pdf(paste0(fig_dir, 'test_voyager_moranI_scdrs_score.pdf'), width=10, height=6)
p
dev.off()

# Local Moran's I
sfe <- colDataUnivariate(sfe, "localmoran",  c("nCount_RNA", "doublet_scores"))

pdf(paste0(fig_dir, 'test_voyager_local_moranI.pdf'), width=10, height=6)
plotSpatialFeature(sfe,c("nCount_RNA", "doublet_scores"))
dev.off()

pdf(paste0(fig_dir, 'test_voyager_local_moranI.pdf'), width=10, height=6)
plotLocalResult(sfe, "localmoran",  c("nCount_RNA", "doublet_scores"), 
                colGeometryName = "centroids",
                divergent = TRUE, diverge_center = 0, ncol = 2)
dev.off()

#-------------------------------------------------------------#
# Try it with Geary's C (not in the tutorial)
#-------------------------------------------------------------#

# Geary's C
sfe <- colDataUnivariate(sfe, type="geary",  c("nCount_RNA", "doublet_scores"))

# LOSH
sfe <- colDataUnivariate(sfe, type="LOSH",  c("nCount_RNA", "doublet_scores"))


colFeatureData(sfe)[c("nCount_RNA", "doublet_scores"),]


#-------------------------------------------------------------#
# Can I run Moran's I on just one cell cluster?
# 
# Yes you can. But you first have to make a new KNN graph after 
# subsetting.
#-------------------------------------------------------------#

# sfe <- colDataUnivariate(sfe, "moran",  c("nCount_RNA", "doublet_scores"), sample_id = 'cell_type')
# colFeatureData(sfe)[ c("nCount_RNA", "doublet_scores"),]

test <- sfe[,sfe$cell_type == 'MG']

# run KNN
foo <- findKNN(reducedDim(test, "SCANVI")[,1:30], k=25, BNPARAM=AnnoyParam())

# Split by row
foo_nb <- asplit(foo$index, 1)
dmat <- 1/foo$distance

# Row normalize the weights
dmat <- sweep(dmat, 1, rowSums(dmat), FUN = "/")
glist <- asplit(dmat, 1)

# Sort based on index
ord <- lapply(foo_nb, order)
foo_nb <- lapply(seq_along(foo_nb), function(i) foo_nb[[i]][ord[[i]]])
class(foo_nb) <- "nb"
glist <- lapply(seq_along(glist), function(i) glist[[i]][ord[[i]]])

listw <- list(style = "W",
              neighbours = foo_nb,
              weights = glist)
class(listw) <- "listw"
attr(listw, "region.id") <- colnames(test)

colGraph(test, "knn") <- listw


test <- colDataUnivariate(test, "moran",  c("nCount_RNA", "doublet_scores", 'scdrs_score'))
colFeatureData(test)[ c("nCount_RNA", "doublet_scores", 'scdrs_score'),]

test <- colDataUnivariate(test, "geary",  c("nCount_RNA", "doublet_scores", 'scdrs_score'))
colFeatureData(test)[ c("nCount_RNA", "doublet_scores", 'scdrs_score'),]

test <- runMoransI(test,  
    features = c('CSF1R', 'SLC17A7', 'MOBP'))
rowData(test)[c('CSF1R', 'SLC17A7', 'MOBP'),]

# run this for all genes:
test <- runMoransI(test, features = rownames(test), BPPARAM = MulticoreParam(2))

p <- plotRowDataHistogram(test, "moran_sample01", bins = 50) +
    scale_color_continuous(breaks = scales::breaks_width(2))


pdf(paste0(fig_dir, 'test_voyager_moranI_hist.pdf'), width=10, height=6)
p
dev.off()

```